# Results {#25-results}

```{r load-libs}
#| message: false

library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
library(broom)
library(report)
library(tools)
# library(ggtext)
# library(glue) now included with tidyverse as str_glue
# library(gridExtra)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r load-data}

car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results"
)

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics"
)
```

```{r def-cust-theme}
base_color <- "#69b3a2"
outline_color <- "grey25"

# Define a custom theme for ggplot
custom_theme <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      # plot.subtitle = element_text(face = "italic"),
      # plot.caption = element_text(color = "grey50"),
      # axis.title.x = element_text(face = "bold"),
      # axis.title.y = element_text(face = "bold"),
      # axis.text = element_text(color = "black"),
      # axis.line = element_line(color = "grey80"),
      # legend.position = "bottom",
      # legend.background = element_rect(fill = "white", colour = "grey50"),
      # legend.key = element_rect(fill = "white", colour = "grey50"),
    )
}

# Set the custom theme as the default theme
theme_set(custom_theme())
```

```{r study-metadata}

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# Convert the trial_datetime to Date type
demographics <- demographics |>
  mutate(
    trial_date = as.Date(
      as.POSIXct(trial_datetime, format = "%Y-%m-%d %H:%M:%S")
    )
  )

# Get the first and last date
first_trial_date <- demographics |>
  summarize(first_date = min(trial_date)) |>
  pull(first_date)

last_trial_date <- demographics |>
  summarize(last_date = max(trial_date)) |>
  pull(last_date)

trial_start_str <- str_glue(
  "{month(first_trial_date)}/{day(first_trial_date)}"
)
trial_end_str <- str_glue(
  "{month(last_trial_date)}/{day(last_trial_date)}"
)
trial_year_str <- str_glue("{year(last_trial_date)}")

# start, end = Feb 10, Apr 27 2023
```

## Study Administration and Participation

This study was administered during the Spring of 2023. Ultimately, `r participant_count` eligible participants were recruited. All completed the Learning and Recall experiments of the first session beginning February 10th. Progress of the conduct of these trials is illustrated by @fig-s1-progress.

```{r fig-s1-progress}
#| fig-cap: "Daily Trial Administration and Completion Percentage"
#| fig-height: 3.5

# Count experiments per day
daily_counts <- demographics |>
  group_by(trial_date) |>
  summarize(daily_count = n()) |>
  arrange(trial_date)

# Calculate the cumulative count of experiments
daily_counts <- daily_counts |>
  mutate(cumulative_count = cumsum(daily_count))

# Calculate remaining experiments
daily_counts <- daily_counts |>
  mutate(
    percentage_complete = (cumulative_count / participant_count) * 100
  )

pc_eng <- as.character(english::english(participant_count))
pc_eng <- str_replace(pc_eng, "^.", str_to_upper)
subtitle <- str_glue("{pc_eng} Learning and Recall Experiments Conducted between {trial_start_str} and {trial_end_str} of {trial_year_str}")

# Plotting
ggplot(daily_counts, aes(x = trial_date)) +
  geom_line(
    aes(y = daily_count, group = 1),
    color = base_color,
    linewidth = 1,
    alpha = 0.5
  ) +
  geom_line(
    aes(y = percentage_complete, group = 1),
    color = base_color,
    linewidth = 1.25,
    alpha = 1
  ) +
  scale_y_continuous(
    name = "Trials Administered",
    sec.axis = sec_axis(~., name = "Percentage Complete"),
    position = "right",
    limits = c(0, 100)
  ) +
  labs(
    title = "First Session Progress",
    subtitle = subtitle,
    x = element_blank(),
    y = element_blank()
  ) +
  custom_theme()
```

Only eight trials were run before March 16th. That period was treated as a pilot study, during which the research team was trained, administration methods were refined, the self-service onboarding system was deployed, and version 2.1 of our IRB application was approved. Most significantly, the intake process was adjusted to collect more demographic data, increase the role of the TLX and SUS instruments, and gather BCS data for a companion study.

Otherwise, only minor tweaks were made during the pilot to streamline the experiment's administration. Although the changes outlined above were not believed to have a material effect on the results, the data collected during the pilot were omitted from further analysis. Following the pilot, steady progress was made until the completion of the study's first session on April 27th.

The second session was conducted during the end of study event on *XXX*. *YYY* of the study's original participants volunteered to attend and complete the Retention experiment.

::: content-hidden
TODO: complete above paragraph about second session.
:::

::: content-hidden
MAYBE: perform analysis to evaluate differences between pilot and main group

Pilot Study Comparison
- Compare the key measures (e.g., performance, workload, usability) between the pilot study (n=8) and main study (n=54) participants
- Use descriptive statistics and visualizations to present these comparisons
- Discuss any notable similarities or differences between the two groups

Statistical Power and Sample Size Considerations
- Present the results of statistical power analyses based on the observed effect sizes and the planned statistical tests
- Discuss the adequacy of the sample size for detecting meaningful effects
- Consider the potential impact of including or excluding the pilot study data on the statistical power and precision of the estimates
- sample size calculation / justification: power analysis, practical considerations, precedent

Decision on Pilot Study Inclusion
- Based on the findings from 5.2.3 and 5.2.4, make a decision on whether to include the pilot study data in the main analysis
- Clearly report your decision and rationale, considering factors such as data comparability, study design modifications, and statistical power
:::

## Analysis of Participant Demographics

### Univariate Analysis

```{r drop-pilot-data}
# drop participants in pilot study (first eight)
demographics <- demographics |>
  filter(!(participant >= 1001 & participant <= 1010))

# drop event data from the pilot study
car_results <- car_results |>
  filter(!(participant >= 1001 & participant <= 1010))
```

```{r def-to-factors}

# safely convert data to ord / cat factors
# create levels / labels as required and identify / warn about mismatches

to_factors <- function(
    column,
    levels = NULL,
    labels = NULL,
    ordered = FALSE) {
  if (is.null(levels)) {
    levels <- unique(na.omit(column))
  }

  if (is.null(labels)) {
    labels <- levels
  }

  mismatched <- setdiff(unique(na.omit(column)), levels)

  if (length(mismatched) > 0) {
    # Append mismatched values to levels and labels with a specific prefix
    bad_levels <- paste("Bad -", mismatched)
    levels <- c(levels, mismatched)
    labels <- c(labels, bad_levels)
    warning(
      "Mismatched or unexpected values incorporated: ",
      paste(bad_levels, collapse = ", ")
    )
  }

  # Proceed to convert to factor using cleaned-up data
  factor_column <- factor(column,
    levels = levels, labels = labels,
    ordered = ordered
  )

  return(factor_column)
}
```

```{r factor-demographics}

# convert all categorical / ordinal data to factors

demo_f <- demographics |>
  mutate(
    # categorical, levels != labels
    # unordered with predefined levels
    gender = to_factors(
      gender,
      levels = c("F", "M", "O"),
      labels = c("Female", "Male", "Non-Binary")
    ),
    race = to_factors(
      race,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "American Indian or Alaska Native",
        "Asian",
        "Black or African American",
        "Native Hawaiian or Other Pacific Islander",
        "White",
        "More than one race",
        "Unknown or not reported"
      )
    ),
    ethnicity = to_factors(
      ethnicity,
      levels = c("1", "2", "3"),
      labels = c(
        "Hispanic or Latino",
        "Not Hispanic or Latino",
        "Unknown or not reported"
      )
    ),
    lenses = to_factors(
      lenses,
      levels = c("G", "C", "N", "Y"),
      labels = c(
        "Glasses",
        "Contacts",
        "No",
        "Yes"
      )
    ),

    # categorical, levels == labels
    # unordered with predefined levels
    language = to_factors(
      language,
      levels = c("English", "Other"),
    ),
    worn = to_factors(
      worn,
      levels = c("Y", "N")
    ),
    color_blind = to_factors(
      color_blind,
      levels = c("Y", "N")
    ),
    condition = to_factors(
      condition,
      levels = c("Y", "N")
    ),

    # ad-hoc
    # categorized in transcription, no predefined levels
    country = to_factors(country),
    college = to_factors(college),
    program = to_factors(program),
    referral = to_factors(referral),

    # ordinal, levels != labels
    # ordered with predefined levels
    education = to_factors(
      education,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "Pre-HS", # No HS diploma
        "HS", # High school diploma or equivalent
        "CND", # Some college but no degree
        "AS",
        "BS",
        "MS",
        "PhD"
      ),
      ordered = TRUE
    ),
    lego = to_factors(
      lego,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "Little to no experience",
        "Some experience",
        "Lots of experience",
        "Expert builder"
      ),
      ordered = TRUE
    ),
    manuf = to_factors(
      manuf,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "No experience",
        "Taken one or more classes",
        "Held a part time or temporary position",
        "One or more year’s experience"
      ),
      ordered = TRUE
    ),
    runner = to_factors(
      runner,
      levels = c("N", "Y", "5", "10", "H", "M", "T"),
      labels = c(
        "None",
        "Yes - Unspecified",
        "5K",
        "10K",
        "Half-Marathon",
        "Marathon",
        "Triathalon"
      ),
      ordered = TRUE
    ),
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    )
  )
```

#### Numeric Data

```{r numeric-data}
nums <- demo_f |>
  select(where(is.numeric))

height_stats <- nums |>
  summarize(
    mean = mean(height, na.rm = TRUE),
    sd = sd(height, na.rm = TRUE),
    min = min(height, na.rm = TRUE),
    max = max(height, na.rm = TRUE),
    pct_missing = sum(is.na(height)) / n() * 100
  )

height_summary <- str_glue(
  "Mean height = {round(height_stats$mean, 1)}, SD = {round(height_stats$sd, 1)}, range: [{height_stats$min}, {height_stats$max}], {round(height_stats$pct_missing, 1)}% missing"
)
```

After excluding data from the pilot study, `r report_participants(nums, age = "age", participants = "participant")` were included in this analysis. Other than age, height (`r height_summary`) was the only numeric data collected. As discussed in @sec-res-challenges, height was first recorded following the pilot study, when its impact on user behavior and results was first observed. Both numeric characteristics are visualized with a combined box and scatter plot in @fig-numeric. Where present, the notches of these box plots show the 95% confidence interval around the median.

```{r fig-numeric}
#| fig-cap: "Distribution of Participant Age and Height"
#| fig-height: 3.5
#| warning: false

# Plotting age
age_plot <- demo_f |>
  filter(!is.na(age)) |>
  ggplot(aes(x = age)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    x = "Age (yr)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting height
height_plot <- demo_f |>
  filter(!is.na(height)) |>
  ggplot(aes(x = height)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    width = 0.55,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    x = "Height (in)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
age_plot + height_plot +
  # plot_layout(ncol = 2, widths = c(1,0.8)) +
  plot_annotation(
    title = "Numeric Sample Characteristics",
    subtitle = "Age reflects the student-oriented nature of the sample"
  )
```

#### Categorical Data

```{r categorical-data}
cats <- demo_f |>
  select(where(~ is.factor(.) && !is.ordered(.))) |>
  select(-c("referral")) |>
  mutate(across(everything(), \(x) fct_lump(x, prop = 0.1)))

# Summary table
cat_tbl <- cats |>
  select(gender:language) |>
  tbl_summary(
    label = setNames(str_to_title(names(cats)), names(cats)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

cat_names <- cats |>
  select(gender:country) |>
  names()

cats_long <- cats |>
  select(gender:country) |>
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "Value"
  ) |>
  drop_na() |>
  group_by(Category, Value) |>
  summarize(Count = n(), .groups = "drop") |>
  group_by(Category) |>
  mutate(Freq = Count / sum(Count))

country_levels <- demo_f |>
  filter(country != "USA") |>
  count(country) |>
  arrange(desc(n))
```

Of the sample's `r inline_text(cat_tbl, language, level = "English", pattern = "{N} participants, {p}% (n={n})")` reported that they mainly speak English at home. As shown in @fig-categorical, the gender composition was relatively balanced, with slightly more males `r inline_text(cat_tbl, gender, level = "Male", pattern = "(n={n}, {p}%)")` than females `r inline_text(cat_tbl, gender, level = "Female", pattern = "(n={n}, {p}%)")`. The racial makeup of the sample was predominantly White `r inline_text(cat_tbl, race, level = "White", pattern = "(n={n}, {p}%)")` and Asian `r inline_text(cat_tbl, race, level = "Asian", pattern = "(n={n}, {p}%)")`, with smaller representations from `r as.character(english::english(n_distinct(demo_f$race) - 2))` other groups. Fully `r inline_text(cat_tbl, ethnicity, level="Not Hispanic or Latino", pattern = "{p}% of participants (n={n})")` reported a non-Hispanic or Latino ethnicity. The majority of participants `r inline_text(cat_tbl, country, level = "USA", pattern = "(n={n}, {p}%)")` were from the United States, though `r as.character(english::english(n_distinct(demo_f$country, na.rm = TRUE) - 1))` other countries were represented, including `r paste(country_levels[1, ]$country, " (n=", country_levels[1, ]$n, ")", sep = "")` and `r paste(country_levels[2, ]$country, " (n=", country_levels[2, ]$n, ")", sep = "")`. `r paste(country_levels$country[3:(nrow(country_levels) - 1)], collapse = ", ")`, and `r tail(country_levels$country, 1)` were also listed as countries of origin by one participant each (n=1). All of this aligns with the study's recruitment focus.

```{r fig-categorical}
#| fig-cap: "Distribution of Categorical Characteristics"
#| fig-height: 3.5
#| results: hide

create_cat_plot <- function(data, category, base_color) {
  # specify alpha ramp end points
  min_alpha <- 0.3
  max_alpha <- 0.8

  # Filter and reorder data
  plot_data <- data |>
    filter(Category == category) |>
    mutate(
      Value = if_else(Value == "Not Hispanic or Latino",
        "Not Hispanic\nor Latino", Value
      ),
      Value = reorder(Value, Count, FUN = sum, decreasing = TRUE)
    )

  # Calculate alpha values based on Count linearly between min_alpha and max_alpha
  alpha_values <- plot_data |>
    mutate(alpha = scales::rescale(Count, to = c(min_alpha, max_alpha))) |>
    pull(alpha)

  # Set unique Value groups for setting colors and alphas
  unique_values <- unique(plot_data$Value)

  # Create the plot
  ggplot(
    plot_data,
    aes(x = Value, y = Count, fill = I(base_color), alpha = Value)
  ) +
    geom_bar(stat = "identity", position = "dodge", width = 0.8, color = outline_color) +
    scale_x_discrete(labels = function(x) tools::toTitleCase(x)) +
    labs(x = tools::toTitleCase(category), y = element_blank()) +
    scale_alpha_manual(values = setNames(alpha_values, unique_values)) +
    guides(fill = "none", alpha = "none") +
    ylim(0, max_count) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
}

# Calculate the maximum count to standardize the y-axis
max_count <- max(cats_long$Count)

# Create all plots
cat_plots <- map(
  cat_names,
  ~ create_cat_plot(cats_long, .x, base_color)
)

# Add y-axis text and ticks to first plot
first_plot <- cat_plots[[1]] +
  theme(
    axis.text.y = element_text(),
    axis.ticks.y = element_line()
  )

# Other plots without y-axis text and ticks, drop first
other_plots <- cat_plots[-1]

# combine!
all_plots <- c(list(first_plot), other_plots)

# Combine all plots
combined_plot <- reduce(all_plots, `+`) +
  plot_layout(ncol = length(all_plots), widths = c(1, 1, 0.667, 0.667)) +
  plot_annotation(
    title = "Categorical Sample Characteristics",
    subtitle = "Value Count for Gender, Race, Ethnicity, and Country"
  )

# Print the combined plot
print(combined_plot)
```

```{r make-vc-table}
vc_tbl <- cats |>
  select(lenses:worn) |>
  filter(!lenses %in% c("No", "Other", "Unknown")) |>
  filter(!(is.na(lenses) & is.na(worn))) |>
  droplevels() |>
  mutate(worn = recode(worn, "Y" = "Yes", "N" = "No")) |>
  tbl_cross(
    row = lenses,
    col = worn,
    label = list(lenses = "Lenses", worn = "Worn During Session"),
    percent = "row",
    missing_text = "NA"
  ) |>
  add_stat_label() |>
  bold_labels()
```

Of the 54 participants, 52% (n=28) reported having a vision prescription. Of those, `r inline_text(vc_tbl, row_level = "Contacts", col_level = "Total")` were for contact lenses , but `r inline_text(vc_tbl, row_level = "Glasses", col_level = "Total")` had glasses. Whereas all contact wearers reported they would wear them during the session, among glasses wearers only seven (47%) planned the same. Five (33%) did not intend to do so, and three (20%) gave no indication. This breakdown is summarized by @tbl-vc. Two participants also reported color-blindness (n=1) or other vision conditions (n=1). No participants indicated any other condition that might affect their performance.

::: content-hidden
MAYBE: make this all inline code; rethink percent by row, col, or cell? format n= etc. as markdown?
:::

```{r tbl-vc}
#| tbl-cap: Vision Correction Table

vc_tbl |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

Following the primary hypothesis testing, additional analyses may explore the differences in key outcomes across these demographic subgroups. That effort will aim to provide additional context and insights into potential underlying factors influencing the findings.

#### Ordinal Data

```{r ord-data}
ords <- demo_f |>
  select(where(~ is.factor(.) && is.ordered(.)))

# Summary table
ords_tbl <- ords |>
  tbl_summary(
    label = setNames(str_to_title(names(ords)), names(ords)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

# ords_tbl

# Helper function to calculate mode
calculate_mode <- function(data, var_name) {
  count_table <- table(data[[var_name]])
  max_freq <- max(count_table)
  # pick first if more than one value has max count
  mode_value <- names(count_table[count_table == max_freq])[1]

  # Convert to an output that includes the level's name (mode_value) and its index
  mode_index <- which(levels(data[[var_name]]) == mode_value)
  return(list(value = mode_value, index = mode_index))
}

# Main function to calculate statistics and return formatted string
calculate_ord_summary_stats <- function(data, var_name) {
  # Ensure the variable is an ordered factor
  if (!is.ordered(data[[var_name]])) {
    stop("The variable must be an ordered factor.")
  }

  # Convert factor to integer codes
  factor_int <- as.integer(data[[var_name]])

  # Calculate median, mode, and IQR
  median_value <- median(factor_int, na.rm = TRUE)
  mode_results <- calculate_mode(data, var_name)
  iqr_value <- IQR(factor_int, na.rm = TRUE)

  # Format and return the result as a string
  result_string <- sprintf(
    "(Median = %.1f; Mode = %d, %s; IQR = %.1f)",
    median_value, mode_results$index, mode_results$value,
    iqr_value
  )
  return(result_string)
}
```

Unlike the data above, the categories analyzed in this section are inherently ordered. Responses for each participant's current Education level `r calculate_ord_summary_stats(ords, "education")` and their experience with Lego `r calculate_ord_summary_stats(ords, "lego")` and Manufacturing `r calculate_ord_summary_stats(ords, "manuf")` are visualized in increasing order by @fig-ordinal. The results are mostly in line with expectations based on the recruitment, though it is disappointing to see that `r inline_text(ords_tbl, lego, level = "Little to no experience", pattern = "{p}% of participants (n={n})")` claimed to have little to no experience with Lego. The great majority were undergraduate students working towards a degree `r inline_text(ords_tbl, education, level = "CND", pattern = "(n={n}, {p}%)")`, with no experience in manufacturing `r inline_text(ords_tbl, manuf, level = "No experience", pattern = "(n={n}, {p}%)")`.

```{r fig-ordinal}
#| fig-cap: "Distribution of Ordinal Characteristics"
#| fig-height: 3.5
#| results: hide

plot_ords <- function(data, var, title) {
  norm_data <- data |>
    count(var_name = {{ var }}) |>
    mutate(percent = n / sum(n) * 100)

  p <- norm_data |>
    ggplot(aes(y = var_name, x = percent)) +
    geom_bar(
      aes(alpha = percent / 100),
      fill = base_color,
      stat = "identity",
      width = 0.75,
      color = outline_color
    ) +
    scale_alpha_continuous(trans = "log10", range = c(0.2, 0.8), guide = "none") +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank()
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(size = 9, face = "bold"),
      legend.position = "none",
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
  return(p)
}


educ_plot <- ords |>
  mutate(education = fct_recode(education,
    College = "CND",
    Associate = "AS",
    Bachelor = "BS",
    Master = "MS",
    Doctorate = "PhD"
  )) |>
  plot_ords(var = education, title = "Current Education Level")

lego_plot <- ords |>
  mutate(lego = fct_recode(lego,
    Little = "Little to no experience",
    Some = "Some experience",
    Lots = "Lots of experience",
    Expert = "Expert builder"
  )) |>
  plot_ords(lego, title = "Lego Experience")

manuf_plot <- ords |>
  mutate(manuf = fct_recode(manuf,
    None = "No experience",
    Training = "Taken one or more classes",
    PartTime = "Held a part time or temporary position",
    Some = "One or more year’s experience"
  )) |>
  plot_ords(manuf, title = "Manufacturing Experience")

# combine using patchwork
educ_plot / (lego_plot + manuf_plot) +
  # plot_layout(ncol = 2) +
  plot_annotation(
    title = "Ordinal Sample Characteristics",
    subtitle = "Reflects the Focus of Recruiting Efforts"
  )

```

### Multivariate Analysis

In order to assess the equivalence of treatment groups and validate the random assignment that is assumed by most analysis to follow, Age, Gender, Lego experience, and Education were compared across treatment groups. These variables were selected for their meaningful variation and potential relevance to the outcomes of interest. Other notable variables, including Ethnicity and Country of Origin were considered less critical due to their skewed distributions and small subgroup sizes.

```{r make-mv-table}

num_cols <- demographics |>
  select(age)

cat_cols <- demo_f |>
  select(c("gender", "lego", "education", "treatment"))

mv_data <- bind_cols(num_cols, cat_cols)

mv_tbl <- mv_data |>
  mutate(across(where(is.factor), ~ fct_drop(.))) |>
  mutate(
    lego = fct_recode(lego,
      "Little" = "Little to no experience",
      "Some" = "Some experience",
      "Lots" = "Lots of experience",
      "Expert" = "Expert builder"
    ),
    education = fct_recode(education,
      "College" = "CND",
      "Associate" = "AS",
      "Bachelor" = "BS",
      "Master" = "MS",
      "Doctorate" = "PhD"
    )
  ) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    label = list(
      "age" ~ "Age",
      "gender" ~ "Gender",
      "lego" ~ "Lego",
      "education" ~ "Education"
    )
  ) |>
  add_p() |>
  # separate_p_footnotes() |>
  # add_stat_label() |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  italicize_levels() |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nN = {n} ({style_percent(p)}%)",
    p.value ~ "**p-val**"
  )

```

@tbl-demo-by-treatment shows the result of this analysis. The Kruskal-Wallis rank sum test was utilized for the numeric variable (Age) to compare distributions across groups, while Fisher's exact test was applied to the categorical and ordinal variables (Gender, Education, and Lego experience) to assess the independence of distributions from treatment assignments.

```{r tbl-demo-by-treatment}
#| tbl-cap: Participant Demographics by Treatment Group

mv_tbl |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )

```

::: content-hidden
MAYBE: wrap gtsummary calls in a function that saves an svg to project
and inserts embedding so that all original formatting is preserved?
could they still be labeled as tables?
:::

The high p-values for Age, `r inline_text(mv_tbl, variable = age, column = "p.value")`; Gender, `r inline_text(mv_tbl, variable = gender, column = "p.value")`; Lego, `r inline_text(mv_tbl, variable = lego, column = "p.value")`; and Education, `r inline_text(mv_tbl, variable = lego, column = "p.value")` indicate that these observed characteristics are not strongly associated with treatment groupings. This result validates the group randomization effort, mitigating the influence of confounding variables, and reducing bias in the data. It is essential to the validity of most of the statistical tests employed by this study, which assume sample independence.

::: content-hidden
MAYBE: further analysis to show that:
- nums
  - height (0.11)
- cats
  - country (0.085)
  - lenses (0.063)
- ords
  - manuf exp (0.046)

may be significant (p-val)
used the following code to test all demographics

``` text
# test significance of treatment groups
trt <- demo_f |>
  select(treatment)

# bind treatment for nums, ords (included in cats)
cd <- bind_cols(ords, trt)

cd |>
  tbl_summary(
      by = "treatment",
      missing = "no",
      statistic = list(
        all_continuous() ~ "{mean} ({sd})"
      )
    ) |>
    add_p()
```
:::

## Learning Phase Results

Only completed tasks, the primary result of interest, are considered in this analysis. While excluding incomplete tasks may limit the generalizability of the findings to scenarios where time constraints are common, it ensures the analysis is based on objective, directly observed data and avoids introducing assumptions or biases that could arise from predicting results for unfinished tasks. 

```{r factor-car-results}

res_f <- car_results |>
  mutate(
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    ),
    car_outcome = to_factors(car_outcome)
  )
```

```{r learn-data}
learn <- res_f |>
  filter(
    phase == 1,
    car_outcome == "complete",
    # hack to remove incorrectly labeled car
    uid != "1048-P1-C5"
  ) |>
  select(
    participant,
    car_elapsed_time,
    uce,
    treatment,
    drop_event_count,
    drop_event_tot_time_adj,
    repair_event_count,
    repair_event_tot_time,
    system_event_count,
    system_event_tot_time
  )

# add tot_sys_adj, tot_usr_adj, and adjusted TCT time to learn
# tot_sys_adj will be used to calculate OEE
# tot_usr_adj just to reduce TCT

learn <- learn |>
  mutate(
    # Replace NA with 0 using coalesce for safe addition
    tot_sys_adj = coalesce(drop_event_tot_time_adj, 0) +
      coalesce(system_event_tot_time, 0),
    tot_usr_adj = coalesce(repair_event_tot_time, 0),
    tot_adj = tot_sys_adj + tot_usr_adj,
    car_elapsed_time_adj = car_elapsed_time - tot_adj
  )

changes <- learn |>
  filter(car_elapsed_time != car_elapsed_time_adj)

total_tasks_completed <- dim(learn)[1]
total_tasks_adjusted <- dim(changes)[1]
```

One final step of data preparation was taken to account for drop-outs and other system events, along with time spent making repairs. This lost time was deducted from measured task times to give a more accurate record of participant performance. `r total_tasks_adjusted` of `r total_tasks_completed` completed tasks (`r round(total_tasks_adjusted / total_tasks_completed * 100, 1)`%) were affected by this adjustment.

::: content-hidden
MAYBE: consider better quantifying the effect of these adjustments
:::

### Descriptive Statistics for Granular Data

```{r make-report}
# report descriptive stats for cet and uce
learn_rpt <- learn |>
  rename(
    "Task Completion Time (sec)" = car_elapsed_time_adj,
    "Uncorrected Error Count" = uce
  ) |>
  select("Task Completion Time (sec)", "Uncorrected Error Count") |>
  report::report() # options not working!

rpt_table <- as.report_table(learn_rpt)
cv_tct <- rpt_table$SD[1] / rpt_table$Mean[1]
cv_uce <- rpt_table$SD[2] / rpt_table$Mean[2]
```

::: content-hidden
TODO: correct data for `1048-P1-C5`; flagged complete with elapsed time of 4.267s
TODO: investigate report::report options
:::

`r learn_rpt`

These statistics provide a high-level summary of performance across all tasks and participants, where each observation corresponds to a single repetition of the Learning task. Both Task Completion Time (TCT) and Uncorrected Error Count (UCE) show considerable variance, as depicted in @fig-learn.

```{r fig-learn}
#| fig-cap: "Distribution of Task Completion Time and Error Rates in Learning"
#| fig-height: 3.5
#| warning: false

# Plotting adjusted car_elapsed_time
cet_plot <- learn |>
  ggplot(aes(x = car_elapsed_time_adj)) +
  geom_histogram(
    aes(y = ..density.. * 65), # hack for relative size / scale
    binwidth = 10,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    title = "Task Completion Time (sec)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting Uncorrected Error Counts
uce_plot <- learn |>
  ggplot(aes(x = uce)) +
  geom_histogram(
    aes(y = ..density.. * 1.5), # hack for relative size / scale
    binwidth = 1,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    title = "Number of Uncorrected Errors per Task",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
cet_plot / uce_plot +
  # plot_layout(ncol = 2, widths = c(1,1)) +
  plot_annotation(
    title = "Learning Phase - Observed Measures",
    subtitle = "Error Count Exhibits Greater Variability than Task Completion Time"
  )
```

Compared to TCT, with a coefficient of variation (CV) of `r round(cv_tct, 2)`, UCE shows substantially greater spread (CV=`r round(cv_uce, 2)`) and the presence of a floor effect. The later could potentially suggest limitations in the task's ability to capture a full range of participant performance. Further investigation is required to determine if this is due to task complexity or other factors.

::: content-hidden
MAYBE: discuss overdispersion of data, where sd^2 / mean > 1; this is true for both TCT and UCE - implications? better metrick than CV? more appropriate later in report, after agg / outliers?
:::

### Aggregated Data

```{r agg-results}
# agg participant results
learn_agg <- learn |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    tot_cars = n(), # Count the number of tasks per participant
    avg_tct_adj = mean(car_elapsed_time_adj, na.rm = TRUE),
    avg_uce = mean(uce, na.rm = TRUE)
  ) |>
  ungroup()
```

::: content-hidden
MAYBE: may need to add back other variables from learn into this agg table
:::

@tbl-agg-results-by-treat presents participant-level metrics, including the number of tasks completed along with the average TCT and UCE per task. In addition to the overall results, the data is broken out by assigned treatment group to allow for comparisons across conditions. Note that the statistics presented are the overall (aka "grand") mean and median values. That is, they are the average or middle values of all individual participant averages. This provides overall summary statistics for each treatment group that reflect the central tendency and variability among the average outcomes reported by participants within each group.

```{r tbl-agg-results-by-treat}
#| tbl-cap: Avg. Participant Performance Data Summarized by Treatment Group

t <- learn_agg |>
  select(-participant) |>
  tbl_summary(
    type = all_continuous() ~ "continuous2",
    digits = all_continuous2() ~ 1,
    label = list(
      avg_tct_adj ~ "Avg TCT per Task (sec)",
      avg_uce ~ "Avg UCE per Task",
      tot_cars ~ "Number of Tasks Completed"
    ),
    by = treatment,
    statistic = list(
      all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} ({p25}, {p75})"
      )
    )
  ) |>
  # bold_labels() |>
  # add_p() |>
  # separate_p_footnotes() |>
  add_overall() |>
  # bold_p(t = 0.05) |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nn = {n} ({style_percent(p)}%)",
    # p.value ~ "**p-val**"
  )

t |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

See @fig-perf-by-treat for a depiction of the grand means for all variables, with 95% confidence intervals for each. Observe that the previously noted variability and positive skew are also present in the grouped data, and some treatments like PWI may have prioritized speed over accuracy, or vice versa (e.g. AR).

```{r make-perf-data}
# Reshape data to long format
long_data <- learn_agg |>
  pivot_longer(
    cols = c(avg_tct_adj, tot_cars, avg_uce),
    names_to = "metric",
    values_to = "value"
  )

# Calculate means, sample size, and SE for each metric within each treatment
metrics_stats <- long_data |>
  group_by(treatment, metric) |>
  summarize(
    mean_value = mean(value, na.rm = TRUE),
    n = n(), # sample size
    se = sd(value, na.rm = TRUE) / sqrt(n()), # standard error
    .groups = "drop"
  )

# Find the highest mean for each metric across all treatments
max_means <- metrics_stats |>
  group_by(metric) |>
  summarize(max_mean = max(mean_value), .groups = "drop")

# Normalize means and calculate CI within their metrics
normalized_metrics_stats <- metrics_stats |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = mean_value / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  )

# Calculate overall mean, SE, and normalize
overall_stats <- long_data |>
  group_by(metric) |>
  summarize(
    overall_mean = mean(value, na.rm = TRUE),
    n = n(),
    se = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = overall_mean / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  ) |>
  mutate(treatment = "ALL")

# Combine with other stats
final_stats <- bind_rows(normalized_metrics_stats, overall_stats)

# Factorize the 'treatment' column
final_stats$treatment <- factor(final_stats$treatment, levels = c("ALL", "PWI", "PAR", "AR", "MR"))

# Factorize and label 'metric' column
final_stats$metric <- factor(
  final_stats$metric,
  levels = c("avg_tct_adj", "avg_uce", "tot_cars"),
  labels = c("Average Task Time", "Average Error Count", "Tasks Completed")
)

# Predefined alpha values by treatment
alpha_values <- c(0.4, 0.55, 0.7, 0.85, 1.0)
names(alpha_values) <- levels(final_stats$treatment)
```


```{r fig-perf-by-treat}
#| fig-cap: "Aggregated Learning Performance Data by Treatment"
#| fig-height: 3.5
#| warning: false

ggplot(final_stats, aes(x = metric, y = normalized_mean, fill = treatment, label = treatment)) +
  geom_col(
    aes(alpha = normalized_mean),
    position = position_dodge(0.9),
    color = outline_color
  ) +
  geom_errorbar(
    aes(ymin = lower_ci, ymax = upper_ci),
    position = position_dodge(0.9),
    width = 0.25,
    alpha = 0.7
  ) +
  geom_text(
    aes(y = 0),
    position = position_dodge(width = 0.9),
    vjust = 1.65, # Adjust based on visual preference
    size = 3,
    alpha = 0.9
  ) +
  scale_fill_manual(values = rep(base_color, length(unique(final_stats$treatment)))) +
  scale_alpha_continuous(transform = "pseudo_log", range = c(0.3, 1.0)) +
  labs(
    title = "Comparison of Normalized Performance Metrics by Treatment",
    subtitle = "Includes Overall and Treatment Groups with 95% Confidence Intervals",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(
      margin = margin(b = 5, unit = "pt"),
      color = "black"
    )
  )

```

#### Assess outliers

Given the level of variation and spread seen above, the presence of outliers was analyzed. Participant-level outliers within treatment groups were the focus, rather than individual data points across the entire dataset. This method helps ensure that outliers are identified relative to the specific context of each treatment, accounting for the corresponding variability.

Four different tests were used to assess each participant as an outlier in their treatment group: (1) Tukey's Fences at 1.5 times the IQR, (2) Tukey's Fences at 3.0 times the IQR, (3) outside the 5th and 95th percentiles, and (4) using the Mahalanobis distance. TCE and UCE were each tested separately using the first three methods. The fourth method is a multivariate test that considers TCE and UCE together. The results are summarized in @tbl-outlier-summary, for all participants that were identified by two or more tests.

> A helpful heuristic is to distinguish between error outliers, interesting outliers, and random outliers (Leys et al. 2019). Error outliers are likely due to human error and should be corrected before data analysis. Interesting outliers are not due to technical error and may be of theoretical interest; it might thus be relevant to investigate them further even though they should be removed from the current analysis of interest. Random outliers are assumed to be due to chance alone and to belong to the correct distribution and, therefore, should be retained.

```{r find-outliers}
# Function to detect outliers in a vector
find_outliers <- function(x, mult) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower_bound <- q1 - mult * iqr
  upper_bound <- q3 + mult * iqr
  x < lower_bound | x > upper_bound
}

# Function to calculate Mahalanobis distance for a group
compute_mahalanobis <- function(df, cols) {
  cov_matrix <- cov(df[, cols], use = "complete.obs")
  center <- colMeans(df[, cols], na.rm = TRUE)
  mahal_dist <- mahalanobis(df[, cols], center, cov_matrix)
  return(mahal_dist)
}

outlier_summary <- learn_agg |>
  group_by(treatment) |>
  mutate(
    # find 1.5 and 3x IQR outliers using Tukey method
    out15_tct = find_outliers(avg_tct_adj, 1.5),
    out15_uce = find_outliers(avg_uce, 1.5),
    out30_tct = find_outliers(avg_tct_adj, 3.0),
    out30_uce = find_outliers(avg_uce, 3.0),
    # find by 5th and 95th percentile
    pct_tct = avg_tct_adj < quantile(avg_tct_adj, 0.05, na.rm = TRUE) |
      avg_tct_adj > quantile(avg_tct_adj, 0.95, na.rm = TRUE),
    pct_uce = avg_uce < quantile(avg_uce, 0.05, na.rm = TRUE) |
      avg_uce > quantile(avg_uce, 0.95, na.rm = TRUE),
    # find by Mahalanobis distance
    mahal_dist = compute_mahalanobis(
      pick(avg_tct_adj, avg_uce),
      c("avg_tct_adj", "avg_uce")
    ),
    md_out = mahal_dist > qchisq(0.975, df = 2)
  ) |>
  ungroup() |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    out15_tct = sum(out15_tct),
    out15_uce = sum(out15_uce),
    out30_tct = sum(out30_tct),
    out30_uce = sum(out30_uce),
    pct_tct = sum(pct_tct),
    pct_uce = sum(pct_uce),
    md_out = sum(md_out),
    .groups = "drop"
  )
```


```{r tbl-outlier-summary}
#| tbl-cap: Summary of Participants Identified by Two or More Outlier Tests
#| warning: false

outlier_summary |>
  rowwise() |>
  mutate(tot_out = sum(c_across(-c(participant, treatment)))) |>
  ungroup() |>
  filter(tot_out > 1) |>
  arrange(desc(tot_out)) |>
  select(participant, treatment, tot_out) |>
  inner_join(learn_agg) |>
  gt()

```

::: content-hidden
TODO: make this table prettier
:::

Further investigation of these seven participants showed that only 1063 experienced systematic problems that might warrant excluding their performance. The other participants were just relatively poor performers, with high TCT and/or UCE. Given the limited amount of data collected per treatment, it was decided to retain all but 1063's data for hypothesis testing.

```{r drop-learn-outliers}
# drop 1063
learn_agg_out <- learn_agg |> filter(participant != 1063)
```

```{r save-learn-data}
# save learn objects with types, etc. as binary data file for external analysis
save(
  learn, 
  learn_agg_out, 
  outline_color,
  base_color,
  file = "learn.RData")
```

### Hypothesis Testing

This first group of hypotheses are designed to test the effect of each
treatment on training outcomes, including average TCT and UCE, along with the learning rate.

#### Average Time per Car

Here we will test:
$$H_{1a}\textrm{: Average time per car varies with treatment.}$$

Specifically, statistical comparison of the average time taken to complete the assembly task under various treatments will identify which instructional approach leads to the quickest task completion. As noted in the original assessment of the learning data, the overall TCT dataset exhibits skewness and kurtosis, suggesting deviations from a normal distribution. To choose the most appropriate statistical method for hypothesis testing, this must be revisited. Standard between-groups hypothesis tests rely on the assumption that the data *within each group* follows a normal distribution. Furthermore, the data has changed. By averaging the observations and eliminating an outlier, the grouped data likely better resembles a normal distribution.

To accurately assess normality of each treatment group, a combination of quantitative and qualitative analysis is required. Statistical tests provide a formal but imperfect measure of the data's shape. Visual confirmation of the characteristic quantile-quantile (Q-Q) plot provides further support for the claim. @tbl-learn-norm-tct tabulates the results of the D'Agostino skewness test, Anscombe-Glynn kurtosis test, and the Shapiro-Wilk normality test. In each case, the null hypothesis ($H_0$) for these tests assumes that the data follows the characteristics of a normal distribution. Therefore, low p-values indicate evidence against $H_0$, suggesting that the data deviates from normality in the tested aspect.

```{r def-tab-norm}
# Function to tabulate normality test results for each treatment group
tabulate_normality_tests <-
  function(data, treatment_col, value_col) {
    levels <- levels(data[[treatment_col]])
    all_levels <- c("ALL", levels)

    results <- map(all_levels, function(level) {
      if (level == "ALL") {
        data_subset <- data |> pull(!!sym(value_col))
      } else {
        data_subset <-
          data |>
          filter((!!sym(treatment_col)) == level) |>
          pull(!!sym(value_col))
      }

      if (length(data_subset) > 0) {
        # Ensure there's data to process
        skew_test <- moments::agostino.test(data_subset)
        kurtosis_test <- moments::anscombe.test(data_subset)
        normality_test <- shapiro.test(data_subset)

        tibble(
          Group = level,
          Skewness = paste(
            round(skew_test$statistic[1], 2),
            round(skew_test$p.value, 3),
            sep = ", "
          ),
          Kurtosis = paste(
            round(kurtosis_test$statistic[1], 2),
            round(kurtosis_test$p.value, 3),
            sep = ", "
          ),
          Normality = paste(
            round(normality_test$statistic, 2),
            round(normality_test$p.value, 3),
            sep = ", "
          )
        )
      } else {
        tibble(
          Group = level,
          Skewness = NA,
          Kurtosis = NA,
          Normality = NA
        )
      }
    }) |> bind_rows()

    # Ensure the desired order, converting Group to a factor
    results <- results |>
      mutate(Group = factor(Group, levels = c("ALL", "MR", "AR", "PWI", "PAR"))) |>
      filter(!is.na(Group)) |>
      arrange(Group)

    results
  }
```


```{r tbl-learn-norm-tct}
#| tbl-cap: "Summary of TCT Test Statistics, p-Values for Normality Assessment"
#|
tct_norm <- tabulate_normality_tests(learn_agg_out, "treatment", "avg_tct_adj")
tct_norm |> gt()
```

From this we see that only PWI fails the normality test (p=0.021), despite showing marginally non-significant skewness (p=0.058). Similarly, MR exhibits significant positive skew (p=0.021) but the Shapiro-Wilk test gives evidence of normality (p=0.084). Both AR and PAR are consistent in confirming normality, though PAR's slightly positive skew is marginally non-significant (p=0.068).

::: content-hidden
TODO: correct hard-coded values in text
:::

To reconcile these findings for MR and PWI, and confirm those for AR and PAR, the set of four Q-Q plots in @fig-qq-h1a were generated. Observations for normal data should fall along or near the line, within the shaded confidence interval. These plots show that PAR and AR both meet the expectations of normality. The MR group mostly aligns with the line, except for a single point in the right tail. This aligns with the skewness and kurtosis test results and provides insufficient evidence to reject normality. On the other hand, the PWI group diverges substantially from the line, with several points leaving the confidence interval. This supports the statistical evidence that the PWI data is not normally distributed.

```{r def-qq-plotter}

# Function to create QQ plots and arrange them in a 2x2 grid
qq_plot_patchwork <-
  function(data,
           x_var,
           by_var) {
    levels <- unique(data[[by_var]])

    if (length(levels) != 4) {
      stop("This script expects exactly 4 levels in the grouping variable.")
    }

    plots <- lapply(levels, function(level) {
      p <- data |>
        filter((!!sym(by_var)) == level) |>
        ggpubr::ggqqplot(
          x = x_var,
          xlab = "",
          ylab = "",
          title = level,
          ggtheme = theme_minimal(base_size = 10) +
            theme(
              plot.margin = margin(0, 0, 0, 0),
              panel.spacing = unit(0, "lines"),
              panel.grid = element_blank(),
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              axis.line = element_blank(),
              plot.title = element_text(hjust = 0.5),
              panel.border = element_rect(
                color = outline_color,
                fill = NA,
                linewidth = 0.5
              ),
            )
        ) |>
        ggpubr::ggpar(ticks = FALSE, tickslab = FALSE)
    })

    # Arrange the plots in a 2x2 grid using patchwork
    plot_layout <-
      (plots[[1]] + plots[[2]]) / (plots[[3]] + plots[[4]])
    
    return(plot_layout)
  }
```

```{r fig-qq-h1a}
#| fig-cap: Q-Q Plots of Learning Times by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(learn_agg_out, x_var = "avg_tct_adj", by_var = "treatment")

p + plot_annotation(title = "Quantile-Quantile Plots of TCT Treatment Data",
                    subtitle = "PWI Data is the Most Uncharacteristic of Normality")

```

Given the minor deviations from normality, particularly for the PWI group, and *considering the robust consistency of results between parametric and non-parametric tests*, it was concluded that non-parametric methods provide the most reliable and straightforward approach for hypothesis testing. This decision *avoids the complexities associated with transformation pipelines*, ensuring robust statistical conclusions while maintaining the interpretability of the original data. Should future analysis require linear regression or related methods (e.g. ANOVA and its variants), the data will be transformed accordingly,

::: content-hidden
MAYBE: try to fix data transformation issues with h1a? see: _temp.qmd
TODO: this has to be updated in light of subsequent modeling efforts
:::

@fig-h1a-plot shows the results of a non-parametric comparison of average TCT within each treatment group. It combines elements of violin, scatter, and box plots to provide a comprehensive portrayal of the data. Treatment differences were analyzed using the Kruskal-Wallis test by ranks, a non-parametric equivalent of ANOVA. Its test statistic, $H$, approximately follows a $\chi^{2}_{k-1}$ distribution, where $k$ is the number of groups. Here, $\chi^{2}_{Kruskal-Wallis}(3)=25.9$, with a p-value of 1.0e-5. These results indicate a statistically significant difference between groups. The effect size of this difference is $\hat{\epsilon}^{2}_{rank}=0.5$, with $CI_{95\%}\:[0.4,1.0]$, indicating a moderate to large treatment effect where 50% of the variance in TCT can be explained by the difference in groups. Together, these results provide strong evidence of overall differences in TCT that are both statistically and practically significant.

::: content-hidden
TODO: fix hard-coded numbers
:::

```{r fig-h1a-plot}
#| fig-cap: Task Completion Time Differences by Treatment
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

h1a_plot <- ggstatsplot::ggbetweenstats(
  data = learn_agg_out,
  x = treatment,
  y = avg_tct_adj,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "TCT Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  results.subtitle = FALSE,
  subtitle = "Significant Pairwise Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 1,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

h1a_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "25.9", italic(p) == 
#     "1.0e-05", widehat(epsilon)["ordinal"]^2 == "0.5", CI["95%"] ~ 
#     "[" * "0.4", "1.0" * "]", italic("n")["obs"] == "53")
```

::: content-hidden
TODO: check all in text values after setting standard seed before all ggbetween stats calls - see comment in code above
:::

Pairwise comparisons were conducted post-hoc using Dunn's non-parametric test for Kruskal-type ranked data. Holm's adjustment for multiple comparisons was preferred over the oft-cited Bonferroni method, which tends to be too conservative. The results show that PWI and PAR are both significantly faster than AR and MR, with all p-values less than 0.01. Statistically, PWI and PAR are both equivalent and faster than AR and MR, which are also equivalent to one another.

```{r test-power-sim}
library(pwr)

# # Function to calculate power for Kruskal-Wallis test via simulation
simulate_kruskal_power <- function(n, k, effect_size, sig_level = 0.05, num_simulations = 1000) {
  p_values <- numeric(num_simulations)
  for (i in 1:num_simulations) {
    # Simulate data for each group with an effect size
    data <- data.frame(
      value = unlist(lapply(1:k, function(x) rnorm(n, mean = x * effect_size))),
      group = gl(k, n)
    )
    # Perform Kruskal-Wallis tes
    p_values[i] <- kruskal.test(value ~ group, data = data)$p.value
  }
  # Proportion of tests rejecting null hypothesis gives power estimate
  power <- mean(p_values < sig_level)
  return(power)
}

# # Function to run post-hoc power analysis based on observed epsilon squared
posthoc_power_analysis <- function(n, k, observed_effect_size, sig_level = 0.05, num_simulations = 1000) {
  # Convert the epsilon squared back to a standard effect size approximation for simulation
  std_effect_size <- sqrt(observed_effect_size / (1 - observed_effect_size))

  # Use simulation to estimate power as before
  power <- simulate_kruskal_power(n, k, std_effect_size, sig_level, num_simulations)
  return(power)
}

k <- length(unique(learn_agg_out$treatment))
n <- round(nrow(learn_agg_out) / k)
epsilon_sq <- 0.5

# Estimate post-hoc power
posthoc_power <- posthoc_power_analysis(n, k, epsilon_sq, sig_level = 0.05, num_simulations = 1000)
```

A post-hoc simulation was run to validate these results by estimating the statistical power. Power analysis measures the effectiveness of a statistical test in detecting true differences when they exist. High power (typically 0.8 or above) suggests that the test is appropriate for the given data and experimental conditions and therefore aligns with the goal of making correct decisions based on the data. The simulation repeatedly applies the Kruskal-Wallis test to data generated based on the observed mean and effect size. The resulting power is the proportion of tests that correctly reject the null hypothesis. The outcome of `r posthoc_power` (`r posthoc_power * 100`% power), while uncommon, suggests that, for the conditions of this experiment (n=53, an average of 13.25 samples in each of the four groups, $\hat{\epsilon}^{2}=0.5$, and $\alpha = 0.05$), the test is extremely effective at detecting the observed differences. The combination of a large effect size and a reasonable sample size is largely attributable to this value, which offers great confidence in the validity of the test results.

*Unfortunately, all of this is moot. The Kruskal-Wallis test has three assumptions that must be met: (1) ordinal or continuous response variable, (2) independent observations in each group, and (3) similar distributions in each group. Task completion time is indeed continuous and previous analysis showed that the random sampling method ensured equivalent treatment distributions. However, the experimental design violates the independent observations assumption due to the repeated measures taken from each participant. As a result, other methods must be employed.*

Mixed effects models are best-suited for repeated measures or nested data structures. Repeated measures ANOVA was also considered but ultimately rejected as it also assumes independence of observations within groups. A simple form of mixed-effects model for this problem could be:

$$
\text{TCT}_{i} = \beta_{0} + \beta_{1} \text{TRT}_{i} + (1 \mid \text{PART}_{i}) + \epsilon_{i}
$$

Where:

- $i$: Index for the observation.
- $\beta_0$: Overall intercept, a baseline TCT for the reference treatment group.
- $\beta_1$: Fixed effect of the treatment (TRT).
- $(1 \mid \text{participant}_i)$: Random effect for each observation, from the participant.
- $\epsilon_{ij}$: Residual error term for each observation.

This formulation allows the model to capture individual variability by allowing for each participant to have a different initial TCT (intercept). This accounts for variation in TCT due to differences between participants (random effects) and within each participant's repeated measures (residuals). Using this approach controls for expected differences among participants. Furthermore, it naturally accounts for the unbalanced nature of the data, where the number of tasks completed differs for each participant. Finally, it also accounts for the inherent non-independence of repeated measures for each participant.

A linear mixed model was fit to predict TCT, using the formulation described above. The `lmer` function from the `{lme4}` package was used to perform the fit using residual maximum likelihood method (REML). Prior to the fit, TCT was log-transformed to improve model performance and diagnostics.

*The resulting model was evaluated using...*

```{r}
library(emmeans)
library(performance)
library(lmerTest)

# create data
learn_h1a <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj)

# mutate
learn_h1a <- learn_h1a |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

glimpse(learn_h1a)
```

Compare model with different transformations of the response to deal with its non-linearity (exponential decay)

```{r}
# model comparison
m1 <- lmer(log(tct) ~ treatment + (1 | participant), data = learn_h1a)
m2 <- lmer(sqrt(tct) ~ treatment + (1 | participant), data = learn_h1a)
m3 <- lmer(inverse(tct) ~ treatment + (1 | participant), data = learn_h1a)

compare_performance(m1, m2, m3, rank = T)
```

Best model uses a simple inverse transformation

```{r}
# plots comparing tct with inverse(tct)

# Create a data frame with the original and inverse transformed TCT values
pd <- learn_h1a %>%
  mutate(inverse_tct = 1 / tct)

# Plot for tct
plot_tct <- ggplot(pd, aes(x = seq, y = tct, color = treatment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "TCT vs. Sequence",
    x = "Sequence",
    y = "Task Completion Time (TCT)"
  ) +
  theme_minimal()

# Plot for inverse(tct)
plot_inverse_tct <- ggplot(pd, aes(x = seq, y = inverse_tct, color = treatment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Inverse(TCT) vs. Sequence",
    x = "Sequence",
    y = "Inverse Task Completion Time (1/TCT)"
  ) +
  theme_minimal()

# Display the plots
p <- plot_tct + plot_inverse_tct

print(p)
```

Check performance:

```{r}
plot(performance::check_model(m3, panel = F))
```

Model summary:

```{r}
summary(m3)
```

Estimated marginal means for pairwise comparison:

```{r}
emmeans_tct <- emmeans::emmeans(
  m3, pairwise ~ treatment, type = "response"
)

summary(emmeans_tct)

```

::: content-hidden
*this text needs to be updated:*

The total explanatory power of the resulting model is substantial (conditional $R^2$ = 0.89, marginal $R^2$ = 0.47). These results indicate that the model explains 89% of the variance in TCT, of which 47% is related to the fixed effects alone. This highlights the importance of individual differences among participants in the data, validating the mixed-effects approach. The model's intercept, corresponding to the AR treatment, is at 0.99 (95% CI [0.62, 1.36], p < .001). Within this model:

  - The effect of treatment MR is statistically non-significant and positive (beta = 0.03, 95% CI [-0.48, 0.54], t(264) = 0.13, p = 0.899; Std. beta = 0.03, 95% CI [-0.48, 0.54])
  - The effect of treatment PWI is statistically significant and negative (beta = -1.54, 95% CI [-2.05, -1.04], t(264) = -6.00, p < .001; Std. beta = -1.54, 95% CI [-2.05, -1.04])
  - The effect of treatment PAR is statistically significant and negative (beta = -1.05, 95% CI [-1.58, -0.53], t(264) = -3.95, p < .001; Std. beta = -1.05, 95% CI [-1.58, -0.53])

:::

These results are visualized in @fig-h1a-fixedeff, where overlapping arrows indicate statistically insignificant differences in TCT between treatments:

```{r fig-h1a-fixedeff}
#| fig-caption: This is the caption
#| fig-height: 3.5

plot(emmeans_tct$emmeans, comparisons = TRUE) +
  labs(
    title = "Estimated Means by Treatment",
    subtitle = "Talk about observed differences",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(
      margin = margin(b = 5, unit = "pt"),
      color = "black"
    )
  )

```

The `{emmeans}` package was used for post-hoc pairwise comparison, generating the following contrasts based on estimated marginal means:

```text
 contrast   estimate      SE   df t.ratio p.value
 AR - MR   -5.44e-05 0.00115 53.5  -0.047  1.0000
 AR - PWI  -6.80e-03 0.00112 48.5  -6.086  <.0001
 AR - PAR  -4.04e-03 0.00116 48.9  -3.470  0.0059
 MR - PWI  -6.75e-03 0.00110 49.2  -6.128  <.0001
 MR - PAR  -3.98e-03 0.00115 49.6  -3.470  0.0058
 PWI - PAR  2.76e-03 0.00112 44.9   2.469  0.0788

Note: contrasts are still on the inverse scale 
Degrees-of-freedom method: kenward-roger 
P value adjustment: tukey method for comparing a family of 4 estimates
```

Significant differences are identified between PWI and both AR and MR, and between PAR and AR and MR. No significant difference was found between PWI and PAR or AR and MR. This matches the previous findings, providing robust statistical evidence that PWI and PAR are both equivalent and faster than AR and MR, which are also equivalent to one another.


```{r eval=FALSE}
# Extract estimated marginal means and confidence intervals
emmeans_df <- as.data.frame(emmeans_tct)

# Extract pairwise comparisons
pairwise_df <- as.data.frame(pairs(emmeans_tct))

# Create a new column in emmeans_df with the back-transformed emmeans and confidence intervals
emmeans_df <- emmeans_df %>%
  mutate(
    tct = inverse_log_shifted(emmean, shift_amount),
    lower.CL_orig = inverse_log_shifted(lower.CL, shift_amount),
    upper.CL_orig = inverse_log_shifted(upper.CL, shift_amount)
  )

# Violin plot of the raw data
p <-
  ggplot(learn_h1a, aes(x = treatment, y = tct, fill = treatment)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_jitter(width = 0.1,
              size = 2,
              alpha = 0.5) +
  stat_summary(
    fun = median,
    geom = "point",
    size = 3,
    color = "red"
  ) +
  theme_minimal() +
  labs(
    title = "TCT Differences by Treatment",
    subtitle = "Violin Plot with Estimated Marginal Means and Pairwise Comparisons",
    x = "Treatment",
    y = "Task Completion Time (TCT)"
  )

# Add estimated marginal means with confidence intervals
p <- p + geom_point(
  data = emmeans_df,
  aes(x = treatment, y = tct),
  size = 3,
  color = "blue"
) +
  geom_errorbar(
    data = emmeans_df,
    aes(x = treatment, ymin = lower.CL_orig, ymax = upper.CL_orig),
    width = 0.2,
    color = "blue"
  )


# Add pairwise comparisons
p <-
  p + geom_text(
    data = pairwise_df,
    aes(
      x = contrast,
      y = max(learn_h1a$tct) + 20,
      label = paste("p =", format(
        p.value, digits = 2, scientific = TRUE
      ))
    ),
    vjust = -0.5,
    color = "black"
  )

# Rotate x-axis labels for better readability
p <- p + theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the plot
print(p)

```

#### Average Error Count per Car

Similar methodology will be applied to testing the second hypothesis associated with the Learning phase. Here again, we are interested in finding statistically significant differences in the average error count under different treatment conditions to identify which instructional method leads to the lowest defect rate. Formally stated:

$$H_{1b}\textrm{: Average error count per car varies with treatment.}$$

```{r learn-norm-uce}
##| tbl-cap: "Summary of UCE Test Statistics, p-Values for Normality Assessment"

uce_norm <- tabulate_normality_tests(learn_agg_out, "treatment", "avg_uce")
# uce_norm |> gt()
```

As seen before, the overall average error count data exhibits floor effects that contribute to a strongly positive skew. The Shapiro-Wilk normality test gives strong evidence that the PAR, AR, and MR groups are non-normal (p<=0.001 in all cases). This is likely due to the observed skewness, which the D'Agostino test indicates is significant in each case (PAR p=0.012, AR p=0.03, MR p=0.017). Only the PWI treatment group passes the Shapiro-Wilk test (p=0.607) with normal-like skewness (p=0.725). The Anscombe-Glynn kurtosis test showed no significant devation from normality for any treatment. These findings are confirmed by @fig-qq-h1b, which shows strong deviations from the normal line for all but the PWI treatment.

::: content-hidden
TODO: correct hard-coded values in text
:::

```{r fig-qq-h1b}
#| fig-cap: Q-Q Plots of Learning Errors by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(learn_agg_out, x_var = "avg_uce", by_var = "treatment")

p + plot_annotation(title = "Quantile-Quantile Plots of UCE Treatment Data",
                    subtitle = "observation")

```


Given the non-normal nature of these distributions, a non-parametric hypothesis test is called for. Furthermore, it is important to note that this is counted data. Despite the fact that means / grand means may appear continuous in this aggregated form, the underlying distribution remains discrete. Traditional parametric tests are not recommended in this case. Kruskal-Wallis and Dunn tests were again employed for this analysis, the results of which are summarized in @fig-h1b-plot.

```{r fig-h1b-plot}
#| fig-cap: Uncorrected Error Count Differences by Treatment
#| fig-height: 5
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

h1b_plot <- ggstatsplot::ggbetweenstats(
  data = learn_agg_out,
  x = treatment,
  y = avg_uce,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "UCE Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  results.subtitle = FALSE,
  subtitle = "Significant Pairwise Differences Identified between PWI and Augmented Treatments",
  digits = 1,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

h1b_res <- ggstatsplot::extract_stats(h1b_plot)
h1b_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "25.5", italic(p) == 
#     "1.2e-05", widehat(epsilon)["ordinal"]^2 == "0.5", CI["95%"] ~ 
#     "[" * "0.3", "1.0" * "]", italic("n")["obs"] == "53")
```

The Kruskal-Wallis test statistic $\chi^{2}_{Kruskal-Wallis}(3)=25.5$, with a p-value of 1.2e-5, indicating a statistically significant difference between groups. Again, the effect size is found to be $\hat{\epsilon}^{2}_{rank}=0.5$, with a slightly wider $CI_{95\%}\:[0.3,1.0]$, reflecting the data's variability. The results of Dunn's Holm-adjusted pairwise comparisons show statistically significant differences between PWI and all other treatments (p<<0.01 in all cases). PAR, AR, and MR treatments are statistically similar. These results formally quantify the obvious reduction in error count experienced with any form of augmented instruction. The power of this test is again 100%, given that the experimental conditions have not changed and the effect size is the same.

::: content-hidden
TODO: fix hard-coded numbers
MAYBE: consider investigating zero-inflated model, e.g. ZIP / ZINB?
MAYBE: reevaluate using mixed effect models to account for participant variation?
:::

#### Learning Rates

$$H_{1c}\textrm{: Learning rates vary with treatment.}$$
Initially tried to fit an exponential decay model to estimate learning rate, $c$. The equation for this model is:

$$y(t) = a + b \times e^{-ct}$$

I was unable to get a solution to converge, so began exploring alternatives, including:

- mixed-effects model
- growth curve analysis
- survival analysis (time to reach standard)

Having no experience with any of these approaches, I decided to start with the most straightforward of the bunch - MEM.

A simple form of mixed-effects model for the data could be:

$$
\text{TCT}_{ij} = \beta_{0} + \beta_{1} \text{TRT}_{i} + \beta_{2} \text{SEQ}_{j} + (1 \mid \text{PART}_{i}) + \epsilon_{ij}
$$

Where:

- $i$ and $j$: Indices for the participant and task sequence
- $\beta_0$: Overall intercept.
- $\beta_1$: Fixed effect of treatment.
- $\beta_2$: Fixed effect of the sequence number.
- $(1 \mid \text{participant}_i)$: Random intercepts for each participant, capturing individual variability. It essentially represents the part of the error term that is specific to each participant.
- $\epsilon_{ij}$: Residual error term for each observation.

Including Interaction Terms

To see if the effect of the sequence number (learning curve) depends on the treatment, an interaction term is included:

$$
\text{TCT}_{ij} = \beta_{0} + \beta_{1} \text{TRT}_{i} + \beta_{2} \text{SEQ}_{j} + \beta_{3} (\text{TRT}_{i} \times \text{SEQ}_{j}) + (1 + \text{SEQ}_{j} \mid \text{PART}_{i}) + \epsilon_{ij}
$$

A nice thing about mixed-effect models is that they inherently account for the unbalanced nature of repeated measures data like ours, where participants complete a different number of tasks. Because participants are treated as random effects this contribution to variability is accounted for.

The dataset for this hypothesis has the form:

```{r make-h1c-data}
# learning rate modeling requires disaggregated data
learn_h1c <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj) |>
  group_by(participant) |>
  mutate(seq = row_number()) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

glimpse(learn_h1c)
```

Each row represents the outcome of a single participant repetition of a task, where:

- participant is a UID
- treatment is instructional method: AR (augmented reality), MR (mixed reality), PAR (projected AR), PWI (paper work instructions)
- seq is the task number
- tct is the task completion time (in seconds)
- uce is the number of uncorrected errors

Here is a plot of smoothed TCT data by treatment:

```{r}
#| warning: false

# Plot smoothed TCT data by treatment
ggplot(learn_h1c,
       aes(
         x = seq,
         y = tct,
         color = treatment,
         group = participant
       )) +
  # geom_line(aes(group = participant), alpha = 0.3) +
  geom_smooth(aes(group = treatment),
              method = "loess",
              se = FALSE,
              size = 1) +
  labs(title = "Learning Curves by Treatment",
       x = "Task Number",
       y = "Task Completion Time (TCT)") +
  theme_minimal()

```

AR, MR appear to have higher slope, but neither reach takt time of 60 seconds.

Here is the output of a model fit based on non-standardized data:

```{r h1c-lmer-model-1}
#| warning: false

#library(lme4)
library(broom.mixed)
library(lmerTest)

# use unordered factor to improve interpretability of results
learn_h1c$treatment <- factor(as.character(learn_h1c$treatment))

# Fit the mixed-effects model with interaction effects
# use the overloaded version of lmer from lmerTest to get CI, p-vals
h1c_mod_lmer <- lmerTest::lmer(
    tct ~ treatment * seq + (1 + seq | participant),
    data = learn_h1c,
    control = lmerControl(optimizer = "bobyqa")
  )

# Summary of the model to understand fixed and random effects
h1c_mod_sum <- summary(h1c_mod_lmer)

h1c_mod_sum

# resid(h1c_mod_lmer)
# coef(h1c_mod_lmer)
# ranef(h1c_mod_lmer)

# Clean output for easier interpretation if needed
# h1c_mod_tsum <- tidy(h1c_mod_lmer, effects = "fixed", conf.int = TRUE)

# verbose narrative report
# report::report(h1c_mod_lmer)

# other display methods and diagnostics
# tbl_regression(h1c_mod_lmer)
# modelsummary::modelsummary(h1c_mod_lmer)
# marginaleffects::avg_comparisons(h1c_mod_lmer)
```

This uses a mixed effects model:

`tct ~ treatment * seq + (1 + seq | participant)`

where Task completion time is a function of:

- fixed effects (predictors):
  - treatment (PWI, PAR, AR, MR)
  - seq (task number)
  - treatment * seq (interaction)
- random effects (account for participant variability):
  - (1 + seq | participant)

The treatment * seq interaction indicates how the relationship between seq and tct changes for different levels of treatment.

The random effects indicate that, in terms of TCT, each participant will have a unique starting point (intercept) and learning rate (slope).

In the results we see that the total variance contributed by random effects is 1977.0 + 79.3 + 228.5 = `r 1977.0 + 79.3 + 228.5`, with participant idiosyncrasies accounting for `r round(1977 / (1977.0 + 79.3 + 228.5) * 100, 1)`% of that. 

These results are based on data that has not been standardized. Results of standard post-hoc model evaluation showed issues with homoscedasticity and multicolinearity, etc. They are kept here for comparison with the output using standardized data.

Here is an alternate model, fit on standardized data:

```{r h1c-lmer-model-2}

# record original means and SDs
mean_tct <- mean(learn_h1c$tct, na.rm = TRUE)
sd_tct <- sd(learn_h1c$tct, na.rm = TRUE)

# standardize data
learn_h1c_std <- learn_h1c |>
  mutate(
    tct = scale(tct) |> as.vector(),
    uce = scale(uce) |> as.vector()
  )

h1c_mod_lmer_std <- lmerTest::lmer(
    tct ~ treatment * seq + (1 + seq | participant),
    data = learn_h1c_std,
#    control = lmerControl(optimizer = "bobyqa")
  )

summary(h1c_mod_lmer_std)
```

These results use a dataset where TCT was standardized in the following manner:

$$z_x = \frac{x - \mu_x}{\sigma_x}$$

Several tests are run to confirm the model quality:

```{r}
## model tests
# normality of residuals 
mr <- resid(h1c_mod_lmer_std)
qqnorm(mr)
qqline(mr)
shapiro.test(mr)
# homoscedaasticity - constant variance of residuals
plot(h1c_mod_lmer_std)
# multicolinearity among predictors
car::vif(h1c_mod_lmer_std)
# R2 goodness of fit
performance::r2_nakagawa(h1c_mod_lmer_std)
```

The test results confirm model validity. All but MR-related coefficients are significant:

```text
                 Estimate Std. Error       df t value Pr(>|t|)    
(Intercept)       2.08412    0.31182 48.34414   6.684 2.18e-08 ***
treatmentMR       0.01964    0.43352 48.42402   0.045 0.964044    
treatmentPAR     -1.66019    0.44385 45.55992  -3.740 0.000512 ***
treatmentPWI     -1.72445    0.42565 44.90851  -4.051 0.000199 ***
seq              -0.51620    0.07129 46.19347  -7.241 3.88e-09 ***
treatmentMR:seq  -0.02100    0.10001 46.70136  -0.210 0.834635    
treatmentPAR:seq  0.32280    0.09651 35.36790   3.345 0.001960 ** 
treatmentPWI:seq  0.27296    0.09225 34.26448   2.959 0.005565 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

All values are expressed in terms of standard deviations of TCT, and can be restored to the original units for easier interpretation. Only the intercept gets the mean TCT value added back, as the others are offsets from that.

$$x = z_x \times \sigma_x + \mu_x$$

Here are the results of reversing the standardization process:

```{r}
# Extract coefficients
coefficients_std <- fixef(h1c_mod_lmer_std)

# Unstandardize coefficients - return to original units for interpretation
coef_unstd <- coefficients_std * sd_tct
coef_unstd["(Intercept)"] <- coef_unstd["(Intercept)"] + mean_tct

print(coef_unstd)
```

These roughly match the values from the model using the unaltered data:

```text
Fixed effects:
                 Estimate Std. Error       df t value Pr(>|t|)    
(Intercept)      185.5695    13.2651  48.3432  13.989  < 2e-16 ***
treatmentMR        0.8357    18.4420  48.4231   0.045 0.964045    
treatmentPAR     -70.6252    18.8816  45.5591  -3.740 0.000512 ***
treatmentPWI     -73.3588    18.1075  44.9077  -4.051 0.000199 ***
seq              -21.9595     3.0327  46.1927  -7.241 3.88e-09 ***
treatmentMR:seq   -0.8931     4.2545  46.7006  -0.210 0.834637    
treatmentPAR:seq  13.7321     4.1056  35.3674   3.345 0.001960 ** 
treatmentPWI:seq  11.6117     3.9243  34.2640   2.959 0.005565 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

From this, we can gather that (all in seconds, rounding to 1 digit):

- starting values of TCT
  - the significant average for AR (aka Intercept) is 185.6
  - MR was insignificantly higher
  - PAR was significantly lower, 185.6 - 70.6 = 115.0
  - PWI was significantly lower, 185.6 - 73.4 = 112.2
- change with each task completed
  - AR's sigificant estimate of -22.0 indicates that, on average, each task was completed 22s faster
  - MR was insignificantly lower
  - PAR was significantly higher, -22.0 + 13.7 = -8.3
  - PWI was significantly higher, -22.0 + 11.6 = -10.4

Interpretation:

- PAR and PWI groups started faster than AR/MR, exhibiting significantly lower TCT
- The TCT decreased significantly faster per task with AR/MR than for PWI or PAR
- MR was not significantly different than AR in either case, suggesting they are equivalent


Continued evaluation by comparing with simpler models:

```{r}
# compare alternative models
# baseline, as formulated above
m1 <- lmerTest::lmer(tct ~ treatment * seq + (1 + seq | participant),
    data = learn_h1c,
    control = lmerControl(optimizer = "bobyqa")
  )

# random intercept only
m2 <- lmerTest::lmer(tct ~ treatment * seq + (1 | participant),
    data = learn_h1c,
    control = lmerControl(optimizer = "bobyqa")
  )

# fixed intercept, random slope
m3 <- lmerTest::lmer(tct ~ treatment * seq + (seq | participant),
    data = learn_h1c,
    control = lmerControl(optimizer = "bobyqa")
  )

AIC(m1, m2, m3)
```

The results show that m1 and m3 are equal, and better than m2. Sticking with m1. This really only tells us that the starting TCT and rates varied by treatment, but we know that there is a relationship between those three. Is the difference coming from the treatment or that relationship?

Goals: try other approaches to model the non-linear nature of learning, account for different initial efficiencies, variable repetition counts, and the dual criteria of speed and accuracy.

First effort to improve understanding: start simple, add UCE as a covariate in the current formulation:

`tct ~ treatment * seq + uce + (1 + seq | participant)`

It does make sense that errors could extend the TCT...

```{r}
# first just try adding uce as a covariate
h1c_mod_lmer_std_uce <- lmerTest::lmer(
  tct ~ treatment * seq + uce + (1 + seq | participant),
  data = learn_h1c_std,
  control = lmerControl(optimizer = "bobyqa")
)

# summary(h1c_mod_lmer_std_uce)

# uce is not significant
broom::tidy(h1c_mod_lmer_std_uce) |>
  filter(term == "uce") |>
  select(p.value)

```

Nope. UCE is not significant. Uncorrected errors go unnoticed and do not impact time?

Also, it is a counted (discrete) value, so modeling it in this way is, in hindsight, not advisable. We'll come back to error analysis later. For now, focus on TCT.

So far we've used linear models but learning curves are naturally exponential in shape. Also, there is a natural tendency of learning rate to depend on the difference between starting point and where the curve naturally flattens out. Here AR and MR start out slower, which may be the primary reason that they improve faster than PWI / PAR.

Second effort: tried using nlme with SSasymp and SSasympOrig to deal with these issues. These are non-linear methods that account for asymptotic behavior. In mixed-effect form they are also naturally robust with repeated measures data (e.g. variable number of tasks completed).

After hours of trying to get something to converge, I gave up. Sample code snippets below:

```R
# never got this to work!!
library(nlme)

data <- learn_h1c_std |>
  mutate(treatment = factor(as.character(treatment)))

gd <- nlme::groupedData(tct ~ seq | participant, data = data)

# Extract simple model coefficients
simple_model <- nls(tct ~ SSasymp(seq, Asym, R0, lrc), data = data)
simple_start_vals <- coef(simple_model)

# Check the levels of the treatment
lc <- length(unique(data$treatment))

start_vals <- c(
  rep(simple_start_vals["Asym"], lc),
  rep(simple_start_vals["R0"], lc),
  rep(simple_start_vals["lrc"], lc)
)

names(start_vals) <- c(paste0("Asym.", unique(data$treatment)),
                       paste0("R0.", unique(data$treatment)),
                       paste0("lrc.", unique(data$treatment)))

# exponential model with SSasymp (self-starting asymptotic regression) function
exp1 <- nlme::nlme(tct ~ SSasymp(seq, Asym, R0, lrc),
                  fixed = Asym + R0 + lrc ~ treatment,
                  random = nlme::pdDiag(Asym ~ 1 | participant),
                  start = start_vals,
                  data = gd)

summary(exp1)

# using SSasympOrig if starting points (R0) differ significantly
exp2 <- nlme::nlme(tct ~ SSasympOrig(seq, Asym, lrc),
                   fixed = Asym + lrc ~ treatment,
                   random = Asym ~ 1 | participant,
                   data = gd,
                   control = nlme::nlmeControl(pnlsTol = 0.1))

summary(exp2)
```

Let's try a different approach, based on change in TCT over time. Essentially find differences in rate of change in TCT. Related to "first differences" in Time Series Analysis, slope comparison in Growth Curves.

```{r}
# first attempt

# Extract Starting Values from the model estimates
# hardcoded - ick!
start_values <- data.frame(
  treatment = c("AR", "MR", "PAR", "PWI"),
  start_tct = c(185.6, 185.6 + 0.8, 185.6 - 70.6, 185.6 - 73.4)
)

# Calculate First Differences including properly handling initial TCT
d <- learn_h1c |>
  left_join(start_values, by = "treatment") |>
  group_by(participant) |>
  mutate(
    tct_diff = ifelse(
      is.na(lag(tct)),
      tct - start_tct,  # For t = 1, calculate difference from estimated TCT(t=0) directly
      tct - lag(tct)    # For subsequent tasks, difference from the previous task
    )
  ) |>
  ungroup()

d
```

At this point, we can simply compare `tct_diff` between groups with non-parametric tests:

```{r}
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

ggstatsplot::ggbetweenstats(
  data = d,
  x = treatment,
  y = tct_diff,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "rate of change in tct by treatment",
  #results.subtitle = FALSE,
  #pairwise.display = "all",
  subtitle = "Significant Pairwise Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 1,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)
```

This shows significant differences between AR/MR and PWI/PAR. Unfortunately, this isn't very informative as it only confirms what we already know:

- high initial TCT is correlated with high TCT differences
- AR and MR have higher initial TCT
- there is a natural limit to TCT that prevents further improvement
- together these create a kind of ceiling and floor effect that limit the value of mean / median statistics for comparison purposes

Instead, we should try to fit a mixed effects model to the differences, for example:

`tct_diff ~ treatment * seq + (1 + seq | participant)`

This would allow us to understand the participant-independent (fixed) effects of treatment and sequence on the changes in task completion time.

One such attempt:

```{r}
# Fit the mixed-effects model
m <- lmerTest::lmer(
  tct_diff ~ treatment * seq + (1 + seq | participant),
  data = d,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m)

```

No significant effects other than intercept (the average initial tct_diff for AR).
Singularity fit due to a lack of variance to model per participant, as shown in Random effects table. Large residuals indicate lots of unexplained variance.

Tried lots of things to address these issues. Here's one attempt to improve the data and simplify the model:

```{r}
# use unordered factor to improve interpretability of results
d$treatment <- factor(as.character(learn_h1c$treatment))

# record original means and SDs
mean_diff <- mean(d$tct_diff, na.rm = TRUE)
sd_diff <- sd(d$tct_diff, na.rm = TRUE)

# standardize data
d2 <- d |>
  mutate(
    tct_diff = scale(tct_diff) |> as.vector()
  )

glimpse(d2)

# Fit the mixed-effects model with interaction effects
# use the overloaded version of lmer from lmerTest to get CI, p-vals
m <- lmerTest::lmer(
    tct_diff ~ treatment + seq + (1 | participant),
    data = d2,
    control = lmerControl(optimizer = "bobyqa")
  )

summary(m)
```

Singularity problems persist. Marginal to low significance in PWI, PAR differences. Lack of participant variance is frustrating as it implies that all 50+ participants experienced the same learning progression, regardless of treatment, personal idiosyncracies, etc. Doesn't make sense.

Potential sources of this discrepancy are issues with data transformation or model specification, or  insufficient participant variability relative to within-participant variability (intraclass correlation, ICC). *The later seems unlikely?*

To attempt to address these issues, further changes were made to:

- eliminate the use of estimated TCT(t=0), which might be concealing some participant variation, and base changes only on lag
- incorporate ICC checks and other model diagnostics including plots (e.g. change in TCT, as shown below)
- varying the model formulation, e.g. adding random slopes `(1 + seq | participant)`, treating SEQ as a random effect only `tct_diff ~ treatment + (1 + seq | participant)`, fixed-effects only
- centering the SEQ data in an attempt to address correlation between the intercept and slope of TCT changes (a waste of time since they are naturally related in learning rates)
- converting change in TCT from time data to proportions of total improvement


```{r}
#| warning: false

ggplot(d, aes(x = seq, y = tct_diff, group = participant, color = treatment)) +
  geom_line(alpha = 0.2) +
  geom_smooth(aes(group = treatment),
              method = "loess",
              se = FALSE,
              size = 1) +
  theme_minimal() +
  labs(title = "Participant Learning Progressions",
       x = "Task Sequence",
       y = "TCT Differences")
```

Here is one such attempt, using simplified implementation of diffs:

```{r}
ds <- learn_h1c |>
  left_join(start_values, by = "treatment") |>
  group_by(participant) |>
  mutate(tct_diff = tct - lag(tct)) |>
  ungroup() |>
  filter(!is.na(tct_diff))

# use unordered factor to improve interpretability of results
ds$treatment <- factor(as.character(ds$treatment))

ms <- lmerTest::lmer(
  tct_diff ~ treatment + seq + (1 + seq | participant),
  data = ds,
  control = lmerControl(optimizer = "bobyqa")
)

summary(ms)

# Calculate ICC
as.numeric(VarCorr(ms)$participant[1]) /
  (as.numeric(VarCorr(ms)$participant[1]) +
     as.numeric(attr(VarCorr(ms), "sc") ^ 2))
```

This shows an improvement in the participant variance, but no significant treatment effects. RE stats indicate that intercept and slope are 100% inversely correlated. ICC of 0.81 suggests substantial participant-level variability.

*Is this the "best" model of all?*

Here's another attempt based on normalized (proportions) `tct_diff` values. In spite of finding marginally significant (p=0.04) overall differences, the Holm-adjusted Dunn comparisons show no significant treatment effects.

```{r}
#| warning: false

# add start times and diffs
d <- learn_h1c |>
  left_join(start_values, by = "treatment") |>
  group_by(participant) |>
  mutate(tct_diff = tct - ifelse(is.na(lag(tct)), start_tct, lag(tct))) |>
  ungroup() |>
  filter(!is.na(tct_diff))

# use progress (proportion) instead of difference
# include Start TCT in Total Improvement Calculation
d <- d |>
  group_by(participant) |>
  mutate(max_tct = max(c(tct, start_tct)),
         min_tct = min(tct),
         tot_imp = max_tct - min_tct) |>
  ungroup()

# normalize tct_diff by tot_imp
d <- d |>
  mutate(tct_diff_norm = tct_diff / tot_imp) |>  # Normalize the differences
  filter(!is.na(tct_diff_norm))   # Ensure there are no NA values

# View the transformed data
glimpse(d)
```

```{r}
#| fig-height: 6

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

ggstatsplot::ggbetweenstats(
  data = d,
  x = treatment,
  y = tct_diff_norm,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "rate of change in tct by treatment",
  #results.subtitle = FALSE,
  pairwise.display = "all",
  subtitle = "Significant Pairwise Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 1,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

```

At this point it seems clear that, if there is a treatment effect at all, it is small. This is not an effort to 'p-hack', I just want to make sure all efforts are made to get a valid finding. We've collected strong statistical evidence that there is little to no treatment effect, but our confidence in that result is not high based on prior expectations (which should not unduly influence interpretation). Some statistical results (e.g. the initial mixed-effect model on raw TCT, its changes) seemed to support the idea of treatment effect, but were influenced by other effects / covariance.

One last shot, fitting fixed effect and mixed effects models on the proportional data, standardized:

```{r}
# use unordered factor to improve interpretability of results
d$treatment <- factor(as.character(d$treatment))

# record original means and SDs
mean_diff <- mean(d$tct_diff, na.rm = TRUE)
sd_diff <- sd(d$tct_diff, na.rm = TRUE)

# standardize data
d <- d |>
  mutate(
    tct_diff = scale(tct_diff) |> as.vector()
  )

glimpse(d)

## try simple baseline model
# Fit a fixed-effects linear model (OLS)
fem <- lm(tct_diff_norm ~ treatment + seq, data = d)

# Summarize the model
summary(fem)

# Fit the mixed-effects model with interaction effects
# use the overloaded version of lmer from lmerTest to get CI, p-vals
h1c_mod_diffs <- lmerTest::lmer(
    tct_diff_norm ~ treatment + seq + (1 + seq | participant),
    data = d,
    control = lmerControl(optimizer = "bobyqa")
  )

summary(h1c_mod_diffs)
```

Poor fit, singularity, no significant treatment effects. Pffft.

Confluence of evidence that, either (1) I don't know what I'm doing, or (2) there is no treatment effect.

Things to consider:
1. We have focused almost exclusively on improvements to TCT as the sole indicator of increasing proficiency. How can error counts be considered? Bivariate analysis or survival models?
2. Should we attempt a piecewise linear model of the data to see if treatment effect can be spotted?

::: content-hidden
MAYBE: learning add. analysis

### Additional Analyses

- what treatment group had the highest percentage of "qualified operators" at the end of the 10-min session?

#### Error Types

#### Task Completion Rate

#### First-Task Performance

#### Qualified Operators

#### Transfer Effectiveness Rating (TER)
:::

## Recall Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Overall Equipment Effectiveness (H2a)

#### Paper Work Instruction References (H2b)

::: content-hidden
TODO: recall add. analysis

### Additional Analyses

- what treatment group had the highest OEE during recall?

#### Efficiency vs. Quality

#### PWI Reference Frequency and Length

#### Error Types and Task Completion Rate

#### Ongoing Learning
:::

## Retention Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Change in OEE (H3a)

::: content-hidden
TODO: retention add. analysis

### Additional Analyses

#### Absolute and Percentage Change in OEE

#### Primary Driver of OEE Change
:::

## Workload, Usability, and User Experience

### NASA Task Load Index Results

### System Usability Scale Results

### Qualitative Feedback Analysis

::: content-hidden
TODO: exploratory analysis

## Exploratory Analyses

### Relationship between TLX, SUS, and Performance

- first task performance in learning as a measure of initial comprehension, intuitiveness of each treatment, effectiveness of instructional design, etc.; may correlate with SUS / TLX

### Impact of Participant Demographics on Performance

### Within-Group Performance Variance

### TLX Components and Performance

### Interactions between Workload, Usability, and Treatment
:::

## Challenges and Unexpected Observations {#sec-res-challenges}

Discussion of any challenges encountered during the conduct of the protocol  Modifications made to the protocol in response to these challenges

demoing AR/MR: The challenge of this choreography was compounded by significant latency in streamed video, which lagged the live action by approximately 2 seconds. With practice

> Technical Issues with Equipment: Discussion of any technical challenges encountered with the HL2 or other devices, and how these were addressed.
>
> Participant Scheduling and No-Shows: Challenges related to participant scheduling, no-shows, or last-minute cancellations, and the strategies developed to mitigate these issues.
>
> Data Collection Consistency: Any issues encountered in maintaining consistency in data collection across different trials and days, and how these were resolved.

limitations

- participant behavior
    -   few participants assigned HMDMR took advantage of the affordance
- analysis
    -   does not account for interactions between affordances

noticed improvising - discourage as occurs

sleeves, long hair problems for HL2

height and "down-lookers"; tracking dropouts;

battery life and performance; brightness and accuracy;

complexity of MR, less reliable tracking, crashes

## Summary of Key Findings
