# Results {#25-results}

```{r load-libs}
#| message: false

library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
library(broom)
library(report)
library(tools)
# library(glue) now included with tidyverse as str_glue
# library(gridExtra)
# library(ggtext)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r load-data}

car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results"
)

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics"
)
```

```{r def-cust-theme}
base_color <- "#69b3a2"

# Define a custom theme for ggplot
custom_theme <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.title.position = 'plot',
      plot.title = element_text(face = "bold"),
      # plot.subtitle = element_text(face = "italic"),
      # plot.caption = element_text(color = "grey50"),
      # axis.title.x = element_text(face = "bold"),
      # axis.title.y = element_text(face = "bold"),
      # axis.text = element_text(color = "black"),
      # axis.line = element_line(color = "grey80"),
      # legend.position = "bottom",
      # legend.background = element_rect(fill = "white", colour = "grey50"),
      # legend.key = element_rect(fill = "white", colour = "grey50"),
    )
}

# Set the custom theme as the default theme
theme_set(custom_theme())
```

```{r study-metadata}

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# Convert the trial_datetime to Date type
demographics <- demographics |>
  mutate(
    trial_date = as.Date(
      as.POSIXct(trial_datetime, format = "%Y-%m-%d %H:%M:%S")
    )
  )

# Get the first and last date
first_trial_date <- demographics |>
  summarize(first_date = min(trial_date)) |>
  pull(first_date)

last_trial_date <- demographics |>
  summarize(last_date = max(trial_date)) |>
  pull(last_date)

trial_start_str <- str_glue("{month(first_trial_date)}/{day(first_trial_date)}")
trial_end_str <- str_glue("{month(last_trial_date)}/{day(last_trial_date)}")
trial_year_str <- str_glue("{year(last_trial_date)}")
# start, end = Feb 10, Apr 27 2023
```

## Study Administration and Participation

This study was administered during the Spring of 2023. Ultimately, `r participant_count` eligible participants were recruited. All completed the Learning and Recall experiments of the first session beginning February 10th. Progress of the conduct of these trials is illustrated by @fig-s1-progress.

```{r fig-s1-progress}
#| fig-cap: "Daily Trial Administration and Completion Percentage"
#| fig-height: 4

# Count experiments per day
daily_counts <- demographics |>
  group_by(trial_date) |>
  summarize(daily_count = n()) |>
  arrange(trial_date)

# Calculate the cumulative count of experiments
daily_counts <- daily_counts |>
  mutate(cumulative_count = cumsum(daily_count))

# Calculate remaining experiments
daily_counts <- daily_counts |>
  mutate(
    percentage_complete = (cumulative_count / participant_count) * 100
  )

pc_eng <- as.character(english::english(participant_count))
pc_eng <- str_replace(pc_eng, "^.", str_to_upper)
subtitle <- str_glue("{pc_eng} Learning and Recall Experiments Conducted between {trial_start_str} and {trial_end_str} of {trial_year_str}")

# Plotting
ggplot(daily_counts, aes(x = trial_date)) +
  geom_line(
    aes(y = daily_count, group = 1), 
    color = base_color,
    linewidth = 1,
    alpha = 0.5
  ) +
  geom_line(
    aes(y = percentage_complete, group = 1), 
    color = base_color,
    linewidth = 1.25,
    alpha = 1
  ) +
  scale_y_continuous(
    name = "Trials Administered",
    sec.axis = sec_axis(~ ., name = "Percentage Complete"),
    position = "right",
    limits = c(0,100)
  ) +
  labs(
    title = "First Session Progress",
    subtitle = subtitle,
    x = element_blank(),
    y = element_blank()
  ) +
  custom_theme()
```

Only eight trials were run before March 16th. That period was treated as a pilot study, during which the research team was trained, administration methods were refined, the self-service onboarding system was deployed, and version 2.1 of our IRB application was approved. Most significantly, the intake process was adjusted to collect more demographic data, increase the role of the TLX and SUS instruments, and gather BCS data for a companion study.

Otherwise, only minor tweaks were made during the pilot to streamline the experiment's administration. Although the changes outlined above were not believed to have a material effect on the results, the data collected during the pilot were omitted from further analysis. Following the pilot, steady progress was made until the completion of the study's first session on April 27th.

The second session was conducted during the end of study event on *XXX*. *YYY* of the study's original participants volunteered to attend and complete the Retention experiment.

::: content-hidden
TODO: complete above paragraph about second session.
:::

::: content-hidden
MAYBE: perform analysis to evaluate differences between pilot and main group

Pilot Study Comparison
- Compare the key measures (e.g., performance, workload, usability) between the pilot study (n=8) and main study (n=54) participants
- Use descriptive statistics and visualizations to present these comparisons
- Discuss any notable similarities or differences between the two groups

Statistical Power and Sample Size Considerations
- Present the results of statistical power analyses based on the observed effect sizes and the planned statistical tests
- Discuss the adequacy of the sample size for detecting meaningful effects
- Consider the potential impact of including or excluding the pilot study data on the statistical power and precision of the estimates
- sample size calculation / justification: power analysis, practical considerations, precedent

Decision on Pilot Study Inclusion
- Based on the findings from 5.2.3 and 5.2.4, make a decision on whether to include the pilot study data in the main analysis
- Clearly report your decision and rationale, considering factors such as data comparability, study design modifications, and statistical power
:::

## Analysis of Participant Demographics

### Univariate Analysis

```{r drop-pilot-data}
# drop participants in pilot study (first eight)
demographics <- demographics |>
  filter(!(participant >= 1001 & participant <= 1010))

# drop event data from the pilot study
car_results <- car_results |>
  filter(!(participant >= 1001 & participant <= 1010))
```

```{r def-to-factors}

# safely convert data to ord / cat factors
# create levels / labels as required and identify / warn about mismatches

to_factors <- function(
    column, 
    levels = NULL, 
    labels = NULL,
    ordered = FALSE
    ) {

  if (is.null(levels)) {
    levels <- unique(na.omit(column))
  }
  
  if (is.null(labels)) {
    labels <- levels
  }
  
  mismatched <- setdiff(unique(na.omit(column)), levels)
  
  if (length(mismatched) > 0) {
    # Append mismatched values to levels and labels with a specific prefix
    bad_levels <- paste("Bad -", mismatched)
    levels <- c(levels, mismatched)
    labels <- c(labels, bad_levels)
    warning(
      "Mismatched or unexpected values incorporated: ", 
      paste(bad_levels, collapse = ", ")
    )
  }

  # Proceed to convert to factor using cleaned-up data
  factor_column <- factor(column, levels = levels, labels = labels, 
                          ordered = ordered)

  return(factor_column)
}
```

```{r factor-demographics}

# convert all categorical / ordinal data to factors

demo_f <- demographics |>
  mutate(
    # categorical, levels != labels
    # unordered with predefined levels
    gender = to_factors(
      gender,
      levels = c("F", "M", "O"),
      labels = c("Female", "Male", "Non-Binary")),
    race = to_factors(
      race,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "American Indian or Alaska Native",
        "Asian",
        "Black or African American",
        "Native Hawaiian or Other Pacific Islander",
        "White",
        "More than one race",
        "Unknown or not reported"
      )
    ),
    ethnicity = to_factors(
      ethnicity,
      levels = c("1", "2", "3"),
      labels = c(
        "Hispanic or Latino",
        "Not Hispanic or Latino",
        "Unknown or not reported"
      )
    ),
    lenses = to_factors(
      lenses,
      levels = c("G", "C", "N", "Y"),
      labels = c(
        "Glasses",
        "Contacts",
        "No",
        "Yes"
      )
    ),
    
    # categorical, levels == labels
    # unordered with predefined levels
    language = to_factors(
      language,
      levels = c("English", "Other"),
    ),
    worn = to_factors(
      worn,
      levels = c("Y", "N")
    ),
    color_blind = to_factors(
      color_blind,
      levels = c("Y", "N")
    ),
    condition = to_factors(
      condition,
      levels = c("Y", "N")
    ),

    # ad-hoc
    # categorized in transcription, no predefined levels
    country = to_factors(country),
    college = to_factors(college),
    program = to_factors(program),
    referral = to_factors(referral),
    treatment = to_factors(treatment),
    
    # ordinal, levels != labels
    # ordered with predefined levels
    education = to_factors(
      education,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "Pre-HS", # No HS diploma
        "HS", # High school diploma or equivalent
        "CND", # Some college but no degree
        "AS",
        "BS",
        "MS",
        "PhD"
      ),
      ordered = TRUE
    ),
    lego = to_factors(
      lego,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "Little to no experience",
        "Some experience",
        "Lots of experience",
        "Expert builder"
      ),
      ordered = TRUE
    ),
    manuf = to_factors(
      manuf,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "No experience",
        "Taken one or more classes",
        "Held a part time or temporary position",
        "One or more year’s experience"
      ),
      ordered = TRUE
    ),
    runner = to_factors(
      runner,
      levels = c("N", "Y", "5", "10", "H", "M", "T"),
      labels = c(
        "None",
        "Yes - Unspecified",
        "5K",
        "10K",
        "Half-Marathon",
        "Marathon",
        "Triathalon"
      ),
    )
  )
```

#### Numeric Data

```{r numeric-data}
nums <- demo_f |>
  select(where(is.numeric))

height_stats <- nums |>
  summarize(
    mean = mean(height, na.rm = TRUE),
    sd = sd(height, na.rm = TRUE),
    min = min(height, na.rm = TRUE),
    max = max(height, na.rm = TRUE),
    pct_missing = sum(is.na(height)) / n() * 100
  )

height_summary <- str_glue(
  "Mean height = {round(height_stats$mean, 1)}, SD = {round(height_stats$sd, 1)}, range: [{height_stats$min}, {height_stats$max}], {round(height_stats$pct_missing, 1)}% missing"
)
```

After excluding data from the pilot study, `r report_participants(nums, age = "age", participants = "participant")` were included in this analysis. Other than age, height (`r height_summary`) was the only numeric data collected. As discussed in @sec-res-challenges, height was first recorded following the pilot study, when its impact on user behavior and results was first observed. Both numeric characteristics are visualized with a combined box and scatter plot in @fig-numeric.

```{r fig-numeric}
#| fig-cap: "Distribution of Participant Age and Height"
#| fig-height: 4
#| warning: false

# Plotting age
age_plot <- demographics |>
  filter(!is.na(age)) |>
  ggplot(aes(x = age)) +
  geom_bar(
    aes(y = ..count../max(..count..)),
    fill = I(base_color),
    alpha = 0.75
  ) + 
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07)
  ) +
  labs(
    x = "Age (yr)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting height
height_plot <- demographics |>
  filter(!is.na(height)) |>
  ggplot(aes(x = height)) +
  geom_bar(
    aes(y = ..count../max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    width = 0.55
  ) + 
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07)
  ) +
  labs(
    x = "Height (in)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
age_plot + height_plot +
  # plot_layout(ncol = 2, widths = c(1,0.8)) +
  plot_annotation(
    title = "Numeric Sample Characteristics",
    subtitle = "Age reflects the student-oriented nature of the sample"
  )
```


#### Categorical Data

```{r categorical-data}
cats <- demo_f |>
  select(where(~is.factor(.) && !is.ordered(.))) |>
  select(-c("referral", "runner")) |>
  mutate(across(everything(), \(x) fct_lump(x, prop = 0.1)))

# Summary table
cat_tbl <- cats |>
  select(gender:language) |>
  tbl_summary(
    label = setNames(str_to_title(names(cats)), names(cats)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

cat_names <- cats |>
  select(gender:country) |>
  names()

cats_long <- cats |>
  select(gender:country) |>
  pivot_longer(
    cols = everything(), 
    names_to = "Category", 
    values_to = "Value"
  ) |>
  drop_na() |>
  group_by(Category, Value) |>
  summarize(Count = n(), .groups = 'drop') |>
  group_by(Category) |>
  mutate(Freq = Count / sum(Count))

country_levels <- demo_f |>
  filter(country != "USA") |>
  count(country) |>
  arrange(desc(n))
```

Of the sample's `r inline_text(cat_tbl, language, level = "English", pattern = "{N} participants, {p}% (n={n})")` reported that they mainly speak English at home. As shown in @fig-categorical, the gender composition was relatively balanced, with slightly more males `r inline_text(cat_tbl, gender, level = "Male", pattern = "(n={n}, {p}%)")` than females `r inline_text(cat_tbl, gender, level = "Female", pattern = "(n={n}, {p}%)")`. The racial makeup of the sample was predominantly White `r inline_text(cat_tbl, race, level = "White", pattern = "(n={n}, {p}%)")` and Asian `r inline_text(cat_tbl, race, level = "Asian", pattern = "(n={n}, {p}%)")`, with smaller representations from `r as.character(english::english(n_distinct(demo_f$race) - 2))` other groups. Fully `r inline_text(cat_tbl, ethnicity, level="Not Hispanic or Latino", pattern = "{p}% of participants (n={n})")` reported a non-Hispanic or Latino ethnicity. The majority of participants `r inline_text(cat_tbl, country, level = "USA", pattern = "(n={n}, {p}%)")` were from the United States, though `r as.character(english::english(n_distinct(demo_f$country, na.rm = TRUE) - 1))` other countries were represented, including `r paste(country_levels[1, ]$country, " (n=", country_levels[1, ]$n, ")", sep = "")` and `r paste(country_levels[2, ]$country, " (n=", country_levels[2, ]$n, ")", sep = "")`. `r paste(country_levels$country[3:(nrow(country_levels) - 1)], collapse = ", ")`, and `r tail(country_levels$country, 1)` were also listed as countries of origin by one participant each (n=1). All of this aligns with the study's recruitment focus.

```{r fig-categorical}
#| fig-cap: "Distribution of Categorical Characteristics"
#| fig-height: 4
#| results: hide

create_cat_plot <- function(data, category, base_color) {
  # specify alpha ramp end points
  min_alpha <- 0.3
  max_alpha <- 0.8
  
  # Filter and reorder data
  plot_data <- data |>
    filter(Category == category) |>
    mutate(
      Value = if_else(Value == "Not Hispanic or Latino", 
                      "Not Hispanic\nor Latino", Value),
      Value = reorder(Value, Count, FUN = sum, decreasing = TRUE)
    )

  # Calculate alpha values based on Count linearly between min_alpha and max_alpha
  alpha_values <- plot_data |>
    mutate(alpha = scales::rescale(Count, to = c(min_alpha, max_alpha))) |>
    pull(alpha)
  
  # Set unique Value groups for setting colors and alphas
  unique_values <- unique(plot_data$Value)
  
  # Create the plot
  ggplot(
    plot_data, 
    aes(x = Value, y = Count, fill = I(base_color), alpha = Value)
  ) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_x_discrete(labels = function(x) tools::toTitleCase(x)) +
  labs(x = tools::toTitleCase(category), y = element_blank()) +
  scale_alpha_manual(values = setNames(alpha_values, unique_values)) +
  guides(fill = "none", alpha = "none") +
  ylim(0, max_count) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )
}

# Calculate the maximum count to standardize the y-axis
max_count <- max(cats_long$Count)

# Create all plots
cat_plots <- map(
  cat_names, 
  ~create_cat_plot(cats_long, .x, base_color)
)

# Add y-axis text and ticks to first plot
first_plot <- cat_plots[[1]] +
  theme(axis.text.y = element_text(),
        axis.ticks.y = element_line())

# Other plots without y-axis text and ticks, drop first
other_plots <- cat_plots[-1]

# combine!
all_plots <- c(list(first_plot), other_plots)

# Combine all plots
combined_plot <- reduce(all_plots, `+`) +
  plot_layout(ncol = length(all_plots), widths = c(1,1,0.667,0.667)) + 
  plot_annotation(
    title = "Categorical Sample Characteristics",
    subtitle = "Value Count for Gender, Race, Ethnicity, and Country"
  )

# Print the combined plot
print(combined_plot)
```

```{r make-vc-table}

vc_tbl <- cats |>
  select(lenses:worn) |>
  filter(!lenses %in% c("No", "Other", "Unknown")) |>
  filter(!(is.na(lenses) & is.na(worn))) |>
  droplevels() |>
  mutate(worn = recode(worn, 'Y' = 'Yes', 'N' = 'No')) |>
  tbl_cross(
    row = lenses,
    col = worn,
    label = list(lenses = "Lenses", worn = "Worn During Session"),
    percent = "row",
    missing_text = "NA"
  ) |>
  add_stat_label() |>
  bold_labels()

```

Of the 54 participants, 52% (n=28) reported having a vision prescription. Of those, `r inline_text(vc_tbl, row_level = "Contacts", col_level = "Total")` were for contact lenses , but `r inline_text(vc_tbl, row_level = "Glasses", col_level = "Total")` had glasses. Whereas all contact wearers reported they would wear them during the session, among glasses wearers only seven (47%) planned to the same. Five (33%) did not intend to do so, and three (20%) gave no indication. This breakdown is summarized by @tbl-vc. Two participants also reported color-blindness (n=1) or other vision conditions (n=1). No participants indicated any other condition that might affect their performance.

::: content-hidden
MAYBE: make this all inline code; rethink percent by row, col, or cell? format n= etc. as markdown?
:::

```{r tbl-vc}
#| tbl-cap: Vision Correction Table

vc_tbl |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )

```

Following the primary hypothesis testing, additional analyses may explore the differences in key outcomes across these demographic subgroups. That effort will aim to provide additional context and insights into potential underlying factors influencing the findings.

#### Ordinal Data

```{r ord-data}
ords <- demo_f |>
  select(where(~is.factor(.) && is.ordered(.)))

# Summary table
ords_tbl <- ords |>
  tbl_summary(
    label = setNames(str_to_title(names(ords)), names(ords)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

# ords_tbl

# Helper function to calculate mode
calculate_mode <- function(data, var_name) {
  count_table <- table(data[[var_name]])
  max_freq <- max(count_table)
  # pick first if more than one value has max count
  mode_value <- names(count_table[count_table == max_freq])[1]
  
  # Convert to an output that includes the level's name (mode_value) and its index
  mode_index <- which(levels(data[[var_name]]) == mode_value)
  return(list(value = mode_value, index = mode_index))
}

# Main function to calculate statistics and return formatted string
calculate_ord_summary_stats <- function(data, var_name) {
  # Ensure the variable is an ordered factor
  if (!is.ordered(data[[var_name]])) {
    stop("The variable must be an ordered factor.")
  }
  
  # Convert factor to integer codes
  factor_int <- as.integer(data[[var_name]])
  
  # Calculate median, mode, and IQR
  median_value <- median(factor_int, na.rm = TRUE)
  mode_results <- calculate_mode(data, var_name)
  iqr_value <- IQR(factor_int, na.rm = TRUE)
  
  # Format and return the result as a string
  result_string <- sprintf("(Median = %.1f; Mode = %d, %s; IQR = %.1f)", 
                           median_value, mode_results$index, mode_results$value, 
                           iqr_value)
  return(result_string)
}
```

Unlike the data above, the categories analyzed in this section are inherently ordered. Responses for each participant's current Education level `r calculate_ord_summary_stats(ords, "education")` and their experience with Lego `r calculate_ord_summary_stats(ords, "lego")` and Manufacturing `r calculate_ord_summary_stats(ords, "manuf")` are visualized in increasing order by @fig-ordinal. The results are mostly in line with expectations based on the recruitment, though it is disappointing to see that `r inline_text(ords_tbl, lego, level = "Little to no experience", pattern = "{p}% of participants (n={n})")` claimed to have little to no experience with Lego. The great majority were undergraduate students working towards a degree `r inline_text(ords_tbl, education, level = "CND", pattern = "(n={n}, {p}%)")`, with no experience in manufacturing `r inline_text(ords_tbl, manuf, level = "No experience", pattern = "(n={n}, {p}%)")`.

```{r fig-ordinal}
#| fig-cap: "Distribution of Ordinal Characteristics"
#| fig-height: 4
#| results: hide

plot_ords <- function(data, var, title) {
  norm_data <- data |>
    count(var_name = {{ var }}) |>
    mutate(percent = n / sum(n) * 100)
  
  p <- norm_data |>
    ggplot(aes(y = var_name, x = percent)) +
    geom_bar(
      aes(alpha = percent / 100),
      fill = base_color,
      stat = "identity", 
      width = 0.75
    ) +
    scale_alpha_continuous(trans = "log10", range = c(0.2, 0.8), guide = "none") +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank()
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(size = 9, face = "bold"),
      legend.position = "none",
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
  return(p)
}


educ_plot <- ords |>
  mutate(education = fct_recode(education, 
                                College = "CND",
                                Associate = "AS",
                                Bachelor = "BS",
                                Master = "MS",
                                Doctorate = "PhD")
  ) |>
  plot_ords(var = education, title = "Current Education Level")

lego_plot <- ords |>
  mutate(lego = fct_recode(lego,
                           Little = "Little to no experience",
                           Some = "Some experience",
                           Lots = "Lots of experience",
                           Expert = "Expert builder")
  ) |>
  plot_ords(lego, title = "Lego Experience")

manuf_plot <- ords |>
  mutate(manuf = fct_recode(manuf,
                           None = "No experience",
                           Training = "Taken one or more classes",
                           PartTime = "Held a part time or temporary position",
                           Some = "One or more year’s experience")
  ) |>
  plot_ords(manuf, title = "Manufacturing Experience")

# combine using patchwork
educ_plot / (lego_plot + manuf_plot) +
  # plot_layout(ncol = 2) +
  plot_annotation(
    title = "Ordinal Sample Characteristics",
    subtitle = "Reflects the Focus of Recruiting Efforts"
  )

```

### Multivariate Analysis

- Multivariate
  - above by treatment?

@tbl-demo-by-treatment shows the



```{r tbl-demo-by-treatment}
#| tbl-cap: Participant Demographics by Treatment Group

# Compute summary statistics
demographics |>
  select(age:gender, treatment) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    label = list("age" ~ "Age", "height" ~ "Height", "gender" ~ "Gender")
  ) |>
  modify_header(
    all_stat_cols() ~ "**{level}** \nN = {n} ({style_percent(p)}%)"
  ) |>
  add_n() |>
#  bold_labels() |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```



::: content-hidden
### Treatment Group Equivalence

- Compare the demographic characteristics across the four treatment groups
- Use statistical tests (e.g., chi-square, ANOVA) to assess the equivalence of the groups
- Discuss any notable differences or similarities between the groups
::: 



## Learning Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Average Time per Car (H1a)

#### Average Error Count per Car (H1b)

#### Learning Rates (H1c)

### Additional Analyses

- what treatment group had the highest percentage of "qualified operators" at the end of the 10-min session?

#### Error Types

#### Task Completion Rate

#### First-Task Performance

#### Qualified Operators

## Recall Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Overall Equipment Effectiveness (H2a)

#### Paper Work Instruction References (H2b)

### Additional Analyses

- what treatment group had the highest OEE during recall?

#### Efficiency vs. Quality

#### PWI Reference Frequency and Length

#### Error Types and Task Completion Rate

#### Ongoing Learning

## Retention Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Change in OEE (H3a)

### Additional Analyses

#### Absolute and Percentage Change in OEE

#### Primary Driver of OEE Change

## Workload, Usability, and User Experience

### NASA Task Load Index Results

### System Usability Scale Results

### Qualitative Feedback Analysis

## Exploratory Analyses

### Relationship between TLX, SUS, and Performance

- first task performance in learning as a measure of initial comprehension, intuitiveness of each treatment, effectiveness of instructional design, etc.; may correlate with SUS / TLX

### Impact of Participant Demographics on Performance

### Within-Group Performance Variance

### TLX Components and Performance

### Interactions between Workload, Usability, and Treatment

## Challenges and Unexpected Observations {sec-res-challenges}

Discussion of any challenges encountered during the conduct of the protocol  Modifications made to the protocol in response to these challenges

demoing AR/MR: The challenge of this choreography was compounded by significant latency in streamed video, which lagged the live action by approximately 2 seconds. With practice

> Technical Issues with Equipment: Discussion of any technical challenges encountered with the HL2 or other devices, and how these were addressed.
>
> Participant Scheduling and No-Shows: Challenges related to participant scheduling, no-shows, or last-minute cancellations, and the strategies developed to mitigate these issues.
>
> Data Collection Consistency: Any issues encountered in maintaining consistency in data collection across different trials and days, and how these were resolved.

limitations

- participant behavior
  -   few participants assigned HMDMR took advantage of the affordance
- analysis
  -   does not account for interactions between affordances

noticed improvising - discourage as occurs

sleeves, long hair problems for HL2

height and "down-lookers"; tracking dropouts;

battery life and performance; brightness and accuracy; 

complexity of MR, less reliable tracking, crashes


## Summary of Key Findings
