# Results {#25-results}

```{r load-libs}
#| message: false

library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
library(broom)
library(report)
library(tools)
# library(ggtext)
# library(glue) now included with tidyverse as str_glue
# library(gridExtra)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r load-data}

car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results"
)

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics"
)
```

```{r def-cust-theme}
base_color <- "#69b3a2"
outline_color <- "grey25"

# Define a custom theme for ggplot
custom_theme <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      # plot.subtitle = element_text(face = "italic"),
      # plot.caption = element_text(color = "grey50"),
      # axis.title.x = element_text(face = "bold"),
      # axis.title.y = element_text(face = "bold"),
      # axis.text = element_text(color = "black"),
      # axis.line = element_line(color = "grey80"),
      # legend.position = "bottom",
      # legend.background = element_rect(fill = "white", colour = "grey50"),
      # legend.key = element_rect(fill = "white", colour = "grey50"),
    )
}

# Set the custom theme as the default theme
theme_set(custom_theme())
```

```{r def-funcs}
# Function to create QQ plots and arrange them in a 2x2 grid
qq_plot_patchwork <-
  function(data,
           x_var,
           by_var) {
    levels <- unique(data[[by_var]])

    if (length(levels) != 4) {
      stop("This script expects exactly 4 levels in the grouping variable.")
    }

    plots <- lapply(levels, function(level) {
      p <- data |>
        filter((!!sym(by_var)) == level) |>
        ggpubr::ggqqplot(
          x = x_var,
          xlab = "",
          ylab = "",
          title = level,
          ggtheme = theme_minimal(base_size = 10) +
            theme(
              plot.margin = margin(0, 0, 0, 0),
              panel.spacing = unit(0, "lines"),
              panel.grid = element_blank(),
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              axis.line = element_blank(),
              plot.title = element_text(hjust = 0.5),
              panel.border = element_rect(
                color = outline_color,
                fill = NA,
                linewidth = 0.5
              ),
            )
        ) |>
        ggpubr::ggpar(ticks = FALSE, tickslab = FALSE)
    })

    # Arrange the plots in a 2x2 grid using patchwork
    plot_layout <-
      (plots[[1]] + plots[[2]]) / (plots[[3]] + plots[[4]])
    
    return(plot_layout)
  }

# convert a p-value into an appropriately formatted string
fmt_p_val <- function(p_value, alpha = 0.05) {
  # # Check if p-value is greater than alpha
  # if (p_value > alpha) {
  #   return(paste0("> ", alpha))
  # }
  
  if (p_value < 0.001) {
    return("p < 0.001")
  }

  # need more precision for 0.01 > p-val > 0.001
  if (p_value < 0.01) {
    return(paste0("p = ", round(p_value, digits = 3)))
  }
  
  return(paste0("p = ", round(p_value, digits = 2)))
}

# Function to calculate power for Kruskal-Wallis test via simulation
simulate_kruskal_power <- function(
    n, k, effect_size, sig_level = 0.05, num_simulations = 1000
    ) {
    set.seed(42)
    p_values <- numeric(num_simulations)
    for (i in 1:num_simulations) {
      # Simulate data for each group with an effect size
      data <- data.frame(value = unlist(lapply(1:k, function(x)
        rnorm(n, mean = x * effect_size))),
        group = gl(k, n))
      # Perform Kruskal-Wallis test
      p_values[i] <- kruskal.test(value ~ group, data = data)$p.value
    }
    # Proportion of tests rejecting null hypothesis gives power estimate
    power <- mean(p_values < sig_level)
    return(power)
  }

# Function to run post-hoc power analysis based on observed epsilon squared
posthoc_power_analysis <- function(
    n, k, observed_effect_size, sig_level = 0.05, num_simulations = 1000
    ) {
    # Convert epsilon squared back to standard effect size approximation for simulation
    std_effect_size <- sqrt(observed_effect_size / (1 - observed_effect_size))
    
    # Use simulation to estimate power as before
    power <- simulate_kruskal_power(
      n, k, std_effect_size, sig_level, num_simulations
    )
    return(power)
}

# convert a column of a table into a string with the format "a, b, and c"
# numbers are rounded to the specified number of digits
col_to_string <- function(data, column, digits = 2) {
  data |>
    mutate({{column}} := round({{column}}, digits)) |>
    pull({{column}}) |>
    str_c(collapse = ", ") |>
    (\(x) sub(", ([^,]+)$", ", and \\1", x))()
}
```

```{r save-shared}
# saving for use elsewhere
save(
  qq_plot_patchwork, 
  fmt_p_val,
  simulate_kruskal_power,
  posthoc_power_analysis,
  col_to_string,
  base_color,
  outline_color,
  custom_theme,
  file = "shared.RData"
)
```


```{r study-metadata}

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# Convert the trial_datetime to Date type
demographics <- demographics |>
  mutate(
    trial_date = as.Date(
      as.POSIXct(trial_datetime, format = "%Y-%m-%d %H:%M:%S")
    )
  )

# Get the first and last date
first_trial_date <- demographics |>
  summarize(first_date = min(trial_date)) |>
  pull(first_date)

last_trial_date <- demographics |>
  summarize(last_date = max(trial_date)) |>
  pull(last_date)

trial_start_str <- str_glue(
  "{month(first_trial_date)}/{day(first_trial_date)}"
)
trial_end_str <- str_glue(
  "{month(last_trial_date)}/{day(last_trial_date)}"
)
trial_year_str <- str_glue("{year(last_trial_date)}")

# start, end = Feb 10, Apr 27 2023
```

## Study Administration and Participation

This study was administered during the Spring of 2023. Ultimately, `r participant_count` eligible participants were recruited. All completed the Learning and Recall experiments of the first session beginning February 10th. Progress of the conduct of these trials is illustrated by @fig-s1-progress.

```{r fig-s1-progress}
#| fig-cap: "Daily Trial Administration and Completion Percentage"
#| fig-height: 3.5

# Count experiments per day
daily_counts <- demographics |>
  group_by(trial_date) |>
  summarize(daily_count = n()) |>
  arrange(trial_date)

# Calculate the cumulative count of experiments
daily_counts <- daily_counts |>
  mutate(cumulative_count = cumsum(daily_count))

# Calculate remaining experiments
daily_counts <- daily_counts |>
  mutate(
    percentage_complete = (cumulative_count / participant_count) * 100
  )

pc_eng <- as.character(english::english(participant_count))
pc_eng <- str_replace(pc_eng, "^.", str_to_upper)
subtitle <- str_glue("{pc_eng} Learning and Recall Experiments Conducted between {trial_start_str} and {trial_end_str} of {trial_year_str}")

# Plotting
ggplot(daily_counts, aes(x = trial_date)) +
  geom_line(
    aes(y = daily_count, group = 1),
    color = base_color,
    linewidth = 1,
    alpha = 0.5
  ) +
  geom_line(
    aes(y = percentage_complete, group = 1),
    color = base_color,
    linewidth = 1.25,
    alpha = 1
  ) +
  scale_y_continuous(
    name = "Trials Administered",
    sec.axis = sec_axis(~., name = "Percentage Complete"),
    position = "right",
    limits = c(0, 100)
  ) +
  labs(
    title = "First Session Progress",
    subtitle = subtitle,
    x = element_blank(),
    y = element_blank()
  ) +
  custom_theme()
```

Only eight trials were run before March 16th. That period was treated as a pilot study, during which the research team was trained, administration methods were refined, the self-service onboarding system was deployed, and version 2.1 of our IRB application was approved. Most significantly, the intake process was adjusted to collect more demographic data, increase the role of the TLX and SUS instruments, and gather BCS data for a companion study.

Otherwise, only minor tweaks were made during the pilot to streamline the experiment's administration. Although the changes outlined above were not believed to have a material effect on the results, the data collected during the pilot were omitted from further analysis. Following the pilot, steady progress was made until the completion of the study's first session on April 27th.

The second session was conducted during the end of study event on *XXX*. *YYY* of the study's original participants volunteered to attend and complete the Retention experiment.

::: content-hidden
TODO: complete above paragraph about second session.
:::

::: content-hidden
MAYBE: perform analysis to evaluate differences between pilot and main group

Pilot Study Comparison
- Compare the key measures (e.g., performance, workload, usability) between the pilot study (n=8) and main study (n=54) participants
- Use descriptive statistics and visualizations to present these comparisons
- Discuss any notable similarities or differences between the two groups

Statistical Power and Sample Size Considerations
- Present the results of statistical power analyses based on the observed effect sizes and the planned statistical tests
- Discuss the adequacy of the sample size for detecting meaningful effects
- Consider the potential impact of including or excluding the pilot study data on the statistical power and precision of the estimates
- sample size calculation / justification: power analysis, practical considerations, precedent

Decision on Pilot Study Inclusion
- Based on the findings from 5.2.3 and 5.2.4, make a decision on whether to include the pilot study data in the main analysis
- Clearly report your decision and rationale, considering factors such as data comparability, study design modifications, and statistical power
:::

## Analysis of Participant Demographics

### Univariate Analysis

```{r drop-pilot-data}
# drop participants in pilot study (first eight)
demographics <- demographics |>
  filter(!(participant >= 1001 & participant <= 1010))

# drop event data from the pilot study
car_results <- car_results |>
  filter(!(participant >= 1001 & participant <= 1010))
```

```{r def-to-factors}

# safely convert data to ord / cat factors
# create levels / labels as required and identify / warn about mismatches

to_factors <- function(
    column,
    levels = NULL,
    labels = NULL,
    ordered = FALSE) {
  if (is.null(levels)) {
    levels <- unique(na.omit(column))
  }

  if (is.null(labels)) {
    labels <- levels
  }

  mismatched <- setdiff(unique(na.omit(column)), levels)

  if (length(mismatched) > 0) {
    # Append mismatched values to levels and labels with a specific prefix
    bad_levels <- paste("Bad -", mismatched)
    levels <- c(levels, mismatched)
    labels <- c(labels, bad_levels)
    warning(
      "Mismatched or unexpected values incorporated: ",
      paste(bad_levels, collapse = ", ")
    )
  }

  # Proceed to convert to factor using cleaned-up data
  factor_column <- factor(column,
    levels = levels, labels = labels,
    ordered = ordered
  )

  return(factor_column)
}
```

```{r factor-demographics}

# convert all categorical / ordinal data to factors

demo_f <- demographics |>
  mutate(
    # categorical, levels != labels
    # unordered with predefined levels
    gender = to_factors(
      gender,
      levels = c("F", "M", "O"),
      labels = c("Female", "Male", "Non-Binary")
    ),
    race = to_factors(
      race,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "American Indian or Alaska Native",
        "Asian",
        "Black or African American",
        "Native Hawaiian or Other Pacific Islander",
        "White",
        "More than one race",
        "Unknown or not reported"
      )
    ),
    ethnicity = to_factors(
      ethnicity,
      levels = c("1", "2", "3"),
      labels = c(
        "Hispanic or Latino",
        "Not Hispanic or Latino",
        "Unknown or not reported"
      )
    ),
    lenses = to_factors(
      lenses,
      levels = c("G", "C", "N", "Y"),
      labels = c(
        "Glasses",
        "Contacts",
        "No",
        "Yes"
      )
    ),

    # categorical, levels == labels
    # unordered with predefined levels
    language = to_factors(
      language,
      levels = c("English", "Other"),
    ),
    worn = to_factors(
      worn,
      levels = c("Y", "N")
    ),
    color_blind = to_factors(
      color_blind,
      levels = c("Y", "N")
    ),
    condition = to_factors(
      condition,
      levels = c("Y", "N")
    ),

    # ad-hoc
    # categorized in transcription, no predefined levels
    country = to_factors(country),
    college = to_factors(college),
    program = to_factors(program),
    referral = to_factors(referral),

    # ordinal, levels != labels
    # ordered with predefined levels
    education = to_factors(
      education,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "Pre-HS", # No HS diploma
        "HS", # High school diploma or equivalent
        "CND", # Some college but no degree
        "AS",
        "BS",
        "MS",
        "PhD"
      ),
      ordered = TRUE
    ),
    lego = to_factors(
      lego,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "Little to no experience",
        "Some experience",
        "Lots of experience",
        "Expert builder"
      ),
      ordered = TRUE
    ),
    manuf = to_factors(
      manuf,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "No experience",
        "Taken one or more classes",
        "Held a part time or temporary position",
        "One or more year’s experience"
      ),
      ordered = TRUE
    ),
    runner = to_factors(
      runner,
      levels = c("N", "Y", "5", "10", "H", "M", "T"),
      labels = c(
        "None",
        "Yes - Unspecified",
        "5K",
        "10K",
        "Half-Marathon",
        "Marathon",
        "Triathalon"
      ),
      ordered = TRUE
    ),
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    )
  )
```

#### Numeric Data

```{r numeric-data}
nums <- demo_f |>
  select(where(is.numeric))

height_stats <- nums |>
  summarize(
    mean = mean(height, na.rm = TRUE),
    sd = sd(height, na.rm = TRUE),
    min = min(height, na.rm = TRUE),
    max = max(height, na.rm = TRUE),
    pct_missing = sum(is.na(height)) / n() * 100
  )

height_summary <- str_glue(
  "Mean height = {round(height_stats$mean, 1)}, SD = {round(height_stats$sd, 1)}, range: [{height_stats$min}, {height_stats$max}], {round(height_stats$pct_missing, 1)}% missing"
)
```

After excluding data from the pilot study, `r report_participants(nums, age = "age", participants = "participant")` were included in this analysis. Other than age, height (`r height_summary`) was the only numeric data collected. As discussed in @sec-res-challenges, height was first recorded following the pilot study, when its impact on user behavior and results was first observed. Both numeric characteristics are visualized with a combined box and scatter plot in @fig-numeric. Where present, the notches of these box plots show the 95% confidence interval around the median.

```{r fig-numeric}
#| fig-cap: "Distribution of Participant Age and Height"
#| fig-height: 3.5
#| warning: false

# Plotting age
age_plot <- demo_f |>
  filter(!is.na(age)) |>
  ggplot(aes(x = age)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    x = "Age (yr)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting height
height_plot <- demo_f |>
  filter(!is.na(height)) |>
  ggplot(aes(x = height)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    width = 0.55,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    x = "Height (in)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
age_plot + height_plot +
  # plot_layout(ncol = 2, widths = c(1,0.8)) +
  plot_annotation(
    title = "Numeric Sample Characteristics",
    subtitle = "Age reflects the student-oriented nature of the sample"
  )
```

#### Categorical Data

```{r categorical-data}
cats <- demo_f |>
  select(where(~ is.factor(.) && !is.ordered(.))) |>
  select(-c("referral")) |>
  mutate(across(everything(), \(x) fct_lump(x, prop = 0.1)))

# Summary table
cat_tbl <- cats |>
  select(gender:language) |>
  tbl_summary(
    label = setNames(str_to_title(names(cats)), names(cats)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

cat_names <- cats |>
  select(gender:country) |>
  names()

cats_long <- cats |>
  select(gender:country) |>
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "Value"
  ) |>
  drop_na() |>
  group_by(Category, Value) |>
  summarize(Count = n(), .groups = "drop") |>
  group_by(Category) |>
  mutate(Freq = Count / sum(Count))

country_levels <- demo_f |>
  filter(country != "USA") |>
  count(country) |>
  arrange(desc(n))
```

Of the sample's `r inline_text(cat_tbl, language, level = "English", pattern = "{N} participants, {p}% (n={n})")` reported that they mainly speak English at home. As shown in @fig-categorical, the gender composition was relatively balanced, with slightly more males `r inline_text(cat_tbl, gender, level = "Male", pattern = "(n={n}, {p}%)")` than females `r inline_text(cat_tbl, gender, level = "Female", pattern = "(n={n}, {p}%)")`. The racial makeup of the sample was predominantly White `r inline_text(cat_tbl, race, level = "White", pattern = "(n={n}, {p}%)")` and Asian `r inline_text(cat_tbl, race, level = "Asian", pattern = "(n={n}, {p}%)")`, with smaller representations from `r as.character(english::english(n_distinct(demo_f$race) - 2))` other groups. Fully `r inline_text(cat_tbl, ethnicity, level="Not Hispanic or Latino", pattern = "{p}% of participants (n={n})")` reported a non-Hispanic or Latino ethnicity. The majority of participants `r inline_text(cat_tbl, country, level = "USA", pattern = "(n={n}, {p}%)")` were from the United States, though `r as.character(english::english(n_distinct(demo_f$country, na.rm = TRUE) - 1))` other countries were represented, including `r paste(country_levels[1, ]$country, " (n=", country_levels[1, ]$n, ")", sep = "")` and `r paste(country_levels[2, ]$country, " (n=", country_levels[2, ]$n, ")", sep = "")`. `r paste(country_levels$country[3:(nrow(country_levels) - 1)], collapse = ", ")`, and `r tail(country_levels$country, 1)` were also listed as countries of origin by one participant each (n=1). All of this aligns with the study's recruitment focus.

```{r fig-categorical}
#| fig-cap: "Distribution of Categorical Characteristics"
#| fig-height: 3.5
#| results: hide

create_cat_plot <- function(data, category, base_color) {
  # specify alpha ramp end points
  min_alpha <- 0.3
  max_alpha <- 0.8

  # Filter and reorder data
  plot_data <- data |>
    filter(Category == category) |>
    mutate(
      Value = if_else(Value == "Not Hispanic or Latino",
        "Not Hispanic\nor Latino", Value
      ),
      Value = reorder(Value, Count, FUN = sum, decreasing = TRUE)
    )

  # Calculate alpha values based on Count linearly between min_alpha and max_alpha
  alpha_values <- plot_data |>
    mutate(alpha = scales::rescale(Count, to = c(min_alpha, max_alpha))) |>
    pull(alpha)

  # Set unique Value groups for setting colors and alphas
  unique_values <- unique(plot_data$Value)

  # Create the plot
  ggplot(
    plot_data,
    aes(x = Value, y = Count, fill = I(base_color), alpha = Value)
  ) +
    geom_bar(stat = "identity", position = "dodge", width = 0.8, color = outline_color) +
    scale_x_discrete(labels = function(x) tools::toTitleCase(x)) +
    labs(x = tools::toTitleCase(category), y = element_blank()) +
    scale_alpha_manual(values = setNames(alpha_values, unique_values)) +
    guides(fill = "none", alpha = "none") +
    ylim(0, max_count) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
}

# Calculate the maximum count to standardize the y-axis
max_count <- max(cats_long$Count)

# Create all plots
cat_plots <- map(
  cat_names,
  ~ create_cat_plot(cats_long, .x, base_color)
)

# Add y-axis text and ticks to first plot
first_plot <- cat_plots[[1]] +
  theme(
    axis.text.y = element_text(),
    axis.ticks.y = element_line()
  )

# Other plots without y-axis text and ticks, drop first
other_plots <- cat_plots[-1]

# combine!
all_plots <- c(list(first_plot), other_plots)

# Combine all plots
combined_plot <- reduce(all_plots, `+`) +
  plot_layout(ncol = length(all_plots), widths = c(1, 1, 0.667, 0.667)) +
  plot_annotation(
    title = "Categorical Sample Characteristics",
    subtitle = "Value Count for Gender, Race, Ethnicity, and Country"
  )

# Print the combined plot
print(combined_plot)
```

```{r make-vc-table}
vc_tbl <- cats |>
  select(lenses:worn) |>
  filter(!lenses %in% c("No", "Other", "Unknown")) |>
  filter(!(is.na(lenses) & is.na(worn))) |>
  droplevels() |>
  mutate(worn = recode(worn, "Y" = "Yes", "N" = "No")) |>
  tbl_cross(
    row = lenses,
    col = worn,
    label = list(lenses = "Lenses", worn = "Worn During Session"),
    percent = "row",
    missing_text = "NA"
  ) |>
  add_stat_label() |>
  bold_labels()
```

Of the 54 participants, 52% (n=28) reported having a vision prescription. Of those, `r inline_text(vc_tbl, row_level = "Contacts", col_level = "Total")` were for contact lenses , but `r inline_text(vc_tbl, row_level = "Glasses", col_level = "Total")` had glasses. Whereas all contact wearers reported they would wear them during the session, among glasses wearers only seven (47%) planned the same. Five (33%) did not intend to do so, and three (20%) gave no indication. This breakdown is summarized by @tbl-vc. Two participants also reported color-blindness (n=1) or other vision conditions (n=1). No participants indicated any other condition that might affect their performance.

::: content-hidden
MAYBE: make this all inline code; rethink percent by row, col, or cell? format n= etc. as markdown?
:::

```{r tbl-vc}
#| tbl-cap: Vision Correction Table

vc_tbl |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

Following the primary hypothesis testing, additional analyses may explore the differences in key outcomes across these demographic subgroups. That effort will aim to provide additional context and insights into potential underlying factors influencing the findings.

#### Ordinal Data

```{r ord-data}
ords <- demo_f |>
  select(where(~ is.factor(.) && is.ordered(.)))

# Summary table
ords_tbl <- ords |>
  tbl_summary(
    label = setNames(str_to_title(names(ords)), names(ords)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

# ords_tbl

# Helper function to calculate mode
calculate_mode <- function(data, var_name) {
  count_table <- table(data[[var_name]])
  max_freq <- max(count_table)
  # pick first if more than one value has max count
  mode_value <- names(count_table[count_table == max_freq])[1]

  # Convert to an output that includes the level's name (mode_value) and its index
  mode_index <- which(levels(data[[var_name]]) == mode_value)
  return(list(value = mode_value, index = mode_index))
}

# Main function to calculate statistics and return formatted string
calculate_ord_summary_stats <- function(data, var_name) {
  # Ensure the variable is an ordered factor
  if (!is.ordered(data[[var_name]])) {
    stop("The variable must be an ordered factor.")
  }

  # Convert factor to integer codes
  factor_int <- as.integer(data[[var_name]])

  # Calculate median, mode, and IQR
  median_value <- median(factor_int, na.rm = TRUE)
  mode_results <- calculate_mode(data, var_name)
  iqr_value <- IQR(factor_int, na.rm = TRUE)

  # Format and return the result as a string
  result_string <- sprintf(
    "(Median = %.1f; Mode = %d, %s; IQR = %.1f)",
    median_value, mode_results$index, mode_results$value,
    iqr_value
  )
  return(result_string)
}
```

Unlike the data above, the categories analyzed in this section are inherently ordered. Responses for each participant's current Education level `r calculate_ord_summary_stats(ords, "education")` and their experience with Lego `r calculate_ord_summary_stats(ords, "lego")` and Manufacturing `r calculate_ord_summary_stats(ords, "manuf")` are visualized in increasing order by @fig-ordinal. The results are mostly in line with expectations based on the recruitment, though it is disappointing to see that `r inline_text(ords_tbl, lego, level = "Little to no experience", pattern = "{p}% of participants (n={n})")` claimed to have little to no experience with Lego. The great majority were undergraduate students working towards a degree `r inline_text(ords_tbl, education, level = "CND", pattern = "(n={n}, {p}%)")`, with no experience in manufacturing `r inline_text(ords_tbl, manuf, level = "No experience", pattern = "(n={n}, {p}%)")`.

```{r fig-ordinal}
#| fig-cap: "Distribution of Ordinal Characteristics"
#| fig-height: 3.5
#| results: hide

plot_ords <- function(data, var, title) {
  norm_data <- data |>
    count(var_name = {{ var }}) |>
    mutate(percent = n / sum(n) * 100)

  p <- norm_data |>
    ggplot(aes(y = var_name, x = percent)) +
    geom_bar(
      aes(alpha = percent / 100),
      fill = base_color,
      stat = "identity",
      width = 0.75,
      color = outline_color
    ) +
    scale_alpha_continuous(trans = "log10", range = c(0.2, 0.8), guide = "none") +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank()
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(size = 9, face = "bold"),
      legend.position = "none",
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
  return(p)
}


educ_plot <- ords |>
  mutate(education = fct_recode(education,
    College = "CND",
    Associate = "AS",
    Bachelor = "BS",
    Master = "MS",
    Doctorate = "PhD"
  )) |>
  plot_ords(var = education, title = "Current Education Level")

lego_plot <- ords |>
  mutate(lego = fct_recode(lego,
    Little = "Little to no experience",
    Some = "Some experience",
    Lots = "Lots of experience",
    Expert = "Expert builder"
  )) |>
  plot_ords(lego, title = "Lego Experience")

manuf_plot <- ords |>
  mutate(manuf = fct_recode(manuf,
    None = "No experience",
    Training = "Taken one or more classes",
    PartTime = "Held a part time or temporary position",
    Some = "One or more year’s experience"
  )) |>
  plot_ords(manuf, title = "Manufacturing Experience")

# combine using patchwork
educ_plot / (lego_plot + manuf_plot) +
  # plot_layout(ncol = 2) +
  plot_annotation(
    title = "Ordinal Sample Characteristics",
    subtitle = "Reflects the Focus of Recruiting Efforts"
  )
```

### Multivariate Analysis

In order to assess the equivalence of treatment groups and validate the random assignment that is assumed by most analysis to follow, Age, Gender, Lego experience, and Education were compared across treatment groups. These variables were selected for their meaningful variation and potential relevance to the outcomes of interest. Other notable variables, including Ethnicity and Country of Origin were considered less critical due to their skewed distributions and small subgroup sizes.

```{r make-mv-table}

num_cols <- demographics |>
  select(age)

cat_cols <- demo_f |>
  select(c("gender", "lego", "education", "treatment"))

mv_data <- bind_cols(num_cols, cat_cols)

mv_tbl <- mv_data |>
  mutate(across(where(is.factor), ~ fct_drop(.))) |>
  mutate(
    lego = fct_recode(lego,
      "Little" = "Little to no experience",
      "Some" = "Some experience",
      "Lots" = "Lots of experience",
      "Expert" = "Expert builder"
    ),
    education = fct_recode(education,
      "College" = "CND",
      "Associate" = "AS",
      "Bachelor" = "BS",
      "Master" = "MS",
      "Doctorate" = "PhD"
    )
  ) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    label = list(
      "age" ~ "Age",
      "gender" ~ "Gender",
      "lego" ~ "Lego",
      "education" ~ "Education"
    )
  ) |>
  add_p() |>
  # separate_p_footnotes() |>
  # add_stat_label() |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  italicize_levels() |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nN = {n} ({style_percent(p)}%)",
    p.value ~ "**p-val**"
  )

```

@tbl-demo-by-treatment shows the result of this analysis. The Kruskal-Wallis rank sum test was utilized for the numeric variable (Age) to compare distributions across groups, while Fisher's exact test was applied to the categorical and ordinal variables (Gender, Education, and Lego experience) to assess the independence of distributions from treatment assignments.

```{r tbl-demo-by-treatment}
#| tbl-cap: Participant Demographics by Treatment Group

mv_tbl |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )

```

::: content-hidden
MAYBE: wrap gtsummary calls in a function that saves an svg to project
and inserts embedding so that all original formatting is preserved?
could they still be labeled as tables?
:::

The high p-values for Age, `r inline_text(mv_tbl, variable = age, column = "p.value")`; Gender, `r inline_text(mv_tbl, variable = gender, column = "p.value")`; Lego, `r inline_text(mv_tbl, variable = lego, column = "p.value")`; and Education, `r inline_text(mv_tbl, variable = lego, column = "p.value")` indicate that these observed characteristics are not strongly associated with treatment groupings. This result validates the group randomization effort, mitigating the influence of confounding variables, and reducing bias in the data. It is essential to the validity of most of the statistical tests employed by this study, which assume sample independence.

::: content-hidden
MAYBE: further analysis to show that:
- nums
  - height (0.11)
- cats
  - country (0.085)
  - lenses (0.063)
- ords
  - manuf exp (0.046)

may be significant (p-val)
used the following code to test all demographics

``` text
# test significance of treatment groups
trt <- demo_f |>
  select(treatment)

# bind treatment for nums, ords (included in cats)
cd <- bind_cols(ords, trt)

cd |>
  tbl_summary(
      by = "treatment",
      missing = "no",
      statistic = list(
        all_continuous() ~ "{mean} ({sd})"
      )
    ) |>
    add_p()
```
:::

## Learning Phase Results

Only completed tasks, the primary result of interest, are considered in this analysis. While excluding incomplete tasks may limit the generalizability of the findings to scenarios where time constraints are common, it ensures the analysis is based on objective, directly observed data and avoids introducing assumptions or biases that could arise from predicting results for unfinished tasks. 

```{r factor-car-results}

res_f <- car_results |>
  mutate(
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    ),
    car_outcome = to_factors(car_outcome)
  )
```

```{r save-results-data}
# saving for use in broken out recall and retention analysis files
save(res_f, file = "res_f.RData")
```

```{r learn-data}
learn <- res_f |>
  filter(
    phase == 1,
    car_outcome == "complete",
    # hack to remove incorrectly labeled car
    uid != "1048-P1-C5"
  ) |>
  select(
    participant,
    car_elapsed_time,
    uce,
    treatment,
    drop_event_count,
    drop_event_tot_time_adj,
    repair_event_count,
    repair_event_tot_time,
    system_event_count,
    system_event_tot_time
  )

# add tot_sys_adj, tot_usr_adj, and adjusted TCT time to learn
# tot_sys_adj will be used to calculate OEE
# tot_usr_adj just to reduce TCT

learn <- learn |>
  mutate(
    # Replace NA with 0 using coalesce for safe addition
    tot_sys_adj = coalesce(drop_event_tot_time_adj, 0) +
      coalesce(system_event_tot_time, 0),
    tot_usr_adj = coalesce(repair_event_tot_time, 0),
    tot_adj = tot_sys_adj + tot_usr_adj,
    car_elapsed_time_adj = car_elapsed_time - tot_adj
  )

changes <- learn |>
  filter(car_elapsed_time != car_elapsed_time_adj)

total_tasks_completed <- dim(learn)[1]
total_tasks_adjusted <- dim(changes)[1]
```

One final step of data preparation was taken to account for drop-outs and other system events, along with time spent making repairs. This lost time was deducted from measured task times to give a more accurate record of participant performance. `r total_tasks_adjusted` of `r total_tasks_completed` completed tasks (`r round(total_tasks_adjusted / total_tasks_completed * 100, 1)`%) were affected by this adjustment.

::: content-hidden
MAYBE: consider better quantifying the effect of these adjustments
:::

### Descriptive Statistics for Granular Data

```{r make-report}
# report descriptive stats for cet and uce
learn_rpt <- learn |>
  rename(
    "Task Completion Time (sec)" = car_elapsed_time_adj,
    "Uncorrected Error Count" = uce
  ) |>
  select("Task Completion Time (sec)", "Uncorrected Error Count") |>
  report::report() # options not working!

rpt_table <- as.report_table(learn_rpt)
cv_tct <- rpt_table$SD[1] / rpt_table$Mean[1]
cv_uce <- rpt_table$SD[2] / rpt_table$Mean[2]
```

::: content-hidden
TODO: correct data for `1048-P1-C5`; flagged complete with elapsed time of 4.267s
TODO: investigate report::report options
:::

`r learn_rpt`

These statistics provide a high-level summary of performance across all tasks and participants, where each observation corresponds to a single repetition of the Learning task. Both Task Completion Time (TCT) and Uncorrected Error Count (UCE) show considerable variance, as depicted in @fig-learn.

```{r fig-learn}
#| fig-cap: "Distribution of Task Completion Time and Error Rates in Learning"
#| fig-height: 3.5
#| warning: false

# Plotting adjusted car_elapsed_time
cet_plot <- learn |>
  ggplot(aes(x = car_elapsed_time_adj)) +
  geom_histogram(
    aes(y = ..density.. * 65), # hack for relative size / scale
    binwidth = 10,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    title = "Task Completion Time (sec)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting Uncorrected Error Counts
uce_plot <- learn |>
  ggplot(aes(x = uce)) +
  geom_histogram(
    aes(y = ..density.. * 1.5), # hack for relative size / scale
    binwidth = 1,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    title = "Number of Uncorrected Errors per Task",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
cet_plot / uce_plot +
  # plot_layout(ncol = 2, widths = c(1,1)) +
  plot_annotation(
    title = "Learning Phase - Observed Measures",
    subtitle = "Error Count Exhibits Greater Variability than Task Completion Time"
  )
```

Compared to TCT, with a coefficient of variation (CV) of `r round(cv_tct, 2)`, UCE shows substantially greater spread (CV=`r round(cv_uce, 2)`) and the presence of a floor effect. The later could potentially suggest limitations in the task's ability to capture a full range of participant performance. Further investigation is required to determine if this is due to task complexity or other factors.

::: content-hidden
MAYBE: discuss overdispersion of data, where sd^2 / mean > 1; this is true for both TCT and UCE - implications? better metrick than CV? more appropriate later in report, after agg / outliers?
:::

### Aggregated Data

```{r agg-results}
# agg participant results
learn_agg <- learn |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    tot_cars = n(), # Count the number of tasks per participant
    avg_tct_adj = mean(car_elapsed_time_adj, na.rm = TRUE),
    avg_uce = mean(uce, na.rm = TRUE)
  ) |>
  ungroup()
```

::: content-hidden
MAYBE: may need to add back other variables from learn into this agg table
:::

@tbl-agg-results-by-treat presents participant-level metrics, including the number of tasks completed along with the average TCT and UCE per task. In addition to the overall results, the data is broken out by assigned treatment group to allow for comparisons across conditions. Note that the statistics presented are the overall (aka "grand") mean and median values. That is, they are the average or middle values of all individual participant averages. This provides overall summary statistics for each treatment group that reflect the central tendency and variability among the average outcomes reported by participants within each group.

```{r tbl-agg-results-by-treat}
#| tbl-cap: Avg. Participant Performance Data Summarized by Treatment Group

t <- learn_agg |>
  select(-participant) |>
  tbl_summary(
    type = all_continuous() ~ "continuous2",
    digits = all_continuous2() ~ 1,
    label = list(
      avg_tct_adj ~ "Avg TCT per Task (sec)",
      avg_uce ~ "Avg UCE per Task",
      tot_cars ~ "Number of Tasks Completed"
    ),
    by = treatment,
    statistic = list(
      all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} ({p25}, {p75})"
      )
    )
  ) |>
  # bold_labels() |>
  # add_p() |>
  # separate_p_footnotes() |>
  add_overall() |>
  # bold_p(t = 0.05) |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nn = {n} ({style_percent(p)}%)",
    # p.value ~ "**p-val**"
  )

t |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

See @fig-perf-by-treat for a depiction of the grand means for all variables, with 95% confidence intervals for each. Observe that the previously noted variability and positive skew are also present in the grouped data, and some treatments like PWI may have prioritized speed over accuracy, or vice versa (e.g. AR).

```{r make-perf-data}
# Reshape data to long format
long_data <- learn_agg |>
  pivot_longer(
    cols = c(avg_tct_adj, tot_cars, avg_uce),
    names_to = "metric",
    values_to = "value"
  )

# Calculate means, sample size, and SE for each metric within each treatment
metrics_stats <- long_data |>
  group_by(treatment, metric) |>
  summarize(
    mean_value = mean(value, na.rm = TRUE),
    n = n(), # sample size
    se = sd(value, na.rm = TRUE) / sqrt(n()), # standard error
    .groups = "drop"
  )

# Find the highest mean for each metric across all treatments
max_means <- metrics_stats |>
  group_by(metric) |>
  summarize(max_mean = max(mean_value), .groups = "drop")

# Normalize means and calculate CI within their metrics
normalized_metrics_stats <- metrics_stats |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = mean_value / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  )

# Calculate overall mean, SE, and normalize
overall_stats <- long_data |>
  group_by(metric) |>
  summarize(
    overall_mean = mean(value, na.rm = TRUE),
    n = n(),
    se = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = overall_mean / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  ) |>
  mutate(treatment = "ALL")

# Combine with other stats
final_stats <- bind_rows(normalized_metrics_stats, overall_stats)

# Factorize the 'treatment' column
final_stats$treatment <- factor(final_stats$treatment, levels = c("ALL", "PWI", "PAR", "AR", "MR"))

# Factorize and label 'metric' column
final_stats$metric <- factor(
  final_stats$metric,
  levels = c("avg_tct_adj", "avg_uce", "tot_cars"),
  labels = c("Average Task Time", "Average Error Count", "Tasks Completed")
)

# Predefined alpha values by treatment
alpha_values <- c(0.4, 0.55, 0.7, 0.85, 1.0)
names(alpha_values) <- levels(final_stats$treatment)
```


```{r fig-perf-by-treat}
#| fig-cap: "Aggregated Learning Performance Data by Treatment"
#| fig-height: 3.5
#| warning: false

ggplot(final_stats, aes(x = metric, y = normalized_mean, fill = treatment, label = treatment)) +
  geom_col(
    aes(alpha = normalized_mean),
    position = position_dodge(0.9),
    color = outline_color
  ) +
  geom_errorbar(
    aes(ymin = lower_ci, ymax = upper_ci),
    position = position_dodge(0.9),
    width = 0.25,
    alpha = 0.7
  ) +
  geom_text(
    aes(y = 0),
    position = position_dodge(width = 0.9),
    vjust = 1.65, # Adjust based on visual preference
    size = 3,
    alpha = 0.9
  ) +
  scale_fill_manual(values = rep(base_color, length(unique(final_stats$treatment)))) +
  scale_alpha_continuous(transform = "pseudo_log", range = c(0.3, 1.0)) +
  labs(
    title = "Comparison of Normalized Performance Metrics by Treatment",
    subtitle = "Includes Overall and Treatment Groups with 95% Confidence Intervals",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(
      margin = margin(b = 5, unit = "pt"),
      color = "black"
    )
  )

```

#### Assess outliers

Given the level of variation and spread seen above, the presence of outliers was analyzed. Participant-level outliers within treatment groups were the focus, rather than individual data points across the entire dataset. This method helps ensure that outliers are identified relative to the specific context of each treatment, accounting for the corresponding variability.

Four different tests were used to assess each participant as an outlier in their treatment group: (1) Tukey's Fences at 1.5 times the IQR, (2) Tukey's Fences at 3.0 times the IQR, (3) outside the 5th and 95th percentiles, and (4) using the Mahalanobis distance. TCE and UCE were each tested separately using the first three methods. The fourth method is a multivariate test that considers TCE and UCE together. The results are summarized in @tbl-outlier-summary, for all participants that were identified by two or more tests.

> A helpful heuristic is to distinguish between error outliers, interesting outliers, and random outliers (Leys et al. 2019). Error outliers are likely due to human error and should be corrected before data analysis. Interesting outliers are not due to technical error and may be of theoretical interest; it might thus be relevant to investigate them further even though they should be removed from the current analysis of interest. Random outliers are assumed to be due to chance alone and to belong to the correct distribution and, therefore, should be retained.

```{r find-outliers}
# Function to detect outliers in a vector
find_outliers <- function(x, mult) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower_bound <- q1 - mult * iqr
  upper_bound <- q3 + mult * iqr
  x < lower_bound | x > upper_bound
}

# Function to calculate Mahalanobis distance for a group
compute_mahalanobis <- function(df, cols) {
  cov_matrix <- cov(df[, cols], use = "complete.obs")
  center <- colMeans(df[, cols], na.rm = TRUE)
  mahal_dist <- mahalanobis(df[, cols], center, cov_matrix)
  return(mahal_dist)
}

outlier_summary <- learn_agg |>
  group_by(treatment) |>
  mutate(
    # find 1.5 and 3x IQR outliers using Tukey method
    out15_tct = find_outliers(avg_tct_adj, 1.5),
    out15_uce = find_outliers(avg_uce, 1.5),
    out30_tct = find_outliers(avg_tct_adj, 3.0),
    out30_uce = find_outliers(avg_uce, 3.0),
    # find by 5th and 95th percentile
    pct_tct = avg_tct_adj < quantile(avg_tct_adj, 0.05, na.rm = TRUE) |
      avg_tct_adj > quantile(avg_tct_adj, 0.95, na.rm = TRUE),
    pct_uce = avg_uce < quantile(avg_uce, 0.05, na.rm = TRUE) |
      avg_uce > quantile(avg_uce, 0.95, na.rm = TRUE),
    # find by Mahalanobis distance
    mahal_dist = compute_mahalanobis(
      pick(avg_tct_adj, avg_uce),
      c("avg_tct_adj", "avg_uce")
    ),
    md_out = mahal_dist > qchisq(0.975, df = 2)
  ) |>
  ungroup() |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    out15_tct = sum(out15_tct),
    out15_uce = sum(out15_uce),
    out30_tct = sum(out30_tct),
    out30_uce = sum(out30_uce),
    pct_tct = sum(pct_tct),
    pct_uce = sum(pct_uce),
    md_out = sum(md_out),
    .groups = "drop"
  )
```


```{r tbl-outlier-summary}
#| tbl-cap: Summary of Participants Identified by Two or More Outlier Tests
#| warning: false

outlier_summary |>
  rowwise() |>
  mutate(tot_out = sum(c_across(-c(participant, treatment)))) |>
  ungroup() |>
  filter(tot_out > 1) |>
  arrange(desc(tot_out)) |>
  select(participant, treatment, tot_out) |>
  inner_join(learn_agg) |>
  gt()

```

::: content-hidden
TODO: make this table prettier
:::

Further investigation of these seven participants showed that only 1063 experienced systematic problems that might warrant excluding their performance. The other participants were just relatively poor performers, with high TCT and/or UCE. Given the limited amount of data collected per treatment, it was decided to retain all but 1063's data for hypothesis testing.

```{r drop-learn-outliers}
# drop 1063
learn_agg_out <- learn_agg |> filter(participant != 1063)
```

```{r save-learn-data}
# save learn objects with types, etc. as binary data file for external analysis
save(
  learn, 
  learn_agg_out, 
  outline_color,
  base_color,
  file = "learn.RData")
```


### $H_{1a}$: Average Time per Car

```{r load-h1a-libs}
library(emmeans)
library(performance)
library(lmerTest)
```

The first hypothesis of the learning phase:
$$H_{1a}\textrm{: Average task completion time varies with treatment}$$

will be tested by comparing the average TCT of each treatment group to understand the magnitude, direction, and significance of differences.

#### Applicable Statistical Methods

Methods based on one-way ANOVA, including Fisher's or Welch's parametric methods and their non-parametric equivalent, Kruskal-Wallis' test by ranks, are commonly used for such comparisons. All assume independence of observations additivity of treatments. The use of aggregated data reduces each participant to a single observation, addressing independence. The experimental design generally ensures the treatments themselves are additive, with independent effects on the response.

Parametric methods also require a normal distribution of data *within each treatment group*. As noted in the original assessment of the learning data, the overall TCT dataset exhibits skewness and kurtosis, suggesting deviations from a normal distribution. This must be revisited with group wise testing of the aggregated data. By averaging the observations and eliminating an outlier, the grouped data may better resemble a normal distribution.

#### Check Model Assumptions

A combination of quantitative and qualitative analysis is required to accurately assess normality of each treatment group. Statistical tests provide a formal but imperfect measure of the data's shape. Visual confirmation of the characteristic quantile-quantile (Q-Q) plot provides further support for the claim. @tbl-learn-norm-tct tabulates the results of the D'Agostino skewness test, Anscombe-Glynn kurtosis test, and the Shapiro-Wilk normality test. In each case, the null hypothesis ($H_0$) for these tests assumes that the data follows the characteristics of a normal distribution. Therefore, low p-values indicate evidence against $H_0$, suggesting that the data deviates from normality in the tested aspect.

```{r def-tab-norm}
# Function to tabulate normality test results for each treatment group
tabulate_normality_tests <-
  function(data, treatment_col, value_col) {
    levels <- levels(data[[treatment_col]])
    all_levels <- c("ALL", levels)

    results <- map(all_levels, function(level) {
      if (level == "ALL") {
        data_subset <- data |> pull(!!sym(value_col))
      } else {
        data_subset <-
          data |>
          filter((!!sym(treatment_col)) == level) |>
          pull(!!sym(value_col))
      }

      if (length(data_subset) > 0) {
        # Ensure there's data to process
        skew_test <- moments::agostino.test(data_subset)
        kurtosis_test <- moments::anscombe.test(data_subset)
        normality_test <- shapiro.test(data_subset)

        tibble(
          Group = level,
          Skewness = paste(
            round(skew_test$statistic[1], 2),
            round(skew_test$p.value, 3),
            sep = ", "
          ),
          Kurtosis = paste(
            round(kurtosis_test$statistic[1], 2),
            round(kurtosis_test$p.value, 3),
            sep = ", "
          ),
          Normality = paste(
            round(normality_test$statistic, 2),
            round(normality_test$p.value, 3),
            sep = ", "
          )
        )
      } else {
        tibble(
          Group = level,
          Skewness = NA,
          Kurtosis = NA,
          Normality = NA
        )
      }
    }) |> bind_rows()

    # Ensure the desired order, converting Group to a factor
    results <- results |>
      mutate(Group = factor(Group, levels = c("ALL", "MR", "AR", "PWI", "PAR"))) |>
      filter(!is.na(Group)) |>
      arrange(Group)

    results
  }
```

```{r tbl-learn-norm-tct}
#| tbl-cap: "Summary of TCT Test Statistics, p-Values for Normality Assessment"
#|
tct_norm <- tabulate_normality_tests(learn_agg_out, "treatment", "avg_tct_adj")
tct_norm |> gt()
```

From this we see that only PWI fails the normality test (p=0.021), despite showing marginally non-significant skewness (p=0.058). Similarly, MR exhibits significant positive skew (p=0.021) but the Shapiro-Wilk test gives evidence of normality (p=0.084). Both AR and PAR are consistent in confirming normality, though PAR's slightly positive skew is marginally non-significant (p=0.068).

::: content-hidden
TODO: eliminate hard-coded values in text
:::

To reconcile these findings for MR and PWI, and confirm those for AR and PAR, the set of four Q-Q plots in @fig-qq-h1a were generated. Observations for normal data should fall along or near the line, within the shaded confidence interval. These plots show that PAR and AR both meet the expectations of normality. The MR group mostly aligns with the line, except for a single point in the right tail. This aligns with the skewness and kurtosis test results and provides insufficient evidence to reject normality. On the other hand, the PWI group diverges substantially from the line, with several points leaving the confidence interval. This supports the statistical evidence that the PWI data is not normally distributed.

```{r fig-qq-h1a}
#| fig-cap: Q-Q Plots of Learning Times by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(learn_agg_out, x_var = "avg_tct_adj", by_var = "treatment")

p + plot_annotation(title = "Quantile-Quantile Plots of TCT Treatment Data",
                    subtitle = "PWI Data is the Most Uncharacteristic of Normality")

```

Given the minor deviations from normality, particularly for the PWI group, a non-parametric approach was adopted. This provides a reliable and straightforward approach for hypothesis testing for data that are non-normal or heteroscedastic (having unequal variances). This decision also avoids the complexities associated with transformation pipelines, ensuring robust statistical results while maintaining the interpretability of the original data.

#### Analysis

@fig-h1a-plot shows the resulting comparison of average TCT within each treatment group. It combines elements of violin, scatter, and box plots to provide a comprehensive portrayal of the data. Treatment differences were analyzed using the Kruskal-Wallis test by ranks. Its test statistic, $H$, approximately follows a $\chi^{2}_{k-1}$ distribution, where $k$ is the number of groups. Here, $\chi^{2}_{Kruskal-Wallis}(3)=25.9$, with a p-value of 1.0e-5. These results indicate a statistically significant difference between groups. The effect size of this difference is $\hat{\epsilon}^{2}_{rank}=0.5$, with $CI_{95\%}\:[0.3,1.0]$, indicating a moderate to large treatment effect where 50% of the variance in TCT can be explained by the difference in groups. Together, these results provide strong evidence of overall differences in TCT that are both statistically and practically significant.

::: content-hidden
TODO: fix hard-coded numbers
:::

```{r fig-h1a-plot}
#| fig-cap: Task Completion Time Differences by Treatment
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

h1a_plot <- ggstatsplot::ggbetweenstats(
  data = learn_agg_out,
  x = treatment,
  y = avg_tct_adj,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "TCT Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

h1a_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "25.9", italic(p) == 
#     "1.0e-05", widehat(epsilon)["ordinal"]^2 == "0.5", CI["95%"] ~ 
#     "[" * "0.3", "1.0" * "]", italic("n")["obs"] == "53")
```

::: content-hidden
TODO: check all in text values after setting standard seed before all ggbetween stats calls - see comment in code above
:::

Pairwise comparisons were conducted post-hoc using Dunn's non-parametric test for Kruskal-type ranked data. Holm's adjustment for multiple comparisons was preferred over the often cited Bonferroni method, which tends to be too conservative. The results show that PWI and PAR are both significantly faster than AR and MR, with all p-values less than 0.01.

```{r test-power-sim}
k <- length(unique(learn_agg_out$treatment))
n <- round(nrow(learn_agg_out) / k)
epsilon_sq <- 0.5

# Estimate post-hoc power
posthoc_power <- posthoc_power_analysis(n, k, epsilon_sq, sig_level = 0.05, num_simulations = 1000)
```

A post-hoc simulation was run to validate these results by estimating the statistical power. Power analysis measures the effectiveness of a statistical test in detecting true differences when they exist. High power (typically 0.8 or above) suggests that the test is appropriate for the given data and experimental conditions, therefore supporting factual decision-making. The simulation repeatedly applies the Kruskal-Wallis test to data generated based on the observed mean and effect size, and resulting power is the proportion of tests that correctly reject the null hypothesis. The outcome of `r posthoc_power` (`r posthoc_power * 100`% power), while uncommon, suggests that, for the conditions of this experiment (n=53, an average of 13.25 samples in each of the four groups, $\hat{\epsilon}^{2}=0.5$, and $\alpha = 0.05$), the test is extremely effective at detecting the observed differences. The combination of a large effect size and a reasonable sample size is largely attributable to this value, which offers great confidence in the validity of the test results.

#### $H_{1a}$ Result

Based on this thorough analysis of the data and validation of the methods, it is determined that sufficient statistical evidence exists to accept $H_{1a}$: average task completion time varies by treatment. Specifically, the PWI and PAR instructional methods are both equivalent and faster than AR and MR, which are also equivalent to one another.

### $H_{1b}$: Learning Rates

::: content-hidden
TODO: this is now H1b - change the methods section to match
:::

```{r prep-h1b-data}
# disaggregated data required for rate of change of tct
# create data
learn_h1b <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj)

# mutate
learn_h1b <- learn_h1b |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

#glimpse(learn_h1b)
```

The second hypothesis of the learning phase:
$$H_{1b}\textrm{: Learning rates vary with treatment}$$

will be tested by comparing how TCT changes with each repetition of the task based on treatment. Decreased TCT is used as a measure of the learning effect, so a change in TCT over time is a proxy for learning rate.

#### Applicable Statistical Methods

It is tempting to assess this using aggregated data as before. For example, the average change in TCT per task completed might give suitable measure. For clarity, this can be expressed as:

$$\textstyle{\frac{1}{N} \sum_{i=1}^{N} \frac{TCT_{i1} - TCT_{ij}}{j}}$$ {#eq-naive}

where $N$ is the total number of participants, $j$ is the number of tasks completed by participant $i$, and $TCT_{ij}$ is the TCT of participant $i$ for task $j$. This metric can be compared by group using the approach employed by $H_{1a}$, as seen in @fig-h1b-naive. The results show that there is a significant difference in groups (p < 0.001), with small to moderate effect size $\hat{\epsilon}^{2}_{rank}=0.2$. Significant pairwise differences were identified between PAR and both AR (p=0.029) and MR (p=0.040).

```{r agg-h1b-data}
pd <- learn_h1b |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    first_tct = first(tct),
    last_tct = last(tct),
    num_tasks = n(),
    measure = (first_tct - last_tct) / num_tasks
  ) |>
  ungroup()
```

```{r fig-h1b-naive}
#| fig-cap: Naive method
#| fig-height: 5
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

p <- ggstatsplot::ggbetweenstats(
  data = pd,
  x = treatment,
  y = measure,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "Avg Change in TCT per Task by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

p

```

While this method offers initial insights and highlights significant differences in learning rates across treatment groups, it has crucial limitations. Aggregating data by participant and within treatments masks both between and within-participant variability, discarding important nuances in learning rates. By ignoring the repeated measures nature of the data it fails to capture how learning rate changes over time, and how that differs between individuals. It also does not account for imbalanced repeated measures, which may lead to biased estimates. By reducing the number of observations, it may also limit the statistical power of the test. Finally, this method does not account for the non-linear nature of change in TCT over time, further limiting the validity of these findings. These simplifying assumptions may be reasonable for a gross estimate of overall differences, e.g., average TCT, but they are not appropriate for an in-depth analysis of the rate and change of learning in the presence of individual and treatment effects.

*note: relatively low effect size and wide CI suggests this may not be adequate*

For analysis of this sort, particularly where repeated measures are nested within each group and participant variation is an important consideration, mixed-effects models are often used. These models provide a thorough and accurate analysis by accounting for individual differences in learning rates and leveraging all available repeated measures data. This approach preserves data granularity, explicitly models the dynamic aspect of learning, and is specifically designed to handle imbalanced data. It increases the statistical power and flexibility to detect effects and interactions in a complex data structure, and it can also account for non-linear response data. Overall, mixed-effects models provide a robust framework for examining the interaction between treatment and learning effect over time.

The general form of a suitable mixed-effects model can be expressed as:

$$
\textstyle{
y_{ij}=(\beta_0+U_{0j})+(\beta_1+U_{1j})(SEQ_{ij})+\epsilon_{ij}
}
$$ {#eq-lme1}

where:

- $i$ and $j$ are the indices for each observation and participant, respectively.
- $\beta_0$ is the overall intercept, a baseline value of $y$.
- $U_{0j}$ is the participant variation on the intercept.
- $\beta_1$ is the overall slope.
- $U_{1j}$ is the participant variation on the slope.
- $SEQ_{ij}$ is the time value associated with each observation $i$ of participant $j$.
- $\epsilon_{ij}$ represents the residuals for each observation---variation not accounted for by the model.

This formulation allows for random intercepts, $(\beta_0+U_{0j})$, and slopes $(\beta_1+U_{1j})$, each based on individual differences per participants $U_j$. In the context of this analysis those can be interpreted as individual starting points (intercept) and learning rates (slope). This is visualized in @fig-h1b-rsri, where the black lines indicate the best fit within each group, $(\beta_0+\beta_1(SEQ_{ij}))$, and colored lines show the same for each participant.

```{r fig-h1b-rsri}
#| fig-cap: Random slopes and intercepts model visualized.
#| fig-height: 4
#| warning: false

ggplot(learn_h1b, aes(x = seq, y = tct)) +
  geom_point(
    aes(fill = as.factor(participant)),
    pch = 21,
    size = 1,
    stroke = 0.5
  ) +
  stat_smooth(
    aes(group = participant, col = as.factor(participant)),
    se = FALSE,
    method = "lm",
    lwd = 0.5
  ) +
  stat_smooth(
    aes(group = treatment),
    col = "black",
    se = FALSE,
    method = "lm",
    lwd = 1
  ) +
  facet_wrap( ~ treatment) +
  scale_x_continuous(breaks = c(0:12)) +
  labs(
    title = "Completion Time in Seconds for each Completed Task",
    subtitle = "Linear fits for participants (color) vary from group averages (black)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_bw(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
    )
```

learning is inherently non-linear, exponential decay
try curvilinear approximation - linear model with quadratic term

```{r fig-h1b-qrsri}
#| fig-cap: Quadratic random slopes and intercepts model visualized.
#| fig-height: 4
#| warning: false

learn_h1b |>
  # filter(treatment == "MR") |>
ggplot(aes(x = seq, y = tct)) +
  geom_point(
    aes(fill = as.factor(participant)),
    pch = 21,
    size = 1,
    stroke = 0.5
  ) +
  stat_smooth(
    aes(group = participant, col = as.factor(participant)),
    formula=y~x+I(x^2),
    se = FALSE,
    method = "lm",
    lwd = 0.5
  ) +
  stat_smooth(
    aes(group = treatment),
    formula=y~x+I(x^2),
    col = "black",
    se = FALSE,
    method = "lm",
    lwd = 1
  ) +
  facet_wrap( ~ treatment) +
  scale_x_continuous(breaks = c(0:12)) +
  labs(
    title = "Completion Time in Seconds for each Completed Task",
    subtitle = "Linear fits for participants (color) vary from group averages (black)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_bw(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
    )
```

#### Unconditional Models of Time

first fit a model that is independent of the predictor (treatment); depends only on time (sequence) - try sequence of models with increasing complexity

have to use Max Likelihood (not REML) in order to compare results

```{r h1b-fit-uncond}
# transform seq to float from 0 to 1
md <- learn_h1b |>
  mutate(seq.fp = (seq - 1) / (max(seq) - 1))

# summary(md$seq.fp)

# random intercept
m1 <- lmer(tct ~ 1 + (1 | participant), data = md, REML = F)
# fixed slope, random intercept
m2 <- lmer(tct ~ 1 + seq.fp + (1 | participant), data = md, REML = F)
# random slope, random intercept
m3 <- lmer(tct ~ 1 + seq.fp + (1 + seq.fp | participant), data = md, REML = F)
# quadratic random slope, random intercept
m4 <- lmer(
    tct ~ 1 + seq.fp + I(seq.fp ^ 2) + (1 + seq.fp + I(seq.fp ^ 2) | participant),
    data = md, REML = F
)
```

note: all time and slope outputs will now be in terms of 0 to 1, where 1 = `r max(learn_h1b$seq)` (should be 11). For example, task 1 is now 0.0909 and a slope of -354 per unit is equal to -354 / 11 = -32.2 per task.

```{r h1b-anova-uncond}
# compare the unconditional models
anova(m1, m2, m3, m4)
```

AIC preferred over Wald's test of deviance (p-val) as it penalizes added parameters to reduce overfitting. Trades parameters (model complexity) for improved fit.

additional comparisons

```{r h1b-comp-uncond}
# warning for m2 = m3
cp <- compare_performance(m1, m2, m3, m4)
display(cp, layout = "vertical")
```

m2 and m3 are equivalent
both methods agree, based on AIC, AICc, BIC, R2c, RMSE, Sigma
best model is m4 - quadratic random slope, random intercept

inspect fit of some participants in selected treatments

```{r h1b-insp-cond}
# use the best model
m <- m4

md$pred <- fitted(m)  # back-transform here if necessary
first10 <- md |>
  filter(treatment == "PAR")

max_tct = max(first10$tct) + 10

ggplot(first10, aes(x = seq.fp, y = tct)) + 
  geom_point(fill="grey", pch=21, size=2, stroke=1.25) + 
  geom_line(aes(group=participant)) + facet_wrap(~participant, ncol=5)+
  scale_x_continuous(name = "Task Number") + 
  scale_y_continuous(name = "TCT",limits=c(0,max_tct))+
  geom_line(aes(y=pred), col=base_color, lwd=1.5) 

```

look at model summary

```{r h1b-sum-uncond}
summary(m)
```

all fixed effects are significant

::: content-hidden

TODO: check model assumptions / validity
these results are not great

```{r eval=FALSE}
ps <- plot(check_model(m, panel = F))
```

```{r eval=FALSE}
binned_residuals(m)
check_outliers(m)
check_overdispersion(m)

check_autocorrelation(m)
check_heteroscedasticity(m)
check_homogeneity(m)
check_normality(m)
check_residuals(m)
```

:::

#### Conditional Models and Hypothesis Testing

Add fixed treatment effect to best fit from above - does it improve the model? Use estimated slopes and intercepts to understand relationship between time, sequence, treatment, and TCT.

using treatment and interactions with `seq` and `seq ^ 2`

```{r h1b-fit-cond}
# conditional models based on m4 (quadratic random slope and random intercept)
# introduce treatment and interactions as fixed effects
cm1 <- lmer(
    tct ~ 
      1 + seq.fp * treatment + I(seq.fp ^ 2) * treatment + 
      (1 + seq.fp + I(seq.fp ^ 2) | participant),
    data = md, REML = F
)

```

does the conditional model differ from / improve fit? testing for treatment effect

```{r}
anova(m, cm1)
```

28 point reduction in AIC, significant improvement

inspect model results

```{r h1b-anova-cond}
anova(cm1)
```

rough interpretation:

- all fixed effects are significant
- TCT changes with task iteration (seq.fp)
- TCT varies by group (treatment)
- Rate of change of TCT changes by task iteration (seq.fp)
- The effect of iteration (seq.fp) differs by group (treatment)
- The rate of change of TCT differs by group (treatment)

From Chet, after confirming above, elaborating:

- Task Iteration Effect $(\text{seq.fp})$: TCT significantly decreases as task iterations increase, indicating learning over time (p < 0.001).
- Treatment Effect: Baseline TCT varies significantly by treatment group, demonstrating that different treatments have distinct initial impacts on task completion time (p < 0.001).
- Non-Linear Task Iteration Effect (Quadratic Term $(I(\text{seq.fp}^2))): The rate of change in TCT is non-linear and significantly changes with task iterations (p < 0.001).
- Interaction Effects:
  - $\text{seq.fp} \times \text{treatment}$: The impact of task iterations on TCT significantly differs across treatment groups (p = 0.005), indicating that treatments influence how quickly participants improve over iterations.
  - $I(\text{seq.fp}^2) \times \text{treatment}$: The non-linear rate of change in TCT also varies significantly by treatment (p = 0.017), suggesting that the treatments affect the curvature of the learning trajectory.
  
These results highlight the importance of both linear and quadratic terms in modeling learning rates and emphasize the varying effects of different treatments on TCT dynamics over time. The significant interactions underscore how treatments modulate not just the initial performance but also the improvement trajectories.

```{r h1b-sum-cond}
summary(cm1)
```

rough interpretation:
- (Intercept), seq.fp, and I(seq.fp^2) are all significant and give the initial TCT value, initial slope, and rate of change of slope for the reference treatment, AR
- treatmentMR/PWI/PAR give the change in intercept from AR; only PWI and PAR are significant ("reliably different")
- seq.fp:treatmentMR/PWI/PAR give the treatment level difference from the initial slope; only interactions with PWI and PAR are significant
- treatmentMR/PWI/PAR:I(seq.fp^2) gives the treatment level difference from the rate of change of slope; only PWI and PAR are signficant

from this we can see that AR and MR are roughly equivalent in all regards, and both PWI and PAR are significantly different from them in all measures. it does not provide sufficient information to ascertain the differences between MR, PWI, and PAR

compared to AR, these differences are significant:
- PWI and PAR start faster (lower baseline intercept)
- PWI and PAR start learning slower (positive seq.fp:treatment values)
- PWI and PAR improve learning faster (negative treatment:I(seq.fp^2) values)

interpretation of above:
- the first suggests PWI and PAR are more intuitive than AR
- the second aligns with the intuition that learning rate is related to the difference in understanding, i.e. lower initial task times give lower learning potential to the asymptote (proficiency), so the rate is lower
- the third observation seems counterintuitive - despite slower initial learning, the PWI and PAR reduce task completion times more per iteration

This counterintuitive trend highlights the deep insights derived from advanced modeling and underscores the importance of considering both initial performance and subsequent learning trajectories when evaluating instructional methods.

The learning for PWI and PAR might follow a sigmoidal curve, where there's slow initial improvement that accelerates rapidly as participants become more familiar with the task. This can happen if the initial hurdles are high, but once overcome, participants experience more significant gains.

PWI and PAR participants might spend more time initially adapting to a new, potentially more complex instruction method. Once they adapt, they leverage this understanding more effectively, leading to faster improvements.

The instructional methods for PWI and PAR might be inherently more effective in the long run, even if they take longer to grasp initially. These methods might impart skills that lead to compounded improvements over time.

I'm reminded of OH's comment: 'learning takes effort' - perhaps the advanced instructional techniques are too helpful, to the point of diminishing the learning. Where PWI and PAR require greater participant engagement in the task, which requires greater effort initially but that investment pays off with better overall understanding.

Early challenges encourage deeper cognitive processing, which fosters a better overall understanding despite the initial slower progress. The concept of "desirable difficulties" in learning theory posits that learning tasks should be challenging enough to promote deeper cognitive engagement but not so difficult that they become discouraging. PWI and PAR might embody this principle by requiring more effort initially, which results in more substantial long-term gains.

Chet's Revised Interpretation:

Initial Performance:

PWI and PAR participants start with lower baseline TCT, suggesting these methods are more intuitive or easier to grasp initially.
Effort and Engagement:

Advanced instructional techniques like AR might not challenge participants enough initially, leading to faster initial gains but slower long-term improvement.
PWI and PAR require greater initial effort, aligning with the participant insight that "learning takes effort." This investment of effort requires deeper cognitive engagement, which ultimately pays off.
Learning Trajectories:

Initial Slow Learning: PWI and PAR participants initially learn slower because more cognitive effort and engagement are required.
Accelerated Improvement: Over time, the effort invested initially by PWI and PAR participants results in faster learning, leading to greater reductions in TCT per iteration.
Integration of Insight:

This model aligns with learning theories suggesting that greater initial effort and engagement foster deeper understanding and long-term gains in proficiency. The comment "learning takes effort" encapsulates the necessity of active engagement, which may explain why participants eventually achieve faster reductions in TCT despite slower initial learning.

Summary:

PWI and PAR methods promote greater initial effort and cognitive engagement, leading to deeper learning and accelerated improvement over time.
While these methods may initially show slower learning rates, the effort invested results in faster long-term TCT reductions, demonstrating the importance of desirable difficulties in instructional design.

Summarized Interpretation with Concept Integration:

Initial Performance:

Lower Baseline TCT: PWI and PAR start with lower initial TCT values, suggesting these methods are more naturally intuitive.
Effort and Engagement:

Desirable Difficulty: These methods incorporate a level of difficulty that requires greater cognitive effort initially, which is essential for deep learning.
Zone of Proximal Development: PWI and PAR engage participants within their ZPD, challenging them sufficiently to enhance skill acquisition over time.
Learning Trajectories:

Initial Slower Learning: Slower initial learning rates for PWI and PAR participants reflect the necessary investment in overcoming early challenges.
Accelerating Improvement: Over time, the learning rate accelerates as participants gain proficiency, aligning with the principles of Flow Theory where optimal challenge and skill balance lead to deeper engagement and faster learning improvements.
Conclusion and Insight:
The integration of educational theories with model outputs provides a comprehensive explanation for the observed learning patterns:

PWI and PAR methods require initial higher efforts (desirable difficulty) but lead to better long-term understanding and faster TCT reductions.
AR and MR methods may offer quicker, easier starts but do not sustain the accelerated improvement seen with PWI and PAR.
Closing Thoughts:

This nuanced interpretation underscores the importance of instructional design that balances initial challenge with long-term learning benefits, leveraging the principles of desirable difficulty, ZPD, and Flow Theory for optimal educational outcomes.

pairwise comparisons of intercept, seq, and seq^2 effect using emmeans

```{r h1b-pw-cond}
em_int <- emmeans::emmeans(cm1, ~ treatment)
em_slope <- emmeans::emmeans(cm1, ~ treatment | seq.fp)
em_quad <- emmeans::emmeans(cm1, ~ treatment | I(seq.fp ^ 2))

pw_int <- pairs(em_int, adjust = "holm")
pw_slope <- pairs(em_slope, adjust = "holm")
pw_quad <- pairs(em_quad, adjust = "holm")

summary(pw_int)
summary(pw_slope)
summary(pw_quad)
```

in each case, all contrasts are significant except for AR-MR and PWI-PAR

visualize predictions...

```{r h1b-plot-cond}
library(ggplot2)

# Predicted values for different seq.fp levels
seq_levels <- unique(md$seq.fp)
predicted_vals <- expand.grid(treatment = levels(md$treatment), seq.fp = seq_levels)
predicted_vals$pred_tct <- predict(cm1, newdata = predicted_vals, re.form = NA)

# Plot with both predicted lines and actual data points
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  geom_point(data = md, aes(x = seq.fp, y = tct, color = treatment, alpha = 0.4)) +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  facet_wrap(~ treatment) +
  theme_minimal()
```

model is overfitting / extrapolating beyond data
obviously TCT should not rapidly increase after 3-6 iterations of the task
BUT - model performance is only a measure of residuals (y error) of predicted vs real data, so the up trend only matters where there is existing data - min impact on AR, max on PWI

sufficient evidence for now to go with this model, understanding its limitations, but let's try other things - limits of model fit undermine model comparisons since we're comparing models of the data now, not the data itself

further attempts to address this with:

- explored inverse, cubic, logarithmic, logistic, and trigonometric terms as alternative to quadratic / spline in the random slopes; random intercepts LME
- attempted a non-linear mixed effects model based on exponential decay
- fit the model with valid subsets of data with balanced iterations, e.g. the first n observations for all participants that completed that many tasks

no success - worse results, convergence problems

but important to get good model fit as that is what we are comparing now, not the data itself - another layer of abstraction

continuing efforts with spline, bayesian methods moved to `h1b_mod_select.qmd`

::: content-hidden

#### Old Material

This formulation captures individual variability by allowing each participant to have a different initial TCT (intercept). It accounts for variations in TCT due to differences between participants (random effects) and within each participant's repeated measures (residuals). This design naturally incorporates all sources of participant-level variation, including imbalances due to the differing number of tasks completed by each participant, thereby addressing the data's inherent non-independence.

Variations of this model were fit using the residual maximum likelihood method (REML), as implemented by the `lmer` function from the `{lme4}` package. To address issues related to the non-linear nature of TCT, which follows an exponential decay pattern commonly observed in learning outcomes, models were fit to response variables transformed by log, square-root, and inverse (i.e., $1/\text{tct}$) functions. The results were then compared using the `compare_performance` function from the `{performance}` package. @tbl-h1b-results summarizes the results.

|Metric         |      `log(tct)` |     `sqrt(tct)` |  `inverse(tct)` |
|:--------------|:---------------:|:---------------:|:---------------:|
|R2 (cond.)     |            0.72 |            0.70 |            0.71 |
|R2 (marg.)     |            0.39 |            0.36 |            0.39 |
|ICC            |            0.54 |            0.53 |            0.52 |
|RMSE           |            0.20 |            1.03 |        2.32e-03 |
|Sigma          |            0.22 |            1.13 |        2.54e-03 |

: Comparison of Model Performance Metrics {#tbl-h1b-results .striped}

The values of ICC and $R^2$ (both conditional and marginal) are comparable across the three transformations. However, the inverse response resulted in substantially lower root mean square error (RMSE) and Sigma ($\sigma$), which represent the average and standard deviation of the residuals, respectively. With prediction errors that are both lower overall and have less variance, the inverse-transformed response, shown in @fig-inverse-tct, results in the best model performance.

```{r fig-inverse-tct}
#| fig-cap: Inverse transformation of the response
#| fig-height: 3.5
#| warning: false

# Create a data frame with the original and inverse transformed TCT values
pd <- learn_h1b |>
  mutate(inverse_tct = 1 / tct)

# Plot for tct
plot_tct <- ggplot(pd, aes(x = seq, y = tct, color = treatment)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "TCT vs. Sequence",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plot for inverse(tct)
plot_inverse_tct <- ggplot(pd, aes(x = seq, y = inverse_tct, color = treatment)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Inverse(TCT) vs. Sequence",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Display the plots
plot_tct + plot_inverse_tct +
  plot_annotation(
    title = "Task Completion Time and Inverse",
    subtitle = "Inverse transformation of TCT gives best model performance"
  )
```

:::

#### Check Model Assumptions

#### Analysis

#### $H_{1b}$ Results


### $H_{1c}$: Average Error Count per Car

```{r prep-h1c-data}
# create data
learn_h1c <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj) |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  summarize(
    treatment = first(treatment),
    avg_uce = mean(uce, na.rm = T)
  )

glimpse(learn_h1c)
```

The third and final hypothesis of the learning phase:
$$H_{1b}\textrm{: Average error count per car varies with treatment}$$ 

tests a different aspect of learning. Lower error rates signal a higher quality of learning outcomes. Combined with the previous measures of efficiency (TCT) and learning rate (change in TCT per task), it provides a balanced view of training performance.

#### Applicable Statistical Methods

as with TCT, UCE data is unbalanced, not independent, probably not normally distributed; additionally, it is not continuous data, but counted values. all of these have implications for the statistical method used

however, none of this invalidates non-parametric statistical tests of aggregated (average UCE per car) data, which addresses the independence issue

#### Analysis

KW non-parametric test conducted, as seen in @fig-h1c-plot.

```{r fig-h1c-plot}
#| fig-cap: Differences in Average Uncorrected Error per Task by Treatment
#| fig-height: 5
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

ggstatsplot::ggbetweenstats(
  data = learn_h1c,
  x = treatment,
  y = avg_uce,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "Avg UCE per Task by Treatment: Kruskal-Wallis non-parametric rank sum test",
  results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between PWI and others using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "25.5", italic(p) == 
#     "1.2e-05", widehat(epsilon)["ordinal"]^2 == "0.5", CI["95%"] ~ 
#     "[" * "0.3", "1.0" * "]", italic("n")["obs"] == "53")
```

Results confirm significant treatment effect. Large differences in the overall mean are observed, along with some floor effects in the augmented treatments. Post-hoc tests confirm the significance of observed differences, showing that the error rate is significantly larger in the PWI condition than in all other treatments.

```{r h1c-validation}
co <- check_outliers(learn_h1c$avg_uce, method = c("zscore_robust", "iqr"))

# use co as boolean mask to get relevant participants
co_res <- learn_h1c[as.vector(co), ]

# tabulates the number of tasks completed by outlier participants
car_counts <- learn |>
  filter(participant %in% co_res$participant) |> 
  group_by(participant) |> 
  summarize(count = n()
)

# col_to_string(co_res, participant)
# col_to_string(co_res, avg_uce, digits = 1)

# check test results without outliers
without_outliers <- learn_h1c[!learn_h1c$participant %in% co_res$participant, ]
kt_h1b_no_outliers <- kruskal.test(avg_uce ~ treatment, data = without_outliers)
```

KW is a robust test that does not assume normality or homogeneity of variances. It can be sensitive to the presence of outliers, three of which were detected in the UCE data based on a combination of IQR and Z-score. Participants `r col_to_string(co_res, participant)` were all assigned to the PWI treatment and averaged `r col_to_string(co_res, avg_uce, digits = 1)` errors per task, respectively. A second Kruskal-Wallis test of the data without those outliers showed the difference remained significant (p=`r fmt_p_val(kt_h1b_no_outliers$p.val)`). Based on this it is determined that the outliers are not impacting the test result.

While this method was deemed insufficient for capturing the nuances of learning rate analysis, it is appropriate for comparing average error counts.

#### $H_{1c}$ Results

$H_{1c}$ is accepted - all augmented instruction methods result in error rates that are drastically lower in magnitude than PWI, the statistical significance of which is confirmed. 
note the floor effects

further analysis using a generalized linear mixed model with poisson or negative binomial distribution for count data could be used to better account for participant variation but that seems unnecessary given the strong differences observed here

::: content-hidden
MAYBE: learning add. analysis

### Additional Analyses

- number of cars completed - to paint a complete picture of individual performance, e.g. 1056 completed 7 cars with an average of 11.3 errors on each - inverse correlation between TCT and UCE?
- error rate of change?!
- balanced scorecard of efficiency, learning rate, quality?
- what treatment group had the highest percentage of "qualified operators" at the end of the 10-min session?

#### Error Types

#### Task Completion Rate

#### First-Task Performance

#### Qualified Operators

#### Transfer Effectiveness Rating (TER)
:::

## Recall Phase Results

### Descriptive Statistics

### $H_{2a}$: Overall Equipment Effectiveness

### $H_{2b}$: Paper Work Instruction References

::: content-hidden
TODO: recall add. analysis

### Additional Analyses

- what treatment group had the highest OEE during recall?

#### Efficiency vs. Quality

#### PWI Reference Frequency and Length

#### Error Types and Task Completion Rate

#### Ongoing Learning
:::

## Retention Phase Results

### Descriptive Statistics

### $H_{3a}$: Change in OEE

::: content-hidden
TODO: retention add. analysis

### Additional Analyses

#### Absolute and Percentage Change in OEE

#### Primary Driver of OEE Change
:::

## Workload, Usability, and User Experience

### NASA Task Load Index Results

### System Usability Scale Results

### Qualitative Feedback Analysis

::: content-hidden
TODO: exploratory analysis

## Exploratory Analyses

### Relationship between TLX, SUS, and Performance

- first task performance in learning as a measure of initial comprehension, intuitiveness of each treatment, effectiveness of instructional design, etc.; may correlate with SUS / TLX

### Impact of Participant Demographics on Performance

### Within-Group Performance Variance

### TLX Components and Performance

### Interactions between Workload, Usability, and Treatment
:::

## Challenges and Unexpected Observations {#sec-res-challenges}

Discussion of any challenges encountered during the conduct of the protocol  Modifications made to the protocol in response to these challenges

demoing AR/MR: The challenge of this choreography was compounded by significant latency in streamed video, which lagged the live action by approximately 2 seconds. With practice

> Technical Issues with Equipment: Discussion of any technical challenges encountered with the HL2 or other devices, and how these were addressed.
>
> Participant Scheduling and No-Shows: Challenges related to participant scheduling, no-shows, or last-minute cancellations, and the strategies developed to mitigate these issues.
>
> Data Collection Consistency: Any issues encountered in maintaining consistency in data collection across different trials and days, and how these were resolved.

limitations

- participant behavior
    -   few participants assigned HMDMR took advantage of the affordance
- analysis
    -   does not account for interactions between affordances

noticed improvising - discourage as occurs

sleeves, long hair problems for HL2

height and "down-lookers"; tracking dropouts;

battery life and performance; brightness and accuracy;

complexity of MR, less reliable tracking, crashes

## Summary of Key Findings
