# Results {#25-results}

```{r load-libs}
#| message: false

library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
library(broom)
library(report)
library(tools)
# library(ggtext)
# library(glue) now included with tidyverse as str_glue
# library(gridExtra)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r load-data}

car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results"
)

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics"
)
```

```{r def-cust-theme}
base_color <- "#69b3a2"
outline_color <- "grey25"

# Define a custom theme for ggplot
custom_theme <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      # plot.subtitle = element_text(face = "italic"),
      # plot.caption = element_text(color = "grey50"),
      # axis.title.x = element_text(face = "bold"),
      # axis.title.y = element_text(face = "bold"),
      # axis.text = element_text(color = "black"),
      # axis.line = element_line(color = "grey80"),
      # legend.position = "bottom",
      # legend.background = element_rect(fill = "white", colour = "grey50"),
      # legend.key = element_rect(fill = "white", colour = "grey50"),
    )
}

# Set the custom theme as the default theme
theme_set(custom_theme())
```

```{r study-metadata}

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# Convert the trial_datetime to Date type
demographics <- demographics |>
  mutate(
    trial_date = as.Date(
      as.POSIXct(trial_datetime, format = "%Y-%m-%d %H:%M:%S")
    )
  )

# Get the first and last date
first_trial_date <- demographics |>
  summarize(first_date = min(trial_date)) |>
  pull(first_date)

last_trial_date <- demographics |>
  summarize(last_date = max(trial_date)) |>
  pull(last_date)

trial_start_str <- str_glue(
  "{month(first_trial_date)}/{day(first_trial_date)}"
)
trial_end_str <- str_glue(
  "{month(last_trial_date)}/{day(last_trial_date)}"
)
trial_year_str <- str_glue("{year(last_trial_date)}")

# start, end = Feb 10, Apr 27 2023
```

## Study Administration and Participation

This study was administered during the Spring of 2023. Ultimately, `r participant_count` eligible participants were recruited. All completed the Learning and Recall experiments of the first session beginning February 10th. Progress of the conduct of these trials is illustrated by @fig-s1-progress.

```{r fig-s1-progress}
#| fig-cap: "Daily Trial Administration and Completion Percentage"
#| fig-height: 3.5

# Count experiments per day
daily_counts <- demographics |>
  group_by(trial_date) |>
  summarize(daily_count = n()) |>
  arrange(trial_date)

# Calculate the cumulative count of experiments
daily_counts <- daily_counts |>
  mutate(cumulative_count = cumsum(daily_count))

# Calculate remaining experiments
daily_counts <- daily_counts |>
  mutate(
    percentage_complete = (cumulative_count / participant_count) * 100
  )

pc_eng <- as.character(english::english(participant_count))
pc_eng <- str_replace(pc_eng, "^.", str_to_upper)
subtitle <- str_glue("{pc_eng} Learning and Recall Experiments Conducted between {trial_start_str} and {trial_end_str} of {trial_year_str}")

# Plotting
ggplot(daily_counts, aes(x = trial_date)) +
  geom_line(
    aes(y = daily_count, group = 1),
    color = base_color,
    linewidth = 1,
    alpha = 0.5
  ) +
  geom_line(
    aes(y = percentage_complete, group = 1),
    color = base_color,
    linewidth = 1.25,
    alpha = 1
  ) +
  scale_y_continuous(
    name = "Trials Administered",
    sec.axis = sec_axis(~., name = "Percentage Complete"),
    position = "right",
    limits = c(0, 100)
  ) +
  labs(
    title = "First Session Progress",
    subtitle = subtitle,
    x = element_blank(),
    y = element_blank()
  ) +
  custom_theme()
```

Only eight trials were run before March 16th. That period was treated as a pilot study, during which the research team was trained, administration methods were refined, the self-service onboarding system was deployed, and version 2.1 of our IRB application was approved. Most significantly, the intake process was adjusted to collect more demographic data, increase the role of the TLX and SUS instruments, and gather BCS data for a companion study.

Otherwise, only minor tweaks were made during the pilot to streamline the experiment's administration. Although the changes outlined above were not believed to have a material effect on the results, the data collected during the pilot were omitted from further analysis. Following the pilot, steady progress was made until the completion of the study's first session on April 27th.

The second session was conducted during the end of study event on *XXX*. *YYY* of the study's original participants volunteered to attend and complete the Retention experiment.

::: content-hidden
TODO: complete above paragraph about second session.
:::

::: content-hidden
MAYBE: perform analysis to evaluate differences between pilot and main group

Pilot Study Comparison
- Compare the key measures (e.g., performance, workload, usability) between the pilot study (n=8) and main study (n=54) participants
- Use descriptive statistics and visualizations to present these comparisons
- Discuss any notable similarities or differences between the two groups

Statistical Power and Sample Size Considerations
- Present the results of statistical power analyses based on the observed effect sizes and the planned statistical tests
- Discuss the adequacy of the sample size for detecting meaningful effects
- Consider the potential impact of including or excluding the pilot study data on the statistical power and precision of the estimates
- sample size calculation / justification: power analysis, practical considerations, precedent

Decision on Pilot Study Inclusion
- Based on the findings from 5.2.3 and 5.2.4, make a decision on whether to include the pilot study data in the main analysis
- Clearly report your decision and rationale, considering factors such as data comparability, study design modifications, and statistical power
:::

## Analysis of Participant Demographics

### Univariate Analysis

```{r drop-pilot-data}
# drop participants in pilot study (first eight)
demographics <- demographics |>
  filter(!(participant >= 1001 & participant <= 1010))

# drop event data from the pilot study
car_results <- car_results |>
  filter(!(participant >= 1001 & participant <= 1010))
```

```{r def-to-factors}

# safely convert data to ord / cat factors
# create levels / labels as required and identify / warn about mismatches

to_factors <- function(
    column,
    levels = NULL,
    labels = NULL,
    ordered = FALSE) {
  if (is.null(levels)) {
    levels <- unique(na.omit(column))
  }

  if (is.null(labels)) {
    labels <- levels
  }

  mismatched <- setdiff(unique(na.omit(column)), levels)

  if (length(mismatched) > 0) {
    # Append mismatched values to levels and labels with a specific prefix
    bad_levels <- paste("Bad -", mismatched)
    levels <- c(levels, mismatched)
    labels <- c(labels, bad_levels)
    warning(
      "Mismatched or unexpected values incorporated: ",
      paste(bad_levels, collapse = ", ")
    )
  }

  # Proceed to convert to factor using cleaned-up data
  factor_column <- factor(column,
    levels = levels, labels = labels,
    ordered = ordered
  )

  return(factor_column)
}
```

```{r factor-demographics}

# convert all categorical / ordinal data to factors

demo_f <- demographics |>
  mutate(
    # categorical, levels != labels
    # unordered with predefined levels
    gender = to_factors(
      gender,
      levels = c("F", "M", "O"),
      labels = c("Female", "Male", "Non-Binary")
    ),
    race = to_factors(
      race,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "American Indian or Alaska Native",
        "Asian",
        "Black or African American",
        "Native Hawaiian or Other Pacific Islander",
        "White",
        "More than one race",
        "Unknown or not reported"
      )
    ),
    ethnicity = to_factors(
      ethnicity,
      levels = c("1", "2", "3"),
      labels = c(
        "Hispanic or Latino",
        "Not Hispanic or Latino",
        "Unknown or not reported"
      )
    ),
    lenses = to_factors(
      lenses,
      levels = c("G", "C", "N", "Y"),
      labels = c(
        "Glasses",
        "Contacts",
        "No",
        "Yes"
      )
    ),

    # categorical, levels == labels
    # unordered with predefined levels
    language = to_factors(
      language,
      levels = c("English", "Other"),
    ),
    worn = to_factors(
      worn,
      levels = c("Y", "N")
    ),
    color_blind = to_factors(
      color_blind,
      levels = c("Y", "N")
    ),
    condition = to_factors(
      condition,
      levels = c("Y", "N")
    ),

    # ad-hoc
    # categorized in transcription, no predefined levels
    country = to_factors(country),
    college = to_factors(college),
    program = to_factors(program),
    referral = to_factors(referral),

    # ordinal, levels != labels
    # ordered with predefined levels
    education = to_factors(
      education,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "Pre-HS", # No HS diploma
        "HS", # High school diploma or equivalent
        "CND", # Some college but no degree
        "AS",
        "BS",
        "MS",
        "PhD"
      ),
      ordered = TRUE
    ),
    lego = to_factors(
      lego,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "Little to no experience",
        "Some experience",
        "Lots of experience",
        "Expert builder"
      ),
      ordered = TRUE
    ),
    manuf = to_factors(
      manuf,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "No experience",
        "Taken one or more classes",
        "Held a part time or temporary position",
        "One or more year’s experience"
      ),
      ordered = TRUE
    ),
    runner = to_factors(
      runner,
      levels = c("N", "Y", "5", "10", "H", "M", "T"),
      labels = c(
        "None",
        "Yes - Unspecified",
        "5K",
        "10K",
        "Half-Marathon",
        "Marathon",
        "Triathalon"
      ),
      ordered = TRUE
    ),
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    )
  )
```

#### Numeric Data

```{r numeric-data}
nums <- demo_f |>
  select(where(is.numeric))

height_stats <- nums |>
  summarize(
    mean = mean(height, na.rm = TRUE),
    sd = sd(height, na.rm = TRUE),
    min = min(height, na.rm = TRUE),
    max = max(height, na.rm = TRUE),
    pct_missing = sum(is.na(height)) / n() * 100
  )

height_summary <- str_glue(
  "Mean height = {round(height_stats$mean, 1)}, SD = {round(height_stats$sd, 1)}, range: [{height_stats$min}, {height_stats$max}], {round(height_stats$pct_missing, 1)}% missing"
)
```

After excluding data from the pilot study, `r report_participants(nums, age = "age", participants = "participant")` were included in this analysis. Other than age, height (`r height_summary`) was the only numeric data collected. As discussed in @sec-res-challenges, height was first recorded following the pilot study, when its impact on user behavior and results was first observed. Both numeric characteristics are visualized with a combined box and scatter plot in @fig-numeric. Where present, the notches of these box plots show the 95% confidence interval around the median.

```{r fig-numeric}
#| fig-cap: "Distribution of Participant Age and Height"
#| fig-height: 3.5
#| warning: false

# Plotting age
age_plot <- demo_f |>
  filter(!is.na(age)) |>
  ggplot(aes(x = age)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    x = "Age (yr)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting height
height_plot <- demo_f |>
  filter(!is.na(height)) |>
  ggplot(aes(x = height)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    width = 0.55,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    x = "Height (in)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
age_plot + height_plot +
  # plot_layout(ncol = 2, widths = c(1,0.8)) +
  plot_annotation(
    title = "Numeric Sample Characteristics",
    subtitle = "Age reflects the student-oriented nature of the sample"
  )
```

#### Categorical Data

```{r categorical-data}
cats <- demo_f |>
  select(where(~ is.factor(.) && !is.ordered(.))) |>
  select(-c("referral")) |>
  mutate(across(everything(), \(x) fct_lump(x, prop = 0.1)))

# Summary table
cat_tbl <- cats |>
  select(gender:language) |>
  tbl_summary(
    label = setNames(str_to_title(names(cats)), names(cats)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

cat_names <- cats |>
  select(gender:country) |>
  names()

cats_long <- cats |>
  select(gender:country) |>
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "Value"
  ) |>
  drop_na() |>
  group_by(Category, Value) |>
  summarize(Count = n(), .groups = "drop") |>
  group_by(Category) |>
  mutate(Freq = Count / sum(Count))

country_levels <- demo_f |>
  filter(country != "USA") |>
  count(country) |>
  arrange(desc(n))
```

Of the sample's `r inline_text(cat_tbl, language, level = "English", pattern = "{N} participants, {p}% (n={n})")` reported that they mainly speak English at home. As shown in @fig-categorical, the gender composition was relatively balanced, with slightly more males `r inline_text(cat_tbl, gender, level = "Male", pattern = "(n={n}, {p}%)")` than females `r inline_text(cat_tbl, gender, level = "Female", pattern = "(n={n}, {p}%)")`. The racial makeup of the sample was predominantly White `r inline_text(cat_tbl, race, level = "White", pattern = "(n={n}, {p}%)")` and Asian `r inline_text(cat_tbl, race, level = "Asian", pattern = "(n={n}, {p}%)")`, with smaller representations from `r as.character(english::english(n_distinct(demo_f$race) - 2))` other groups. Fully `r inline_text(cat_tbl, ethnicity, level="Not Hispanic or Latino", pattern = "{p}% of participants (n={n})")` reported a non-Hispanic or Latino ethnicity. The majority of participants `r inline_text(cat_tbl, country, level = "USA", pattern = "(n={n}, {p}%)")` were from the United States, though `r as.character(english::english(n_distinct(demo_f$country, na.rm = TRUE) - 1))` other countries were represented, including `r paste(country_levels[1, ]$country, " (n=", country_levels[1, ]$n, ")", sep = "")` and `r paste(country_levels[2, ]$country, " (n=", country_levels[2, ]$n, ")", sep = "")`. `r paste(country_levels$country[3:(nrow(country_levels) - 1)], collapse = ", ")`, and `r tail(country_levels$country, 1)` were also listed as countries of origin by one participant each (n=1). All of this aligns with the study's recruitment focus.

```{r fig-categorical}
#| fig-cap: "Distribution of Categorical Characteristics"
#| fig-height: 3.5
#| results: hide

create_cat_plot <- function(data, category, base_color) {
  # specify alpha ramp end points
  min_alpha <- 0.3
  max_alpha <- 0.8

  # Filter and reorder data
  plot_data <- data |>
    filter(Category == category) |>
    mutate(
      Value = if_else(Value == "Not Hispanic or Latino",
        "Not Hispanic\nor Latino", Value
      ),
      Value = reorder(Value, Count, FUN = sum, decreasing = TRUE)
    )

  # Calculate alpha values based on Count linearly between min_alpha and max_alpha
  alpha_values <- plot_data |>
    mutate(alpha = scales::rescale(Count, to = c(min_alpha, max_alpha))) |>
    pull(alpha)

  # Set unique Value groups for setting colors and alphas
  unique_values <- unique(plot_data$Value)

  # Create the plot
  ggplot(
    plot_data,
    aes(x = Value, y = Count, fill = I(base_color), alpha = Value)
  ) +
    geom_bar(stat = "identity", position = "dodge", width = 0.8, color = outline_color) +
    scale_x_discrete(labels = function(x) tools::toTitleCase(x)) +
    labs(x = tools::toTitleCase(category), y = element_blank()) +
    scale_alpha_manual(values = setNames(alpha_values, unique_values)) +
    guides(fill = "none", alpha = "none") +
    ylim(0, max_count) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
}

# Calculate the maximum count to standardize the y-axis
max_count <- max(cats_long$Count)

# Create all plots
cat_plots <- map(
  cat_names,
  ~ create_cat_plot(cats_long, .x, base_color)
)

# Add y-axis text and ticks to first plot
first_plot <- cat_plots[[1]] +
  theme(
    axis.text.y = element_text(),
    axis.ticks.y = element_line()
  )

# Other plots without y-axis text and ticks, drop first
other_plots <- cat_plots[-1]

# combine!
all_plots <- c(list(first_plot), other_plots)

# Combine all plots
combined_plot <- reduce(all_plots, `+`) +
  plot_layout(ncol = length(all_plots), widths = c(1, 1, 0.667, 0.667)) +
  plot_annotation(
    title = "Categorical Sample Characteristics",
    subtitle = "Value Count for Gender, Race, Ethnicity, and Country"
  )

# Print the combined plot
print(combined_plot)
```

```{r make-vc-table}
vc_tbl <- cats |>
  select(lenses:worn) |>
  filter(!lenses %in% c("No", "Other", "Unknown")) |>
  filter(!(is.na(lenses) & is.na(worn))) |>
  droplevels() |>
  mutate(worn = recode(worn, "Y" = "Yes", "N" = "No")) |>
  tbl_cross(
    row = lenses,
    col = worn,
    label = list(lenses = "Lenses", worn = "Worn During Session"),
    percent = "row",
    missing_text = "NA"
  ) |>
  add_stat_label() |>
  bold_labels()
```

Of the 54 participants, 52% (n=28) reported having a vision prescription. Of those, `r inline_text(vc_tbl, row_level = "Contacts", col_level = "Total")` were for contact lenses , but `r inline_text(vc_tbl, row_level = "Glasses", col_level = "Total")` had glasses. Whereas all contact wearers reported they would wear them during the session, among glasses wearers only seven (47%) planned the same. Five (33%) did not intend to do so, and three (20%) gave no indication. This breakdown is summarized by @tbl-vc. Two participants also reported color-blindness (n=1) or other vision conditions (n=1). No participants indicated any other condition that might affect their performance.

::: content-hidden
MAYBE: make this all inline code; rethink percent by row, col, or cell? format n= etc. as markdown?
:::

```{r tbl-vc}
#| tbl-cap: Vision Correction Table

vc_tbl |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

Following the primary hypothesis testing, additional analyses may explore the differences in key outcomes across these demographic subgroups. That effort will aim to provide additional context and insights into potential underlying factors influencing the findings.

#### Ordinal Data

```{r ord-data}
ords <- demo_f |>
  select(where(~ is.factor(.) && is.ordered(.)))

# Summary table
ords_tbl <- ords |>
  tbl_summary(
    label = setNames(str_to_title(names(ords)), names(ords)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

# ords_tbl

# Helper function to calculate mode
calculate_mode <- function(data, var_name) {
  count_table <- table(data[[var_name]])
  max_freq <- max(count_table)
  # pick first if more than one value has max count
  mode_value <- names(count_table[count_table == max_freq])[1]

  # Convert to an output that includes the level's name (mode_value) and its index
  mode_index <- which(levels(data[[var_name]]) == mode_value)
  return(list(value = mode_value, index = mode_index))
}

# Main function to calculate statistics and return formatted string
calculate_ord_summary_stats <- function(data, var_name) {
  # Ensure the variable is an ordered factor
  if (!is.ordered(data[[var_name]])) {
    stop("The variable must be an ordered factor.")
  }

  # Convert factor to integer codes
  factor_int <- as.integer(data[[var_name]])

  # Calculate median, mode, and IQR
  median_value <- median(factor_int, na.rm = TRUE)
  mode_results <- calculate_mode(data, var_name)
  iqr_value <- IQR(factor_int, na.rm = TRUE)

  # Format and return the result as a string
  result_string <- sprintf(
    "(Median = %.1f; Mode = %d, %s; IQR = %.1f)",
    median_value, mode_results$index, mode_results$value,
    iqr_value
  )
  return(result_string)
}
```

Unlike the data above, the categories analyzed in this section are inherently ordered. Responses for each participant's current Education level `r calculate_ord_summary_stats(ords, "education")` and their experience with Lego `r calculate_ord_summary_stats(ords, "lego")` and Manufacturing `r calculate_ord_summary_stats(ords, "manuf")` are visualized in increasing order by @fig-ordinal. The results are mostly in line with expectations based on the recruitment, though it is disappointing to see that `r inline_text(ords_tbl, lego, level = "Little to no experience", pattern = "{p}% of participants (n={n})")` claimed to have little to no experience with Lego. The great majority were undergraduate students working towards a degree `r inline_text(ords_tbl, education, level = "CND", pattern = "(n={n}, {p}%)")`, with no experience in manufacturing `r inline_text(ords_tbl, manuf, level = "No experience", pattern = "(n={n}, {p}%)")`.

```{r fig-ordinal}
#| fig-cap: "Distribution of Ordinal Characteristics"
#| fig-height: 3.5
#| results: hide

plot_ords <- function(data, var, title) {
  norm_data <- data |>
    count(var_name = {{ var }}) |>
    mutate(percent = n / sum(n) * 100)

  p <- norm_data |>
    ggplot(aes(y = var_name, x = percent)) +
    geom_bar(
      aes(alpha = percent / 100),
      fill = base_color,
      stat = "identity",
      width = 0.75,
      color = outline_color
    ) +
    scale_alpha_continuous(trans = "log10", range = c(0.2, 0.8), guide = "none") +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank()
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(size = 9, face = "bold"),
      legend.position = "none",
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
  return(p)
}


educ_plot <- ords |>
  mutate(education = fct_recode(education,
    College = "CND",
    Associate = "AS",
    Bachelor = "BS",
    Master = "MS",
    Doctorate = "PhD"
  )) |>
  plot_ords(var = education, title = "Current Education Level")

lego_plot <- ords |>
  mutate(lego = fct_recode(lego,
    Little = "Little to no experience",
    Some = "Some experience",
    Lots = "Lots of experience",
    Expert = "Expert builder"
  )) |>
  plot_ords(lego, title = "Lego Experience")

manuf_plot <- ords |>
  mutate(manuf = fct_recode(manuf,
    None = "No experience",
    Training = "Taken one or more classes",
    PartTime = "Held a part time or temporary position",
    Some = "One or more year’s experience"
  )) |>
  plot_ords(manuf, title = "Manufacturing Experience")

# combine using patchwork
educ_plot / (lego_plot + manuf_plot) +
  # plot_layout(ncol = 2) +
  plot_annotation(
    title = "Ordinal Sample Characteristics",
    subtitle = "Reflects the Focus of Recruiting Efforts"
  )

```

### Multivariate Analysis

In order to assess the equivalence of treatment groups and validate the random assignment that is assumed by most analysis to follow, Age, Gender, Lego experience, and Education were compared across treatment groups. These variables were selected for their meaningful variation and potential relevance to the outcomes of interest. Other notable variables, including Ethnicity and Country of Origin were considered less critical due to their skewed distributions and small subgroup sizes.

```{r make-mv-table}

num_cols <- demographics |>
  select(age)

cat_cols <- demo_f |>
  select(c("gender", "lego", "education", "treatment"))

mv_data <- bind_cols(num_cols, cat_cols)

mv_tbl <- mv_data |>
  mutate(across(where(is.factor), ~ fct_drop(.))) |>
  mutate(
    lego = fct_recode(lego,
      "Little" = "Little to no experience",
      "Some" = "Some experience",
      "Lots" = "Lots of experience",
      "Expert" = "Expert builder"
    ),
    education = fct_recode(education,
      "College" = "CND",
      "Associate" = "AS",
      "Bachelor" = "BS",
      "Master" = "MS",
      "Doctorate" = "PhD"
    )
  ) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    label = list(
      "age" ~ "Age",
      "gender" ~ "Gender",
      "lego" ~ "Lego",
      "education" ~ "Education"
    )
  ) |>
  add_p() |>
  # separate_p_footnotes() |>
  # add_stat_label() |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  italicize_levels() |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nN = {n} ({style_percent(p)}%)",
    p.value ~ "**p-val**"
  )

```

@tbl-demo-by-treatment shows the result of this analysis. The Kruskal-Wallis rank sum test was utilized for the numeric variable (Age) to compare distributions across groups, while Fisher's exact test was applied to the categorical and ordinal variables (Gender, Education, and Lego experience) to assess the independence of distributions from treatment assignments.

```{r tbl-demo-by-treatment}
#| tbl-cap: Participant Demographics by Treatment Group

mv_tbl |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )

```

::: content-hidden
MAYBE: wrap gtsummary calls in a function that saves an svg to project
and inserts embedding so that all original formatting is preserved?
could they still be labeled as tables?
:::

The high p-values for Age, `r inline_text(mv_tbl, variable = age, column = "p.value")`; Gender, `r inline_text(mv_tbl, variable = gender, column = "p.value")`; Lego, `r inline_text(mv_tbl, variable = lego, column = "p.value")`; and Education, `r inline_text(mv_tbl, variable = lego, column = "p.value")` indicate that these observed characteristics are not strongly associated with treatment groupings. This result validates the group randomization effort, mitigating the influence of confounding variables, and reducing bias in the data. It is essential to the validity of most of the statistical tests employed by this study, which assume sample independence.

::: content-hidden
MAYBE: further analysis to show that:
- nums
  - height (0.11)
- cats
  - country (0.085)
  - lenses (0.063)
- ords
  - manuf exp (0.046)

may be significant (p-val)
used the following code to test all demographics

``` text
# test significance of treatment groups
trt <- demo_f |>
  select(treatment)

# bind treatment for nums, ords (included in cats)
cd <- bind_cols(ords, trt)

cd |>
  tbl_summary(
      by = "treatment",
      missing = "no",
      statistic = list(
        all_continuous() ~ "{mean} ({sd})"
      )
    ) |>
    add_p()
```
:::

## Learning Phase Results

Only completed tasks, the primary result of interest, are considered in this analysis. While excluding incomplete tasks may limit the generalizability of the findings to scenarios where time constraints are common, it ensures the analysis is based on objective, directly observed data and avoids introducing assumptions or biases that could arise from predicting results for unfinished tasks. 

```{r factor-car-results}

res_f <- car_results |>
  mutate(
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    ),
    car_outcome = to_factors(car_outcome)
  )
```

```{r learn-data}
learn <- res_f |>
  filter(
    phase == 1,
    car_outcome == "complete",
    # hack to remove incorrectly labeled car
    uid != "1048-P1-C5"
  ) |>
  select(
    participant,
    car_elapsed_time,
    uce,
    treatment,
    drop_event_count,
    drop_event_tot_time_adj,
    repair_event_count,
    repair_event_tot_time,
    system_event_count,
    system_event_tot_time
  )

# add tot_sys_adj, tot_usr_adj, and adjusted TCT time to learn
# tot_sys_adj will be used to calculate OEE
# tot_usr_adj just to reduce TCT

learn <- learn |>
  mutate(
    # Replace NA with 0 using coalesce for safe addition
    tot_sys_adj = coalesce(drop_event_tot_time_adj, 0) +
      coalesce(system_event_tot_time, 0),
    tot_usr_adj = coalesce(repair_event_tot_time, 0),
    tot_adj = tot_sys_adj + tot_usr_adj,
    car_elapsed_time_adj = car_elapsed_time - tot_adj
  )

changes <- learn |>
  filter(car_elapsed_time != car_elapsed_time_adj)

total_tasks_completed <- dim(learn)[1]
total_tasks_adjusted <- dim(changes)[1]
```

One final step of data preparation was taken to account for drop-outs and other system events, along with time spent making repairs. This lost time was deducted from measured task times to give a more accurate record of participant performance. `r total_tasks_adjusted` of `r total_tasks_completed` completed tasks (`r round(total_tasks_adjusted / total_tasks_completed * 100, 1)`%) were affected by this adjustment.

::: content-hidden
MAYBE: consider better quantifying the effect of these adjustments
:::

### Descriptive Statistics for Granular Data

```{r make-report}
# report descriptive stats for cet and uce
learn_rpt <- learn |>
  rename(
    "Task Completion Time (sec)" = car_elapsed_time_adj,
    "Uncorrected Error Count" = uce
  ) |>
  select("Task Completion Time (sec)", "Uncorrected Error Count") |>
  report::report() # options not working!

rpt_table <- as.report_table(learn_rpt)
cv_tct <- rpt_table$SD[1] / rpt_table$Mean[1]
cv_uce <- rpt_table$SD[2] / rpt_table$Mean[2]
```

::: content-hidden
TODO: correct data for `1048-P1-C5`; flagged complete with elapsed time of 4.267s
TODO: investigate report::report options
:::

`r learn_rpt`

These statistics provide a high-level summary of performance across all tasks and participants, where each observation corresponds to a single repetition of the Learning task. Both Task Completion Time (TCT) and Uncorrected Error Count (UCE) show considerable variance, as depicted in @fig-learn.

```{r fig-learn}
#| fig-cap: "Distribution of Task Completion Time and Error Rates in Learning"
#| fig-height: 3.5
#| warning: false

# Plotting adjusted car_elapsed_time
cet_plot <- learn |>
  ggplot(aes(x = car_elapsed_time_adj)) +
  geom_histogram(
    aes(y = ..density.. * 65), # hack for relative size / scale
    binwidth = 10,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    title = "Task Completion Time (sec)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting Uncorrected Error Counts
uce_plot <- learn |>
  ggplot(aes(x = uce)) +
  geom_histogram(
    aes(y = ..density.. * 1.5), # hack for relative size / scale
    binwidth = 1,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    title = "Number of Uncorrected Errors per Task",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
cet_plot / uce_plot +
  # plot_layout(ncol = 2, widths = c(1,1)) +
  plot_annotation(
    title = "Learning Phase - Observed Measures",
    subtitle = "Error Count Exhibits Greater Variability than Task Completion Time"
  )
```

Compared to TCT, with a coefficient of variation (CV) of `r round(cv_tct, 2)`, UCE shows substantially greater spread (CV=`r round(cv_uce, 2)`) and the presence of a floor effect. The later could potentially suggest limitations in the task's ability to capture a full range of participant performance. Further investigation is required to determine if this is due to task complexity or other factors.

### Aggregated Data

```{r agg-results}
# agg participant results
learn_agg <- learn |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    tot_cars = n(), # Count the number of tasks per participant
    avg_tct_adj = mean(car_elapsed_time_adj, na.rm = TRUE),
    avg_uce = mean(uce, na.rm = TRUE)
  ) |>
  ungroup()
```

::: content-hidden
MAYBE: may need to add back other variables from learn into this agg table
:::

@tbl-agg-results-by-treat presents participant-level metrics, including the number of tasks completed along with the average TCT and UCE per task. In addition to the overall results, the data is broken out by assigned treatment group to allow for comparisons across conditions. Note that the statistics presented are the overall (aka "grand") mean and median values. That is, they are the average or middle values of all individual participant averages. This provides overall summary statistics for each treatment group that reflect the central tendency and variability among the average outcomes reported by participants within each group.

```{r tbl-agg-results-by-treat}
#| tbl-cap: Avg. Participant Performance Data Summarized by Treatment Group

t <- learn_agg |>
  select(-participant) |>
  tbl_summary(
    type = all_continuous() ~ "continuous2",
    digits = all_continuous2() ~ 1,
    label = list(
      avg_tct_adj ~ "Avg TCT per Task (sec)",
      avg_uce ~ "Avg UCE per Task",
      tot_cars ~ "Number of Tasks Completed"
    ),
    by = treatment,
    statistic = list(
      all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} ({p25}, {p75})"
      )
    )
  ) |>
  # bold_labels() |>
  # add_p() |>
  # separate_p_footnotes() |>
  add_overall() |>
  # bold_p(t = 0.05) |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nn = {n} ({style_percent(p)}%)",
    # p.value ~ "**p-val**"
  )

t |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    #linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

See @fig-perf-by-treat for a depiction of the grand means for all variables, with 95% confidence intervals for each. Observe that the previously noted variability and positive skew are also present in the grouped data, and some treatments like PWI may have prioritized speed over accuracy, or vice versa (e.g. AR).

```{r make-perf-data}
# Reshape data to long format
long_data <- learn_agg |>
  pivot_longer(
    cols = c(avg_tct_adj, tot_cars, avg_uce),
    names_to = "metric",
    values_to = "value"
  )

# Calculate means, sample size, and SE for each metric within each treatment
metrics_stats <- long_data |>
  group_by(treatment, metric) |>
  summarize(
    mean_value = mean(value, na.rm = TRUE),
    n = n(), # sample size
    se = sd(value, na.rm = TRUE) / sqrt(n()), # standard error
    .groups = "drop"
  )

# Find the highest mean for each metric across all treatments
max_means <- metrics_stats |>
  group_by(metric) |>
  summarize(max_mean = max(mean_value), .groups = "drop")

# Normalize means and calculate CI within their metrics
normalized_metrics_stats <- metrics_stats |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = mean_value / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  )

# Calculate overall mean, SE, and normalize
overall_stats <- long_data |>
  group_by(metric) |>
  summarize(
    overall_mean = mean(value, na.rm = TRUE),
    n = n(),
    se = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = overall_mean / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  ) |>
  mutate(treatment = "ALL")

# Combine with other stats
final_stats <- bind_rows(normalized_metrics_stats, overall_stats)

# Factorize the 'treatment' column
final_stats$treatment <- factor(final_stats$treatment, levels = c("ALL", "PWI", "PAR", "AR", "MR"))

# Factorize and label 'metric' column
final_stats$metric <- factor(
  final_stats$metric,
  levels = c("avg_tct_adj", "avg_uce", "tot_cars"),
  labels = c("Average Task Time", "Average Error Count", "Tasks Completed")
)

# Predefined alpha values by treatment
alpha_values <- c(0.4, 0.55, 0.7, 0.85, 1.0)
names(alpha_values) <- levels(final_stats$treatment)
```


```{r fig-perf-by-treat}
#| fig-cap: "Aggregated Learning Performance Data by Treatment"
#| fig-height: 3.5
#| warning: false

ggplot(final_stats, aes(x = metric, y = normalized_mean, fill = treatment, label = treatment)) +
  geom_col(
    aes(alpha = normalized_mean),
    position = position_dodge(0.9),
    color = outline_color
  ) +
  geom_errorbar(
    aes(ymin = lower_ci, ymax = upper_ci),
    position = position_dodge(0.9),
    width = 0.25,
    alpha = 0.7
  ) +
  geom_text(
    aes(y = 0),
    position = position_dodge(width = 0.9),
    vjust = 1.65, # Adjust based on visual preference
    size = 3,
    alpha = 0.9
  ) +
  scale_fill_manual(values = rep(base_color, length(unique(final_stats$treatment)))) +
  scale_alpha_continuous(transform = "pseudo_log", range = c(0.3, 1.0)) +
  labs(
    title = "Comparison of Normalized Performance Metrics by Treatment",
    subtitle = "Includes Overall and Treatment Groups with 95% Confidence Intervals",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(
      margin = margin(b = 5, unit = "pt"),
      color = "black"
    )
  )

```

#### Assess outliers

Given the level of variation and spread seen above, the presence of outliers was analyzed. Participant-level outliers within treatment groups were the focus, rather than individual data points across the entire dataset. This method helps ensure that outliers are identified relative to the specific context of each treatment, accounting for the corresponding variability.

Four different tests were used to assess each participant as an outlier in their treatment group: (1) Tukey's Fences at 1.5 times the IQR, (2) Tukey's Fences at 3.0 times the IQR, (3) outside the 5th and 95th percentiles, and (4) using the Mahalanobis distance. TCE and UCE were each tested separately using the first three methods. The fourth method is a multivariate test that considers TCE and UCE together. The results are summarized in @tbl-outlier-summary, for all participants that were identified by two or more tests.

```{r find-outliers}
# Function to detect outliers in a vector
find_outliers <- function(x, mult) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower_bound <- q1 - mult * iqr
  upper_bound <- q3 + mult * iqr
  x < lower_bound | x > upper_bound
}

# Function to calculate Mahalanobis distance for a group
compute_mahalanobis <- function(df, cols) {
  cov_matrix <- cov(df[, cols], use = "complete.obs")
  center <- colMeans(df[, cols], na.rm = TRUE)
  mahal_dist <- mahalanobis(df[, cols], center, cov_matrix)
  return(mahal_dist)
}

outlier_summary <- learn_agg |>
  group_by(treatment) |>
  mutate(
    # find 1.5 and 3x IQR outliers using Tukey method
    out15_tct = find_outliers(avg_tct_adj, 1.5),
    out15_uce = find_outliers(avg_uce, 1.5),
    out30_tct = find_outliers(avg_tct_adj, 3.0),
    out30_uce = find_outliers(avg_uce, 3.0),
    # find by 5th and 95th percentile
    pct_tct = avg_tct_adj < quantile(avg_tct_adj, 0.05, na.rm = TRUE) |
      avg_tct_adj > quantile(avg_tct_adj, 0.95, na.rm = TRUE),
    pct_uce = avg_uce < quantile(avg_uce, 0.05, na.rm = TRUE) |
      avg_uce > quantile(avg_uce, 0.95, na.rm = TRUE),
    # find by Mahalanobis distance
    mahal_dist = compute_mahalanobis(
      pick(avg_tct_adj, avg_uce),
      c("avg_tct_adj", "avg_uce")
    ),
    md_out = mahal_dist > qchisq(0.975, df = 2)
  ) |>
  ungroup() |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    out15_tct = sum(out15_tct),
    out15_uce = sum(out15_uce),
    out30_tct = sum(out30_tct),
    out30_uce = sum(out30_uce),
    pct_tct = sum(pct_tct),
    pct_uce = sum(pct_uce),
    md_out = sum(md_out),
    .groups = "drop"
  )
```


```{r tbl-outlier-summary}
#| tbl-cap: Summary of Participants Identified by Two or More Outlier Tests
#| warning: false

outlier_summary |>
  rowwise() |>
  mutate(tot_out = sum(c_across(-c(participant, treatment)))) |>
  ungroup() |>
  filter(tot_out > 1) |>
  arrange(desc(tot_out)) |>
  select(participant, treatment, tot_out) |>
  inner_join(learn_agg) |>
  gt()

```

::: content-hidden
TODO: make this table prettier
:::

Further investigation of these seven participants showed that only 1063 experienced systematic problems that might warrant excluding their performance. The other participants were just relatively poor performers, with high TCT and/or UCE. Given the limited amount of data collected per treatment, it was decided to retain all but 1063's data for hypothesis testing.

```{r drop-learn-outliers}
# drop 1063
learn_agg_out <- learn_agg |> filter(treatment != 1063)
```

### Hypothesis Testing

This first group of hypotheses are designed to test the effect of each
treatment on training outcomes, including average TCT and UCE, along with the learning rate.

#### Average Time per Car

Here we will test:
$$H_{1a}\textrm{: Average time per car varies with treatment}$$

As noted in the original assessment of the learning data, the overall TCT dataset exhibits skewness and kurtosis, suggesting deviations from a normal distribution. To choose the most appropriate statistical method for hypothesis testing, this must be revisited. Standard between-groups hypothesis tests rely on the assumption that the data *within each group* follows a normal distribution. Furthermore, the data has changed. By averaging the observations and eliminating an outlier, the grouped data likely better resembles a normal distribution.

To accurately assess normality of each treatment group, a combination of quantitative and qualitative analysis is required. Statistical tests provide a formal but imperfect measure of the data's shape. Visual confirmation of the characteristic quantile-quantile (Q-Q) plot provides further support for the claim.

@tbl-learn-norm tabulates the results of the D'Agostino skewness test, Anscombe-Glynn kurtosis test, and the Shapiro-Wilk normality test. In each case, the null hypothesis ($H_0$) for these tests assumes that the data follows the characteristics of a normal distribution. Therefore, low p-values indicate evidence against $H_0$, suggesting that the data deviates from normality in the tested aspect.

```{r def-tab-norm}
# Function to tabulate normality test results for each treatment group
tabulate_normality_tests <-
  function(data, treatment_col, value_col) {
    levels <- levels(data[[treatment_col]])
    all_levels <- c("ALL", levels)

    results <- map(all_levels, function(level) {
      if (level == "ALL") {
        data_subset <- data |> pull(!!sym(value_col))
      } else {
        data_subset <-
          data |>
          filter((!!sym(treatment_col)) == level) |>
          pull(!!sym(value_col))
      }

      if (length(data_subset) > 0) {
        # Ensure there's data to process
        skew_test <- moments::agostino.test(data_subset)
        kurtosis_test <- moments::anscombe.test(data_subset)
        normality_test <- shapiro.test(data_subset)

        tibble(
          Group = level,
          Skewness = paste(
            round(skew_test$statistic[1], 2),
            round(skew_test$p.value, 3),
            sep = ", "
          ),
          Kurtosis = paste(
            round(kurtosis_test$statistic[1], 2),
            round(kurtosis_test$p.value, 3),
            sep = ", "
          ),
          Normality = paste(
            round(normality_test$statistic, 2),
            round(normality_test$p.value, 3),
            sep = ", "
          )
        )
      } else {
        tibble(
          Group = level,
          Skewness = NA,
          Kurtosis = NA,
          Normality = NA
        )
      }
    }) |> bind_rows()

    # Ensure the desired order, converting Group to a factor
    results <- results |>
      mutate(Group = factor(Group, levels = c("ALL", "MR", "AR", "PWI", "PAR"))) |>
      filter(!is.na(Group)) |>
      arrange(Group)

    results
  }
```


```{r tbl-learn-norm}
#| tbl-cap: "Summary of Test Statistics, p-Values for Normality Assessment"
#|
result_table <- tabulate_normality_tests(learn_agg_out, "treatment", "avg_tct_adj")
result_table |> gt()
```

From this we see that only PWI fails the normality test (p=0.02), despite showing only marginally non-significant skewness (p=0.06). Similarly, MR exhibits significant positive skew (p=0.02) but the Shapiro-Wilk test gives evidence of normality (p=0.08). Both AR and PAR are relatively consistent in confirming normality, though AR's slightly positive skew shows borderline significance (1.03, p=0.05).

::: content-hidden
TODO: correct hard-coded values in text
:::

To reconcile these findings for MR and PWI, and confirm those for AR and PAR, the set of four Q-Q plots in @fig-qq-h1a were generated. Observations for normal  data should fall along or near the line, within the shaded confidence interval. These plots show that PAR, AR, and MR all meet the expectations of normality, save for a point in the right tail of AR and MR. This aligns with the skewness and kurtosis test results. PWI, on the other hand, diverges from the line enough to leave the confidence interval, supporting the statistical evidence that it is not normally distributed.

```{r fig-qq-h1a}
#| fig-cap: Q-Q Plots of Learning Data by Treatment
#| fig-height: 3.5


# Function to create QQ plots and arrange them in a 2x2 grid
qq_plot_patchwork <-
  function(data,
           x_var,
           by_var) {
    levels <- unique(data[[by_var]])

    if (length(levels) != 4) {
      stop("This script expects exactly 4 levels in the grouping variable.")
    }

    plots <- lapply(levels, function(level) {
      p <- data |>
        filter((!!sym(by_var)) == level) |>
        ggpubr::ggqqplot(
          x = x_var,
          xlab = "",
          ylab = "",
          title = level,
          ggtheme = theme_minimal(base_size = 10) +
            theme(
              plot.margin = margin(0, 0, 0, 0),
              panel.spacing = unit(0, "lines"),
              panel.grid = element_blank(),
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              axis.line = element_blank(),
              plot.title = element_text(hjust = 0.5),
              panel.border = element_rect(
                color = outline_color,
                fill = NA,
                linewidth = 0.5
              ),
            )
        ) |>
        ggpubr::ggpar(ticks = FALSE, tickslab = FALSE)
    })

    # Arrange the plots in a 2x2 grid using patchwork
    plot_layout <-
      (plots[[1]] + plots[[2]]) / (plots[[3]] + plots[[4]]) +
      plot_annotation(title = "Quantile-Quantile Plots of Treatment Data",
                      subtitle = "PWI Data is the Most Uncharacteristic")
    
    return(plot_layout)
  }

qq_plot_patchwork(learn_agg_out, x_var = "avg_tct_adj", by_var = "treatment")


```

Given the minor deviations from normality, particularly for the PWI group, and considering the robust consistency of results between parametric and non-parametric tests, it was concluded that non-parametric methods provide the most reliable and straightforward approach for hypothesis testing. This decision avoids the complexities associated with transformation pipelines, ensuring robust statistical conclusions while maintaining the interpretability of the original data. Should future analysis require linear regression or related methods (e.g. ANOVA and its variants), the data will be transformed accordingly,

::: content-hidden
MAYBE: try to fix data transformation issues with h1a? see: _temp.qmd
:::


See @fig-h1a-plot ...

```{r fig-h1a-plot}
#| fig-cap: Task Completion Time Differences by Treatment
#| fig-height: 6
#| warning: false

h1a_plot <- ggstatsplot::ggbetweenstats(
  data = learn_agg_out,
  x = treatment,
  y = avg_tct_adj,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "TCT Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  results.subtitle = FALSE,
  subtitle = "Significant Pairwise Differences Identified using Dunn's Test, Holm-adjusted post-hoc",
  digits = 1,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

h1a_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "26.8", italic(p) == 
#     "6.4e-06", widehat(epsilon)["ordinal"]^2 == "0.5", CI["95%"] ~ 
#     "[" * "0.4", "1.0" * "]", italic("n")["obs"] == "54")
```

r ggstatsplot::extract_stats(h1a_plot)$pairwise_comparisons_data

non-parametric
centrality: median -> datawizard::describe_distribution()
between-subjects one-way ANOVA: Kruskal-Wallis one-way ANOVA -> stats::kruskal.test()
why chi-squared?
effect size estimation: rank epsilon squared -> effectsize::rank_epsilon_squared()
pairwise between-subjects test: Dunn Test -> PMCMRpluss::kwAllPairsDunnTest()
Holm's adjustment?

There are ss overall differences in TCT by treatment (KW)
Pairwise Dunn test shows PWI and PAR are both ss different from AR and MR but neither PWI / PAR (faster group) nor AR / MR (slower group) are ss different from one another
Effect size of 0.5 (moderate to large) indicates that proportion of the variance in TCT can be explained by treatment groups.
95% CI of 0.3 to 1.0 suggests at least moderate effect of treatment, but could explain up to 100% of the variance.
ES + CI -> moderate to large practical significance

```{r test-power-sim}
library(pwr)

# # Function to calculate power for Kruskal-Wallis test via simulation
simulate_kruskal_power <- function(n, k, effect_size, sig_level = 0.05, num_simulations = 1000) {
  p_values <- numeric(num_simulations)
  for (i in 1:num_simulations) {
    # Simulate data for each group with an effect size
    data <- data.frame(
      value = unlist(lapply(1:k, function(x) rnorm(n, mean = x * effect_size))),
      group = gl(k, n)
    )
    # Perform Kruskal-Wallis tes
    p_values[i] <- kruskal.test(value ~ group, data = data)$p.value
  }
  # Proportion of tests rejecting null hypothesis gives power estimate
  power <- mean(p_values < sig_level)
  return(power)
}

# # Function to run post-hoc power analysis based on observed epsilon squared
posthoc_power_analysis <- function(n, k, observed_effect_size, sig_level = 0.05, num_simulations = 1000) {
  # Convert the epsilon squared back to a standard effect size approximation for simulation
  std_effect_size <- sqrt(observed_effect_size / (1 - observed_effect_size))

  # Use simulation to estimate power as before
  power <- simulate_kruskal_power(n, k, std_effect_size, sig_level, num_simulations)
  return(power)
}

# Number of groups
k <- length(unique(learn_agg_out$treatment))

# Sample size per group
n <- round(nrow(learn_agg_out) / k)

epsilon_sq <- 0.51

# # Estimate post-hoc power
posthoc_power <- posthoc_power_analysis(n, k, epsilon_sq, sig_level = 0.05, num_simulations = 1000)
# print(posthoc_power)

# # Optional: Run with a smaller effect size to compare
epsilon_sq_smaller <- 0.2
posthoc_power_smaller <- posthoc_power_analysis(n, k, epsilon_sq_smaller, sig_level = 0.05, num_simulations = 1000)
# print(posthoc_power_smaller)
```



#### Average Error Count per Car (H1b)

$H_{1b}\textrm{: Average error count per car varies with treatment}$.

Note that elapsed time is a continuous numerical value, but error count
is discrete. This difference necessitates different approaches for
hypothesis testing and the underlying assumptions that must be checked.

#### Learning Rates (H1c)

$H_{1c}\textrm{: Learning rates vary with treatment}$.

::: content-hidden
TODO: learning add. analysis

### Additional Analyses

- what treatment group had the highest percentage of "qualified operators" at the end of the 10-min session?

#### Error Types

#### Task Completion Rate

#### First-Task Performance

#### Qualified Operators
:::

## Recall Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Overall Equipment Effectiveness (H2a)

#### Paper Work Instruction References (H2b)

::: content-hidden
TODO: recall add. analysis

### Additional Analyses

- what treatment group had the highest OEE during recall?

#### Efficiency vs. Quality

#### PWI Reference Frequency and Length

#### Error Types and Task Completion Rate

#### Ongoing Learning
:::

## Retention Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Change in OEE (H3a)

::: content-hidden
TODO: retention add. analysis

### Additional Analyses

#### Absolute and Percentage Change in OEE

#### Primary Driver of OEE Change
:::

## Workload, Usability, and User Experience

### NASA Task Load Index Results

### System Usability Scale Results

### Qualitative Feedback Analysis

::: content-hidden
TODO: exploratory analysis

## Exploratory Analyses

### Relationship between TLX, SUS, and Performance

- first task performance in learning as a measure of initial comprehension, intuitiveness of each treatment, effectiveness of instructional design, etc.; may correlate with SUS / TLX

### Impact of Participant Demographics on Performance

### Within-Group Performance Variance

### TLX Components and Performance

### Interactions between Workload, Usability, and Treatment
:::

## Challenges and Unexpected Observations {#sec-res-challenges}

Discussion of any challenges encountered during the conduct of the protocol  Modifications made to the protocol in response to these challenges

demoing AR/MR: The challenge of this choreography was compounded by significant latency in streamed video, which lagged the live action by approximately 2 seconds. With practice

> Technical Issues with Equipment: Discussion of any technical challenges encountered with the HL2 or other devices, and how these were addressed.
>
> Participant Scheduling and No-Shows: Challenges related to participant scheduling, no-shows, or last-minute cancellations, and the strategies developed to mitigate these issues.
>
> Data Collection Consistency: Any issues encountered in maintaining consistency in data collection across different trials and days, and how these were resolved.

limitations

- participant behavior
    -   few participants assigned HMDMR took advantage of the affordance
- analysis
    -   does not account for interactions between affordances

noticed improvising - discourage as occurs

sleeves, long hair problems for HL2

height and "down-lookers"; tracking dropouts;

battery life and performance; brightness and accuracy;

complexity of MR, less reliable tracking, crashes

## Summary of Key Findings
