# Results {#25-results}

```{r}
#| label: load-libs
#| output: false

library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
library(broom)
# library(glue) now included with tidyverse as str_glue
# library(gridExtra)
# library(ggtext)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r}
#| label: load-data
#| echo: false

car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results")

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics")
```

```{r}
#| label: define-functions


```

## Pilot Study

::: content-hidden
If a pilot study was conducted prior to the main study, provide details on its purpose, design, and outcomes, and how it informed the main study.
:::


### Challenges and Modifications

### Key Findings and Implications


## Participant Demographics and Sample Characteristics

```{r study-metadata}

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# result 62 matches table row count and number of appointments

# Convert the trial_datetime to Date type
demographics <- demographics |>
  mutate(trial_date = as.Date(as.POSIXct(trial_datetime, format = "%Y-%m-%d %H:%M:%S")))

# Get the first and last date
first_trial_date <- demographics |> 
  summarize(first_date = min(trial_date)) |>
  pull(first_date)

last_trial_date <- demographics |>
  summarize(last_date = max(trial_date)) |>
  pull(last_date)

```

- Experiments conducted between `r first_trial_date` and `r last_trial_date`
- Number of participants: `r participant_count`

As seen in @fig-sess1-progress.

```{r fig-sess1-progress}
#| fig-cap: "Daily Trial Administration and Completion Percentage"

# Count experiments per day
daily_counts <- demographics |>
  group_by(trial_date) |>
  summarize(daily_count = n()) |>
  arrange(trial_date)

# Calculate the cumulative count of experiments
daily_counts <- daily_counts |>
  mutate(cumulative_count = cumsum(daily_count))

# Calculate remaining experiments
total_experiments <- nrow(demographics)
daily_counts <- daily_counts |>
  mutate(remaining = total_experiments - cumulative_count, 
         percentage_remaining = (remaining / total_experiments) * 100)

trial_start_str <- str_glue("{month(first_trial_date)}/{day(first_trial_date)}")
trial_end_str <- str_glue("{month(last_trial_date)}/{day(last_trial_date)}")
trial_year_str <- str_glue("{year(last_trial_date)}")

subtitle <- str_glue("{participant_count} Learning and Recall Experiments Conducted between {trial_start_str} and {trial_end_str} of {trial_year_str}")

# Plotting
ggplot(daily_counts, aes(x = trial_date)) +
  geom_line(aes(y = daily_count, group = 1), color = "blue") +
  geom_line(aes(y = remaining, group = 1), color = "red") +
  labs(title = "First Session Progress",
       subtitle = subtitle,
       x = "Date",
       y = "Trials Administered each Day") +
  scale_y_continuous(
    sec.axis = sec_axis(~ (1 - . / total_experiments) * 100, name = "Percentage Complete")
  ) +
  theme_minimal(
    base_size = 12,
  ) +
  theme(
    plot.title.position = 'plot',
  )
```


### Overall Sample Characteristics

```{r fig-demographics}
#| fig-cap: "Participant Demographics"
#| warning: false

base_color <- "#69b3a2"

# Creating box plots for 'age'
age_plot <- demographics |> 
  ggplot(aes(x = "Age (years)", y = age)) +
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.5, color = base_color) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(
    base_size = 14,
  )

# Creating box plots for 'height'
height_plot <- demographics |> 
  ggplot(aes(x = "Height (feet)", y = height)) +
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.5, color = base_color) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(
    base_size = 14,
    )

# Bar chart for 'gender'
gender_plot <- demographics |>
  ggplot(aes(x = gender, fill = gender, alpha = gender)) +
  geom_bar(width = 0.4) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  scale_fill_manual(values = c("F" = base_color, "M" = base_color, "O" = base_color)) +
  scale_alpha_manual(values = c("F" = 0.3, "M" = 0.6, "O" = 1.0)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

### combine using patchwork
age_plot + height_plot + gender_plot +
  plot_annotation(
    title = "Age, Height, and Gender of Participants",
    subtitle = "Age reflects student-oriented nature of the sample"
  ) +
  theme_minimal(base_size = 14)
```

- Present descriptive statistics and visualizations for the entire sample (N=62)
- Discuss key demographic variables, such as age, gender, education, and prior experience

### Treatment Group Equivalence

```{r}
#| label: tbl-demo-by-treatment

# Compute summary statistics
demographics |>
  select(age:gender, treatment) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    label = "age" ~ "Age"
  ) |>
  modify_header(all_stat_cols() ~ "**{level}**<br>N = {n} ({style_percent(p)}%)") |>
  add_n() |>
  bold_labels() |>
  as_gt()
# tab_source_note("Data updated September 19, 2023")

```


- Compare the demographic characteristics across the four treatment groups
- Use statistical tests (e.g., chi-square, ANOVA) to assess the equivalence of the groups
- Discuss any notable differences or similarities between the groups

### Pilot Study Comparison

- Compare the key measures (e.g., performance, workload, usability) between the pilot study (n=8) and main study (n=54) participants
- Use descriptive statistics and visualizations to present these comparisons
- Discuss any notable similarities or differences between the two groups

### Statistical Power and Sample Size Considerations

- Present the results of statistical power analyses based on the observed effect sizes and the planned statistical tests
- Discuss the adequacy of the sample size for detecting meaningful effects
- Consider the potential impact of including or excluding the pilot study data on the statistical power and precision of the estimates
- sample size calculation / justification: power analysis, practical considerations, precedent

### Decision on Pilot Study Inclusion

- Based on the findings from 5.2.3 and 5.2.4, make a decision on whether to include the pilot study data in the main analysis
- Clearly report your decision and rationale, considering factors such as data comparability, study design modifications, and statistical power


## Learning Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Average Time per Car (H1a)

#### Average Error Count per Car (H1b)

#### Learning Rates (H1c)

### Additional Analyses

- what treatment group had the highest percentage of "qualified operators" at the end of the 10-min session?

#### Error Types

#### Task Completion Rate

#### First-Task Performance

#### Qualified Operators

## Recall Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Overall Equipment Effectiveness (H2a)

#### Paper Work Instruction References (H2b)

### Additional Analyses

- what treatment group had the highest OEE during recall?

#### Efficiency vs. Quality

#### PWI Reference Frequency and Length

#### Error Types and Task Completion Rate

#### Ongoing Learning

## Retention Phase Results

### Descriptive Statistics

### Hypothesis Testing

#### Change in OEE (H3a)

### Additional Analyses

#### Absolute and Percentage Change in OEE

#### Primary Driver of OEE Change

## Workload, Usability, and User Experience

### NASA Task Load Index Results

### System Usability Scale Results

### Qualitative Feedback Analysis

## Exploratory Analyses

### Relationship between TLX, SUS, and Performance

- first task performance in learning as a measure of initial comprehension, intuitiveness of each treatment, effectiveness of instructional design, etc.; may correlate with SUS / TLX

### Impact of Participant Demographics on Performance

### Within-Group Performance Variance

### TLX Components and Performance

### Interactions between Workload, Usability, and Treatment

## Challenges and Unexpected Observations

Discussion of any challenges encountered during the conduct of the protocol  Modifications made to the protocol in response to these challenges

demoing AR/MR: The challenge of this choreography was compounded by significant latency in streamed video, which lagged the live action by approximately 2 seconds. With practice

> Technical Issues with Equipment: Discussion of any technical challenges encountered with the HL2 or other devices, and how these were addressed.
>
> Participant Scheduling and No-Shows: Challenges related to participant scheduling, no-shows, or last-minute cancellations, and the strategies developed to mitigate these issues.
>
> Data Collection Consistency: Any issues encountered in maintaining consistency in data collection across different trials and days, and how these were resolved.

limitations

- participant behavior
  -   few participants assigned HMDMR took advantage of the affordance
- analysis
  -   does not account for interactions between affordances

noticed improvising - discourage as occurs

sleeves, long hair problems for HL2

height and "down-lookers"; tracking dropouts;

battery life and performance; brightness and accuracy; 

complexity of MR, less reliable tracking, crashes


## Summary of Key Findings
