# H1a Task Completion Time {#25a-tct}

```{r load-libs}
#| warning: false

library(tidyverse)
library(pwr)
library(emmeans)
library(performance)
library(lmerTest)
library(gt)
library(patchwork)
library(moments)
library(ggpubr)
library(ggstatsplot)
```

```{r load-data}
# Load objects from the file
load("learn.RData")
```

#### Average Time per Car

Here we will test:
$$H_{1a}\textrm{: Average time per car varies with treatment.}$$

Specifically, statistical comparison of the average time taken to complete the assembly task under various treatments will identify which instructional approach leads to the quickest task completion. As noted in the original assessment of the learning data, the overall TCT dataset exhibits skewness and kurtosis, suggesting deviations from a normal distribution. To choose the most appropriate statistical method for hypothesis testing, this must be revisited. Standard between-groups hypothesis tests rely on the assumption that the data *within each group* follows a normal distribution. Furthermore, the data has changed. By averaging the observations and eliminating an outlier, the grouped data likely better resembles a normal distribution.

To accurately assess normality of each treatment group, a combination of quantitative and qualitative analysis is required. Statistical tests provide a formal but imperfect measure of the data's shape. Visual confirmation of the characteristic quantile-quantile (Q-Q) plot provides further support for the claim. @tbl-learn-norm-tct tabulates the results of the D'Agostino skewness test, Anscombe-Glynn kurtosis test, and the Shapiro-Wilk normality test. In each case, the null hypothesis ($H_0$) for these tests assumes that the data follows the characteristics of a normal distribution. Therefore, low p-values indicate evidence against $H_0$, suggesting that the data deviates from normality in the tested aspect.

```{r def-tab-norm}
# Function to tabulate normality test results for each treatment group
tabulate_normality_tests <-
  function(data, treatment_col, value_col) {
    levels <- levels(data[[treatment_col]])
    all_levels <- c("ALL", levels)

    results <- map(all_levels, function(level) {
      if (level == "ALL") {
        data_subset <- data |> pull(!!sym(value_col))
      } else {
        data_subset <-
          data |>
          filter((!!sym(treatment_col)) == level) |>
          pull(!!sym(value_col))
      }

      if (length(data_subset) > 0) {
        # Ensure there's data to process
        skew_test <- moments::agostino.test(data_subset)
        kurtosis_test <- moments::anscombe.test(data_subset)
        normality_test <- shapiro.test(data_subset)

        tibble(
          Group = level,
          Skewness = paste(
            round(skew_test$statistic[1], 2),
            round(skew_test$p.value, 3),
            sep = ", "
          ),
          Kurtosis = paste(
            round(kurtosis_test$statistic[1], 2),
            round(kurtosis_test$p.value, 3),
            sep = ", "
          ),
          Normality = paste(
            round(normality_test$statistic, 2),
            round(normality_test$p.value, 3),
            sep = ", "
          )
        )
      } else {
        tibble(
          Group = level,
          Skewness = NA,
          Kurtosis = NA,
          Normality = NA
        )
      }
    }) |> bind_rows()

    # Ensure the desired order, converting Group to a factor
    results <- results |>
      mutate(Group = factor(Group, levels = c("ALL", "MR", "AR", "PWI", "PAR"))) |>
      filter(!is.na(Group)) |>
      arrange(Group)

    results
  }
```


```{r tbl-learn-norm-tct}
#| tbl-cap: "Summary of TCT Test Statistics, p-Values for Normality Assessment"
#|
tct_norm <- tabulate_normality_tests(learn_agg_out, "treatment", "avg_tct_adj")
tct_norm |> gt()
```

From this we see that only PWI fails the normality test (p=0.021), despite showing marginally non-significant skewness (p=0.058). Similarly, MR exhibits significant positive skew (p=0.021) but the Shapiro-Wilk test gives evidence of normality (p=0.084). Both AR and PAR are consistent in confirming normality, though PAR's slightly positive skew is marginally non-significant (p=0.068).

::: content-hidden
TODO: eliminate hard-coded values in text
:::

To reconcile these findings for MR and PWI, and confirm those for AR and PAR, the set of four Q-Q plots in @fig-qq-h1a were generated. Observations for normal data should fall along or near the line, within the shaded confidence interval. These plots show that PAR and AR both meet the expectations of normality. The MR group mostly aligns with the line, except for a single point in the right tail. This aligns with the skewness and kurtosis test results and provides insufficient evidence to reject normality. On the other hand, the PWI group diverges substantially from the line, with several points leaving the confidence interval. This supports the statistical evidence that the PWI data is not normally distributed.

```{r def-qq-plotter}

# Function to create QQ plots and arrange them in a 2x2 grid
qq_plot_patchwork <-
  function(data,
           x_var,
           by_var) {
    levels <- unique(data[[by_var]])

    if (length(levels) != 4) {
      stop("This script expects exactly 4 levels in the grouping variable.")
    }

    plots <- lapply(levels, function(level) {
      p <- data |>
        filter((!!sym(by_var)) == level) |>
        ggpubr::ggqqplot(
          x = x_var,
          xlab = "",
          ylab = "",
          title = level,
          ggtheme = theme_minimal(base_size = 10) +
            theme(
              plot.margin = margin(0, 0, 0, 0),
              panel.spacing = unit(0, "lines"),
              panel.grid = element_blank(),
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              axis.line = element_blank(),
              plot.title = element_text(hjust = 0.5),
              panel.border = element_rect(
                color = outline_color,
                fill = NA,
                linewidth = 0.5
              ),
            )
        ) |>
        ggpubr::ggpar(ticks = FALSE, tickslab = FALSE)
    })

    # Arrange the plots in a 2x2 grid using patchwork
    plot_layout <-
      (plots[[1]] + plots[[2]]) / (plots[[3]] + plots[[4]])
    
    return(plot_layout)
  }
```

```{r fig-qq-h1a}
#| fig-cap: Q-Q Plots of Learning Times by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(learn_agg_out, x_var = "avg_tct_adj", by_var = "treatment")

p + plot_annotation(title = "Quantile-Quantile Plots of TCT Treatment Data",
                    subtitle = "PWI Data is the Most Uncharacteristic of Normality")

```

Given the minor deviations from normality, particularly for the PWI group, and *considering the robust consistency of results between parametric and non-parametric tests*, it was concluded that non-parametric methods provide the most reliable and straightforward approach for hypothesis testing. This decision *avoids the complexities associated with transformation pipelines*, ensuring robust statistical conclusions while maintaining the interpretability of the original data. Should future analysis require linear regression or related methods (e.g. ANOVA and its variants), the data will be transformed accordingly,

::: content-hidden
MAYBE: try to fix data transformation issues with h1a? see: _temp.qmd
TODO: this has to be updated in light of subsequent modeling efforts
:::

@fig-h1a-plot shows the results of a non-parametric comparison of average TCT within each treatment group. It combines elements of violin, scatter, and box plots to provide a comprehensive portrayal of the data. Treatment differences were analyzed using the Kruskal-Wallis test by ranks, a non-parametric equivalent of ANOVA. Its test statistic, $H$, approximately follows a $\chi^{2}_{k-1}$ distribution, where $k$ is the number of groups. Here, $\chi^{2}_{Kruskal-Wallis}(3)=25.9$, with a p-value of 1.0e-5. These results indicate a statistically significant difference between groups. The effect size of this difference is $\hat{\epsilon}^{2}_{rank}=0.5$, with $CI_{95\%}\:[0.4,1.0]$, indicating a moderate to large treatment effect where 50% of the variance in TCT can be explained by the difference in groups. Together, these results provide strong evidence of overall differences in TCT that are both statistically and practically significant.

::: content-hidden
TODO: fix hard-coded numbers
:::

```{r fig-h1a-plot}
#| fig-cap: Task Completion Time Differences by Treatment
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

h1a_plot <- ggstatsplot::ggbetweenstats(
  data = learn_agg_out,
  x = treatment,
  y = avg_tct_adj,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "TCT Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  results.subtitle = FALSE,
  subtitle = "Significant Pairwise Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 1,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

h1a_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "25.9", italic(p) == 
#     "1.0e-05", widehat(epsilon)["ordinal"]^2 == "0.5", CI["95%"] ~ 
#     "[" * "0.4", "1.0" * "]", italic("n")["obs"] == "53")
```

::: content-hidden
TODO: check all in text values after setting standard seed before all ggbetween stats calls - see comment in code above
:::

Pairwise comparisons were conducted post-hoc using Dunn's non-parametric test for Kruskal-type ranked data. Holm's adjustment for multiple comparisons was preferred over the oft-cited Bonferroni method, which tends to be too conservative. The results show that PWI and PAR are both significantly faster than AR and MR, with all p-values less than 0.01. Statistically, PWI and PAR are both equivalent and faster than AR and MR, which are also equivalent to one another.

```{r test-power-sim}
# # Function to calculate power for Kruskal-Wallis test via simulation
simulate_kruskal_power <- function(n, k, effect_size, sig_level = 0.05, num_simulations = 1000) {
  p_values <- numeric(num_simulations)
  for (i in 1:num_simulations) {
    # Simulate data for each group with an effect size
    data <- data.frame(
      value = unlist(lapply(1:k, function(x) rnorm(n, mean = x * effect_size))),
      group = gl(k, n)
    )
    # Perform Kruskal-Wallis tes
    p_values[i] <- kruskal.test(value ~ group, data = data)$p.value
  }
  # Proportion of tests rejecting null hypothesis gives power estimate
  power <- mean(p_values < sig_level)
  return(power)
}

# # Function to run post-hoc power analysis based on observed epsilon squared
posthoc_power_analysis <- function(n, k, observed_effect_size, sig_level = 0.05, num_simulations = 1000) {
  # Convert the epsilon squared back to a standard effect size approximation for simulation
  std_effect_size <- sqrt(observed_effect_size / (1 - observed_effect_size))

  # Use simulation to estimate power as before
  power <- simulate_kruskal_power(n, k, std_effect_size, sig_level, num_simulations)
  return(power)
}

k <- length(unique(learn_agg_out$treatment))
n <- round(nrow(learn_agg_out) / k)
epsilon_sq <- 0.5

# Estimate post-hoc power
posthoc_power <- posthoc_power_analysis(n, k, epsilon_sq, sig_level = 0.05, num_simulations = 1000)
```

A post-hoc simulation was run to validate these results by estimating the statistical power. Power analysis measures the effectiveness of a statistical test in detecting true differences when they exist. High power (typically 0.8 or above) suggests that the test is appropriate for the given data and experimental conditions and therefore aligns with the goal of making correct decisions based on the data. The simulation repeatedly applies the Kruskal-Wallis test to data generated based on the observed mean and effect size. The resulting power is the proportion of tests that correctly reject the null hypothesis. The outcome of `r posthoc_power` (`r posthoc_power * 100`% power), while uncommon, suggests that, for the conditions of this experiment (n=53, an average of 13.25 samples in each of the four groups, $\hat{\epsilon}^{2}=0.5$, and $\alpha = 0.05$), the test is extremely effective at detecting the observed differences. The combination of a large effect size and a reasonable sample size is largely attributable to this value, which offers great confidence in the validity of the test results.

*Unfortunately, all of this is moot. The Kruskal-Wallis test has three assumptions that must be met: (1) ordinal or continuous response variable, (2) independent observations in each group, and (3) similar distributions in each group. Task completion time is indeed continuous and previous analysis showed that the random sampling method ensured equivalent treatment distributions. However, the experimental design violates the independent observations assumption due to the repeated measures taken from each participant. As a result, other methods must be employed.*

Mixed effects models are best-suited for repeated measures or nested data structures. Repeated measures ANOVA was also considered but ultimately rejected as it also assumes independence of observations within groups. A simple form of mixed-effects model for this problem could be:

$$
\text{TCT}_{i} = \beta_{0} + \beta_{1} \text{TRT}_{i} + (1 \mid \text{PART}_{i}) + \epsilon_{i}
$$

Where:

- $i$: Index for the observation.
- $\beta_0$: Overall intercept, a baseline TCT for the reference treatment group.
- $\beta_1$: Fixed effect of the treatment (TRT).
- $(1 \mid \text{participant}_i)$: Random effect for each observation, from the participant.
- $\epsilon_{ij}$: Residual error term for each observation.

This formulation allows the model to capture individual variability by allowing for each participant to have a different initial TCT (intercept). This accounts for variation in TCT due to differences between participants (random effects) and within each participant's repeated measures (residuals). Using this approach controls for expected differences among participants. Furthermore, it naturally accounts for the unbalanced nature of the data, where the number of tasks completed differs for each participant. Finally, it also accounts for the inherent non-independence of repeated measures for each participant.

A linear mixed model was fit to predict TCT, using the formulation described above. The `lmer` function from the `{lme4}` package was used to perform the fit using residual maximum likelihood method (REML). Prior to the fit, TCT was log-transformed to improve model performance and diagnostics.

*The resulting model was evaluated using...*

```{r}

# create data
learn_h1a <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj)

# mutate
learn_h1a <- learn_h1a |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

glimpse(learn_h1a)
```

Compare model with different transformations of the response to deal with its non-linearity (exponential decay)

```{r}
# model comparison
m1 <- lmer(log(tct) ~ treatment + (1 | participant), data = learn_h1a)
m2 <- lmer(sqrt(tct) ~ treatment + (1 | participant), data = learn_h1a)
m3 <- lmer(inverse(tct) ~ treatment + (1 | participant), data = learn_h1a)

compare_performance(m1, m2, m3, rank = T)
```

Best model uses a simple inverse transformation

```{r}
# plots comparing tct with inverse(tct)

# Create a data frame with the original and inverse transformed TCT values
pd <- learn_h1a %>%
  mutate(inverse_tct = 1 / tct)

# Plot for tct
plot_tct <- ggplot(pd, aes(x = seq, y = tct, color = treatment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "TCT vs. Sequence",
    x = "Sequence",
    y = "Task Completion Time (TCT)"
  ) +
  theme_minimal()

# Plot for inverse(tct)
plot_inverse_tct <- ggplot(pd, aes(x = seq, y = inverse_tct, color = treatment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Inverse(TCT) vs. Sequence",
    x = "Sequence",
    y = "Inverse Task Completion Time (1/TCT)"
  ) +
  theme_minimal()

# Display the plots
p <- plot_tct + plot_inverse_tct

print(p)
```

Check performance:

```{r}
plot(performance::check_model(m3, panel = F))
```

Model summary:

```{r}
summary(m3)
```

Estimated marginal means for pairwise comparison:

```{r}
emmeans_tct <- emmeans::emmeans(
  m3, pairwise ~ treatment, type = "response"
)

summary(emmeans_tct)

```

::: content-hidden
*this text needs to be updated:*

The total explanatory power of the resulting model is substantial (conditional $R^2$ = 0.89, marginal $R^2$ = 0.47). These results indicate that the model explains 89% of the variance in TCT, of which 47% is related to the fixed effects alone. This highlights the importance of individual differences among participants in the data, validating the mixed-effects approach. The model's intercept, corresponding to the AR treatment, is at 0.99 (95% CI [0.62, 1.36], p < .001). Within this model:

  - The effect of treatment MR is statistically non-significant and positive (beta = 0.03, 95% CI [-0.48, 0.54], t(264) = 0.13, p = 0.899; Std. beta = 0.03, 95% CI [-0.48, 0.54])
  - The effect of treatment PWI is statistically significant and negative (beta = -1.54, 95% CI [-2.05, -1.04], t(264) = -6.00, p < .001; Std. beta = -1.54, 95% CI [-2.05, -1.04])
  - The effect of treatment PAR is statistically significant and negative (beta = -1.05, 95% CI [-1.58, -0.53], t(264) = -3.95, p < .001; Std. beta = -1.05, 95% CI [-1.58, -0.53])

:::

These results are visualized in @fig-h1a-fixedeff, where overlapping arrows indicate statistically insignificant differences in TCT between treatments:

```{r fig-h1a-fixedeff}
#| fig-caption: This is the caption
#| fig-height: 3.5

plot(emmeans_tct$emmeans, comparisons = TRUE) +
  labs(
    title = "Estimated Means by Treatment",
    subtitle = "Talk about observed differences",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(
      margin = margin(b = 5, unit = "pt"),
      color = "black"
    )
  )

```

The `{emmeans}` package was used for post-hoc pairwise comparison, generating the following contrasts based on estimated marginal means:

```text
 contrast   estimate      SE   df t.ratio p.value
 AR - MR   -5.44e-05 0.00115 53.5  -0.047  1.0000
 AR - PWI  -6.80e-03 0.00112 48.5  -6.086  <.0001
 AR - PAR  -4.04e-03 0.00116 48.9  -3.470  0.0059
 MR - PWI  -6.75e-03 0.00110 49.2  -6.128  <.0001
 MR - PAR  -3.98e-03 0.00115 49.6  -3.470  0.0058
 PWI - PAR  2.76e-03 0.00112 44.9   2.469  0.0788

Note: contrasts are still on the inverse scale 
Degrees-of-freedom method: kenward-roger 
P value adjustment: tukey method for comparing a family of 4 estimates
```

Significant differences are identified between PWI and both AR and MR, and between PAR and AR and MR. No significant difference was found between PWI and PAR or AR and MR. This matches the previous findings, providing robust statistical evidence that PWI and PAR are both equivalent and faster than AR and MR, which are also equivalent to one another.


```{r eval=FALSE}
# Extract estimated marginal means and confidence intervals
emmeans_df <- as.data.frame(emmeans_tct)

# Extract pairwise comparisons
pairwise_df <- as.data.frame(pairs(emmeans_tct))

# Create a new column in emmeans_df with the back-transformed emmeans and confidence intervals
emmeans_df <- emmeans_df %>%
  mutate(
    tct = inverse_log_shifted(emmean, shift_amount),
    lower.CL_orig = inverse_log_shifted(lower.CL, shift_amount),
    upper.CL_orig = inverse_log_shifted(upper.CL, shift_amount)
  )

# Violin plot of the raw data
p <-
  ggplot(learn_h1a, aes(x = treatment, y = tct, fill = treatment)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_jitter(width = 0.1,
              size = 2,
              alpha = 0.5) +
  stat_summary(
    fun = median,
    geom = "point",
    size = 3,
    color = "red"
  ) +
  theme_minimal() +
  labs(
    title = "TCT Differences by Treatment",
    subtitle = "Violin Plot with Estimated Marginal Means and Pairwise Comparisons",
    x = "Treatment",
    y = "Task Completion Time (TCT)"
  )

# Add estimated marginal means with confidence intervals
p <- p + geom_point(
  data = emmeans_df,
  aes(x = treatment, y = tct),
  size = 3,
  color = "blue"
) +
  geom_errorbar(
    data = emmeans_df,
    aes(x = treatment, ymin = lower.CL_orig, ymax = upper.CL_orig),
    width = 0.2,
    color = "blue"
  )


# Add pairwise comparisons
p <-
  p + geom_text(
    data = pairwise_df,
    aes(
      x = contrast,
      y = max(learn_h1a$tct) + 20,
      label = paste("p =", format(
        p.value, digits = 2, scientific = TRUE
      ))
    ),
    vjust = -0.5,
    color = "black"
  )

# Rotate x-axis labels for better readability
p <- p + theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the plot
print(p)

```

```{r}
save(
  tabulate_normality_tests, 
  qq_plot_patchwork,
  simulate_kruskal_power,
  posthoc_power_analysis,
  file = "h1a.RData")
```

