# H1a Task Completion Time {#25a-tct}

```{r load-libs}
#| warning: false

library(tidyverse)
library(pwr)
library(emmeans)
library(performance)
library(lmerTest)
library(gt)
library(patchwork)
library(moments)
library(ggpubr)
library(ggstatsplot)
```

```{r load-data}
# Load objects from the file
load("learn.RData")
```

#### Average Time per Car





*Unfortunately, all of this is moot. The Kruskal-Wallis test has three assumptions that must be met: (1) ordinal or continuous response variable, (2) independent observations in each group, and (3) similar distributions in each group. Task completion time is indeed continuous and previous analysis showed that the random sampling method ensured equivalent treatment distributions. However, the experimental design violates the independent observations assumption due to the repeated measures taken from each participant. As a result, other methods must be employed.*


*The resulting model was evaluated using...*

```{r}

# create data
learn_h1a <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj)

# mutate
learn_h1a <- learn_h1a |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

glimpse(learn_h1a)
```

Compare model with different transformations of the response to deal with its non-linearity (exponential decay)

```{r}
# model comparison
m1 <- lmer(log(tct) ~ treatment + (1 | participant), data = learn_h1a)
m2 <- lmer(sqrt(tct) ~ treatment + (1 | participant), data = learn_h1a)
m3 <- lmer(inverse(tct) ~ treatment + (1 | participant), data = learn_h1a)

compare_performance(m1, m2, m3, rank = T)
```

Best model uses a simple inverse transformation

```{r}
# plots comparing tct with inverse(tct)

# Create a data frame with the original and inverse transformed TCT values
pd <- learn_h1a %>%
  mutate(inverse_tct = 1 / tct)

# Plot for tct
plot_tct <- ggplot(pd, aes(x = seq, y = tct, color = treatment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "TCT vs. Sequence",
    x = "Sequence",
    y = "Task Completion Time (TCT)"
  ) +
  theme_minimal()

# Plot for inverse(tct)
plot_inverse_tct <- ggplot(pd, aes(x = seq, y = inverse_tct, color = treatment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Inverse(TCT) vs. Sequence",
    x = "Sequence",
    y = "Inverse Task Completion Time (1/TCT)"
  ) +
  theme_minimal()

# Display the plots
p <- plot_tct + plot_inverse_tct

print(p)
```

Check performance:

```{r}
plot(performance::check_model(m3, panel = F))
```

Model summary:

```{r}
summary(m3)
```

Estimated marginal means for pairwise comparison:

```{r}
emmeans_tct <- emmeans::emmeans(
  m3, pairwise ~ treatment, type = "response"
)

summary(emmeans_tct)

```

::: content-hidden
*this text needs to be updated:*

The total explanatory power of the resulting model is substantial (conditional $R^2$ = 0.89, marginal $R^2$ = 0.47). These results indicate that the model explains 89% of the variance in TCT, of which 47% is related to the fixed effects alone. This highlights the importance of individual differences among participants in the data, validating the mixed-effects approach. The model's intercept, corresponding to the AR treatment, is at 0.99 (95% CI [0.62, 1.36], p < .001). Within this model:

  - The effect of treatment MR is statistically non-significant and positive (beta = 0.03, 95% CI [-0.48, 0.54], t(264) = 0.13, p = 0.899; Std. beta = 0.03, 95% CI [-0.48, 0.54])
  - The effect of treatment PWI is statistically significant and negative (beta = -1.54, 95% CI [-2.05, -1.04], t(264) = -6.00, p < .001; Std. beta = -1.54, 95% CI [-2.05, -1.04])
  - The effect of treatment PAR is statistically significant and negative (beta = -1.05, 95% CI [-1.58, -0.53], t(264) = -3.95, p < .001; Std. beta = -1.05, 95% CI [-1.58, -0.53])

:::

These results are visualized in @fig-h1a-fixedeff, where overlapping arrows indicate statistically insignificant differences in TCT between treatments:

```{r fig-h1a-fixedeff}
#| fig-caption: This is the caption
#| fig-height: 3.5

plot(emmeans_tct$emmeans, comparisons = TRUE) +
  labs(
    title = "Estimated Means by Treatment",
    subtitle = "Talk about observed differences",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(
      margin = margin(b = 5, unit = "pt"),
      color = "black"
    )
  )

```

The `{emmeans}` package was used for post-hoc pairwise comparison, generating the following contrasts based on estimated marginal means:

```text
 contrast   estimate      SE   df t.ratio p.value
 AR - MR   -5.44e-05 0.00115 53.5  -0.047  1.0000
 AR - PWI  -6.80e-03 0.00112 48.5  -6.086  <.0001
 AR - PAR  -4.04e-03 0.00116 48.9  -3.470  0.0059
 MR - PWI  -6.75e-03 0.00110 49.2  -6.128  <.0001
 MR - PAR  -3.98e-03 0.00115 49.6  -3.470  0.0058
 PWI - PAR  2.76e-03 0.00112 44.9   2.469  0.0788

Note: contrasts are still on the inverse scale 
Degrees-of-freedom method: kenward-roger 
P value adjustment: tukey method for comparing a family of 4 estimates
```

Significant differences are identified between PWI and both AR and MR, and between PAR and AR and MR. No significant difference was found between PWI and PAR or AR and MR. This matches the previous findings, providing robust statistical evidence that PWI and PAR are both equivalent and faster than AR and MR, which are also equivalent to one another.


```{r eval=FALSE}
# Extract estimated marginal means and confidence intervals
emmeans_df <- as.data.frame(emmeans_tct)

# Extract pairwise comparisons
pairwise_df <- as.data.frame(pairs(emmeans_tct))

# Create a new column in emmeans_df with the back-transformed emmeans and confidence intervals
emmeans_df <- emmeans_df %>%
  mutate(
    tct = inverse_log_shifted(emmean, shift_amount),
    lower.CL_orig = inverse_log_shifted(lower.CL, shift_amount),
    upper.CL_orig = inverse_log_shifted(upper.CL, shift_amount)
  )

# Violin plot of the raw data
p <-
  ggplot(learn_h1a, aes(x = treatment, y = tct, fill = treatment)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_jitter(width = 0.1,
              size = 2,
              alpha = 0.5) +
  stat_summary(
    fun = median,
    geom = "point",
    size = 3,
    color = "red"
  ) +
  theme_minimal() +
  labs(
    title = "TCT Differences by Treatment",
    subtitle = "Violin Plot with Estimated Marginal Means and Pairwise Comparisons",
    x = "Treatment",
    y = "Task Completion Time (TCT)"
  )

# Add estimated marginal means with confidence intervals
p <- p + geom_point(
  data = emmeans_df,
  aes(x = treatment, y = tct),
  size = 3,
  color = "blue"
) +
  geom_errorbar(
    data = emmeans_df,
    aes(x = treatment, ymin = lower.CL_orig, ymax = upper.CL_orig),
    width = 0.2,
    color = "blue"
  )


# Add pairwise comparisons
p <-
  p + geom_text(
    data = pairwise_df,
    aes(
      x = contrast,
      y = max(learn_h1a$tct) + 20,
      label = paste("p =", format(
        p.value, digits = 2, scientific = TRUE
      ))
    ),
    vjust = -0.5,
    color = "black"
  )

# Rotate x-axis labels for better readability
p <- p + theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the plot
print(p)

```

```{r}
save(
  tabulate_normality_tests, 
  qq_plot_patchwork,
  simulate_kruskal_power,
  posthoc_power_analysis,
  file = "h1a.RData")
```

