```{r load-libs}
#| warning: false
#| message: false

library(tidyverse)
library(ggstatsplot)
library(lmerTest)
library(report)
library(patchwork)
library(car)
library(boot)
library(gt)
library(paletteer)
library(english)
library(glue)
library(modelsummary)
library(tictoc)
library(parallel)
library(doParallel)
library(foreach)

# library(pwr)
# library(emmeans)
# library(performance)
# library(moments)
# library(ggpubr)
```

```{r load-stuff}
# load results data
load("res_f.RData")

# load shared objects
load("shared.RData")
```

```{r define-constants}
TREATMENTS <- levels(res_f$treatment)

ORDERED_PAIRS <- c()
for (i in 1:(length(TREATMENTS) - 1)) {
  for (j in (i + 1):length(TREATMENTS)) {
    ORDERED_PAIRS <- c(ORDERED_PAIRS, paste(TREATMENTS[i], "-", TREATMENTS[j]))
  }
}

# set global flag variable used for logging current statistic
CUR_STAT <- NULL
```

```{r def-boot-funcs}
# Hodges-Lehmann Estimator - stat_func option
hl_estimator <- function(x) {
  n <- length(x)
  pairwise_means <- outer(x, x, FUN = function(a, b) (a + b) / 2)
  median(pairwise_means[lower.tri(pairwise_means, diag = TRUE)])
}

# Generalized function to perform bootstrap and get results
bootstrap_stat_diff <- function(data, group1, group2, column, stat_func) {
  # cat("Starting bootstrap_stat_diff for", group1, "vs", group2, "\n")
  
  stat_diff <- function(data, indices, group1, group2, column, stat_func) {
    # static_counter <- 0
    # static_counter <- static_counter + 1
    # cat("stat_diff called for the", static_counter, "time for this resample\n")

    d <- data[indices, ]
    g1_data <- d[d$treatment == group1, column, drop = TRUE]
    g2_data <- d[d$treatment == group2, column, drop = TRUE]
    
    # cat(group1, "-", paste(sprintf("%.2f", g1_data), collapse = " "), "\n")
    # cat(group2, "-", paste(sprintf("%.2f", g2_data), collapse = " "), "\n")

    # cat(
    #   "Subset sizes: ", group1, "=", length(g1_data), ", ",
    #   group2, "=", length(g2_data), "\n"
    # )
    
    stat_func(g1_data) - stat_func(g2_data)
  }

  # Define statistic function for bootstrap
  stat_wrapper <- function(data, indices) {
    stat_diff(data, indices, group1, group2, column, stat_func)
  }

  set.seed(42)  # Set seed right before boot call
  
  # browser()

  results <- boot(
    data = data,
    statistic = stat_wrapper,
    R = 10000,
    strata = data$treatment,
    # parallel = "multicore"
  )
  
  # cat("\nBootstrap results for", group1, "vs", group2, ":\n")
  # print(results)

  # Get bootstrap confidence intervals
  ci <- boot.ci(results, type = "perc")
  
  # cat("\n\nConfidence intervals for", group1, "vs", group2, ":\n")
  # print(ci)

  # Create a summary of bootstrap results
  summary <- list(
    original = results$t0,
    bias = mean(results$t) - results$t0,
    std_error = sd(results$t),
    ci_lower = ci$percent[4],
    ci_upper = ci$percent[5]
  )

  # cat("Returning summary for", group1, "vs", group2, "\n")
  return(summary)
}

# Function to perform multiple replications of the bootstrap
multiple_bootstrap <- function(data, group1, group2, column, stat_func, num_reps) {
  # cat("Starting multiple_bootstrap for", group1, "and", group2, "\n")
  
  # # original parallel implementation, results in invalid results
  # all_results <- foreach(
  #     i = 1:num_reps,
  #     .combine = list,
  #     .packages = c("boot", "dplyr"),
  #     .errorhandling = "remove"  # was stop
  #   ) %dopar% {
  #     # cat("Starting replication", i, "\n")
  #     tryCatch({
  #       result <- bootstrap_stat_diff(data, group1, group2, column, stat_func)
  #       # cat("Completed replication", i, "\n")
  #       result
  #     }, error = function(e) {
  #       cat("*** Error in replication", i, ":", conditionMessage(e), "\n")
  #       NULL
  #     })
  #   }
  
  # cat("Num reps =", num_reps, "\n")
  
  # working non-parallel implementation
  all_results <- vector("list", num_reps)  # Initialize an empty list of the correct length

  for (i in 1:num_reps) {
    # cat("\nStarting replication", i, "\n")
    tryCatch({
      result <- bootstrap_stat_diff(data, group1, group2, column, stat_func)
      all_results[[i]] <- result
    }, error = function(e) {
      cat("*** Error in replication", i, ":", conditionMessage(e), "\n")
      all_results[[i]] <- NULL
    })
  }

  cat("Completed all replications. Number of results:", length(all_results), "\n")
  
  # Check for errors
  errors <- sapply(all_results, is.null)
  if (any(errors)) {
    cat("*** Errors occurred during bootstrap for the following replications:\n")
    cat(paste("Replication", which(errors), "\n", collapse = ""))
  }
  
  # Process successful results
  results_list <- lapply(seq_along(all_results), function(i) {
    res <- all_results[[i]]
    if (is.null(res)) {
      cat("*** Warning: Missing result for replication", i, "\n")
      return(NULL)
    }

    if (!is.list(res) || !all(c("original", "bias", "std_error", "ci_lower", "ci_upper") %in% names(res))) {
      cat("*** Warning: Invalid result for replication", i, "\n")
      return(NULL)
    }

    list(
      Replication = i,
      original = res$original,
      bias = res$bias,
      std_error = res$std_error,
      ci_lower = res$ci_lower,
      ci_upper = res$ci_upper
    )
  })

  results_list <- results_list[!sapply(results_list, is.null)]

  if (length(results_list) == 0) {
    stop("*** No valid results were produced. ***")
  }

  cat("Processed", length(results_list), "valid results.\n")

  return(do.call(rbind, results_list))
}

# Generalized pairwise tests function to capture results for each replication
bootstrap_pairs <- function(data, column, stat_func, num_reps = 5) {
  set.seed(42)
  cat("+++ Starting bootstrap for", CUR_STAT, "of", column, "...\n")
  msg = paste("+++ Bootstrap for", column, CUR_STAT, "complete")
  tic(msg)
  results <- data.frame(
    Group1 = character(),
    Group2 = character(),
    Replication = numeric(),
    Mean_Difference = numeric(),
    Bias = numeric(),
    Std_Error = numeric(),
    CI_Lower = numeric(),
    CI_Upper = numeric(),
    Sig = character(),
    stringsAsFactors = FALSE
  )

  for (i in 1:(length(TREATMENTS) - 1)) {
    for (j in (i + 1):length(TREATMENTS)) {
      group1 <- TREATMENTS[i]
      group2 <- TREATMENTS[j]
      
      cat("Starting bootstrap for", group1, "vs", group2, "\n")

      # Perform bootstrap test
      test_results <- multiple_bootstrap(data, group1, group2, column, stat_func, num_reps)
      
      # cat("Completed bootstrap for", group1, "vs", group2, "\n")

      # Add results to the dataframe
      for (k in 1:nrow(test_results)) {
        rep_result <- test_results[k, ]
        results <- rbind(results, data.frame(
          Group1 = group1,
          Group2 = group2,
          Replication = rep_result$Replication,
          Mean_Difference = rep_result$original,
          Bias = rep_result$bias,
          Std_Error = rep_result$std_error,
          CI_Lower = rep_result$ci_lower,
          CI_Upper = rep_result$ci_upper,
          Sig = ifelse(rep_result$ci_lower > 0 | rep_result$ci_upper < 0, "*", "")
        ))
      }
    }
  }

  toc()
  return(results)
}

# Function to summarize pairwise results
summarize_pairwise_results <- function(pairwise_results) {
    result_summary <- pairwise_results |>
    filter(is.finite(Bias) & is.finite(Std_Error)) |>
    group_by(Group1, Group2) |>
    # summarize across all replications
    summarise(
      Diff = mean(Mean_Difference, na.rm = TRUE),
      Diff_L = mean(CI_Lower, na.rm = TRUE),
      Diff_U = mean(CI_Upper, na.rm = TRUE),
      Bias_mean = mean(Bias, na.rm = TRUE),
      Bias_sd = sd(Bias, na.rm = TRUE),
      SE_mean = mean(Std_Error, na.rm = TRUE),
      SE_sd = sd(Std_Error, na.rm = TRUE),
      Sig = ifelse(
        mean(CI_Lower > 0, na.rm = TRUE) | mean(CI_Upper < 0, na.rm = TRUE), "*", ""
      ),
      .groups = "drop"
    ) |>
    mutate(
      Pair = paste(Group1, Group2, sep = " - "),
      Pair = factor(Pair, levels = ORDERED_PAIRS)
    ) |>
    arrange(Pair) |>
    select(Pair, Diff:Sig)
  
  return(result_summary)
}

# Wrapper function to run bootstrap pairs for each stat function and combine the results
run_bootstrap_tests <- function(data, column, stat_funcs) {
  results <- map_dfr(names(stat_funcs), function(stat_name) {
    stat_func <- stat_funcs[[stat_name]]
    
    # set global for current stat
    assign("CUR_STAT", stat_name, envir = .GlobalEnv)
    
    # Perform bootstrap pairs
    pairwise_results <- bootstrap_pairs(data, column = column, stat_func = stat_func)

    # Summarize the results
    summary_results <- summarize_pairwise_results(pairwise_results)

    # Add the statistic column
    summary_results <- summary_results |> 
      mutate(Stat = stat_name)

    return(summary_results)
  })
  
  # Ensure 'Stat' column correctly placed
  results <- results |> 
    select(Pair, Stat, Diff:Sig)
  
  return(results)
}

create_boot_summary_table <- function(data, title, subtitle) {
  data |>
    filter(Sig == "*") |> 
    select(-Sig) |> 
    gt() |>
    tab_header(
      title = title,
      subtitle = subtitle
    ) |>
    cols_label(
      Pair = "Comparison",
      Diff = "OEE Difference",
      Diff_L = "Lower CI",
      Diff_U = "Upper CI",
      Bias_mean = "Bias",
      Bias_sd = "sd_Bias",
      SE_mean = "Std Error",
      SE_sd = "sd_SE"
    ) |>
    cols_merge(
      columns = c(Bias_mean, Bias_sd),
      pattern = "{1} ({2})"
    ) |>
    cols_merge(
      columns = c(SE_mean, SE_sd),
      pattern = "{1} ({2})"
    ) |>
    cols_merge(
      columns = c(Diff_L, Diff_U),
      pattern = "[{1}, {2}]"
    ) |>
    cols_label(
      .list = c(
        "Diff" = "Mean",
        "Diff_L" = "95% CI"
      )
    ) |>
    fmt_number(
      decimals = 3
    ) |>
    cols_align(align = "center") |>
    tab_options(
      table.font.size = "smaller",
      column_labels.font.weight = "bold",
      heading.title.font.weight = "bolder",
      heading.subtitle.font.size = px(13)
    ) |> 
    tab_footnote(
      footnote = "Significance indicated by confidence intervals that do not include zero."
    )
}

```

```{r recall-data}

# drop outlier from source
res_f <- res_f |>
  filter(participant != 1063)

recall <- res_f |>
  filter(
    phase == 2,
    car_outcome == "complete",
  ) |>
  select(
    participant,
    car_elapsed_time,
    uce,
    treatment,
    drop_event_count,
    drop_event_tot_time_adj,
    repair_event_count,
    repair_event_tot_time,
    system_event_count,
    system_event_tot_time,
    pwi_event_count,
    pwi_event_tot_time
  )

# add tot_sys_adj, tot_usr_adj, and adjusted TCT time to learn
# tot_sys_adj will be used to calculate OEE
# tot_usr_adj just to reduce TCT

recall <- recall |>
  mutate(
    # Replace NA with 0 using coalesce for safe addition
    tot_sys_adj = coalesce(drop_event_tot_time_adj, 0) +
      coalesce(system_event_tot_time, 0),
    tot_usr_adj = coalesce(repair_event_tot_time, 0),
    tot_adj = tot_sys_adj + tot_usr_adj,
    car_elapsed_time_adj = car_elapsed_time - tot_adj,
    pwi_event_count = replace_na(pwi_event_count, 0),
    pwi_event_tot_time = replace_na(pwi_event_tot_time, 0)
  )


changes <- recall |>
  filter(car_elapsed_time != car_elapsed_time_adj)

# only 1057 had deductions for system events; none for user events
# what were the source of these in recall, which is PWI only?
# hololens menu popping up inadvertantly - lost time dismissing it

total_tasks_completed <- nrow(recall) # 212
total_tasks_adjusted <- nrow(changes) # 2

recall <- recall |>
  select(
    participant,
    treatment,
    car_elapsed_time_adj,
    uce,
    pwi_event_count,
    pwi_event_tot_time,
    tot_sys_adj,
    tot_usr_adj
  )

save(recall, file = "recall.RData")
```

```{r make-h2a-data}
# calculate OEE from filtered res_f data
oee_data <- res_f |>
  filter(
    phase == 2,
    car_outcome == "complete",
  ) |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    session_start = min(car_start_time, na.rm = TRUE),
    session_end = max(car_end_time, na.rm = TRUE),
    session_duration = session_end - session_start,
    units_produced = n(),
    units_defective = sum(uce > 0, na.rm = TRUE),
    tot_sys_adj = sum(
      coalesce(drop_event_tot_time_adj, 0)
      + coalesce(system_event_tot_time, 0),
      na.rm = TRUE
    ),
    prod = units_produced * 60 / session_duration,
    qual = (units_produced - units_defective) / units_produced,
    avail = (session_duration - tot_sys_adj) / session_duration,
    oee_100 = 1 * prod * qual,
    oee_85 = 0.85 * prod * qual,
    oee_avail = avail * prod * qual
  )
```

```{r make-h2b-data}
recall_agg <- recall |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    tot_pwi_count = sum(pwi_event_count, na.rm = TRUE),
    tot_pwi_time = sum(pwi_event_tot_time, na.rm = TRUE),
    task_num = n()
  ) |> 
  mutate(
    avg_pwi_count = tot_pwi_count / task_num,
    avg_pwi_time = tot_pwi_time / task_num,
    pwi_time_per_ref_agg = ifelse(avg_pwi_count > 0, avg_pwi_time / avg_pwi_count, 0)
  )
```

## Recall Phase Results

As with the learning phase results, only completed tasks are considered for recall analysis. Paper work instructions were available for reference, and participants wore the HL2 only for recording purposes. This almost entirely eliminated system-related interruptions, though participant 1057 inadvertently triggered the HL2 menu twice during their trial. The time lost to those events was deducted from measured task times as before. `r total_tasks_completed` observations were recorded during this phase, as expected for `r length(unique(recall$participant))` participants, each with four iterations.

### $H_{2a}$: Overall Equipment Effectiveness

The first hypothesis of the recall phase:
$$H_{2a}\textrm{: OEE varies with treatment}$$

is designed to determine if the instructional method has an impact on overall performance after the initial training period. As described in @sec-calc-meas, OEE is the product of performance (task rate relative to takt time), quality (percentage of units produced without defects), and availability (percentage of system up time). For the purposes of this analysis, 100% availability is assumed. Takt time is set to 60 seconds in accordance with the line policy and instructional design.

#### Descriptive Statistics

```{r make-h2a-report}
# report descriptive stats for cet and uce
h2a_rpt <- oee_data |>
  select(prod, qual, oee_100) |>
  rename(
    "Productivity" = prod,
    "Quality" = qual,
    "Overall Equipment Effectiveness" = oee_100
  ) |>
  report::report() # options not working!

# Extract and split the string by the pattern "\n - ", remove first line
h2a_rpt_list <- strsplit(h2a_rpt[1], "\n  - ")[[1]][-1]

# Add hyphen to first element and rejoin
h2a_rpt_list[1] <- paste0("- ", h2a_rpt_list[1])
h2a_rpt_list <- paste(h2a_rpt_list, collapse = "\n- ")
h2a_rpt_list <- gsub(", 0 missing", "", h2a_rpt_list)
h2a_rpt_list <- gsub("n = 54, ", "", h2a_rpt_list)

# cat(h2a_rpt_list)
```

The calculated measures are summarized below:

`r h2a_rpt_list`

As this is aggregated data, there is one observation for each participant, and no missing values. Productivity and quality are slightly skewed right (0.20) and left (-0.25), respectively. Kurtosis measures suggest that productivity is close to normal (1.28), but quality has a flatter distribution (-1.96). Productivity exhibits relatively low variability and similar mean (0.84) and median (0.82), suggesting consistency among participants. For quality, the median (1.00) and its absolute deviation (MAD = 0.00) show that many participants achieved perfect quality during recall. However, the range (0, 1) and high standard deviation (0.49) relative to its mean (0.56) indicates considerable variability. In fact, several participants produced *only* defective assemblies. Note that quality is essentially a discrete variable, as it can only take on values of $n/4$, where $n = 0, 1, \dots, 4$, corresponding to each of the four task iterations.

OEE shows a notably asymmetric distribution with a mean of 0.48 and a higher median of 0.62, indicating a significant number of low values dragging down the mean. A high concentration of zeros are observed in the OEE histogram, @fig-oee-hist, suggesting that many participants experience zero effectiveness due to high defect rates. The standard deviation (sd = 0.43) and median absolute deviation (MAD = 0.58) reflect substantial variability in effectiveness across participants. Despite the approximately symmetric skewness (-0.07), the range (0 to 1.14) indicates diverse levels of effectiveness, from complete failures to high performance. These findings highlight the variability in participants' abilities to achieve consistent and effective output.

```{r fig-oee-hist}
#| fig-cap: Histogram of Overall Equipment Effectiveness in Recall
#| fig-height: 3.5
#| warning: false
#| message: false

oee_data |>
  ggplot(aes(x = oee_100)) +
  geom_histogram(
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color,
    binwidth = 0.1
  ) +
  labs(
    title = "Distribution of OEE Values",
    subtitle = "High defect rates have an obvious effect on effectiveness",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
  )
```

```{r oee-tests}
st_oee_treat <- by(oee_data$oee_100, oee_data$treatment, shapiro.test)

# no outliers are detected (output nrow = 0)
outliers <- oee_data |>
  group_by(treatment) |>
  mutate(z_oee = (oee_100 - mean(oee_100, na.rm = TRUE)) / sd(oee_100, na.rm = TRUE)) |>
  filter(abs(z_oee) > 2.5) |>
  nrow()

lt_oee_treat <- leveneTest(oee_100 ~ treatment, data = oee_data)
pv <- lt_oee_treat$`Pr(>F)`[1]
```

Results from the Shapiro-Wilk test provide further evidence against the normality hypothesis for OEE, both overall (p `r fmt_p_val(shapiro.test(oee_data$oee_100)$p.value)`) and by treatment (p < 0.05 in all cases). Q-Q plots in @fig-qq-h2a suggest that OEE is close to normally distributed for all augmented treatments, but PWI is clearly not. No outliers were detected in the OEE data based on the Z-score test. Finally, Levene's test indicated no significant heterogeneity of variance between treatment groups (`r fmt_p_val(pv)`). These findings again dictate a non-parametric approach to testing $H_{2a}$.

```{r fig-qq-h2a}
#| fig-cap: Q-Q Plots of OEE by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(oee_data, x_var = "oee_100", by_var = "treatment")

p + plot_annotation(
  title = "Quantile-Quantile Plots of OEE Treatment Data",
  subtitle = "PWI effectiveness is not normally distributed"
)
```

#### Analysis

```{r test-oee-power}
k <- length(unique(oee_data$treatment))
n <- round(nrow(oee_data) / k)
epsilon_sq <- 0.19

# Estimate post-hoc power
oee_power <- posthoc_power_analysis(
  n, k, epsilon_sq,
  sig_level = 0.05, num_simulations = 1000
)
```

In @fig-h2a-plot, group-wise significant differences are found (p = 0.02), with with small to moderate effect size $\hat{\epsilon}^{2}_{rank}=0.19$ and a broad $CI_{95\%}\:[0.08,1.0]$. Pair-wise tests show only PWI and PAR are significantly different (also p = 0.02). A post-hoc simulation was performed to assess the power of the analysis. The result of `r round(oee_power * 100, 1)`% indicates the analysis had sufficient power to detect significant differences, which lends further support to the reliability of this outcome.

```{r fig-h2a-plot}
#| fig-cap: $H_{2a}$ Test - OEE Differences by Treatment
#| fig-height: 5
#| warning: false 
#| message: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

h2a_plot <- ggstatsplot::ggbetweenstats(
  data = oee_data,
  x = treatment,
  y = oee_100,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "OEE Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between PWI and PAR Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

h2a_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "9.79", italic(p) ==
#     "0.02", widehat(epsilon)["ordinal"]^2 == "0.19", CI["95%"] ~
#     "[" * "0.08", "1.00" * "]", italic("n")["obs"] == "53")
```

```{r oee_summary}
ost <- oee_data |>
  group_by(treatment) |>
  summarize(
    mean = mean(oee_100, na.rm = T),
    sd = sd(oee_100, na.rm = T),
    med = median(oee_100, na.rm = T),
    iqr = IQR(oee_100, na.rm = T),
    min = min(oee_100, na.rm = T),
    max = max(oee_100, na.rm = T),
  ) |>
  mutate(
    cv = sd / mean,
    rng = max - min,
    iqr_ratio = iqr / rng
  )

# Extract the iqr_ratio for PAR
pr <- function(trt) {
  iqr_r <- ost |>
    filter(treatment == trt) |>
    pull(iqr_ratio)
  return(round(iqr_r, 2))
}
```

::: content-hidden
TODO: add table of oee data summarized by treatment
:::

This unexpected result indicates that, despite *substantial observed differences*, the PWI, AR, and MR treatments have *statistically equivalent* OEE. High within-group variability likely contributes to this finding by obscuring genuine differences in group means. Kruskal-Wallis test is somewhat vulnerable to this effect due to its rank-based design and underlying assumptions. The ratio of interquartile range to overall range for PAR (`r pr("PAR")`), AR (`r pr("AR")`), and MR (`r pr("MR")`) show that the central 50% of observations in these treatments is spread over more than half of its overall range. This wide variation is easily observed in the box plots of @fig-h2a-plot.

To further investigate the observed variability and validate these unexpected findings, bootstrap resampling analysis was employed. This flexible technique samples with replacement to construct robust estimates of the distribution for the statistics of interest (e.g. mean or median) and their confidence intervals. Bootstrapping relies on the empirical distribution of the data to accurately capture variability with few assumptions. Additionally, the replication process effectively increases the sample size and statistical power of the analysis, enhancing the accuracy and reliability of its results.

A bootstrap analysis was used to estimate pairwise differences in three measures of centrality for OEE--mean, median, and the Hodges-Lehmann estimator. The later is a non-parametric approach to estimate population median shifts that is robust to outliers. It provides a consistent estimate of the true shift in medians, even when the underlying distributions are dissimilar, by considering the differences between all possible observation pairs. This process was replicated five times, each with 10,000 resamples, and the results were averaged to determine grand means and related statistics. The significant results of this complementary analysis, where the 95% confidence interval does not include zero, are tabulated in @tbl-bootstrap-oee.

```{r bootstrap-oee, eval=FALSE}
# registerDoParallel(cores = parallel::detectCores())

stat_funcs <- list(
  Mean = mean,
  Median = median,
  HL_Est = hl_estimator
)

h2a_oee_boot_res <- run_bootstrap_tests(
  oee_data,
  column = "oee_100", 
  stat_funcs = stat_funcs
)

h2a_oee_boot_res
save(h2a_oee_boot_res, file = "h2a_oee_boot_res.RData")
```

```{r tbl-boot-oee}
#| tbl-cap: $H_{2a}$ Bootstrap Analysis of Pairwise OEE Differences

load("h2a_oee_boot_res.RData")

title <- "Significant OEE Differences in Centrality: Bootstrapped Estimates"
subtitle <- "Five Replications of 10,000 Resamples for Mean, Median, and the Hodges-Lehmann Estimator"


create_boot_summary_table(h2a_oee_boot_res, title = title, subtitle = subtitle)
```

Compared to the initial findings, bootstrapping identifies a number of significant pairwise differences. Even when testing OEE median values, as done with Kruskal-Wallis, this method finds both PWI-PAR and PWI-AR significant. The mean tests, which are more sensitive to variance and outliers, also include PWI-MR. Hodges-Lehmann matches the mean findings, despite its reduced sensitivity. 

This result supports the previous findings of significant overall differences in OEE, as well as between the PWI and PAR conditions for all measures of centrality. In addition, the bootstrap analysis, with its robust estimation methods, provided strong evidence of significant differences between PWI and both the AR (for all measures) and MR groups (except for median). These results show that the OEE of PWI is substantially lower than all other treatments, with average estimated differences of 0.687 (PAR), 0.585 (AR), and 0.307 (MR). The negligible bias values and low standard error reported across all replicates reinforce the reliability of these findings.

::: content-hidden
MAYBE: look into factors underlying OEE

#### Additional Analysis

does PWI time play a significant role in variation? with / without?
look at trends in recall - still learning?

look at differences in OEE components by treatment

```{r fig-oee-components, eval=FALSE}
# Plot of Productivity by Treatment
plot_prod <- ggplot(oee_data, aes(x = treatment, y = prod)) +
  geom_boxplot() +
  labs(title = "Productivity by Treatment", y = "Productivity", x = "Treatment")

# Plot of Quality by Treatment
plot_qual <- ggplot(oee_data, aes(x = treatment, y = qual)) +
  geom_boxplot() +
  labs(title = "Quality by Treatment", y = "Quality", x = "Treatment")

# Combine the plots
plot_combined <- plot_prod + plot_qual
print(plot_combined)
```

as seen with H1c, PWI quality is very low due to the number of errors
on average, each participant in the PWI treatment makes XXX UCEs, compared to a, b, c for the other treatments

are the differences in productivity, quality significantly different?

```{r fig-oee-prod, eval=FALSE}
#| fig-cap: Productivity Differences by Treatment
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

prod_plot <- ggstatsplot::ggbetweenstats(
  data = oee_data,
  x = treatment,
  y = prod,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "Productivity Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

prod_plot
```

```{r fig-oee-qual, eval=FALSE}
#| fig-cap: Quality Differences by Treatment
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

qual_plot <- ggstatsplot::ggbetweenstats(
  data = oee_data,
  x = treatment,
  y = qual,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "Quality Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

qual_plot
```

:::

#### $H_{2a}$ Result

As both approaches identified a significant treatment effect on OEE, there is ample evidence to accept $H_{2a}$: OEE does vary with instructional method. However, the conflicting pairwise analysis results must be reconciled. While the Kruskal-Wallis test only identified significant differences between the PWI and PAR treatments, subsequent bootstrapped analysis revealed additional differences between PWI and the AR and MR treatments using a variety of measures. Considering the bootstrap method's advantages when faced with high between-group variance, limited sample size, and non-normally distributed data, the consistent differences and robust confidence intervals it produced provide compelling evidence that the observed differences are reliable. Ultimately, it is determined that PAR, AR, and MR treatments all result in statistically significant and practically meaningful improvements in OEE compared to PWI, which was hampered by a high defect rate.

### $H_{2b}$: Paper Work Instruction References

```{r h2b-data}
recall_h2b <- recall |>
  mutate(
    pwi_time_per_ref_all = ifelse(
      pwi_event_count > 0,
      pwi_event_tot_time / pwi_event_count,
      0
    )
  )

recall_h2b_agg <- recall_agg
```

The second hypothesis of the recall phase assesses an alternative measure of training effectiveness:
$$H_{2b}\textrm{: PWI reliance varies with treatment}$$

Reliance is quantified by the number of times each participant refers to the paper work instructions (PWI count), and the duration of each event (PWI time). PWI count is considered a measure of the *frequency* of the participant's task uncertainty, while PWI time is associated with the *degree* of uncertainty experienced.

#### Descriptive Statistics

```{r make-h2b-report}
#| warning: false

# report descriptive stats for pwi count and pwi time (for each car, n = 212)
h2b_rpt <- recall_h2b |>
  select(pwi_event_count, pwi_event_tot_time, pwi_time_per_ref_all) |>
  rename(
    "PWI Count per Task" = pwi_event_count,
    "Total PWI Time per Task" = pwi_event_tot_time,
    "PWI Avg Time per Reference" = pwi_time_per_ref_all
  ) |>
  report::report() # options not working!

# Extract and split the string by the pattern "\n - ", remove first line
h2b_rpt_list <- strsplit(h2b_rpt[1], "\n  - ")[[1]][-1]

# Add hyphen to first element and rejoin
h2b_rpt_list[1] <- paste0("- ", h2b_rpt_list[1])
h2b_rpt_list <- paste(h2b_rpt_list, collapse = "\n- ")
h2b_rpt_list <- gsub(", 0 missing", "", h2b_rpt_list)
h2b_rpt_list <- gsub("n = 212, ", "", h2b_rpt_list)

# convert back into list of strings
h2b_rpt_list <- strsplit(h2b_rpt_list, "\n")[[1]]

# Calculate and test Spearman's correlation
h2b_cor_test_spearman <- cor.test(recall_h2b$pwi_event_count, recall_h2b$pwi_event_tot_time, method = "spearman")

# print(h2b_rpt_list[1:2])
```

::: content-hidden
MAYBE: replace with table, e.g.:

```r
datasummary(
  Heading("Participant Avg Reference Count per Task") * avg_pwi_count +
    Heading("Participant Avg Reference Duration per Task") * avg_pwi_time +
    Heading("Overall Avg Duration per PWI Reference") * pwi_time_per_ref_agg ~
    Mean + SD + Min + Median + Max,
  # title = "this is the title",
  # notes = c("this is a note"),
  data = recall_h2b_agg,
  align = "lccccc",
  output = "kableExtra"
)
```
:::

Overall statistics for the measures of interest (n = `r nrow(recall_h2b)`) are summarized below:

`r paste(h2b_rpt_list[1:2], collapse = "\n")`

Despite low mean values for count and total time (1.47 and 2.86, respectively), both exhibit substantial variability, with standard deviations of 4.12 and 9.43. High skewness (4.51 for count, 5.64 for time) and kurtosis (23.27, 36.73) confirm their strong right skew and heavy tails. Together with a median and MAD of zero for both measures, these statistics suggest that many participants rarely referred to the instructions, but a small number relied on them heavily.

The statistical similarity between count and time suggests a high correlation, confirmed by Spearman's rank correlation test (rho = `r round(h2b_cor_test_spearman$estimate,2)`, `r fmt_p_val(h2b_cor_test_spearman$p.value)`). This is natural and expected for two dependent measures, but must be accounted for. To create a stable measure of overall reliance for each participant, a composite measure was created by averaging the time per PWI reference across all four tasks. This approach reduces the impact of task-specific factors and eliminates the correlation between reference count and duration without diminishing either aspect of task uncertainty. Using a single, composite measure simplifies the analysis and interpretation of the results by reducing dimensionality. It also enables a traditional univariate approach to the comparison, which benefits from reductions to both within-- and between--participant variance due to averaging. The distribution of average time per PWI reference is visualized in @fig-pwi-hist.

```{r fig-pwi-hist}
#| fig-cap: Histogram of Composite PWI Measure in Recall
#| fig-height: 3.5
#| warning: false
#| message: false

# # calculate bin widths based on theory... don't think any work well
# x <- recall_h2b_agg$pwi_time_per_ref_agg
# bin_width_fd <- 2 * IQR(x, na.rm = TRUE) / (length(x[!is.na(x)])^(1/3))
# bin_width_sturges <- (max(x) - min(x)) / (log2(length(x)) + 1)
# bin_width_scott <- 3.5 * sd(x) / (length(x)^(1/3))

recall_h2b_agg |>
  ggplot(aes(x = pwi_time_per_ref_agg)) +
  geom_histogram(
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color,
    binwidth = 0.5
  ) +
  labs(
    title = "Average Time per PWI Reference in Recall",
    subtitle = "Reliance is strongly right-skewed with high variability and long tails",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
  )
```

```{r h2b_tests}
st_rely_all <- shapiro.test(recall_h2b_agg$pwi_time_per_ref_agg)
st_rely_treat <- by(recall_h2b_agg$pwi_time_per_ref_agg, recall_h2b_agg$treatment, shapiro.test)

lt_rely_treat <- leveneTest(pwi_time_per_ref_agg ~ treatment, data = recall_h2b_agg)
lt_pv <- lt_rely_treat$`Pr(>F)`[1]
```

From the descriptive statistics and visualization, this data does not appear normal. To confirm, the Shapiro-Wilk test was administered for the overall data (`r fmt_p_val(st_rely_all$p.value)`) and within each treatment group. The results show that neither PWI (`r fmt_p_val(st_rely_treat$PWI$p.value)`) nor PAR (`r fmt_p_val(st_rely_treat$PAR$p.value)`) are normally distributed, but AR is (`r fmt_p_val(st_rely_treat$AR$p.value)`), and MR shows marginal evidence (`r fmt_p_val(st_rely_treat$MR$p.value)`) of non-normality. Visual inspection of the Q-Q plots for each treatment, provided in @fig-qq-h2b, supports these findings. Levene's test (`r fmt_p_val(lt_pv)`) confirms equal variance by treatment.

```{r fig-qq-h2b}
#| fig-cap: Q-Q Plots of PWI Reliance by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(recall_h2b_agg, x_var = "pwi_time_per_ref_agg", by_var = "treatment")

p + plot_annotation(
  title = "Quantile-Quantile Plots of PWI Treatment Data",
  subtitle = "subtitle..."
)
```

```{r drop-h2b-outliers}
outliers <- recall_h2b_agg |>
  group_by(treatment) |>
  mutate(
    z_rely = abs((pwi_time_per_ref_agg - mean(pwi_time_per_ref_agg, na.rm = T))
    / sd(pwi_time_per_ref_agg, na.rm = T)),
    z_count = abs((avg_pwi_count - mean(avg_pwi_count, na.rm = T))
    / sd(avg_pwi_count, na.rm = T)),
    z_time = abs((avg_pwi_time - mean(avg_pwi_time, na.rm = T))
    / sd(avg_pwi_time, na.rm = T))
  ) |>
  filter(z_rely > 2.5 | z_count > 2.5 | z_time > 2.5) |>
  # filter(z_rely > 3.0 | z_count > 3.0 | z_time > 3.0) |>
  ungroup()

recall_h2b_agg_out <- recall_h2b_agg |>
  anti_join(outliers, by = "participant")
```

Outliers were identified within each treatment group for all measures of task uncertainty. Data for `r as.character(english(nrow(outliers)))` participants with Z-scores of 2.5 or higher are summarized in @tbl-h2b-outliers. These participants were removed from the primary analysis to ensure the data represents typical participant behavior and to maintain the integrity of the comparative analysis.

```{r tbl-h2b-outliers}
#| tbl-cap: Outliers in Measures of Reliance
#| warning: false

c_range <- c(2.5, max(outliers |> select(starts_with("z_")), na.rm = T))
# Define the custom subset of the palette (e.g., first half of the red_material palette)
full_palette <- ggsci::pal_material("deep-orange")(10)
subset_palette <- full_palette[5:10]

h2b_out_tbl <- outliers |>
  select(participant, treatment, z_count, z_time, z_rely) |>
  gt() |>
  tab_header(
    title = "|Z-scores| for Reliance Metrics of Participant Outliers",
    subtitle = "Participants with |Z-score| >= 2.5 in reliance components"
  ) |>
  cols_label(
    participant = "Participant",
    treatment = "Treatment",
    z_count = "Count", # "Avg PWI Count / Task"
    z_time = "Time", # "Avg Total PWI Time / Task"
    z_rely = "Reliance" # "Avg Time / PWI Ref"
  ) |>
  fmt_number(
    columns = c(z_rely, z_count, z_time),
    decimals = 2
  ) |>
  cols_align(align = "center") |>
  tab_options(
    table.font.size = "smaller",
    # table.width = pct(90),
    column_labels.font.weight = "bold",
    heading.title.font.weight = "bolder",
    heading.subtitle.font.size = px(13)
  )
# data_color(
#   columns = starts_with("z_"),
#   method = "numeric",
#   domain = c_range,
#   palette = subset_palette,
#   # reverse = T,
#   na_color = "transparent"
# )

h2b_out_tbl

# Calculate mean PWI time per reference with and without outliers
overall_mean_with_outliers <- mean(recall_h2b_agg$pwi_time_per_ref_agg, na.rm = TRUE)
overall_mean_without_outliers <- mean(recall_h2b_agg_out$pwi_time_per_ref_agg, na.rm = TRUE)
```

::: content-hidden
TODO: standardize outlier test methodology, including z-scores
MAYBE: fix table coloring for PDF?
:::

```{r h2b-test-out}
# does removing outliers change the normality or levene test results?

st_rely_all_out <- shapiro.test(recall_h2b_agg_out$pwi_time_per_ref_agg)
st_rely_treat_out <- by(recall_h2b_agg_out$pwi_time_per_ref_agg, recall_h2b_agg_out$treatment, shapiro.test)

lt_rely_treat_out <- leveneTest(pwi_time_per_ref_agg ~ treatment, data = recall_h2b_agg_out)
lt_pv_out <- lt_rely_treat_out$`Pr(>F)`[1]

h2b_group_sizes <- summary(recall_h2b_agg_out$treatment)
h2b_gs_txt <- glue_collapse(glue("{h2b_group_sizes} {names(h2b_group_sizes)}"), ", ", last = ", and ")

```

```{r h2b-test-out-check, eval=FALSE}
# Store the results in a data frame
h2b_compare_out <- data.frame(
  Test = c(
    "Levene",
    "SW: All",
    "SW: PWI",
    "SW: PAR",
    "SW: AR",
    "SW: MR"
  ),
  p_all = c(
    fmt_p_val(lt_pv),
    fmt_p_val(st_rely_all$p.value),
    fmt_p_val(st_rely_treat$PWI$p.value),
    fmt_p_val(st_rely_treat$PAR$p.value),
    fmt_p_val(st_rely_treat$AR$p.value),
    fmt_p_val(st_rely_treat$MR$p.value)
  ),
  p_out = c(
    fmt_p_val(lt_pv_out),
    fmt_p_val(st_rely_all_out$p.value),
    fmt_p_val(st_rely_treat_out$PWI$p.value),
    fmt_p_val(st_rely_treat_out$PAR$p.value),
    fmt_p_val(st_rely_treat_out$AR$p.value),
    fmt_p_val(st_rely_treat_out$MR$p.value)
  )
)

h2b_compare_out

# as of 6/10 6:40pm...only change in significance at 0.05: Levene test
#      Test     p_all     p_out
# 1  Levene  p = 0.95  p = 0.01 becomes significant, variance differs w/out outliers
# 2 SW: All p < 0.001 p < 0.001 n/c, not normal
# 3 SW: PWI p < 0.001 p < 0.001 n/c, not normal
# 4 SW: PAR p < 0.001 p = 0.004 n/c, not normal in either case
# 5  SW: AR  p = 0.16  p = 0.12 n/c, normal
# 6  SW: MR  p = 0.05  p = 0.05 n/c, borderline significance, no MR obs dropped
```

```{r h2b-comp-out-overall, eval=FALSE}
# does removing the outliers significantly change the mean of the overall data?

# Wilcoxon signed-rank test
overall_wilcoxon <- wilcox.test(
  recall_h2b_agg$pwi_time_per_ref_agg,
  recall_h2b_agg_out$pwi_time_per_ref_agg,
  exact = FALSE
)

print(overall_wilcoxon)

# p-value = 0.5391 n/c, no significant difference in overall data
```

```{r h2b-comp-out-treat, eval=FALSE}
# does removing the outliers significantly change the mean of any treatment group?

# Function to compare means within each treatment group using Wilcoxon rank-sum test (unpaired)
compare_by_treatment <- function(treatment) {
  with_outliers <- recall_h2b_agg |> filter(treatment == !!treatment)
  without_outliers <- recall_h2b_agg_out |> filter(treatment == !!treatment)

  mean_with_outliers <- mean(with_outliers$pwi_time_per_ref_agg, na.rm = TRUE)
  mean_without_outliers <- mean(without_outliers$pwi_time_per_ref_agg, na.rm = TRUE)

  test_result <- wilcox.test(
    with_outliers$pwi_time_per_ref_agg, without_outliers$pwi_time_per_ref_agg,
    exact = FALSE
  )

  list(
    treatment = treatment,
    mean_with_outliers = mean_with_outliers,
    mean_without_outliers = mean_without_outliers,
    p_value = test_result$p.value,
    test_result = test_result
  )
}

# Apply the comparison function to each treatment group
comparison_results <- lapply(TREATMENTS, compare_by_treatment)

# Print results
for (result in comparison_results) {
  cat("Treatment:", result$treatment, "\n")
  cat("  Mean with outliers:", result$mean_with_outliers, "\n")
  cat("  Mean without outliers:", result$mean_without_outliers, "\n")
  cat("  p-value:", result$p_value, "\n")
  print(result$test_result)
  cat("\n")
}

# all p-values > 0.493 (PAR) n/c, not significantly difference within any treatment
```

Several checks were then repeated to assess the impact of outlier removal. The Shapiro-Wilk test showed no change in significance for data overall or by treatment group. While Levene's test found significant difference in variance between groups (`r fmt_p_val(lt_pv_out)`) after the change, Kruskall-Wallis does not assume otherwise. A Wilcoxon signed-rank test was applied to compare overall and group means with and without outliers. No significant differences were found in average PWI time per reference in the the full data set or by treatment group (p > 0.40 for all cases). It is determined that, despite a reduction in the overall mean value from `r round(overall_mean_with_outliers, 2)` to `r round(overall_mean_without_outliers, 2)`, the character of the underlying data is not significantly changed by the outlier removal. Given that most treatments are not normally distributed, and variance differs between groups, a non-parametric approach is chosen for testing $H_{2b}$. The total number of observations (n = `r nrow(recall_h2b_agg_out)`) and group sizes (`r h2b_gs_txt`) remain sufficient for these methods, though statistical power is reduced.

#### Analysis

```{r test-pwi-power}
k <- length(unique(recall_h2b_agg_out$treatment))
n <- round(nrow(recall_h2b_agg_out) / k)
epsilon_sq <- 0.17

# Estimate post-hoc power
pwi_power <- posthoc_power_analysis(
  n, k, epsilon_sq,
  sig_level = 0.05, num_simulations = 1000
)
```

Per @fig-h2b-out-kw, despite a nominal p-value (0.05) the Kruskal-Wallis test fails to identify significant between-group differences for the average PWI reference duration. A small effect size $\hat{\epsilon}^{2}_{rank}=0.17$ with broad $CI_{95\%}\:[0.07,1.0]$, along with a post-hoc power simulation (`r round(pwi_power * 100, 1)`%) all support this result.

```{r fig-h2b-out-kw}
#| fig-cap: $H_{2b}$ Test - Average Time per PWI Reference by Treatment
#| fig-height: 4
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

pwi_rely_plot <- ggstatsplot::ggbetweenstats(
  data = recall_h2b_agg_out,
  x = treatment,
  y = pwi_time_per_ref_agg,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "PWI Reliance by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "No Significant Differences Identified between Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

pwi_rely_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "7.8097", italic(p) ==
#     "0.0501", widehat(epsilon)["ordinal"]^2 == "0.1662", CI["95%"] ~
#     "[" * "0.0730", "1.0000" * "]", italic("n")["obs"] == "48")
```

Given the marginal p-value and high variance data, bootstrap analysis was again performed to provide complementary insights. As before, five replications of 10,000 resamples were conducted for all pairs using mean, median, and the Hodges-Lehmann estimator. Significant results are summarized in @tbl-boot-pwi. As with the Kurskal-Wallis test, no significant differences were identified using median, but the mean and Hodges-Lehmann estimator both identified 95% CI based differences in PWI-AR, PWI-MR, and PAR-AR pairs. A significant difference in mean was also indicated for PAR-MR.

```{r bootstrap-pwi, eval=FALSE}
stat_funcs <- list(
  Mean = mean,
  Median = median,
  HL_Est = hl_estimator
)

h2b_pwi_boot_res <- run_bootstrap_tests(
  recall_h2b_agg_out,
  column = "pwi_time_per_ref_agg", 
  stat_funcs = stat_funcs
)

h2b_pwi_boot_res
save(h2b_pwi_boot_res, file = "h2b_pwi_boot_res.RData")
```

```{r tbl-boot-pwi}
#| tbl-cap: $H_{2b}$ Bootstrap Analysis of Pairwise PWI Differences

load("h2b_pwi_boot_res.RData")
title <- "Significant Reliance Differences in Centrality: Bootstrapped Estimates"
subtitle <- "Five Replications of 10,000 Resamples for Mean, Median, and the Hodges-Lehmann Estimator"
create_boot_summary_table(h2b_pwi_boot_res, title = title, subtitle = subtitle)
```

On average, AR and MR reduce reliance by 0.686 and 0.612 seconds per reference, respectively, relative to the PWI instructional method. This finding is statistically and practically meaningful, given that the overall average reference duration is `r round(mean(recall_h2b_agg_out$pwi_time_per_ref_agg), 2)` seconds. In summary, the bootstrap analysis finds AR and MR are equivalent to one another but reduce reliance more than than PAR and PWI, which are themselves equivalent. The negligible bias values and low standard error reported across all replicates once again reinforce the reliability of these findings.

#### $H_{2b}$ Result

The marginal p-value returned by the Kruskall-Wallis comparison of median reliance, together with consistent and robust estimates of the differences provided by bootstrap analysis give sufficient evidence to accept $H_{2b}$: reliance does vary with instructional method. However, the conflicting results of pairwise comparisons must be interpreted with caution. Like other results in this study, they suggest that AR/MR are superior to PWI/PAR but the limited data and high variation make this somewhat inconclusive. Ultimately, it is determined that statistically significant differences in reliance exist between treatments, but specific pair-wise differences are not clear enough to declare.

::: content-hidden
#### Additional Analysis

?underlying factors - time and count?

additional tables / visuals to consider:

- Box plots or violin plots comparing the distribution of productivity and quality scores across treatments. These plots would complement the OEE comparison and provide a more detailed understanding of how the instructional methods impact these individual components of OEE.
- A stacked bar chart showing the proportion of participants in each treatment group who experienced different levels of PWI reliance (e.g., no reliance, low reliance, high reliance). This chart would help visualize the distribution of reliance levels within and between treatments.
- A table summarizing the key descriptive statistics (mean, median, standard deviation, etc.) for OEE, productivity, quality, and PWI reliance by treatment group. This table would provide a quick reference for comparing the central tendencies and variability of the measures across treatments.
- A line chart displaying the average PWI reference duration per task iteration for each treatment group. This chart would help visualize any trends in reliance over time and potentially reveal differences in learning rates between treatments.
- A forest plot displaying the effect sizes and confidence intervals for the pairwise comparisons of OEE and PWI reliance between treatments. This plot would provide a concise visual summary of the magnitude and uncertainty of the differences between instructional methods.

#### Observations

- PWI confident, regardless of competence
- PAR seems to be the best mix of confidence and competence - would it be better than simple error checking / prevention?
- need for bootstrapping suggests more data is needed to get less nuanced results

:::
