# Results {#25-results}

```{r load-libs}
#| message: false

library(tidyverse)
library(readxl)
```

```{r load-data}

car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results"
)

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics"
)
```

```{r def-cust-theme}
base_color <- "#69b3a2"

# Define a custom theme for ggplot
custom_theme <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      # plot.subtitle = element_text(face = "italic"),
      # plot.caption = element_text(color = "grey50"),
      # axis.title.x = element_text(face = "bold"),
      # axis.title.y = element_text(face = "bold"),
      # axis.text = element_text(color = "black"),
      # axis.line = element_line(color = "grey80"),
      # legend.position = "bottom",
      # legend.background = element_rect(fill = "white", colour = "grey50"),
      # legend.key = element_rect(fill = "white", colour = "grey50"),
    )
}

# Set the custom theme as the default theme
theme_set(custom_theme())
```

```{r study-metadata}

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# Convert the trial_datetime to Date type
demographics <- demographics |>
  mutate(
    trial_date = as.Date(
      as.POSIXct(trial_datetime, format = "%Y-%m-%d %H:%M:%S")
    )
  )

# Get the first and last date
first_trial_date <- demographics |>
  summarize(first_date = min(trial_date)) |>
  pull(first_date)

last_trial_date <- demographics |>
  summarize(last_date = max(trial_date)) |>
  pull(last_date)

trial_start_str <- str_glue(
  "{month(first_trial_date)}/{day(first_trial_date)}"
)
trial_end_str <- str_glue(
  "{month(last_trial_date)}/{day(last_trial_date)}"
)
trial_year_str <- str_glue("{year(last_trial_date)}")

# start, end = Feb 10, Apr 27 2023
```


```{r fig-s1-progress}
#| fig-cap: "Daily Trial Administration and Completion Percentage"
#| fig-height: 4

# Count experiments per day
daily_counts <- demographics |>
  group_by(trial_date) |>
  summarize(daily_count = n()) |>
  arrange(trial_date)

# Calculate the cumulative count of experiments
daily_counts <- daily_counts |>
  mutate(cumulative_count = cumsum(daily_count))

# Calculate remaining experiments
daily_counts <- daily_counts |>
  mutate(
    percentage_complete = (cumulative_count / participant_count) * 100
  )

pc_eng <- as.character(english::english(participant_count))
pc_eng <- str_replace(pc_eng, "^.", str_to_upper)
subtitle <- str_glue("{pc_eng} Learning and Recall Experiments Conducted between {trial_start_str} and {trial_end_str} of {trial_year_str}")

# Plotting
ggplot(daily_counts, aes(x = trial_date)) +
  geom_line(
    aes(y = daily_count, group = 1),
    color = base_color,
    linewidth = 1,
    alpha = 0.5
  ) +
  geom_line(
    aes(y = percentage_complete, group = 1),
    color = base_color,
    linewidth = 1.25,
    alpha = 1
  ) +
  scale_y_continuous(
    name = "Trials Administered",
    sec.axis = sec_axis(~., name = "Percentage Complete"),
    position = "right",
    limits = c(0, 100)
  ) +
  labs(
    title = "First Session Progress",
    subtitle = subtitle,
    x = element_blank(),
    y = element_blank()
  ) +
  custom_theme()
```

Only eight trials were run before March 16th. That period was treated as a pilot study, during which the research team was trained, administration methods were refined, the self-service onboarding system was deployed, and version 2.1 of our IRB application was approved. Most significantly, the intake process was adjusted to collect more demographic data, increase the role of the TLX and SUS instruments, and gather BCS data for a companion study.

Otherwise, only minor tweaks were made during the pilot to streamline the experiment's administration. Although the changes outlined above were not believed to have a material effect on the results, the data collected during the pilot were omitted from further analysis. Following the pilot, steady progress was made until the completion of the study's first session on April 27th.

The second session was conducted during the end of study event on *XXX*. *YYY* of the study's original participants volunteered to attend and complete the Retention experiment.

::: content-hidden
TODO: complete above paragraph about second session.
:::

::: content-hidden
MAYBE: perform analysis to evaluate differences between pilot and main group

Pilot Study Comparison
- Compare the key measures (e.g., performance, workload, usability) between the pilot study (n=8) and main study (n=54) participants
- Use descriptive statistics and visualizations to present these comparisons
- Discuss any notable similarities or differences between the two groups

Statistical Power and Sample Size Considerations
- Present the results of statistical power analyses based on the observed effect sizes and the planned statistical tests
- Discuss the adequacy of the sample size for detecting meaningful effects
- Consider the potential impact of including or excluding the pilot study data on the statistical power and precision of the estimates
- sample size calculation / justification: power analysis, practical considerations, precedent

Decision on Pilot Study Inclusion
- Based on the findings from 5.2.3 and 5.2.4, make a decision on whether to include the pilot study data in the main analysis
- Clearly report your decision and rationale, considering factors such as data comparability, study design modifications, and statistical power
:::

## Analysis of Participant Demographics

### Univariate Analysis

```{r drop-pilot-data}
# drop participants in pilot study (first eight)
demographics <- demographics |>
  filter(!(participant >= 1001 & participant <= 1010))

# drop event data from the pilot study
car_results <- car_results |>
  filter(!(participant >= 1001 & participant <= 1010))
```

```{r def-to-factors}

# safely convert data to ord / cat factors
# create levels / labels as required and identify / warn about mismatches

to_factors <- function(
    column,
    levels = NULL,
    labels = NULL,
    ordered = FALSE) {
  if (is.null(levels)) {
    levels <- unique(na.omit(column))
  }

  if (is.null(labels)) {
    labels <- levels
  }

  mismatched <- setdiff(unique(na.omit(column)), levels)

  if (length(mismatched) > 0) {
    # Append mismatched values to levels and labels with a specific prefix
    bad_levels <- paste("Bad -", mismatched)
    levels <- c(levels, mismatched)
    labels <- c(labels, bad_levels)
    warning(
      "Mismatched or unexpected values incorporated: ",
      paste(bad_levels, collapse = ", ")
    )
  }

  # Proceed to convert to factor using cleaned-up data
  factor_column <- factor(column,
    levels = levels, labels = labels,
    ordered = ordered
  )

  return(factor_column)
}
```

