---
title: "R Notebook"
output: html_notebook
---

Naming:
demog_xls = participant demographics (clean)
outcomes = one observation per car, e.g. completed, retired, etc. (clean)
times = one observation per timed event, e.g. car, breakage, etc. (clean)
results = joined data, one observation per outcome; data from relevant rows of times are combined into a list

Load the XLSX file

```{r}
### load libraries
library(tidyverse)
library(readxl)
```

```{r}
### read and clean the demographic data

raw_demog_xls <- read_xlsx(
  "./data/source/i1_raw_data.xlsx",
  na = c("", "N/A"),
  sheet = "DEMO")

demog_xls <- raw_demog_xls |>
  janitor::clean_names() |>
  # only use time information from trial_time column 
  mutate(trial_time = hms::as_hms(trial_time)) |>
  # combine date and time columns as single datetime and drop original cols
  mutate(trial_datetime = trial_date + trial_time) |>
  select(-c("trial_date", "trial_time")) |>
  # move numeric cols to the front
  select(all_of(c("participant", "age", "height")), everything()) |>
  # convert most cols to factors
  mutate(across(-("participant":"height" | "notes"), as.factor))
```

```{r}
### read and clean the car error data

raw_exp_xls <- read_xlsx(
  "./data/source/i1_raw_data.xlsx",
  sheet = "OUTCOMES")

outcomes <- raw_exp_xls |>
  janitor::clean_names() |>
    mutate(
    result = case_when(
      tolower(flag) == "c" ~ "comment",
      tolower(flag) == "i" ~ "incomplete",
      is.na(flag) ~ "complete",
      tolower(flag) == "r" ~ "retired"
    ),
    result = as.factor(result),
    flag = as.factor(flag)
  ) |>
  rename(e_car = car) |>
  replace_na(list(ce = 0, uce = 0, sel = 0, pos = 0, rot = 0, count = 0)) |>
  group_by(participant, phase, result) |>
  mutate(
    number = ifelse(result == "comment", 0, cumsum(result == result))
  ) |>
  ungroup() |>
  group_by(participant, phase) |>
  mutate(
    sequence = row_number() - 1  # have comments at sequence 0
  ) |>
  ungroup()

outcomes <- outcomes |>
  select(
    participant,
    phase,
    sequence,
    flag,
    result,
    number,
    e_car,
    ce,
    uce,
    sel,
    pos,
    rot,
    count,
    notes
  )

```

```{r}
### read and clean the car time data

col_names = c(
  "participant",
  "phase",
  "event",
  "start_time",
  "duration",
  "note"
)

raw_times_csv <- read_csv("./data/csv/i1_times_v2.csv", col_names=col_names)

times <- raw_times_csv |>
  janitor::clean_names() |>
  mutate(
    description = event,
    event = str_extract(event, "[a-zA-Z]+"), # Extract the event type
    event = as.factor(tolower(event))
  ) |>
  group_by(participant, phase, event) |>
  mutate(
    number = row_number() # sequence number for each event type
  ) |>
  ungroup() |>
  group_by(participant, phase) |>
  mutate(
    sequence = row_number()
  ) |>
  ungroup()

times <- times |>
  select(
    participant,
    phase,
    sequence,
    description,
    event,
    number,
    start_time,
    duration,
    note
  )

```

Flag parent and child events in `times`.

```{r}
times <- times |>
  arrange(participant, phase, sequence) |>
  group_by(participant, phase) |>
  mutate(
    end_time = start_time + duration,
    row_within_group = row_number(),
    type = case_when(
      event == "car" ~ "parent",  # all car events are parents
      event == "drop" ~ "child",  # all drop events are children
      phase == 2 &
        event == "pwi" ~ "child",  # all pwi event is a child
      row_within_group == 1 ~ "parent",  # 1st event in group is a parent
      TRUE ~ map_chr(row_within_group,  # otherwise assign based on times
                     ~ ifelse(
                       !any(start_time[1:(.x - 1)] <= start_time[.x] &
                              end_time[1:(.x - 1)] >= start_time[.x]),
                       "parent",
                       "child"
                     ))
    )
  ) |>
  ungroup()

```

Modify data after careful review. Done here instead of in the source data so the changes and rationale can be spelled out.

```{r}

ignore_list <- list(
  c(1018, 1, 2),  # trying to acquire tracking between cars
  c(1021, 1, 5),  # dismissing HL UI b/w C4 and 5
  c(1021, 1, 6),  # delayed restart of timer to account for lost clock time
  c(1026, 1, 3),  # UI dropout b/w C1 and 2
  c(1026, 1, 4),  # system crash between cars
  c(1035, 1, 6),  # HL menu popped on, ignored by subject
  c(1035, 1, 7),  # stopped timer to dismiss HL menu
  c(1060, 1, 5),  # stopped timer to reset HL system after breakage
  c(1061, 1, 4)   # stopped timer to correct axle stuck in fixture
)

# Utility function
should_ignore <- function(participant, phase, sequence) {
  any(map_lgl(ignore_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Update the data
times <- times |>
  rowwise() |>
  mutate(type = if_else(should_ignore(participant, phase, sequence), "ignore", type)) |>
  ungroup()

```

1. Change the start time of sequence 6 so that it equals the start time of sequence 7 for participant 1018, phase 1.
2. Make sequence 6 a type "child".
3. Reverse the order of the sequences.

```{r}
seq_7_start_time <- times |>
  filter(participant == 1018, phase == 1, sequence == 7) |>
  pull(start_time)

# Now, make the changes as described above to correct (1018, 1, 6)
times <- times |>
  mutate(
    # Change the start time for sequence 6
    start_time = if_else(
      participant == 1018 &
        phase == 1 & sequence == 6,
      seq_7_start_time,
      start_time
    ),
    # Correct duration of that event
    duration = if_else(participant == 1018 &
                         phase == 1 & sequence == 6, 13.07, duration),
    # Change type for sequence 6
    type = if_else(participant == 1018 &
                     phase == 1 & sequence == 6, "child", type)
  ) |>
  # Change the sequence order
  rowwise() |>
  mutate(sequence = if_else(
    participant == 1018 & phase == 1 & sequence == 6, 7L,
    if_else(participant == 1018 &
              phase == 1 & sequence == 7, 6L, sequence)
  )) |>
  ungroup()

```

Correct issues in the data identified during the first join process and subsequent review of data.

```{r}
# several cars were recorded in the times table as incomplete, but not progress
# was made, therefore they weren't recorded in the outcome table
# mark these for ignore
ignore_list <- list(
  c(1005, 1, 4),  # last car not started
  c(1008, 1, 8),  # last car not started
  c(1014, 2, 1),  # first car defect, not started
  c(1036, 1, 6),  # last car not started
  c(1047, 1, 5)   # duplicate subclip in video annotation
)

# Utility function, duplicated from above
should_ignore <- function(participant, phase, sequence) {
  any(map_lgl(ignore_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the type to `ignore` for listed cars
times <- times |>
  rowwise() |>
  mutate(type = if_else(should_ignore(participant, phase, sequence), "ignore", type)) |>
  ungroup()

# Patch times data for (1033, 1, 7) incomplete - was one block short
times <- times |>
  mutate(event = as.character(event)) |>  # Convert to character
  mutate(event = ifelse(participant == 1033 &
                         phase == 1 & sequence == 7, "incomplete", event)) |>
  mutate(event = as.factor(event))  # Convert back to factor if necessary


# Patch times data for 1044, 1, 4 - used start time from prior event
condition <- times$participant == 1044 & times$phase == 1 & times$sequence == 4

times <- times |>
  mutate(
    start_time = ifelse(condition, 584, start_time),
    duration = ifelse(condition, 18.033, duration),
    type = ifelse(condition, "parent", type)
  )

# Patch outcomes data for (1033, 1) to insert a car between 3 and 5
# Move 4 to 5
outcomes <- outcomes |>
  mutate(sequence = if_else(participant == 1033 & phase == 1 & sequence == 4, 5, sequence))

# Create a copy of row 3 as 4
new_record <- outcomes |>
  filter(participant == 1033, phase == 1, sequence == 3) |>
  mutate(sequence = 4)

# Append the new record to the outcomes table and resort the result
outcomes <- bind_rows(outcomes, new_record) |>
  arrange(participant, phase, sequence)

# Drop rows in outcomes corresponding to (1019, 1, 8 and 9) and (1021, 1, 7),
# due to: duplicate (c&p error?), no progress, no progress
outcomes <- outcomes |>
  filter(
    !(participant == 1019 & phase == 1 & sequence %in% c(8, 9)) & 
    !(participant == 1021 & phase == 1 & sequence == 7)
  )

```

Clean up tables.

```{r}
# combine the description (original event name) and note fields
times <- times |>
  mutate(note = if_else(note == "N/A", 
                      paste0("(", description, ")"), 
                      paste0("(", description, ") ", note)))

# assign a UID for each time event, e.g. 1001-P1-C1-T0
times <- times |>
  filter(type != "ignore") |>
  group_by(participant, phase) |>
  mutate(parent_seq = cumsum(type == "parent")) |>
  group_by(participant, phase, parent_seq) |>
  mutate(child_seq = if_else(type == "child", cumsum(type == "child"), 0)) |>
  ungroup() |>
  mutate(UIDT = paste0(
    participant,
    "-P", phase,
    "-C", parent_seq,
    "-T", sprintf("%02d", child_seq)
  ))

# drop some fields and reorder others
times <- times |>
  select(
    participant,
    phase,
    sequence,
    UIDT,
    event,
    type,
    start_time,
    duration,
    end_time,
    note
  ) |>
  glimpse()

# assign a UID for each outcome, e.g. 1001-P1-C1, and choose / order fields
outcomes <- outcomes |>
  group_by(participant, phase) |>
  mutate(UIDO = paste(participant, 
                     paste0("P", phase), 
                     paste0("C", sequence), 
                     sep = "-")) |>
  ungroup() |>
  select(
    participant,
    phase,
    sequence,
    UIDO,
    result,
    ce,
    uce,
    sel,
    pos,
    rot,
    count,
    notes
  )

```

Clean up notes in outcomes. Break out error codes into their own column. Add the treatment column from demog_xls.

```{r}
# Initial separation based on " :: " separator
outcomes <- outcomes |>
  separate(notes, into = c("potential_error_codes", "comments"), sep = " :: ", fill = "right")

outcomes <- outcomes |>
  rowwise() |>
  mutate(
    error_codes = ifelse(
      str_detect(
        potential_error_codes,
        "^[A-Z0-9]+( \\([0-9]+(\\+[0-9]+)?\\))?(, [A-Z0-9]+( \\([0-9]+(\\+[0-9]+)?\\))?|; [A-Z0-9]+( \\([0-9]+(\\+[0-9]+)?\\))?)*$"
      ),
      potential_error_codes,
      NA
    ),
    comments = ifelse(is.na(error_codes), potential_error_codes, comments)
  ) |>
  ungroup() |>
  select(-potential_error_codes)

# add treatment field from demog_xls and reorder columns
outcomes <- outcomes |>
  left_join(select(demog_xls, participant, treatment), by = "participant") |>
  select(
    participant,
    treatment,
    phase:count,
    error_codes,
    outcome_comment = comments
  )

write.csv(outcomes, "test/outcomes.csv", row.names = FALSE)
```

Second cleaning pass, after improved join and review of child events. Three classes of event:
1. Drop events
2. PWI events
3. All other types (e.g. reset, system, tracking, video, etc.)

First, I reviewed the third type, "other." In doing so, I identified three more general classes:
- Ignore: events that occured between cars and therefore had no impact on analysis
- System: delays introduced into the parent car build due to system-related issues, e.g. HL menu appears / dismissing, crashes and resets, model tracking acquisition
- Repair: delays introduced into the parent car build due to the participant's decision to make repairs rather than retire the car

In total there were 20 such events, all of which are corrected in the following code block.

```{r}
## re-classify misc event types as ignore, system, or repair

# IGNORE: patch times to ignore events that will have no impact on analysis
ignore_list <- list(
  c(1006, 1, 3),  # video recording drop, no performance impact
  c(1007, 1, 3),  # video recording drop, no performance impact
  c(1007, 1, 6),  # video recording drop, no performance impact
  c(1030, 1, 2),  # time lost to clock after car 1 marked incomplete due to crash
  c(1041, 1, 4)   # car 3 finished prior to breakage, so this occurred b/w cars
)

# Utility function, duplicated from above
should_ignore <- function(participant, phase, sequence) {
  any(map_lgl(ignore_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the event to `ignore` for listed cars
times <- times |>
  rowwise() |>
  mutate(event = if_else(should_ignore(participant, phase, sequence), "ignore", event)) |>
  ungroup()


# REPAIR: patch times to classify some events as time spent repairing a car
patch_list <- list(
  c(1011, 1, 2),
  c(1011, 1, 5),
  c(1014, 1, 2),
  c(1063, 1, 5)
)

# Utility function, duplicated from above
should_patch <- function(participant, phase, sequence) {
  any(map_lgl(patch_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the event to `reapair` for listed cars
times <- times |>
  rowwise() |>
  mutate(event = if_else(should_patch(participant, phase, sequence), "repair", event)) |>
  ungroup()

  
# SYSTEM: patch times to classify some events as time lost to system issues
patch_list <- list(
  c(1018, 1, 4), # trying to acquire tracking during HMDMR
  c(1018, 1, 7), # trying to acquire tracking during HMDMR
  c(1026, 1, 6), # interrupted by HL menu and time to deactivate it
  c(1039, 1, 2), # system drop and crash on the clock during car 1
  c(1039, 1, 3), # resetting system during car 1
  c(1055, 1, 8), # during car 3 clip
  c(1057, 1, 4), # dismiss HL menu during car 3 clip
  c(1057, 2, 2), # dismiss HL menu during car 1 clip
  c(1057, 2, 5), # dismiss HL menu during car 3 clip
  c(1063, 1, 13), # system drop and crash on the clock during car 2
  c(1063, 1, 14) # system reset during car 2 clip
)

# Utility function, duplicated from above
should_patch <- function(participant, phase, sequence) {
  any(map_lgl(patch_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the event to `repair` for listed cars
times <- times |>
  rowwise() |>
  mutate(event = if_else(should_patch(participant, phase, sequence), "system", event)) |>
  ungroup()

```

Went through all drop events to assess the percentage of its duration that impacted performance.  Created a table of the results, where percentatges range from 0.0 (no impact) to 1.0 (full).

Now add adj_drop time to times table.

```{r}
raw_adj_drops <- read_xlsx("./data/source/adjusted_drop_events.xlsx") |>
  janitor::clean_names()

times <- times |>
  left_join(raw_adj_drops, by = c("participant" = "part", "phase", "sequence" = "t_seq")) |>
  mutate(
    adj_drop = ifelse(
      event == "drop", 
      duration * coalesce(how_much, 1),  # If 'how_much' is missing, we assume a default value of 1
      NA_real_
    )
  ) |>
  rename(drop_adj_perc = how_much) |>
  select(-note, everything(), -during_car, -impact_perf, note) |>
  filter(event != "ignore") |>
  glimpse()

# add treatment field from demog_xls and reorder columns
times <- times |>
  left_join(select(demog_xls, participant, treatment), by = "participant") |>
  select(
    participant,
    treatment,
    phase:adj_drop,
    event_comment = note
  )

write.csv(times, "test/times.csv", row.names = FALSE)

```

Prepare for and execute join, test results.

```{r}
joined <- times |>
  mutate(UIDT_sub = str_replace(UIDT, "-T\\d{1,2}$", "")) |>
  full_join(outcomes, by = c("UIDT_sub" = "UIDO"))

write.csv(joined, "test/joined.csv", row.names = FALSE)
```

Generate two new columns for each type of child event (repair, system, pwi, drop):
- event_ct: This counts the number of child events of each type per parent.
- event_time: This sums the durations of child events of each type per parent.
- drop events result in normal event columns plus an additional col for adj_drop_time

```{r}
# Define the function to generate event summaries
# keep original and adjusted drop times
generate_event_summary <- function(data, event_type) {
  
  # If the event type is "drop", we'll handle it specially
  if (event_type == "drop") {
    
    # For count and original duration
    summary_colname <- "drop_event_ct"
    time_colname <- "drop_event_time"
    
    # For adjusted duration
    adj_time_colname <- "adj_drop_time"
    
    data_subset <- data |> filter(event == event_type)
    
    # Summarize for original and adjusted duration
    result <- data_subset |>
      group_by(UIDT_sub) |>
      summarise(
        !!summary_colname := n(),
        !!time_colname := sum(duration, na.rm = TRUE),
        !!adj_time_colname := sum(ifelse(is.na(adj_drop), duration, adj_drop), na.rm = TRUE)
      ) |>
      ungroup()
    return(result)
    
  } else { # original logic for other event types
    
    time_col <- "duration"
    summary_colname <- paste0(event_type, "_event_ct")
    time_colname <- paste0(event_type, "_event_time")
    
    data |>
      filter(event == event_type) |>
      group_by(UIDT_sub) |>
      summarise(
        !!summary_colname := n(),
        !!time_colname := sum(get(time_col), na.rm = TRUE)
      ) |>
      ungroup()
  }
}


# Use the function for each event type
event_types <- c("drop", "repair", "system", "pwi")
summaries <- lapply(event_types, generate_event_summary, data = joined)

# Merge the summaries back with the original dataset
joined_augmented <- joined
for (summary in summaries) {
  joined_augmented <- left_join(joined_augmented, summary, by = "UIDT_sub")
}

write.csv(joined_augmented, "test/joined_aug.csv", row.names = FALSE)

```

Automated tests.
Check that the aggregated data for each parent matches the sum of its child rows for all event types.

```{r}
# Create a function to perform tests for each event type
perform_tests <- function(event_type) {
  # Corresponding column names for aggregated counts and durations
  event_ct_col <- paste0(event_type, "_event_ct")
  event_time_col <- paste0(event_type, "_event_time")
  
  # Duration column in the original dataset
  if (event_type == "drop") {
    duration_col_original <- "duration"
    duration_col_adjusted <- "adj_drop"
  } else {
    duration_col_original <- "duration"
  }
  
  # Test 1: Compare total number of child events for original and aggregated data
  total_original <- sum(joined$event == event_type, na.rm = TRUE)
  total_aggregated <- joined_augmented |>
    filter(type == "parent") |>
    summarize(total = sum(!!sym(event_ct_col), na.rm = TRUE)) |>
    pull(total)
  
  stopifnot(total_original == total_aggregated)
  
  # Test 2: Compare total duration for original and aggregated data
  if (event_type == "drop") {
    # Compare original duration with drop_event_time
    total_duration_original <- joined |>
      filter(event == event_type) |>
      summarise(total_duration = sum(!!sym(duration_col_original), na.rm = TRUE)) |>
      pull(total_duration)
    
    total_duration_aggregated <- joined_augmented |>
      filter(type == "parent") |>
      summarise(total_duration = sum(!!sym(event_time_col), na.rm = TRUE)) |>
      pull(total_duration)
    
    stopifnot(total_duration_original == total_duration_aggregated)
    
    # Compare adjusted drop time with adj_drop_time
    total_duration_original_adj <- joined |>
      filter(event == event_type) |>
      summarise(total_duration = sum(!!sym(duration_col_adjusted), na.rm = TRUE)) |>
      pull(total_duration)
    
    total_duration_aggregated_adj <- joined_augmented |>
      filter(type == "parent") |>
      summarise(total_duration = sum(adj_drop_time, na.rm = TRUE)) |>
      pull(total_duration)
    
    stopifnot(total_duration_original_adj == total_duration_aggregated_adj)
    
  } else {
    # Original logic for other event types
    total_duration_original <- joined |>
      filter(event == event_type) |>
      summarise(total_duration = sum(!!sym(duration_col_original), na.rm = TRUE)) |>
      pull(total_duration)
    
    total_duration_aggregated <- joined_augmented |>
      filter(type == "parent") |>
      summarise(total_duration = sum(!!sym(event_time_col), na.rm = TRUE)) |>
      pull(total_duration)
    
    stopifnot(total_duration_original == total_duration_aggregated)
  }
}

# List of event types
event_types <- c("drop", "pwi", "repair", "system")

# Apply the function to each event type
invisible(lapply(event_types, perform_tests))
```


```{r}
# clean up joined_augmented table

combined <- joined_augmented |>
  filter(type == "parent") |>
  select(
    uid = UIDT_sub,
    participant = participant.y,
    treatment = treatment.x,
    phase = phase.y,
    sequence = sequence.y,
    car_outcome = result,
    car_start_time = start_time,
    car_end_time = end_time,
    car_elapsed_time = duration,
    ce:count,
    drop_event_count = drop_event_ct,
    drop_event_tot_time = drop_event_time,
    drop_event_tot_time_adj = adj_drop_time,
    repair_event_count = repair_event_ct,
    repair_event_tot_time = repair_event_time,
    system_event_count = system_event_ct,
    system_event_tot_time = system_event_time,
    pwi_event_count = pwi_event_ct,
    pwi_event_tot_time = pwi_event_time,
    error_codes
  )

write.csv(combined, "test/combined.csv", row.names = FALSE)
```

Get TLX data in a tidy format.

```{r}
# load all excel tabs into a combined tidy tibble
target_sheets <- paste0("TLX-", 0:2)

tlx_data_tidy <- map_dfr(
  target_sheets,
  ~ read_excel(
    "./data/source/i1_raw_data.xlsx",
    na = c("", "N/A"),
    sheet = .x
  ) |>
    select(-Notes) |>
    gather(key = "Question", value = "Response", -Participant) |>
    separate(Question, into = c("q_type", "q_num"), sep = "(?<=^[SR])(?=\\d+$)") |>
    mutate(
      q_num = as.numeric(q_num),
      q_type = factor(q_type, levels = c("S", "R"))
    ) |>
    arrange(Participant, q_type, q_num) |>
    # unite(Question, Letter, Number, sep = "") |>
    mutate(phase = which(target_sheets == .x) - 1) |>
    janitor::clean_names() |>
    rename(value = response) |>
    select(participant, phase, everything())
)

```

Rescale the R responses for the first 5 participants. Their scores were recorded using an older version of the scale, with a range of 1 to 7 instead of the 100 point scale used for others. This is corrected by uniformly scaling all values using the formula:

$$\text{transformed_value} = \frac{\text{original_value} - 1}{7 - 1} \times 100$$

```{r}
tlx_data_tidy <- tlx_data_tidy |>
  mutate(value = ifelse(
    q_type == "R" & participant %in% 1001:1007,
    ceiling((value - 1) / (7 - 1) * 100),  # all results rounded up
    value
  ))

```

Score the TLX:
1. Count responses to pair-wise source of workload comparisons to get a *weight* for each source, `source_weight`. (aka factor weights or raw ratings, 0-5)
2. Find the associated *rating* for each source, `source_rating`. (aka magnitude of load or factor ratings, 0-100)
3. Calculate the *workload* score for each source, `source_workload = source_weight * source_rating` (0-500).
4. Calculate the total task *workload* score, `task_workload = sum(source_workload) / 15` (0-100)

Because the total source weights is always 15, the average is 3. Thus the average source workload is 300, with a range of 0-500. The total of source workloads is always 1500, and the task workload is therefore in the range 0-100.

First, count the source weights.

```{r}
## Tally the Sources of Workload Data (weight) for each participant
# Define the mapping between S columns and sources
s_to_sources <- list(
  S1 = c("EF", "PF"),
  S2 = c("TD", "FF"),
  S3 = c("PD", "PF"),
  S4 = c("TD", "MD"),
  S5 = c("MD", "PD"),
  S6 = c("TD", "EF"),
  S7 = c("PD", "FF"),
  S8 = c("FF", "EF"),
  S9 = c("PF", "MD"),
  S10 = c("EF", "PD"),
  S11 = c("PF", "FF"),
  S12 = c("PD", "TD"),
  S13 = c("PF", "TD"),
  S14 = c("MD", "EF"),
  S15 = c("FF", "MD")
)

workload_factors <- c("MD", "PD", "TD", "PF", "EF", "FF")

tlx_source_weights <- tlx_data_tidy |>
  filter(q_type == "S") |>
  group_by(participant, phase) |>
  rowwise() |>
  mutate(choice = s_to_sources[[paste0(q_type, q_num)]][value]) |>
  ungroup() |>
  count(participant, phase, choice) |>
  pivot_wider(names_from = choice, values_from = n, values_fill = 0) |>
  arrange(participant, phase) |>
  select(participant, phase, all_of(workload_factors))

# confirm all rows add to 15
test <- tlx_source_weights |>
  rowwise() |>
  mutate(total_weight = sum(c_across(all_of(workload_factors)))) |>
  ungroup()

all(test$total_weight == 15)

# convert the result to long format like others
tlx_source_weights <- tlx_source_weights |>
  pivot_longer(
    cols = starts_with(workload_factors),
    names_to = "source",
    values_to = "value"
  ) |>
  mutate(type = "source weight") |>
  select(participant, phase, source, type, value)

```

Calculate `source_rating` and `source_workload` for each source.
Then calculate `task_workload` for each (participant, phase).
Combine all tlx tables into one.

```{r}
# create empty data grid
expanded_data <- expand_grid(
  participant = unique(tlx_data_tidy$participant),
  phase = 0:2,
  workload_index = 1:length(workload_factors)
) 

# calculate source ratings
tlx_source_ratings <- expanded_data |>
  left_join(tlx_data_tidy, by = c("participant", "phase", "workload_index" = "q_num")) |>
  filter(q_type == "R") |>
  mutate(
    source = workload_factors[workload_index],
    type = "source rating" # ,
    # value = response
  ) |>
  select(participant, phase, source, type, value)

# calculate source workloads
tlx_source_workloads <- expanded_data |>
  left_join(tlx_data_tidy, by = c("participant", "phase", "workload_index" = "q_num")) |>
  filter(q_type == "R") |>
  mutate(source = workload_factors[workload_index]) |>
  left_join(tlx_source_weights, by = c("participant", "phase", "source")) |>
  rowwise() |>
  mutate(
    value = value.x * value.y,
    type = "source workload"
  ) |>
  ungroup() |>
  select(participant, phase, source, type, value)

# calculate task workloads
tlx_task_workloads <- tlx_source_workloads |>
  group_by(participant, phase) |>
  summarise(value = sum(value) / 15, .groups = "drop") |>
  mutate(
    source = "combined",
    type = "task workload"
  )

tlx_scores <- bind_rows(
  tlx_source_weights,
  tlx_source_ratings,
  tlx_source_workloads,
  tlx_task_workloads
)

# add treatment field from demog_xls and reorder columns
type_order = c("source weight", "source rating", "source workload", "task workload")

tlx_scores <- tlx_scores |>
  left_join(select(demog_xls, participant, treatment), by = "participant") |>
  select(
    participant,
    treatment,
    everything()
  ) |>
  arrange(participant, treatment, phase, match(type, type_order), match(source, workload_factors))

write.csv(tlx_scores, "test/tlx_scores.csv", row.names = FALSE)
```

Process the System Usability Scale Data

```{r}
# load all excel tabs into a combined tidy tibble
target_sheets <- paste0("SUS-", 1:2)

sus_data <- map_dfr(
  target_sheets,
  ~ read_excel(
    "./data/source/i1_raw_data.xlsx",
    na = c("", "N/A"),
    sheet = .x
  ) |>
  mutate(phase = which(target_sheets == .x)) |>
  janitor::clean_names() |>
  select(participant, phase, everything(), -notes)
)

sus_data_tidy <- sus_data |>
  pivot_longer(cols = starts_with("q"),
               names_to = "question",
               values_to = "value") |>
  mutate(question = as.numeric(str_extract(question, "\\d+"))) |>
  arrange(participant, phase, question)
```

calculate System Usability Scores (sus) for each (participant, phase) using this method:

To calculate the SUS score, first sum the score contributions from each item. Each item's score contribution will range from 0 to 4. For items 1,3,5,7,and 9 the score contribution is the scale position minus 1. For items 2,4,6,8 and 10, the contribution is 5 minus the scale position. Multiply the sum of the scores by 2.5 to obtain the overall value of SU.

```{r}
# add column for score contribution based on question number
sus_data_tidy <- sus_data_tidy %>%
  mutate(score_contribution = case_when(
    question %in% c(1, 3, 5, 7, 9) ~ value - 1,
    question %in% c(2, 4, 6, 8, 10) ~ 5 - value,
    TRUE ~ NA_real_  # this line is just a fail-safe
  ))

# calculate scores for each (participant, phase)
sus_scores <- sus_data_tidy |>
  group_by(participant, phase) %>%
  summarise(sus = sum(score_contribution) * 2.5, .groups = "drop")

sus_scores <- sus_scores |>
  left_join(select(demog_xls, participant, treatment), by = "participant") |>
  select(
    participant,
    treatment,
    everything()
  ) |>
  arrange(participant, treatment, phase)
```

Process the Behaviorial Control Survey Data.
Just loading as VB has already performed the analysis and I'm not using this data.

```{r}


bcs_data <- read_excel(
    "./data/source/i1_raw_data.xlsx",
    na = c("", "N/A"),
    sheet = "BCS"
  )

bcs_data_tidy <- bcs_data |>
  janitor::clean_names() |>
  pivot_longer(cols = starts_with("q"),
               names_to = "question",
               values_to = "value") |>
  mutate(question = as.numeric(str_extract(question, "\\d+"))) |>
  arrange(participant, question)
```

Export combined XLSX file.

```{r}
library(writexl)

write_xlsx(
  list(
    "demographics" = demog_xls,
    "car outcomes" = outcomes,
    "car times" = times,
    "car results" = combined,
    "bcs raw data" = bcs_data_tidy,
    "system usability scores" = sus_scores,
    "tlx scores" = tlx_scores
  ),
  "./data/combined_results.xlsx"
)
```

TODO:
- update data dictionary and sync to box, let Victoria know
- collect all text data (observation notes in T and O; comment lines in O)
- add phase 3
- use combined data to update report generator
- analysis!!

Try to recreate the visualization from the original TLX instructions.
Needs some work to account for sources that don't contribute (weight / rating) and fix spacing.

```{r}
library(RColorBrewer)

participant_id <- 1001  # Replace with the desired participant ID
phase_num <- 2  # Replace with the desired phase number

# Filter data
filtered_data <- tlx_scores %>%
  filter(participant == participant_id, phase == phase_num)

# Create source weight and source rating datasets
source_weight_data <- filtered_data %>% filter(type == "source weight")
source_rating_data <- filtered_data %>% filter(type == "source rating")

# Join the datasets
combined_data <- left_join(source_weight_data, source_rating_data, by = c("participant", "treatment", "phase", "source"))

# Assigning a numeric value to each source based on the order in workload_factors
combined_data$source_num <- as.numeric(factor(combined_data$source, levels = workload_factors))

# Adding cumulative sum of source weights for bar positioning
combined_data$cumsum_weight <- ave(combined_data$value.x, combined_data$participant, FUN = cumsum) - (0.5 * combined_data$value.x)

# Plotting using ggplot
ggplot(combined_data, aes(x = cumsum_weight - 0.5 * value.x, y = value.y, width = value.x, fill = source)) + 
  geom_col(position = "identity", colour = "black") + 
  scale_fill_brewer(palette = "Set3", breaks = workload_factors) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(name = "Source Weights", limits = c(0, 15),
                     breaks = combined_data$cumsum_weight - 0.5 * combined_data$value.x,
                     labels = combined_data$source) + 
  labs(title = paste("Task Workload Score for Participant", participant_id, "during Phase", phase_num),
       y = "Source Rating") +
  theme_minimal() +
  theme(axis.title.x = element_text(vjust = -0.5))


```

Notes:
- Observations when thinking about how to organize the data from here... how will it be used?
  - Child events will be handled differently based on analysis goals.
    1. Events that affect individual (build times) or system (down time) performance. These will be used to apply corrections and/or calculate OEE. Ultimately these will adjust the outcome data used for the primary analysis.
    2. PWI references will be aggregated per car (total number and duration of PWI events). This will aid in the assessment of recall.
    3. Other analysis. The event data can be used to support other analysis, e.g. the relationship between number of drops or total system down time and SUS / TLX results.
- Observations while reviewing drop events
  - consider adding a step to sort all time events by start and diff results with current times.csv
  - some issues where events start b/w cars, get associated with prior car, not next (e.g. 1050,1,4)
  - or events that overlap two cars
  - what about when a drop happens during breakage? double-charging for the time
- Ignore error codes for incomplete cars (no counts recorded), 1003.1.6 and 1053.1.4
