---
title: "R Notebook"
output: html_notebook
---

Naming:
demog_xls = participant demographics (clean)
outcomes = one observation per car, e.g. completed, retired, etc. (clean)
times = one observation per timed event, e.g. car, breakage, etc. (clean)
results = joined data, one observation per outcome; data from relevant rows of times are combined into a list

Load the XLSX file

```{r}
### load libraries
library(tidyverse)
library(readxl)
```

```{r}
### read and clean the demographic data

raw_demog_xls <- read_xlsx(
  "./data/source/i1_raw_data.xlsx",
  na = c("", "N/A"),
  sheet = "DEMO")

demog_xls <- raw_demog_xls |>
  janitor::clean_names() |>
  # only use time information from trial_time column 
  mutate(trial_time = hms::as_hms(trial_time)) |>
  # combine date and time columns as single datetime and drop original cols
  mutate(trial_datetime = trial_date + trial_time) |>
  select(-c("trial_date", "trial_time")) |>
  # move numeric cols to the front
  select(all_of(c("participant", "age", "height")), everything()) |>
  # convert most cols to factors
  mutate(across(-("participant":"height" | "notes"), as.factor)) |>
  glimpse()
```

```{r}
### read and clean the car error data

raw_exp_xls <- read_xlsx(
  "./data/source/i1_raw_data.xlsx",
  sheet = "OUTCOMES")

outcomes <- raw_exp_xls |>
  janitor::clean_names() |>
    mutate(
    result = case_when(
      tolower(flag) == "c" ~ "comment",
      tolower(flag) == "i" ~ "incomplete",
      is.na(flag) ~ "complete",
      tolower(flag) == "r" ~ "retired"
    ),
    result = as.factor(result),
    flag = as.factor(flag)
  ) |>
  rename(e_car = car) |>
  replace_na(list(ce = 0, uce = 0, sel = 0, pos = 0, rot = 0, count = 0)) |>
  group_by(participant, phase, result) |>
  mutate(
    number = ifelse(result == "comment", 0, cumsum(result == result))
  ) |>
  ungroup() |>
  group_by(participant, phase) |>
  mutate(
    sequence = row_number() - 1  # have comments at sequence 0
  ) |>
  ungroup()

outcomes <- outcomes |>
  select(
    participant,
    phase,
    sequence,
    flag,
    result,
    number,
    e_car,
    ce,
    uce,
    sel,
    pos,
    rot,
    count,
    notes
  ) |>
  glimpse()

write.csv(outcomes, "test/outcomes.csv", row.names = FALSE)
```

Check that errors add up:

```{r}
all(outcomes$uce == outcomes$sel + outcomes$pos + outcomes$rot)
```

```{r}
### read and clean the car time data

col_names = c(
  "participant",
  "phase",
  "event",
  "start_time",
  "duration",
  "note"
)

raw_times_csv <- read_csv("./data/csv/i1_times_v2.csv", col_names=col_names)

times <- raw_times_csv |>
  janitor::clean_names() |>
  mutate(
    description = event,
    event = str_extract(event, "[a-zA-Z]+"), # Extract the event type
    event = as.factor(tolower(event))
  ) |>
  group_by(participant, phase, event) |>
  mutate(
    number = row_number() # sequence number for each event type
  ) |>
  ungroup() |>
  group_by(participant, phase) |>
  mutate(
    sequence = row_number()
  ) |>
  ungroup()

times <- times |>
  select(
    participant,
    phase,
    sequence,
    description,
    event,
    number,
    start_time,
    duration,
    note
  ) |>
  glimpse()

```


```{r}
library(gt)
times <- times |>
  arrange(participant, phase, sequence) |>
  group_by(participant, phase) |>
  mutate(
    end_time = start_time + duration,
    row_within_group = row_number(),
    type = case_when(
      event == "car" ~ "parent",  # all car events are parents
      event == "drop" ~ "child",  # all drop events are children
      phase == 2 &
        event == "pwi" ~ "child",  # all pwi event is a child
      row_within_group == 1 ~ "parent",  # 1st event in group is a parent
      TRUE ~ map_chr(row_within_group,  # otherwise assign based on times
                     ~ ifelse(
                       !any(start_time[1:(.x - 1)] <= start_time[.x] &
                              end_time[1:(.x - 1)] >= start_time[.x]),
                       "parent",
                       "child"
                     ))
    )
  ) |>
  ungroup()

```

Modify data after careful review. Done here instead of in the source data so the changes and rationale can be spelled out.

```{r}

ignore_list <- list(
  c(1018, 1, 2),  # trying to acquire tracking between cars
  c(1021, 1, 5),  # dismissing HL UI b/w C4 and 5
  c(1021, 1, 6),  # delayed restart of timer to account for lost clock time
  c(1026, 1, 3),  # UI dropout b/w C1 and 2
  c(1026, 1, 4),  # system crash between cars
  c(1035, 1, 6),  # HL menu popped on, ignored by subject
  c(1035, 1, 7),  # stopped timer to dismiss HL menu
  c(1060, 1, 5),  # stopped timer to reset HL system after breakage
  c(1061, 1, 4)   # stopped timer to correct axle stuck in fixture
)

# Utility function
should_ignore <- function(participant, phase, sequence) {
  any(map_lgl(ignore_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Update the data
times <- times |>
  rowwise() |>
  mutate(type = if_else(should_ignore(participant, phase, sequence), "ignore", type)) |>
  ungroup()

```

1. Change the start time of sequence 6 so that it equals the start time of sequence 7 for participant 1018, phase 1.
2. Make sequence 6 a type "child".
3. Reverse the order of the sequences.

```{r}
seq_7_start_time <- times |>
  filter(participant == 1018, phase == 1, sequence == 7) %>%
  pull(start_time)

# Now, make the changes as described
times <- times |>
  mutate(
    # Change the start time for sequence 6
    start_time = if_else(
      participant == 1018 &
        phase == 1 & sequence == 6,
      seq_7_start_time,
      start_time
    ),
    # Change type for sequence 6
    type = if_else(participant == 1018 &
                     phase == 1 & sequence == 6, "child", type)
  ) |>
  # Change the sequence order
  rowwise() |>
  mutate(sequence = if_else(
    participant == 1018 & phase == 1 & sequence == 6, 7L,
    if_else(participant == 1018 &
              phase == 1 & sequence == 7, 6L, sequence)
  )) |>
  ungroup()

write.csv(times, "test/times.csv", row.names = FALSE)
```


# Stop Here
This needs to be reworked after adding data to times.

```{r}
# library(gt)

# add temporary match_condition for the join
# need to match rows where results$result = "complete" and times$event = "car"
outcomes <- outcomes |>
  mutate(match_condition = result == "complete")

times <- times |>
  mutate(match_condition = event == "car")

# full join to preserve all rows from both tables
results <- outcomes |>
  full_join(
    times,
    by = c("participant", "phase", "number", "match_condition")
    ) |>
  rename(
    e_notes = notes,
    t_notes = note,
#    t_flag = event,
    e_flag = flag
    ) |>
  mutate(e_flag = as.factor(e_flag))

# write.csv(combined, "data/combined.csv", row.names = FALSE)

results <- results |>
  left_join(
    demog_xls |> select(participant, treatment),
    by = "participant"
  )

# needs updating...

# reorder columns
# results <- results %>%
#   select(
#     participant,
#     treatment,
#     phase,
#     sequence,
#     description,
#     event,
#     number,
#     start_time,
#     duration,
#     e_flag,
#     ce,
#     uce,
#     sel,
#     pos,
#     rot,
#     count,
#     e_notes,
#     t_notes
#   )

# write.csv(combined, "combined.csv", row.names = FALSE)

# output table
# gt(car_data)
```


```{r}
### play with plots (EDA)

library(gtExtras)

exp_xls |>
  filter(phase == 1) |>
  gt_plt_summary()
```

