---
title: "R Notebook"
output: html_notebook
---

Naming:
demog_xls = participant demographics (clean)
outcomes = one observation per car, e.g. completed, retired, etc. (clean)
times = one observation per timed event, e.g. car, breakage, etc. (clean)
results = joined data, one observation per outcome; data from relevant rows of times are combined into a list

Load the XLSX file

```{r}
### load libraries
library(tidyverse)
library(readxl)
```

```{r}
### read and clean the demographic data

raw_demog_xls <- read_xlsx(
  "./data/source/i1_raw_data.xlsx",
  na = c("", "N/A"),
  sheet = "DEMO")

demog_xls <- raw_demog_xls |>
  janitor::clean_names() |>
  # only use time information from trial_time column 
  mutate(trial_time = hms::as_hms(trial_time)) |>
  # combine date and time columns as single datetime and drop original cols
  mutate(trial_datetime = trial_date + trial_time) |>
  select(-c("trial_date", "trial_time")) |>
  # move numeric cols to the front
  select(all_of(c("participant", "age", "height")), everything()) |>
  # convert most cols to factors
  mutate(across(-("participant":"height" | "notes"), as.factor)) |>
  glimpse()
```

```{r}
### read and clean the car error data

raw_exp_xls <- read_xlsx(
  "./data/source/i1_raw_data.xlsx",
  sheet = "OUTCOMES")

outcomes <- raw_exp_xls |>
  janitor::clean_names() |>
    mutate(
    result = case_when(
      tolower(flag) == "c" ~ "comment",
      tolower(flag) == "i" ~ "incomplete",
      is.na(flag) ~ "complete",
      tolower(flag) == "r" ~ "retired"
    ),
    result = as.factor(result),
    flag = as.factor(flag)
  ) |>
  rename(e_car = car) |>
  replace_na(list(ce = 0, uce = 0, sel = 0, pos = 0, rot = 0, count = 0)) |>
  group_by(participant, phase, result) |>
  mutate(
    number = ifelse(result == "comment", 0, cumsum(result == result))
  ) |>
  ungroup() |>
  group_by(participant, phase) |>
  mutate(
    sequence = row_number() - 1  # have comments at sequence 0
  ) |>
  ungroup()

outcomes <- outcomes |>
  select(
    participant,
    phase,
    sequence,
    flag,
    result,
    number,
    e_car,
    ce,
    uce,
    sel,
    pos,
    rot,
    count,
    notes
  ) |>
  glimpse()

```

Check that errors add up:

```{r}
all(outcomes$uce == outcomes$sel + outcomes$pos + outcomes$rot)
```

```{r}
### read and clean the car time data

col_names = c(
  "participant",
  "phase",
  "event",
  "start_time",
  "duration",
  "note"
)

raw_times_csv <- read_csv("./data/csv/i1_times_v2.csv", col_names=col_names)

times <- raw_times_csv |>
  janitor::clean_names() |>
  mutate(
    description = event,
    event = str_extract(event, "[a-zA-Z]+"), # Extract the event type
    event = as.factor(tolower(event))
  ) |>
  group_by(participant, phase, event) |>
  mutate(
    number = row_number() # sequence number for each event type
  ) |>
  ungroup() |>
  group_by(participant, phase) |>
  mutate(
    sequence = row_number()
  ) |>
  ungroup()

times <- times |>
  select(
    participant,
    phase,
    sequence,
    description,
    event,
    number,
    start_time,
    duration,
    note
  ) |>
  glimpse()

```

Flag parent and child events in `times`.

```{r}
times <- times |>
  arrange(participant, phase, sequence) |>
  group_by(participant, phase) |>
  mutate(
    end_time = start_time + duration,
    row_within_group = row_number(),
    type = case_when(
      event == "car" ~ "parent",  # all car events are parents
      event == "drop" ~ "child",  # all drop events are children
      phase == 2 &
        event == "pwi" ~ "child",  # all pwi event is a child
      row_within_group == 1 ~ "parent",  # 1st event in group is a parent
      TRUE ~ map_chr(row_within_group,  # otherwise assign based on times
                     ~ ifelse(
                       !any(start_time[1:(.x - 1)] <= start_time[.x] &
                              end_time[1:(.x - 1)] >= start_time[.x]),
                       "parent",
                       "child"
                     ))
    )
  ) |>
  ungroup()

```

Modify data after careful review. Done here instead of in the source data so the changes and rationale can be spelled out.

```{r}

ignore_list <- list(
  c(1018, 1, 2),  # trying to acquire tracking between cars
  c(1021, 1, 5),  # dismissing HL UI b/w C4 and 5
  c(1021, 1, 6),  # delayed restart of timer to account for lost clock time
  c(1026, 1, 3),  # UI dropout b/w C1 and 2
  c(1026, 1, 4),  # system crash between cars
  c(1035, 1, 6),  # HL menu popped on, ignored by subject
  c(1035, 1, 7),  # stopped timer to dismiss HL menu
  c(1060, 1, 5),  # stopped timer to reset HL system after breakage
  c(1061, 1, 4)   # stopped timer to correct axle stuck in fixture
)

# Utility function
should_ignore <- function(participant, phase, sequence) {
  any(map_lgl(ignore_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Update the data
times <- times |>
  rowwise() |>
  mutate(type = if_else(should_ignore(participant, phase, sequence), "ignore", type)) |>
  ungroup()

```

1. Change the start time of sequence 6 so that it equals the start time of sequence 7 for participant 1018, phase 1.
2. Make sequence 6 a type "child".
3. Reverse the order of the sequences.

```{r}
seq_7_start_time <- times |>
  filter(participant == 1018, phase == 1, sequence == 7) |>
  pull(start_time)

# Now, make the changes as described above to correct (1018, 1, 6)
times <- times |>
  mutate(
    # Change the start time for sequence 6
    start_time = if_else(
      participant == 1018 &
        phase == 1 & sequence == 6,
      seq_7_start_time,
      start_time
    ),
    # Correct duration of that event
    duration = if_else(participant == 1018 &
                         phase == 1 & sequence == 6, 13.07, duration),
    # Change type for sequence 6
    type = if_else(participant == 1018 &
                     phase == 1 & sequence == 6, "child", type)
  ) |>
  # Change the sequence order
  rowwise() |>
  mutate(sequence = if_else(
    participant == 1018 & phase == 1 & sequence == 6, 7L,
    if_else(participant == 1018 &
              phase == 1 & sequence == 7, 6L, sequence)
  )) |>
  ungroup()

```

Correct issues in the data identified during the first join process and subsequent review of data.

```{r}
# several cars were recorded in the times table as incomplete, but not progress
# was made, therefore they weren't recorded in the outcome table
# mark these for ignore
ignore_list <- list(
  c(1005, 1, 4),  # last car not started
  c(1008, 1, 8),  # last car not started
  c(1014, 2, 1),  # first car defect, not started
  c(1036, 1, 6),  # last car not started
  c(1047, 1, 5)   # duplicate subclip in video annotation
)

# Utility function, duplicated from above
should_ignore <- function(participant, phase, sequence) {
  any(map_lgl(ignore_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the type to `ignore` for listed cars
times <- times |>
  rowwise() |>
  mutate(type = if_else(should_ignore(participant, phase, sequence), "ignore", type)) |>
  ungroup()

# Patch times data for (1033, 1, 7) incomplete - was one block short
times <- times |>
  mutate(event = as.character(event)) |>  # Convert to character
  mutate(event = ifelse(participant == 1033 &
                         phase == 1 & sequence == 7, "incomplete", event)) |>
  mutate(event = as.factor(event))  # Convert back to factor if necessary


# Patch times data for 1044, 1, 4 - used start time from prior event
condition <- times$participant == 1044 & times$phase == 1 & times$sequence == 4

times <- times |>
  mutate(
    start_time = ifelse(condition, 584, start_time),
    duration = ifelse(condition, 18.033, duration),
    type = ifelse(condition, "parent", type)
  )

# Patch outcomes data for (1033, 1) to insert a car between 3 and 5
# Move 4 to 5
outcomes <- outcomes |>
  mutate(sequence = if_else(participant == 1033 & phase == 1 & sequence == 4, 5, sequence))

# Create a copy of row 3 as 4
new_record <- outcomes |>
  filter(participant == 1033, phase == 1, sequence == 3) |>
  mutate(sequence = 4)

# Append the new record to the outcomes table and resort the result
outcomes <- bind_rows(outcomes, new_record) |>
  arrange(participant, phase, sequence)

# Drop rows in outcomes corresponding to (1019, 1, 8 and 9) and (1021, 1, 7),
# due to: duplicate (c&p error?), no progress, no progress
outcomes <- outcomes |>
  filter(
    !(participant == 1019 & phase == 1 & sequence %in% c(8, 9)) & 
    !(participant == 1021 & phase == 1 & sequence == 7)
  )

```

Clean up tables.

```{r}
# combine the description (original event name) and note fields
times <- times |>
  mutate(note = if_else(note == "N/A", 
                      paste0("(", description, ")"), 
                      paste0("(", description, ") ", note)))

# assign a UID for each time event, e.g. 1001-P1-C1-T0
times <- times |>
  filter(type != "ignore") |>
  group_by(participant, phase) |>
  mutate(parent_seq = cumsum(type == "parent")) |>
  group_by(participant, phase, parent_seq) |>
  mutate(child_seq = if_else(type == "child", cumsum(type == "child"), 0)) |>
  ungroup() |>
  mutate(UIDT = paste0(
    participant,
    "-P", phase,
    "-C", parent_seq,
    "-T", sprintf("%02d", child_seq)
  ))

# drop some fields and reorder others
times <- times |>
  select(
    participant,
    phase,
    sequence,
    UIDT,
    event,
    type,
    start_time,
    duration,
    end_time,
    note
  ) |>
  glimpse()

# assign a UID for each outcome, e.g. 1001-P1-C1, and choose / order fields
outcomes <- outcomes |>
  group_by(participant, phase) |>
  mutate(UIDO = paste(participant, 
                     paste0("P", phase), 
                     paste0("C", sequence), 
                     sep = "-")) |>
  ungroup() |>
  select(
    participant,
    phase,
    sequence,
    UIDO,
    result,
    ce,
    uce,
    sel,
    pos,
    rot,
    count,
    notes
  ) |>
  glimpse()

# write.csv(times, "test/times.csv", row.names = FALSE)
# write.csv(outcomes, "test/outcomes.csv", row.names = FALSE)
```

## Observations when thinking about how to organize the data from here... how will it be used?

Child events will be handled differently based on analysis goals.
1. Events that affect individual (build times) or system (down time) performance. These will be used to apply corrections and/or calculate OEE. Ultimately these will adjust the outcome data used for the primary analysis.
2. PWI references will be aggregated per car (total number and duration of PWI events). This will aid in the assessment of recall.
3. Other analysis. The event data can be used to support other analysis, e.g. the relationship between number of drops or total system down time and SUS / TLX results.


Second cleaning pass, after improved join and review of child events. Three classes of event:
1. Drop events
2. PWI events
3. All other types (e.g. reset, system, tracking, video, etc.)

First, I reviewed the third type, "other." In doing so, I identified three more general classes:
- Ignore: events that occured between cars and therefore had no impact on analysis
- System: delays introduced into the parent car build due to system-related issues, e.g. HL menu appears / dismissing, crashes and resets, model tracking acquisition
- Repair: delays introduced into the parent car build due to the participant's decision to make repairs rather than retire the car

In total there were 20 such events, all of which are corrected in the following code block.

```{r}
## re-classify misc event types as ignore, system, or repair

# IGNORE: patch times to ignore events that will have no impact on analysis
ignore_list <- list(
  c(1006, 1, 3),  # video recording drop, no performance impact
  c(1007, 1, 3),  # video recording drop, no performance impact
  c(1007, 1, 6),  # video recording drop, no performance impact
  c(1030, 1, 2),  # time lost to clock after car 1 marked incomplete due to crash
  c(1041, 1, 4)   # car 3 finished prior to breakage, so this occurred b/w cars
)

# Utility function, duplicated from above
should_ignore <- function(participant, phase, sequence) {
  any(map_lgl(ignore_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the event to `ignore` for listed cars
times <- times |>
  rowwise() |>
  mutate(event = if_else(should_ignore(participant, phase, sequence), "ignore", event)) |>
  ungroup()


# REPAIR: patch times to classify some events as time spent repairing a car
patch_list <- list(
  c(1011, 1, 2),
  c(1011, 1, 5),
  c(1014, 1, 2),
  c(1063, 1, 5)
)

# Utility function, duplicated from above
should_patch <- function(participant, phase, sequence) {
  any(map_lgl(patch_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the event to `reapair` for listed cars
times <- times |>
  rowwise() |>
  mutate(event = if_else(should_patch(participant, phase, sequence), "repair", event)) |>
  ungroup()

  
# SYSTEM: patch times to classify some events as time lost to system issues
patch_list <- list(
  c(1018, 1, 4), # trying to acquire tracking during HMDMR
  c(1018, 1, 7), # trying to acquire tracking during HMDMR
  c(1026, 1, 6), # interrupted by HL menu and time to deactivate it
  c(1039, 1, 2), # system drop and crash on the clock during car 1
  c(1039, 1, 3), # resetting system during car 1
  c(1055, 1, 8), # during car 3 clip
  c(1057, 1, 4), # dismiss HL menu during car 3 clip
  c(1057, 2, 2), # dismiss HL menu during car 1 clip
  c(1057, 2, 5), # dismiss HL menu during car 3 clip
  c(1063, 1, 13), # system drop and crash on the clock during car 2
  c(1063, 1, 14) # system reset during car 2 clip
)

# Utility function, duplicated from above
should_patch <- function(participant, phase, sequence) {
  any(map_lgl(patch_list, ~ identical(c(participant, phase, sequence), .x)))
}

# Patch times data to set the event to `repair` for listed cars
times <- times |>
  rowwise() |>
  mutate(event = if_else(should_patch(participant, phase, sequence), "system", event)) |>
  ungroup()

```

Went through all drop events to assess the percentage of its duration that impacted performance.  Created a table of the results, where percentatges range from 0.0 (no impact) to 1.0 (full).

Now add adj_drop time to times table.

```{r}
raw_adj_drops <- read_xlsx("./data/source/adjusted_drop_events.xlsx") |>
  janitor::clean_names()

times <- times |>
  left_join(raw_adj_drops, by = c("participant" = "part", "phase", "sequence" = "t_seq")) |>
  mutate(
    adj_drop = ifelse(
      event == "drop", 
      duration * coalesce(how_much, 1),  # If 'how_much' is missing, we assume a default value of 1
      NA_real_
    )
  ) |>
  rename(drop_adj_perc = how_much) |>
  select(-note, everything(), -during_car, -impact_perf, note) |>
  filter(event != "ignore") |>
  glimpse()

write.csv(times, "test/times.csv", row.names = FALSE)
```

```{r}
# test for duplicate rows in the times table
times |> 
  filter(event == "drop") |> 
  group_by(UIDT) |> 
  tally(sort = TRUE) |> 
  filter(n > 1)

```

Notes from reviewing drop events:
- consider adding a step to sort all time events by start and diff results with current times.csv
- some issues where event times start between cars, get associated with prior car, should be with next (e.g. 1050,1,4)
- or overlap two cars
- what about when a drop happens during breakage? double-charging for the time


Prepare for and execute join, test results.

```{r}
joined <- times |>
  mutate(UIDT_sub = str_replace(UIDT, "-T\\d{1,2}$", "")) |>
  full_join(outcomes, by = c("UIDT_sub" = "UIDO"))

write.csv(joined, "test/joined.csv", row.names = FALSE)
```

Generate two new columns for each type of child event (repair, system, pwi, drop):
- event_ct: This counts the number of child events of each type per parent.
- event_time: This sums the durations of child events of each type per parent.
  - For drop, this will sum the values in adj_drop instead of duration.

```{r}
# 
# Define the function to generate event summaries
generate_event_summary <- function(data, event_type) {
  time_col <- ifelse(event_type == "drop", "adj_drop", "duration")
  
  summary_colname <- paste0(event_type, "_event_ct")
  time_colname <- paste0(event_type, "_event_time")
  
  data %>%
    filter(event == event_type) %>%
    group_by(UIDT_sub) %>%
    summarise(!!summary_colname := n(),
              !!time_colname := sum(get(time_col), na.rm = TRUE)) %>%
    ungroup()
}

# Use the function for each event type
event_types <- c("drop", "repair", "system", "pwi")
summaries <- lapply(event_types, generate_event_summary, data = joined)

# Merge the summaries back with the original dataset
joined_augmented <- joined
for (summary in summaries) {
  joined_augmented <- left_join(joined_augmented, summary, by = "UIDT_sub")
}

write.csv(joined_augmented, "test/joined_aug.csv", row.names = FALSE)

```

Automated tests.
Check that the aggregated data for each parent matches the sum of its child rows for all event types.

```{r, echo=FALSE, results='hide', message=TRUE}
# Create a function to perform tests for each event type
perform_tests <- function(event_type) {
  
  # Corresponding column names for aggregated counts and durations
  event_ct_col <- paste0(event_type, "_event_ct")
  event_time_col <- paste0(event_type, "_event_time")
  
  # Duration column in the original dataset
  duration_col <- ifelse(event_type == "drop", "adj_drop", "duration")
  
  # Test 1: Compare total number of child events for original and aggregated data
  total_original <- sum(joined$event == event_type, na.rm = TRUE)
  total_aggregated <- joined_augmented %>%
    filter(type == "parent") %>%
    summarize(total = sum(!!sym(event_ct_col), na.rm = TRUE)) %>%
    pull(total)
  
  stopifnot(total_original == total_aggregated)
  
  # Test 2: Compare total duration for original and aggregated data
  total_duration_original <- joined %>%
    filter(event == event_type) %>%
    summarise(total_duration = sum(!!sym(duration_col), na.rm = TRUE)) %>%
    pull(total_duration)
  
  total_duration_aggregated <- joined_augmented %>%
    filter(type == "parent") %>%
    summarise(total_duration = sum(!!sym(event_time_col), na.rm = TRUE)) %>%
    pull(total_duration)
  
  stopifnot(total_duration_original == total_duration_aggregated)
  
  message(paste0("All tests passed for event type: ", event_type))
}

# List of event types
event_types <- c("drop", "pwi", "repair", "system")

# Apply the function to each event type
invisible(lapply(event_types, perform_tests))

```


