---
title: "Initial Analysis"
output: html_notebook
---


## Setup Code

```{r setup, echo=FALSE, results='hide'}
# Set defaults for knitr output
knitr::opts_chunk$set(
  echo = TRUE,         # include code (folded by default, see above)
  results = 'markup',  # code output included as markup (default)
  message = FALSE,     # hide messages
  # include = no used, controls inclusion of both code and output (T/F);
  # combination of echo and results settings
  warning = TRUE       # do not hide warnings
)
```

```{r load libraries, echo=FALSE}
library(tidyverse)
library(readxl)
library(skimr)
```

```{r load data, echo=FALSE}
car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results")

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics")
```

*Loaded libraries and data.*


## Questions to Answer

- Is there a significant difference in phase 2 results for different treatments?
  - result: average car time
  - result: average car error count
- Can we assess learning rates by treatment type from phase 1 results?
- Can we uncover relationships with underlying affordances of treatments?
- Do demographics affect results?


## Descriptive Statistics

```{r}
# number of participants
participant_count <- demographics |>
  pull(participant) |>
  n_distinct()
```

Number of participants: `r participant_count`
