---
title: "Initial Analysis"
output:
  html_notebook:
    code_folding: hide
---

```{markdown, include=FALSE}

# ChatGPT Initializer

Working with two data tables: car_results and demographics, with common columns participant and treatment.

`car_results`: 620 rows, 25 columns. Combines `Car Outcomes` and `Car Times` data. Columns: `participant`(numeric), `treatment`(PWI, PAR, etc.), `phase`(1=learn, 2=recall, 3=retention), `sequence`(order within participant, phase group), `car_outcome`(Comment, Complete, Incomplete, Retired), `ce`(corrected errors count), `uce`(uncorrected errors count), `sel`(uncorrected errors - wrong parts), `pos`(uncorrected errors - wrong position), `rot`(uncorrected errors - wrong rotation), `count`(parts installed for incomplete cars), `error_codes`(error details), `car_start_time`(seconds from video start), `car_end_time`(seconds from video start), `car_elapsed_time`(event duration in seconds), `drop_event_count`(count of drop events), `drop_event_tot_time`(total time of drop events), `drop_event_tot_time_adj`(time-adjusted total of drop events), `repair_event_count`(count of repair events), `repair_event_tot_time`(total time of repair events), `system_event_count`(count of system events), `system_event_tot_time`(total time of system events), `pwi_event_count`(PWI references count in phase 2), `pwi_event_tot_time`(total time of PWI references).

`demographics`: 62 rows, 23 columns. Summarizes participant characteristics. Columns: `participant`(numeric), `age`(years, numeric), `height`(inches, numeric), `gender`(F, M, O), `race`(1 to 7 coded), `ethnicity`(1 to 3 coded), `country`(text), `language`(English, Other), `education`(1 to 7 coded), `college`(text), `program`(text), `lego`(1 to 4 coded), `manuf`(1 to 4 coded), `lenses`(G, C, N), `worn`(Y, N), `color_blind`(Y, N), `condition`(Y, N), `runner`(5, 10, H, M, T, Y, N), `referral`(Class codes, GSE, FLYER, SCREEN, FF), `recruit`(AU, IND), `treatment`(PWI, PAR, AR, MR), `notes`(text), `trial_datetime`(datetime).
```


## Setup Code

```{r setup, echo=FALSE, results='hide'}
# Set defaults for knitr output
knitr::opts_chunk$set(
  echo = TRUE,         # include code (folded by default, see above)
  results = 'markup',  # code output included as markup (default)
  message = FALSE,     # hide messages
  # include = not used, controls inclusion of both code and output (T/F);
  # combination of echo and results settings
  warning = TRUE,      # do not hide warnings
  fig.width = 7,
  fig.height = 5
)
```

```{r load libraries, echo=FALSE}
library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
# library(gridExtra)
# library(ggtext)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r load data, echo=FALSE}
car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results")

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics")
```

Loaded libraries and data.

## Questions to Answer

- Is there a significant difference in phase 2 results for different treatments?
  - result: average car time
  - result: average car error count
- Can we assess learning rates by treatment type from phase 1 results?
- Can we uncover relationships with underlying affordances of treatments?
- Do demographics affect results?


## Descriptive Statistics

```{r}
# number of participants
participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# result 62 matches table row count and number of appointments (trial_datetime)
```

Data updated September 19, 2023

Number of participants: `r participant_count`


```{r demographic box plots, warning=FALSE}

# options(repr.plot.width = 10, repr.plot.height = 5)

base_color <- "#69b3a2"

# Creating box plots for 'age'
age_plot <- demographics |> 
  ggplot(aes(x = "Age (years)", y = age)) +
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.5, color = base_color) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(
    base_size = 14,
  )

# Creating box plots for 'height'
height_plot <- demographics |> 
  ggplot(aes(x = "Height (feet)", y = height)) +
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.5, color = base_color) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(
    base_size = 14,
    )

# Bar chart for 'gender'
gender_plot <- demographics |>
  ggplot(aes(x = gender, fill = gender, alpha = gender)) +
  geom_bar(width = 0.4) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  scale_fill_manual(values = c("F" = base_color, "M" = base_color, "O" = base_color)) +
  scale_alpha_manual(values = c("F" = 0.3, "M" = 0.6, "O" = 1.0)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

### combine using patchwork
age_plot + height_plot + gender_plot +
  plot_annotation(
    title = "Participant Demographics: Age, Height, and Gender",
    subtitle = "Age reflects student-oriented nature of the sample"
  )

```


```{r}
# Compute summary statistics
demographics |>
  select(age:gender, treatment) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    label = "age" ~ "Age"
  ) |>
  modify_header(all_stat_cols() ~ "**{level}**<br>N = {n} ({style_percent(p)}%)") |>
  add_n() |>
  bold_labels() |>
  as_gt()
# tab_source_note("Data updated September 19, 2023")

```

```{r}
car_results |>
  filter(phase == 2) |>
  ggplot(aes(x = treatment, y = car_elapsed_time)) +
  geom_boxplot() +
  labs(
    title = "Car Elapsed Time by Treatment for Phase 2",
    x = "Treatment",
    y = "Car Elapsed Time (seconds)"
  )
    
```

```{r}
phase_two_data <- car_results |> filter(phase == 2)

anova_result <- aov(car_elapsed_time ~ treatment, data = phase_two_data)

anova_summary <- summary(anova_result)
print(anova_summary)
```


```{r}
# If significant, conduct a post hoc test
p_value = anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  post_hoc <- TukeyHSD(anova_result)
  print(post_hoc)
} else {
  print("not significant")
}
```

```{r}
plot(post_hoc)
```





