---
title: "Initial Analysis"
format:
  html: 
    code-fold: true
    code-tools: true
    code-link: true
    code-line-numbers: true
    toc: true
    toc-expand: 2
    number-sections: true
    number-depth: 3
    css: custom.css
execute: 
  echo: true          # include code (folded by default, see above)
  results: 'markup'   # code output included as markup (default)
  message: false      # hide messages
  warning: true       # do not hide warnings
  fig.width: 7
  fig.height: 5
  # include: not used, controls inclusion of both code and output (T/F);
  # combination of echo and results settings
---

## Notes {.hide .unnumbered .unlisted}

### ChatGPT Initializer

Working with two data tables: car_results and demographics, with common columns participant and treatment.

`car_results`: 620 rows, 25 columns. Combines `Car Outcomes` and `Car Times` data. Columns: `participant`(numeric), `treatment`(PWI, PAR, etc.), `phase`(1=learn, 2=recall, 3=retention), `sequence`(order within participant, phase group), `car_outcome`(Comment, Complete, Incomplete, Retired), `ce`(corrected errors count), `uce`(uncorrected errors count), `sel`(uncorrected errors - wrong parts), `pos`(uncorrected errors - wrong position), `rot`(uncorrected errors - wrong rotation), `count`(parts installed for incomplete cars), `error_codes`(error details), `car_start_time`(seconds from video start), `car_end_time`(seconds from video start), `car_elapsed_time`(event duration in seconds), `drop_event_count`(count of drop events), `drop_event_tot_time`(total time of drop events), `drop_event_tot_time_adj`(time-adjusted total of drop events), `repair_event_count`(count of repair events), `repair_event_tot_time`(total time of repair events), `system_event_count`(count of system events), `system_event_tot_time`(total time of system events), `pwi_event_count`(PWI references count in phase 2), `pwi_event_tot_time`(total time of PWI references).

`demographics`: 62 rows, 23 columns. Summarizes participant characteristics. Columns: `participant`(numeric), `age`(years, numeric), `height`(inches, numeric), `gender`(F, M, O), `race`(1 to 7 coded), `ethnicity`(1 to 3 coded), `country`(text), `language`(English, Other), `education`(1 to 7 coded), `college`(text), `program`(text), `lego`(1 to 4 coded), `manuf`(1 to 4 coded), `lenses`(G, C, N), `worn`(Y, N), `color_blind`(Y, N), `condition`(Y, N), `runner`(5, 10, H, M, T, Y, N), `referral`(Class codes, GSE, FLYER, SCREEN, FF), `recruit`(AU, IND), `treatment`(PWI, PAR, AR, MR), `notes`(text), `trial_datetime`(datetime).


## Change Log

- 9/29 - first revision shared with Dr. Harris
- 10/1 - update for Drs. Harris and Evans
  - treatment effects only look at completed cars now, ignoring retired (broken) and incomplete (time expired) trials; this did not have a noticable effect on the results, but more accurately reflects the intention
  - added hypotheses, expectations, and observations throughout
  - migrated to Quarto format
  - added to-do items

::: {.callout-note collapse=false}
# **To Do Items**
- Misc
  - [x] Convert `{r load data, echo=FALSE}` format to Quarto style
  - [ ] Add plot labels to code blocks
  - [ ] Add tests for assumptions of ANOVA, etc.
  - [ ] Refactor ANOVA, Tukey, and Assumptions into function
:::

### Questions to Answer

- Is there a significant difference in phase 2 results for different treatments?
  - result: average car time
  - result: average car error count
- Can we assess learning rates by treatment type from phase 1 results?
- Can we uncover relationships with underlying affordances of treatments?
- Do demographics affect results?

## Setup Code

```{r}
#| label: load-libs
#| output: false

library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
# library(gridExtra)
# library(ggtext)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r}
#| label: load-data
#| echo: false

car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results")

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics")
```

Loaded libraries and data.

## Descriptive Statistics

```{r}
#| label: num-participants

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# result 62 matches table row count and number of appointments (trial_datetime)
```

Data updated September 19, 2023

Number of participants: `r participant_count`

::: {.callout-note collapse=false}
# **To Do Items**
- Descriptive Statistics
  - [ ] Make the date based on file timestamp.
:::

```{r}
#| label: demo-box-plots
#| warning: false

base_color <- "#69b3a2"

# Creating box plots for 'age'
age_plot <- demographics |> 
  ggplot(aes(x = "Age (years)", y = age)) +
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.5, color = base_color) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(
    base_size = 14,
  )

# Creating box plots for 'height'
height_plot <- demographics |> 
  ggplot(aes(x = "Height (feet)", y = height)) +
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.5, color = base_color) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(
    base_size = 14,
    )

# Bar chart for 'gender'
gender_plot <- demographics |>
  ggplot(aes(x = gender, fill = gender, alpha = gender)) +
  geom_bar(width = 0.4) +
  labs(
    x = element_blank(),
    y = element_blank()
  ) +
  scale_fill_manual(values = c("F" = base_color, "M" = base_color, "O" = base_color)) +
  scale_alpha_manual(values = c("F" = 0.3, "M" = 0.6, "O" = 1.0)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

### combine using patchwork
age_plot + height_plot + gender_plot +
  plot_annotation(
    title = "Participant Demographics: Age, Height, and Gender",
    subtitle = "Age reflects student-oriented nature of the sample"
  )

```

### Demographics by Treatment

```{r}
#| label: demo-by-treatment

# Compute summary statistics
demographics |>
  select(age:gender, treatment) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    label = "age" ~ "Age"
  ) |>
  modify_header(all_stat_cols() ~ "**{level}**<br>N = {n} ({style_percent(p)}%)") |>
  add_n() |>
  bold_labels() |>
  as_gt()
# tab_source_note("Data updated September 19, 2023")

```

## Treatment Effects

### Elapsed Time

**Hypothesis:** elapsed times will differ

**Expectation:** PAR will be faster than PWI (in-context instruction). This benefit will be mitigated by some inconsistency in gesture recognition. AR and MR will take progressively longer than PAR, owing to similar benefits but more complex and less reliable interfaces.

Results in line with expectations. Surprised to see mean time with AR the highest. Note increased variance in results in the treatment progression. Surprised that PAR and PWI are not significantly different.

::: {.callout-note collapse=false}
# **To Do Items**
- Elapsed Time
  - [ ] Compare with adjusted versions - significant differences?
  - [ ] How would this differ if PAR used a discrete switch rather than a gesture sensor?
:::

```{r}
#| label: boxplot-et-by-treat

car_results |>
  filter(phase == 2, car_outcome == "complete") |>
  ggplot(aes(x = treatment, y = car_elapsed_time)) +
  geom_boxplot() +
  labs(
    title = "Car Elapsed Time by Treatment for Phase 2",
    x = "Treatment",
    y = "Car Elapsed Time (seconds)"
  )
    
```

```{r}
#| label: anova-et-by-treat

phase_two_data <- car_results |> filter(phase == 2, car_outcome == "complete")

time_anova_result <- aov(car_elapsed_time ~ treatment, data = phase_two_data)

time_anova_summary <- summary(time_anova_result)
print(time_anova_summary)
```

```{r}
#| label: tukey-et-by-treat

p_value = time_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  time_post_hoc <- TukeyHSD(time_anova_result)
  print(time_post_hoc)
} else {
  print("not significant")
}
```

```{r}
#| label: plot-tukey-et

plot(time_post_hoc)
```

### Uncorrected Error Count

**Hypothesis:** error count will differ

**Expectation:** PWI will be highest, other treatments will show decreasing error counts in order (PWI, AR, MR) due to multimodal learning. Improvements in AR and especially MR may be confounded by added complexity / reduced reliability of added technology.

Results meet expectations. Note that the PWI are clearly problematic as we observed. Though the other treatments are designed around the same instructions, each afford a certain amount of built-in error checking - the operator sees where the part should go - that help avoid making errors and/or learning to complete the task incorrectly.

Surprised to see AR error count so low. Task isn't complex enough to show lower limit.

*ChatGPT: You might say that the results exhibit "floor effects," where the lower limit of the measure (in this case, zero) constrains the range of possible scores, especially for certain treatments. This can indicate that the measure may not be sensitive enough to detect differences at the lower end of the performance spectrum for these treatments.*

::: {.callout-note collapse=false}
# **To Do Items**
- Uncorrected Error Count
  - [ ] Error types? Patterns?
:::

```{r}
#| label: boxplot-uce-by-treat

car_results |>
  filter(phase == 2, car_outcome == "complete") |>
  ggplot(aes(x = treatment, y = uce)) +
  geom_boxplot() +
  labs(
    title = "Uncorrected Error Count by Treatment for Phase 2",
    x = "Treatment",
    y = "Count"
  )
    
```

```{r}
#| label: anova-uce-by-treat

error_anova_result <- aov(uce ~ treatment, data = phase_two_data)

error_anova_summary <- summary(error_anova_result)
print(error_anova_summary)
```

```{r}
#| label: tukey-uce-by-treat

p_value = error_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  error_post_hoc <- TukeyHSD(error_anova_result)
  print(error_post_hoc)
} else {
  print("not significant")
}
```

```{r}
#| label: plot-tukey-uce

plot(error_post_hoc)
```

## PWI Reference Count

::: {.callout-note collapse=false}
# **To Do Items**
- Treatment Effects
  - [ ] Add PWI Reference Count analysis
  - [ ] Add PWI Reference Duration analysis
:::

How many times did participants have to refer to the paper work instructions during phase 2, by treatment?

```{r}
#| label: boxplot-pwi-ref-by-treat

car_results |>
  filter(phase == 2, car_outcome == "complete") |>
  ggplot(aes(x = treatment, y = uce)) +
  geom_boxplot() +
  labs(
    title = "Uncorrected Error Count by Treatment for Phase 2",
    x = "Treatment",
    y = "Count"
  )
    
```

## PWI Reference Duration

How much total time did participants spend referring to the paper work instructions during phase 2, by treatment?

## Explore Difference in UX by Treatment

Load System Usability and TLX scores.

```{r}
#| label: load-sus-and-tlx
#| echo: false

sus_scores <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "system usability scores")

tlx_scores <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "tlx scores")
```

### System Usability Scores (SUS)

**Hypothesis:** There will be differences in system usability.

**Expectations:** Decreasing usability owing to increased complexity: PWI > PAR > AR > MR.

Surprised to see no significant difference in perceived usability between treatments.

::: {.callout-note collapse=false}
# **To Do Items**
- System Usability Scores
  - [ ] Convert chart to box plot?
  - [ ] Compare PWI treatment across phase 1 and 2?
  - [ ] Phase 2 PWI could serve as baseline for system usability. Use that for comparisons?
:::

```{r}
#| label: barplot-sus-by-treat

sus_summary <- sus_scores |>
  filter(phase == 1) |>
  group_by(treatment) |>
  summarise(
    mean_value = mean(sus, na.rm = TRUE),
    sd_value = sd(sus, na.rm = TRUE)
  )

ggplot(sus_summary, aes(x = treatment, y = mean_value)) +
  geom_bar(stat = "identity", position = "dodge", fill = base_color) +
  geom_errorbar(
    aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
    width = 0.25, position = position_dodge(0.9)
  ) +
  labs(
    title = "SUS by Treatment in Phase 1",
    y = "Mean SUS",
    x = "Treatment"
  ) +
  theme_minimal()

```

```{r}
#| label: anova-sus-by-treat

sus_anova_result <- aov(sus ~ treatment, data = sus_scores)
sus_anova_summary <- summary(sus_anova_result)
print(sus_anova_summary)
```

```{r}
#| label: tukey-sus-by-treat

p_value = sus_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  post_hoc <- TukeyHSD(sus_anova_result)
  print(sus_post_hoc)
} else {
  print("not significant")
}
```

### Task Load Index Scores (TLX)

**Hypothesis:** There will be differences in perceived workload.

**Expectations:** Variable workload based on a combination of instructional quality and interface complexity: PWI high, PWI medium, AR medium, MR high

Surprised to see very even workload across treatments and no significant difference. Suggests other factors have larger effect across all treatment? Experiment design? Time constraints?

::: {.callout-note collapse=false}
# **To Do Items**
- Task Load Index
  - [ ] Convert chart to box plot?
  - [ ] Compare PWI treatment across phase 1 and 2?
  - [ ] Phase 2 PWI could serve as baseline for task workload. Use that for comparisons?
  - [ ] Underlying factors of TLX?
:::

```{r}
#| label: barplot-tlx-by-treat

tlx_task_workload_summary <- tlx_scores |>
  filter(type == "task workload", phase == 1) |>
  group_by(treatment) |>
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )

ggplot(tlx_task_workload_summary, aes(x = treatment, y = mean_value)) +
  geom_bar(stat = "identity", position = "dodge", fill = base_color) +
  geom_errorbar(
    aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
    width = 0.25, position = position_dodge(0.9)
  ) +
  labs(
    title = "Task Workload by Treatment in Phase 1",
    y = "Mean Task Workload",
    x = "Treatment"
  ) +
  theme_minimal()

```

```{r}
#| label: anova-tlx-by-treat

tlx_anova_result <- aov(value ~ treatment, data = tlx_scores)
tlx_anova_summary <- summary(tlx_anova_result)
print(tlx_anova_summary)
```

```{r}
#| label: tukey-tlx-ty-treat

p_value = tlx_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  post_hoc <- TukeyHSD(tlx_anova_result)
  print(tlx_post_hoc)
} else {
  print("not significant")
}
```

## Visualize Car Time Trends in Phase 1

```{r}
#| label: car-time-trends

# Filter data for completed cars
completed_cars <- car_results |> 
  filter(car_outcome == "Complete")

# Create car_number column
completed_cars <- completed_cars |> 
  group_by(participant, treatment) |> 
  mutate(car_number = row_number()) |> 
  ungroup()

# # Plot data
# ggplot(data = completed_cars, aes(x = car_number, y = car_elapsed_time)) +
#   geom_point() +
#   facet_wrap(~treatment, scales = "free_x") +
#   labs(
#     title = "Car Completion Times by Treatment",
#     x = "Car Number",
#     y = "Completion Time (seconds)"
#   )

```

