---
title: "Initial Analysis"
output:
  html_notebook:
    code_folding: hide
---

## Setup Code

```{r setup, echo=FALSE, results='hide'}
# Set defaults for knitr output
knitr::opts_chunk$set(
  echo = TRUE,         # include code (folded by default, see above)
  results = 'markup',  # code output included as markup (default)
  message = FALSE,     # hide messages
  # include = not used, controls inclusion of both code and output (T/F);
  # combination of echo and results settings
  warning = TRUE       # do not hide warnings
)
```

```{r load libraries, echo=FALSE}
library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(gridExtra)
# library(grid)
```

```{r load data, echo=FALSE}
car_results <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "car results")

demographics <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "demographics")
```

Loaded libraries and data.


## Questions to Answer

- Is there a significant difference in phase 2 results for different treatments?
  - result: average car time
  - result: average car error count
- Can we assess learning rates by treatment type from phase 1 results?
- Can we uncover relationships with underlying affordances of treatments?
- Do demographics affect results?


## Descriptive Statistics

```{r}
# number of participants
participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# result 62 matches table row count and number of appointments (trial_datetime)
```

Number of participants: `r participant_count`

```{r}
# Compute summary statistics
demographics |>
  select(age:gender, treatment) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    label = "age" ~ "Age"
  ) |>
  modify_header(all_stat_cols() ~ "**{level}**<br>N = {n} ({style_percent(p)}%)") |>
  add_n() |>
  bold_labels() |>
  as_gt() |>
  tab_source_note("Data updated September 19, 2023")

```

```{r demographic histograms}
# Participant age range histogram
demographics |> 
  filter(!is.na(age)) |>
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "#69b3a2", color = "white", alpha = 0.8) +
  labs(title = "Age Distribution",
       x = "Age",
       y = "Count") +
  theme_minimal()

# Participant height range histogram
demographics |> 
  filter(!is.na(height)) |>
  mutate(height_ft = height / 12) |>
  ggplot(aes(x = height_ft)) +
  geom_histogram(binwidth = 0.1, fill = "#69b3a2", color = "white", alpha = 0.8) +
  labs(title = "Height Distribution",
       x = "Height (ft)",
       y = "Count") +
  theme_minimal()
```

```{r demographic box plots, warning=FALSE}
# Creating box plots for 'age'
age_plot <- demographics |> 
  ggplot(aes(x = factor(1), y = age)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5, color = "blue") +
  labs(x = "Age", y = "Age (years)")

# Creating box plots for 'height'
height_plot <- demographics |> 
  ggplot(aes(x = factor(1), y = height)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5, color = "blue") +
  labs(x = "Height", y = "Height (inches)")

# Combine and display the plots side by side
age_plot + height_plot + plot_layout(ncol = 2)

# print(combined_plot)
```

