# H1b Learning Rates {#25b-lr}

```{r load-libs}
#| warning: false

library(here)
library(tidyverse)
library(pwr)
library(emmeans)
library(performance)
library(lmerTest)
library(gt)
library(patchwork)
library(moments)
library(ggpubr)
library(ggstatsplot)
```

```{r load-data}
# Load data objects from the file
load(here("rdata", "learn.RData"))

# load functions from h1a
load(here("rdata", "h1a.RData"))
```


```{r}
# use the same method as in h1a to prep data

# create data
learn_h1b <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj)

# mutate
learn_h1b <- learn_h1b |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

glimpse(learn_h1b)
```

#### Learning Rates - First attempt at LME model formulation

I was unable to get a solution to converge, so began exploring alternatives, including:

- mixed-effects model
- growth curve analysis
- survival analysis (time to reach standard)

Having no experience with any of these approaches, I decided to start with the most straightforward of the bunch - MEM.

To see if the effect of the sequence number (learning curve) depends on the treatment, an interaction term is included:

$$
\text{TCT}_{ij} = \beta_{0} + \beta_{1} \text{TRT}_{i} + \beta_{2} \text{SEQ}_{j} + \beta_{3} (\text{TRT}_{i} \times \text{SEQ}_{j}) + (1 + \text{SEQ}_{j} \mid \text{PART}_{i}) + \epsilon_{ij}
$$

A nice thing about mixed-effect models is that they inherently account for the unbalanced nature of repeated measures data like ours, where participants complete a different number of tasks. Because participants are treated as random effects this contribution to variability is accounted for.

The dataset for this hypothesis has the form:

```{r make-h1b-data}
# learning rate modeling requires disaggregated data
learn_h1b <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj) |>
  group_by(participant) |>
  mutate(seq = row_number()) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

glimpse(learn_h1b)
```

Each row represents the outcome of a single participant repetition of a task, where:

- participant is a UID
- treatment is instructional method: AR (augmented reality), MR (mixed reality), PAR (projected AR), PWI (paper work instructions)
- seq is the task number
- tct is the task completion time (in seconds)
- uce is the number of uncorrected errors

Here is a plot of smoothed TCT data by treatment:

```{r}
#| warning: false

# Plot smoothed TCT data by treatment
ggplot(
  learn_h1b,
  aes(
    x = seq,
    y = tct,
    color = treatment,
    group = participant
  )
) +
  # geom_line(aes(group = participant), alpha = 0.3) +
  geom_smooth(aes(group = treatment),
    method = "loess",
    se = FALSE,
    size = 1
  ) +
  labs(
    title = "Learning Curves by Treatment",
    x = "Task Number",
    y = "Task Completion Time (TCT)"
  ) +
  theme_minimal()

```

AR, MR appear to have higher slope, but neither reach takt time of 60 seconds.

Here is the output of a model fit based on non-standardized data:

```{r h1b-lmer-model-1}
#| warning: false

# library(lme4)
library(broom.mixed)
library(lmerTest)

# use unordered factor to improve interpretability of results
learn_h1b$treatment <- factor(as.character(learn_h1b$treatment))

# Fit the mixed-effects model with interaction effects
# use the overloaded version of lmer from lmerTest to get CI, p-vals
h1b_mod_lmer <- lmerTest::lmer(
  tct ~ treatment * seq + (1 + seq | participant),
  data = learn_h1b,
  control = lmerControl(optimizer = "bobyqa")
)

# Summary of the model to understand fixed and random effects
h1b_mod_sum <- summary(h1b_mod_lmer)

h1b_mod_sum

# resid(h1b_mod_lmer)
# coef(h1b_mod_lmer)
# ranef(h1b_mod_lmer)

# Clean output for easier interpretation if needed
# h1b_mod_tsum <- tidy(h1b_mod_lmer, effects = "fixed", conf.int = TRUE)

# verbose narrative report
# report::report(h1b_mod_lmer)

# other display methods and diagnostics
# tbl_regression(h1b_mod_lmer)
# modelsummary::modelsummary(h1b_mod_lmer)
# marginaleffects::avg_comparisons(h1b_mod_lmer)
```

This uses a mixed effects model:

`tct ~ treatment * seq + (1 + seq | participant)`

where Task completion time is a function of:

- fixed effects (predictors):
  - treatment (PWI, PAR, AR, MR)
  - seq (task number)
  - treatment * seq (interaction)
- random effects (account for participant variability):
  - (1 + seq | participant)

The treatment * seq interaction indicates how the relationship between seq and tct changes for different levels of treatment.

The random effects indicate that, in terms of TCT, each participant will have a unique starting point (intercept) and learning rate (slope).

In the results we see that the total variance contributed by random effects is 1977.0 + 79.3 + 228.5 = `r 1977.0 + 79.3 + 228.5`, with participant idiosyncrasies accounting for `r round(1977 / (1977.0 + 79.3 + 228.5) * 100, 1)`% of that. 

These results are based on data that has not been standardized. Results of standard post-hoc model evaluation showed issues with homoscedasticity and multicolinearity, etc. They are kept here for comparison with the output using standardized data.

Here is an alternate model, fit on standardized data:

```{r h1b-lmer-model-2}

# record original means and SDs
mean_tct <- mean(learn_h1b$tct, na.rm = TRUE)
sd_tct <- sd(learn_h1b$tct, na.rm = TRUE)

# standardize data
learn_h1b_std <- learn_h1b |>
  mutate(
    tct = scale(tct) |> as.vector(),
    uce = scale(uce) |> as.vector()
  )

h1b_mod_lmer_std <- lmerTest::lmer(
  tct ~ treatment * seq + (1 + seq | participant),
  data = learn_h1b_std,
  #    control = lmerControl(optimizer = "bobyqa")
)

summary(h1b_mod_lmer_std)
```

These results use a dataset where TCT was standardized in the following manner:

$$z_x = \frac{x - \mu_x}{\sigma_x}$$

Several tests are run to confirm the model quality:

```{r}
## model tests
# normality of residuals
mr <- resid(h1b_mod_lmer_std)
qqnorm(mr)
qqline(mr)
shapiro.test(mr)
# homoscedaasticity - constant variance of residuals
plot(h1b_mod_lmer_std)
# multicolinearity among predictors
car::vif(h1b_mod_lmer_std)
# R2 goodness of fit
performance::r2_nakagawa(h1b_mod_lmer_std)
```

The test results confirm model validity. All but MR-related coefficients are significant:

```text
                 Estimate Std. Error       df t value Pr(>|t|)    
(Intercept)       2.08412    0.31182 48.34414   6.684 2.18e-08 ***
treatmentMR       0.01964    0.43352 48.42402   0.045 0.964044    
treatmentPAR     -1.66019    0.44385 45.55992  -3.740 0.000512 ***
treatmentPWI     -1.72445    0.42565 44.90851  -4.051 0.000199 ***
seq              -0.51620    0.07129 46.19347  -7.241 3.88e-09 ***
treatmentMR:seq  -0.02100    0.10001 46.70136  -0.210 0.834635    
treatmentPAR:seq  0.32280    0.09651 35.36790   3.345 0.001960 ** 
treatmentPWI:seq  0.27296    0.09225 34.26448   2.959 0.005565 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

All values are expressed in terms of standard deviations of TCT, and can be restored to the original units for easier interpretation. Only the intercept gets the mean TCT value added back, as the others are offsets from that.

$$x = z_x \times \sigma_x + \mu_x$$

Here are the results of reversing the standardization process:

```{r}
# Extract coefficients
coefficients_std <- fixef(h1b_mod_lmer_std)

# Unstandardize coefficients - return to original units for interpretation
coef_unstd <- coefficients_std * sd_tct
coef_unstd["(Intercept)"] <- coef_unstd["(Intercept)"] + mean_tct

print(coef_unstd)
```

These roughly match the values from the model using the unaltered data:

```text
Fixed effects:
                 Estimate Std. Error       df t value Pr(>|t|)    
(Intercept)      185.5695    13.2651  48.3432  13.989  < 2e-16 ***
treatmentMR        0.8357    18.4420  48.4231   0.045 0.964045    
treatmentPAR     -70.6252    18.8816  45.5591  -3.740 0.000512 ***
treatmentPWI     -73.3588    18.1075  44.9077  -4.051 0.000199 ***
seq              -21.9595     3.0327  46.1927  -7.241 3.88e-09 ***
treatmentMR:seq   -0.8931     4.2545  46.7006  -0.210 0.834637    
treatmentPAR:seq  13.7321     4.1056  35.3674   3.345 0.001960 ** 
treatmentPWI:seq  11.6117     3.9243  34.2640   2.959 0.005565 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

From this, we can gather that (all in seconds, rounding to 1 digit):

- starting values of TCT
  - the significant average for AR (aka Intercept) is 185.6
  - MR was insignificantly higher
  - PAR was significantly lower, 185.6 - 70.6 = 115.0
  - PWI was significantly lower, 185.6 - 73.4 = 112.2
- change with each task completed
  - AR's sigificant estimate of -22.0 indicates that, on average, each task was completed 22s faster
  - MR was insignificantly lower
  - PAR was significantly higher, -22.0 + 13.7 = -8.3
  - PWI was significantly higher, -22.0 + 11.6 = -10.4

Interpretation:

- PAR and PWI groups started faster than AR/MR, exhibiting significantly lower TCT
- The TCT decreased significantly faster per task with AR/MR than for PWI or PAR
- MR was not significantly different than AR in either case, suggesting they are equivalent


Continued evaluation by comparing with simpler models:

```{r}
# compare alternative models
# baseline, as formulated above
m1 <- lmerTest::lmer(tct ~ treatment * seq + (1 + seq | participant),
  data = learn_h1b,
  control = lmerControl(optimizer = "bobyqa")
)

# random intercept only
m2 <- lmerTest::lmer(tct ~ treatment * seq + (1 | participant),
  data = learn_h1b,
  control = lmerControl(optimizer = "bobyqa")
)

# fixed intercept, random slope
m3 <- lmerTest::lmer(tct ~ treatment * seq + (seq | participant),
  data = learn_h1b,
  control = lmerControl(optimizer = "bobyqa")
)

AIC(m1, m2, m3)
```

The results show that m1 and m3 are equal, and better than m2. Sticking with m1. This really only tells us that the starting TCT and rates varied by treatment, but we know that there is a relationship between those three. Is the difference coming from the treatment or that relationship?

Goals: try other approaches to model the non-linear nature of learning, account for different initial efficiencies, variable repetition counts, and the dual criteria of speed and accuracy.

First effort to improve understanding: start simple, add UCE as a covariate in the current formulation:

`tct ~ treatment * seq + uce + (1 + seq | participant)`

It does make sense that errors could extend the TCT...

```{r}
# first just try adding uce as a covariate
h1b_mod_lmer_std_uce <- lmerTest::lmer(
  tct ~ treatment * seq + uce + (1 + seq | participant),
  data = learn_h1b_std,
  control = lmerControl(optimizer = "bobyqa")
)

# summary(h1b_mod_lmer_std_uce)

# uce is not significant
broom::tidy(h1b_mod_lmer_std_uce) |>
  filter(term == "uce") |>
  select(p.value)

```

Nope. UCE is not significant. Uncorrected errors go unnoticed and do not impact time?

Also, it is a counted (discrete) value, so modeling it in this way is, in hindsight, not advisable. We'll come back to error analysis later. For now, focus on TCT.

So far we've used linear models but learning curves are naturally exponential in shape. Also, there is a natural tendency of learning rate to depend on the difference between starting point and where the curve naturally flattens out. Here AR and MR start out slower, which may be the primary reason that they improve faster than PWI / PAR.

Second effort: tried using nlme with SSasymp and SSasympOrig to deal with these issues. These are non-linear methods that account for asymptotic behavior. In mixed-effect form they are also naturally robust with repeated measures data (e.g. variable number of tasks completed).

After hours of trying to get something to converge, I gave up. Sample code snippets below:

```R
# never got this to work!!
library(nlme)

data <- learn_h1b_std |>
  mutate(treatment = factor(as.character(treatment)))

gd <- nlme::groupedData(tct ~ seq | participant, data = data)

# Extract simple model coefficients
simple_model <- nls(tct ~ SSasymp(seq, Asym, R0, lrc), data = data)
simple_start_vals <- coef(simple_model)

# Check the levels of the treatment
lc <- length(unique(data$treatment))

start_vals <- c(
  rep(simple_start_vals["Asym"], lc),
  rep(simple_start_vals["R0"], lc),
  rep(simple_start_vals["lrc"], lc)
)

names(start_vals) <- c(paste0("Asym.", unique(data$treatment)),
                       paste0("R0.", unique(data$treatment)),
                       paste0("lrc.", unique(data$treatment)))

# exponential model with SSasymp (self-starting asymptotic regression) function
exp1 <- nlme::nlme(tct ~ SSasymp(seq, Asym, R0, lrc),
                  fixed = Asym + R0 + lrc ~ treatment,
                  random = nlme::pdDiag(Asym ~ 1 | participant),
                  start = start_vals,
                  data = gd)

summary(exp1)

# using SSasympOrig if starting points (R0) differ significantly
exp2 <- nlme::nlme(tct ~ SSasympOrig(seq, Asym, lrc),
                   fixed = Asym + lrc ~ treatment,
                   random = Asym ~ 1 | participant,
                   data = gd,
                   control = nlme::nlmeControl(pnlsTol = 0.1))

summary(exp2)
```

Let's try a different approach, based on change in TCT over time. Essentially find differences in rate of change in TCT. Related to "first differences" in Time Series Analysis, slope comparison in Growth Curves.

```{r}
# first attempt

# Extract Starting Values from the model estimates
# hardcoded - ick!
start_values <- data.frame(
  treatment = c("AR", "MR", "PAR", "PWI"),
  start_tct = c(185.6, 185.6 + 0.8, 185.6 - 70.6, 185.6 - 73.4)
)

# Calculate First Differences including properly handling initial TCT
d <- learn_h1b |>
  left_join(start_values, by = "treatment") |>
  group_by(participant) |>
  mutate(
    tct_diff = ifelse(
      is.na(lag(tct)),
      tct - start_tct, # For t = 1, calculate difference from estimated TCT(t=0) directly
      tct - lag(tct) # For subsequent tasks, difference from the previous task
    )
  ) |>
  ungroup()

d
```

At this point, we can simply compare `tct_diff` between groups with non-parametric tests:

```{r}
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

ggstatsplot::ggbetweenstats(
  data = d,
  x = treatment,
  y = tct_diff,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "rate of change in tct by treatment",
  # results.subtitle = FALSE,
  # pairwise.display = "all",
  subtitle = "Significant Pairwise Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)
```

This shows significant differences between AR/MR and PWI/PAR. Unfortunately, this isn't very informative as it only confirms what we already know:

- high initial TCT is correlated with high TCT differences
- AR and MR have higher initial TCT
- there is a natural limit to TCT that prevents further improvement
- together these create a kind of ceiling and floor effect that limit the value of mean / median statistics for comparison purposes

Instead, we should try to fit a mixed effects model to the differences, for example:

`tct_diff ~ treatment * seq + (1 + seq | participant)`

This would allow us to understand the participant-independent (fixed) effects of treatment and sequence on the changes in task completion time.

One such attempt:

```{r}
# Fit the mixed-effects model
m <- lmerTest::lmer(
  tct_diff ~ treatment * seq + (1 + seq | participant),
  data = d,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m)

```

No significant effects other than intercept (the average initial tct_diff for AR).
Singularity fit due to a lack of variance to model per participant, as shown in Random effects table. Large residuals indicate lots of unexplained variance.

Tried lots of things to address these issues. Here's one attempt to improve the data and simplify the model:

```{r}
# use unordered factor to improve interpretability of results
d$treatment <- factor(as.character(learn_h1b$treatment))

# record original means and SDs
mean_diff <- mean(d$tct_diff, na.rm = TRUE)
sd_diff <- sd(d$tct_diff, na.rm = TRUE)

# standardize data
d2 <- d |>
  mutate(
    tct_diff = scale(tct_diff) |> as.vector()
  )

glimpse(d2)

# Fit the mixed-effects model with interaction effects
# use the overloaded version of lmer from lmerTest to get CI, p-vals
m <- lmerTest::lmer(
  tct_diff ~ treatment + seq + (1 | participant),
  data = d2,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m)
```

Singularity problems persist. Marginal to low significance in PWI, PAR differences. Lack of participant variance is frustrating as it implies that all 50+ participants experienced the same learning progression, regardless of treatment, personal idiosyncracies, etc. Doesn't make sense.

Potential sources of this discrepancy are issues with data transformation or model specification, or  insufficient participant variability relative to within-participant variability (intraclass correlation, ICC). *The latter seems unlikely?*

To attempt to address these issues, further changes were made to:

- eliminate the use of estimated TCT(t=0), which might be concealing some participant variation, and base changes only on lag
- incorporate ICC checks and other model diagnostics including plots (e.g. change in TCT, as shown below)
- varying the model formulation, e.g. adding random slopes `(1 + seq | participant)`, treating SEQ as a random effect only `tct_diff ~ treatment + (1 + seq | participant)`, fixed-effects only
- centering the SEQ data in an attempt to address correlation between the intercept and slope of TCT changes (a waste of time since they are naturally related in learning rates)
- converting change in TCT from time data to proportions of total improvement


```{r}
#| warning: false

ggplot(d, aes(x = seq, y = tct_diff, group = participant, color = treatment)) +
  geom_line(alpha = 0.2) +
  geom_smooth(aes(group = treatment),
    method = "loess",
    se = FALSE,
    size = 1
  ) +
  theme_minimal() +
  labs(
    title = "Participant Learning Progressions",
    x = "Task Sequence",
    y = "TCT Differences"
  )
```

Here is one such attempt, using simplified implementation of diffs:

```{r}
ds <- learn_h1b |>
  left_join(start_values, by = "treatment") |>
  group_by(participant) |>
  mutate(tct_diff = tct - lag(tct)) |>
  ungroup() |>
  filter(!is.na(tct_diff))

# use unordered factor to improve interpretability of results
ds$treatment <- factor(as.character(ds$treatment))

ms <- lmerTest::lmer(
  tct_diff ~ treatment + seq + (1 + seq | participant),
  data = ds,
  control = lmerControl(optimizer = "bobyqa")
)

summary(ms)

# Calculate ICC
as.numeric(VarCorr(ms)$participant[1]) /
  (as.numeric(VarCorr(ms)$participant[1]) +
    as.numeric(attr(VarCorr(ms), "sc")^2))
```

This shows an improvement in the participant variance, but no significant treatment effects. RE stats indicate that intercept and slope are 100% inversely correlated. ICC of 0.81 suggests substantial participant-level variability.

*Is this the "best" model of all?*

Here's another attempt based on normalized (proportions) `tct_diff` values. In spite of finding marginally significant (p=0.04) overall differences, the Holm-adjusted Dunn comparisons show no significant treatment effects.

```{r}
#| warning: false

# add start times and diffs
d <- learn_h1b |>
  left_join(start_values, by = "treatment") |>
  group_by(participant) |>
  mutate(tct_diff = tct - ifelse(is.na(lag(tct)), start_tct, lag(tct))) |>
  ungroup() |>
  filter(!is.na(tct_diff))

# use progress (proportion) instead of difference
# include Start TCT in Total Improvement Calculation
d <- d |>
  group_by(participant) |>
  mutate(
    max_tct = max(c(tct, start_tct)),
    min_tct = min(tct),
    tot_imp = max_tct - min_tct
  ) |>
  ungroup()

# normalize tct_diff by tot_imp
d <- d |>
  mutate(tct_diff_norm = tct_diff / tot_imp) |> # Normalize the differences
  filter(!is.na(tct_diff_norm)) # Ensure there are no NA values

# View the transformed data
glimpse(d)
```

```{r}
#| fig-height: 6

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

ggstatsplot::ggbetweenstats(
  data = d,
  x = treatment,
  y = tct_diff_norm,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "rate of change in tct by treatment",
  # results.subtitle = FALSE,
  pairwise.display = "all",
  subtitle = "Significant Pairwise Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

```

At this point it seems clear that, if there is a treatment effect at all, it is small. This is not an effort to 'p-hack', I just want to make sure all efforts are made to get a valid finding. We've collected strong statistical evidence that there is little to no treatment effect, but our confidence in that result is not high based on prior expectations (which should not unduly influence interpretation). Some statistical results (e.g. the initial mixed-effect model on raw TCT, its changes) seemed to support the idea of treatment effect, but were influenced by other effects / covariance.

One last shot, fitting fixed effect and mixed effects models on the proportional data, standardized:

```{r}
# use unordered factor to improve interpretability of results
d$treatment <- factor(as.character(d$treatment))

# record original means and SDs
mean_diff <- mean(d$tct_diff, na.rm = TRUE)
sd_diff <- sd(d$tct_diff, na.rm = TRUE)

# standardize data
d <- d |>
  mutate(
    tct_diff = scale(tct_diff) |> as.vector()
  )

glimpse(d)

## try simple baseline model
# Fit a fixed-effects linear model (OLS)
fem <- lm(tct_diff_norm ~ treatment + seq, data = d)

# Summarize the model
summary(fem)

# Fit the mixed-effects model with interaction effects
# use the overloaded version of lmer from lmerTest to get CI, p-vals
h1b_mod_diffs <- lmerTest::lmer(
  tct_diff_norm ~ treatment + seq + (1 + seq | participant),
  data = d,
  control = lmerControl(optimizer = "bobyqa")
)

summary(h1b_mod_diffs)
```

Poor fit, singularity, no significant treatment effects. Pffft.

Confluence of evidence that, either (1) I don't know what I'm doing, or (2) there is no treatment effect.

Things to consider:
1. We have focused almost exclusively on improvements to TCT as the sole indicator of increasing proficiency. How can error counts be considered? Bivariate analysis or survival models?
2. Should we attempt a piecewise linear model of the data to see if treatment effect can be spotted?

#### Final Attempt

Current thinking is in `25-results.qmd`. I was trying to improve the model to address the upswing / extrapolation, without much success. So I moved those attempts here to maybe come back to later...

curve based on quadratic component `seq.fp ^ 2`
need more logarithmic characteristics

```{r}
sl <- sequence(11) / 11

pd <- tibble(seq.fp = sl) |>
  mutate(
    ls = 0.25 * log(seq.fp),
    t = atan(sl),
    es = exp(seq.fp),
    ss = seq.fp^2
  )

pd |>
  ggplot() +
  geom_line(aes(x = seq.fp, y = t, color = "blue"))
# geom_line(aes(x = seq.fp, y = es, color = "red")) +
# geom_line(aes(x = seq.fp, y = ss, color = "green"))

```

```{r}
# Custom transformation function using arctangent
# custom_arctan_transform <- function(x) {
#   scaled_x <- (x - 0.25) * 6 # Adjust the scaling factor to control the steepness and range
#   transformed <- sin(scaled_x)
#   return(transformed)
# }

lt <- function(x, L = 1, k = 10, x0 = 0.5, b = 0.5) {
  b - L / (1 + exp(-k * (x - x0)))
}

# Test the function with a sequence of values
x_values <- seq(0, 1, length.out = 100)
transformed_values <- lt(x_values, L = 0.5, k = 10, x0 = 0)

df <- data.frame(x = x_values, y = transformed_values)
ggplot(df, aes(x = x, y = y)) +
  geom_line() +
  labs(
    title = "transformation tester",
    x = "x",
    y = "Transformed x"
  ) +
  theme_minimal()

```

stop automatic builds here - the next cell crashes

```{r eval=FALSE}
# conditional models based on m4 (quadratic random slope and random intercept)
# introduce treatment and interactions as fixed effects
cm2 <- lmer(
  tct ~
    1 + seq.fp * treatment + I(lt(seq.fp)) * treatment +
    (1 + seq.fp + I(lt(seq.fp)) | participant),
  data = md, REML = F
)

cm3 <- lmer(
  tct ~
    1 + seq.fp * treatment +
    (1 + seq.fp | participant),
  data = md, REML = F
)

cm <- cm3

anova(cm1, cm2, cm3)
compare_performance(cm1, cm2, cm3)
anova(cm)
summary(cm)

# Predicted values for different seq.fp levels
seq_levels <- unique(md$seq.fp)
predicted_vals <- expand.grid(treatment = levels(md$treatment), seq.fp = seq_levels)
predicted_vals$pred_tct <- predict(cm, newdata = predicted_vals, re.form = NA)

# Plot with both predicted lines and actual data points
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  geom_point(data = md, aes(x = seq.fp, y = tct, color = treatment, alpha = 0.4)) +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  facet_wrap(~treatment) +
  theme_minimal()
```




most treatments only average 3-4 iterations

```{r eval=FALSE}
# Step 1: Calculate the maximum seq.fp for each participant within each treatment group
max_seq_per_participant <- md |>
  group_by(treatment, participant) |>
  summarize(max_seq_fp = max(seq.fp), .groups = "drop")

# Step 2: Calculate the mean of these maximum values for each treatment group
avg_max_seq_per_treatment <- max_seq_per_participant |>
  group_by(treatment) |>
  summarize(avg_max_seq = mean(max_seq_fp), .groups = "drop")

# Output the result
print(avg_max_seq_per_treatment)

```



tried using a model fitted using all the data, but evaluating its performance only on the valid part (e.g. thru seq.fp = 0.4) - didn't find a way to use anova / summary for assessment

tried using splines (didn't fit) and polynomial forms. 2nd order polynomial is equivalent to explicitly stated terms (cm1). 3rd order poly is too complex. both poly formulations resulted in singularities. continue with cm1

should I train on a subset of the data? iterations 1-3, or 1-4?
try a range and compare the results?

```{r eval=FALSE}
# # Create subsets with different maximum iteration sequences
# subsets <- list(
#   md1 = md |> filter(seq <= 3),
#   md2 = md |> filter(seq <= 4),
#   md3 = md |> filter(seq <= 5),
#   md4 = md |> filter(seq <= 6),
#   md5 = md |> filter(seq <= 7)
# )

# Determine the number of iterations each participant has completed
participant_iterations <- md |>
  group_by(participant) |>
  summarize(max_iterations = max(seq), .groups = "drop")

# Function to filter data sets by minimum number of iterations
subset_by_iterations <- function(data, min_iterations) {
  eligible_participants <- participant_iterations |>
    filter(max_iterations >= min_iterations) |>
    pull(participant)
  return(data |> filter(participant %in% eligible_participants & seq <= min_iterations))
}

# Create subsets for different iteration thresholds
subsets <- list(
  md1 = subset_by_iterations(md, 3),
  md2 = subset_by_iterations(md, 4),
  md3 = subset_by_iterations(md, 5),
  md4 = subset_by_iterations(md, 6),
  md5 = subset_by_iterations(md, 7),
  md6 = subset_by_iterations(md, 8),
  md7 = subset_by_iterations(md, 9),
  md8 = subset_by_iterations(md, 10),
  md9 = subset_by_iterations(md, 11)
)

# Function to summarize the number of participants in each group for a given subset
summarize_participants <- function(subset_data) {
  subset_data |>
    group_by(treatment) |>
    summarize(num_participants = n_distinct(participant), .groups = "drop")
}

# Create a list to store the summary tables
summary_tables <- list()
for (i in 1:9) {
  subset_data <- subsets[[paste0("md", i)]]
  summary_tables[[paste0("md", i)]] <- summarize_participants(subset_data) |>
    mutate(subset = paste0("md", i))
}

```

only md1 thru 3 are viable - have all treatments represented
md1 is relatively balanced, the groups in 2 and 3 are dominated by PWI and PAR
md1 won't fit - num of observations less than number of random effects

```{r eval=FALSE}
models <- list()
for (i in 2:3) {
  subset_data <- subsets[[paste0("md", i)]]
  print(paste0("fitting model: ", i))
  tryCatch(
    {
      models[[paste0("rcm", i)]] <- lmer(
        tct ~
          1 + seq.fp * treatment + I(seq.fp^2) * treatment +
          (1 + seq.fp + I(seq.fp^2) | participant),
        data = subset_data, REML = F
      )
    },
    error = function(e) {
      print(paste0("Error fitting model ", i, ": ", e$message))
      models[[paste0("rcm", i)]] <- NULL
    }
  )
}

library(broom.mixed)

model_summaries <- lapply(models, glance)
model_summaries_df <- do.call(rbind, model_summaries)
model_summaries_df$dataset <- c("md2", "md3")
print(model_summaries_df)
```


```{r eval=FALSE}
m <- models$rcm3
anova(m)
summary(m)
r2(m)

em_int <- emmeans::emmeans(m, ~treatment)
em_slope <- emmeans::emmeans(m, ~ treatment | seq.fp)
em_quad <- emmeans::emmeans(m, ~ treatment | I(seq.fp^2))

pw_int <- pairs(em_int, adjust = "holm")
pw_slope <- pairs(em_slope, adjust = "holm")
pw_quad <- pairs(em_quad, adjust = "holm")

summary(pw_int)
summary(pw_slope)
summary(pw_quad)
```

```{r eval=FALSE}
md3 <- subsets$md3

# Predicted values for different seq.fp levels
seq_levels <- unique(md3$seq.fp)
predicted_vals <- expand.grid(treatment = levels(md3$treatment), seq.fp = seq_levels)
predicted_vals$pred_tct <- predict(m, newdata = predicted_vals, re.form = NA)

# Plot
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  theme_minimal()

```

#### Exponential Model

Initially tried to fit an exponential decay model to estimate learning rate, $\delta$. The equation for this model is:

$$y(t) = \alpha + (\beta)e^{\delta t}$$

where:

- $\alpha$ is the asymptote
- $\beta$ is the total change between starting point and asymptote
- $\delta$ is the rate of change
- $t$ is time

```{r eval=FALSE}
library(nlme)
set.seed(42)

em <- nlme(tct ~ a + b * exp(d * seq.fp),
  data = md,
  fixed = a + b + d ~ 1,
  random = list(a ~ 1, d ~ 1),
  groups = ~participant,
  start = c(60, -60, -1),
  na.action = na.omit
)
```

Doesn't converge.

#### Linear Splines

Try to use splines to control overfitting.

```{r eval=FALSE}
md$spline02 <- ifelse(
  md$seq.fp > 0.2,
  md$seq.fp - 0.2,
  0
)

plot(y = md$spline02, x = md$seq.fp)

cms <- lmer(
  tct ~
    1 + seq.fp * treatment + spline02 * treatment +
    (1 + seq.fp | participant),
  data = md, REML = F
)

# cm1 <- lmer(
#     tct ~
#       1 + seq.fp * treatment + I(seq.fp ^ 2) * treatment +
#       (1 + seq.fp + I(seq.fp ^ 2) | participant),
#     data = md, REML = F
# )

anova(cms)
summary(cms)
r2(cms)

# Predicted values for different seq.fp levels
seq_levels <- unique(md$seq.fp)
predicted_vals <- expand.grid(treatment = levels(md$treatment), seq.fp = seq_levels, spline02 = md$spline02)
predicted_vals$pred_tct <- predict(cms, newdata = predicted_vals, re.form = NA)

# Plot
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  theme_minimal()
```

