try nonlinear methods that are theoretically aligned with learning

Each model captures different types of decay or growth behavior based on a starting or ending point, expected magnitude of change, and a rate parameter.

```{r}

# exponential decay with asmptote
m8e <- nlme::nlme(tct ~ a + b * exp(-c * seq.fp), 
                    fixed = a + b + c ~ 1,
                    random = a + b ~ 1 | participant,
                    data = mdf, start = c(a = 50, b = 100, c = 2))

# hyperbolic function
m9h <- nlme::nlme(tct ~ a + b / (1 + c * seq.fp), 
                    fixed = a + b + c ~ 1,
                    random = a + b ~ 1 | participant,
                    data = mdf, start = c(a = 50, b = 100, c = 2))

# Gompertz function
m10g <- nlme::nlme(tct ~ a * exp(-b * exp(-c * seq.fp)), 
                     fixed = a + b + c ~ 1,
                     random = a + b ~ 1 | participant,
                     data = mdf, start = c(a = 150, b = 2, c = 5))

anova(m8e, m9h, m10g)
```

all are more parsimonious, and aligned with learning theory, m8e performs best

visually compare model predictions

```{r}
# Create a data frame for predictions
new_data <- expand.grid(seq.fp = seq(0, 1, length.out = 100),
                        participant = unique(mdf$participant)[1])

# Add predictions from each model
new_data$pred_exp <- predict(m8e, newdata = new_data, level = 0)
new_data$pred_hyp <- predict(m9h, newdata = new_data, level = 0)
new_data$pred_gomp <- predict(m10g, newdata = new_data, level = 0)

# Reshape the data for plotting
plot_data <- new_data %>%
  tidyr::pivot_longer(cols = starts_with("pred_"),
                      names_to = "model",
                      values_to = "prediction") %>%
  mutate(model = factor(model, 
                        levels = c("pred_exp", "pred_hyp", "pred_gomp"),
                        labels = c("Exponential", "Hyperbolic", "Gompertz")))

# Create the plot
ggplot(plot_data, aes(x = seq.fp, y = prediction, color = model)) +
  geom_line(size = 1) +
  geom_point(data = mdf, aes(x = seq.fp, y = tct), color = "black", size = 1, alpha = 0.3) +
  scale_color_manual(values = c("Exponential" = "#1b9e77", 
                                "Hyperbolic" = "#d95f02", 
                                "Gompertz" = "#7570b3")) +
  labs(x = "Sequence Progress (seq.fp)",
       y = "Predicted Task Completion Time (TCT)",
       color = "Model",
       title = "Predicted Learning Curves by Model") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

here the Gompertz model is preferable - its asymptote better reflects expectations (takt time) and fits the existing data better

The Gompertz function models growth or decay with a sigmoid shape, commonly used for population growth and similar phenomena:

( a ): The upper asymptote where ( tct ) starts.
( b ): A scale parameter affecting the displacement of the curve.
( c ): Growth rate parameter affecting how quickly the variable changes.

the sigmoidal nature of the Gompertz function provides it with flexibility to fit the early rapid learning followed by by a deceleration and a plateau more accurately. due to double-exponential term

compare model diagnostics

```{r}
# Function for model diagnostics of nlme models
model_diagnostics_nlme <- function(model) {
  # Extract normalized residuals
  residuals_model <- residuals(model, type = "normalized")

  # 1. Shapiro-Wilk test for normality
  shapiro_test <- shapiro.test(residuals_model)
  shapiro_result <- tibble(
    name = 'shapiro',
    test_stat = shapiro_test$statistic,
    p_value = shapiro_test$p.value
  )

  # 2. Breusch-Pagan test for heteroscedasticity using fitted values
  fit_values <- fitted(model)
  bp_test <- lm(I(residuals_model^2) ~ fit_values)
  bp_statistic <- summary(bp_test)$r.squared * length(residuals_model)
  bp_p_value <- 1 - pchisq(bp_statistic, df = 1)
  hetero_result <- tibble(
    name = 'heterosced',
    test_stat = bp_statistic,
    p_value = bp_p_value
  )

  # 3. Autocorrelation function for residuals
  acf_test <- acf(residuals_model, plot = FALSE)
  acf_result <- tibble(
    name = 'acf1',
    test_stat = acf_test$acf[2],  # Lag 1 autocorrelation
    p_value = NA  # Not commonly provided for acf
  )

  # 4. Outliers Detection using standardized residuals
  std_residuals <- residuals_model / sd(residuals_model)
  outliers <- sum(abs(std_residuals) > 2.5)  # Count of residuals > 2.5 SDs
  out_result <- tibble(
    name = 'out',
    test_stat = outliers,
    p_value = NA  # No p-value for simple outlier count
  )

  # Combine all results into one tibble
  results <- bind_rows(shapiro_result, hetero_result, acf_result, out_result)

  return(results)
}

model_diagnostics_nlme(m8e)
model_diagnostics_nlme(m9h)
model_diagnostics_nlme(m10g)

```

m10g is normal with lower acf1 and fewer outliers but exhibits heteroscedasticity

```{r}
plot(m10g)
```

validates heteroscedasticity and outliers

check fit for participants in all treatments

```{r h1b-insp-cond}
# Select the first 5 participants from each treatment
m10g_sample <- mdf |>
  mutate(pred = fitted(m10g)) |> 
  group_by(treatment) |>
  arrange(participant) |>
  filter(participant %in% unique(participant)[1:5]) |>
  ungroup() |> 
  mutate(facet_label = paste(treatment, participant, sep = " - "))

# Set a consistent maximum for y-axis
max_tct <- max(m10g_sample$tct) + 20
```


```{r fig-h1b-rlm-pred}
#| fig-cap: "Visualizing RLM model fit"

ggplot(m10g_sample, aes(x = seq.fp, y = tct)) + 
  geom_point(fill = "grey", pch = 21, size = 2, stroke = 1.25) + 
  geom_line(aes(group = participant)) + 
  geom_line(aes(y = pred, color = treatment), lwd = 1.5) +
  facet_wrap(~ facet_label, ncol = 5) +
  scale_y_continuous(limits = c(0, max_tct)) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "Model Fit by Participant and Treatment",
    subtitle = "RLM predictions for the first five participants in each treatment group"
  )
```

This visualization allows for a comparison between observed data and model predictions across different treatments and participants, and helps confirm that a good fit is obtained for a wide range of observation counts. The variation in TCT is evident here, with participants assigned to the PWI and PAR groups completing substantially more task iterations than their AR / MR peers.

```{r}
m10g_sum <- summary(m10g)
m10g_sum$tTable
```

Random Effects Variability:

Parameter (b): Shows significant variability among participants (StdDev = 0.3544453), indicating that the scale of decay differs across individuals.
Parameter (a): Very small standard deviation (1.682909e-04), suggesting that the upper asymptote (baseline performance) is relatively consistent among participants.
Correlation: There is a moderate positive correlation (0.648) between the random effects of (a) and (b), indicating that participants with higher asymptotes also tend to have higher scales of decay.

Fixed Effects Significance:

All Fixed Effects (a, b, c): The fixed effects for (a), (b), and (c) are highly significant (p-values very close to 0), suggesting that they are important in explaining the variance in task completion time (TCT).
Intercept (a): 48.03332 suggests a relatively high starting point for TCT.
Scale (b): -1.03164 indicates a decaying factor influence.
Rate (c): 2.06771 reflects a rapid rate of decay initially, contributing to the overall learning curve.

Residual Analysis:

Residual Distribution: The distribution of standardized within-group residuals shows no extreme outliers and appears reasonably symmetric (-2.77 to 3.13), indicating a good fit and no major systematic deviations in predictions.

all values are in the original scale of TCT (seconds), measured on the scale of seq.fp, which is the [0,1] transformation of seq, which ranged from 1 to 11

These results can be expressed as:

$$
\small{
\text{TCT}_{ij} = 48.03 \cdot \exp\left(1.03 \cdot \exp\left(-2.07 \cdot \text{SC}_{ij}\right)\right) + \text{RE}_{ij} + \varepsilon_{ij}
}
$$ {#eq-m10q-form}

$TCT_{ij}$ is the TCT in seconds for participant $j$ on task $i$. SC is the continuous transform of the discrete task number, $\text{SC} = (\text{SEQ} - 1) / (\max(\text{SEQ}) - 1)$. In this equation, $\text{RE}_{ij}$ represents the random effects of the m10g model, which take the form:

$$
\small{
\text{RE}_{ij} = U_{aj} + U_{bj} \cdot \text{SC}_{ij}
}
$$
where $U_{aj}$ and $U_{bj}$ are the random effects associated with parameters $a$ and $b$, respectively, for participant $j$. $U_{aj}$ represents the random intercept and $U_{bj}$ the random slope for each participant. These random effects are assumed to follow a normal distribution, $U_{kj} \sim \mathcal{N}(0, \sigma_{k}^2)$ for $k \in a, b$, and $\sigma_{k}$ are the coefficients for each component of the random effect.

#### Conditional Models and Hypothesis Testing

```{r}
m10gc1 <- nlme::nlme(tct ~ a * exp(-b * exp(-c * seq.fp)), 
                   fixed = list(a ~ treatment, b ~ 1, c ~ 1),
                   random = a + b ~ 1 | participant,
                   data = mdf, 
                   start = c(a = rep(150, 4), b = 2, c = 5),
                   control = list(
                     opt = "optim",
                     optMethod = "BFGS",
                     nlme::nlmeControl(maxIter = 1000, msMaxIter = 1000)
                    )
                  )

m10gc2 <- nlme::nlme(tct ~ a * exp(-b * exp(-c * seq.fp)), 
                   fixed = list(a ~ treatment, b ~ treatment, c ~ 1),
                   random = a + b ~ 1 | participant,
                   data = mdf, 
                   start = c(a = rep(150, 4), b = rep(2, 4), c = 5))

m10gc3 <- nlme::nlme(tct ~ a * exp(-b * exp(-c * seq.fp)), 
                   fixed = list(a ~ treatment, b ~ treatment, c ~ treatment),
                   random = a + b ~ 1 | participant,
                   data = mdf, 
                   start = c(a = rep(150, 4), b = rep(2, 4), c = rep(5, 4)))

```

```{r}
m10g_brms <- brm(
  bf(tct ~ a * exp(-b * exp(-c * seq.fp)),
     a ~ 1 + (1|participant),
     b ~ 1 + (1|participant),
     c ~ 1,
     nl = TRUE),
  data = mdf,
  family = gaussian(),
  prior = c(
    prior(normal(150, 50), nlpar = "a"),
    prior(normal(2, 1), nlpar = "b"),
    prior(normal(5, 2), nlpar = "c"),
    prior(cauchy(0, 20), class = "sd", nlpar = "a"),
    prior(cauchy(0, 20), class = "sd", nlpar = "b"),
    prior(cauchy(0, 20), class = "sigma")
  ),
  control = list(adapt_delta = 0.95),
  cores = 4,
  chains = 4,
  iter = 4000,
  warmup = 2000
)

```


```{r}
library(brms)
m_brms <- brm(bf(tct ~ a * exp(-b * exp(-c * seq.fp)),
                 a ~ treatment + (1|participant),
                 b ~ 1,
                 c ~ 1,
                 nl = TRUE),
              data = mdf,
              family = gaussian())

```