
#### linear model validation and finalizing

validate model m7...

```{r}
residuals_m7 <- residuals(m7)

# Shapiro-Wilk test for normality
m7_st <- shapiro.test(residuals_m7)
m7_st_pv <- m7_st$p.value

# Custom function for heteroscedasticity test
hetero_test <- function(model) {
  res <- residuals(model)
  fit <- fitted(model)
  aux_model <- lm(I(res^2) ~ fit)
  bp_statistic <- summary(aux_model)$r.squared * length(res)
  p_value <- 1 - pchisq(bp_statistic, df = 1)
  return(list(statistic = bp_statistic, p.value = p_value))
}

# Apply the test
m7_het <- hetero_test(m7)
m7_het_pv <- m7_het$p.value

# DHARMa diagnostics
sim_m7 <- DHARMa::simulateResiduals(m7)

invisible(capture.output({
  m7_dharma <- DHARMa::testResiduals(sim_m7, plot = FALSE)
}))

# Extract individual test results
m7_ks <- m7_dharma$uniformity
m7_disp <- m7_dharma$dispersion
m7_out <- m7_dharma$outliers

m7_ks_pv <- m7_ks$p.value
m7_disp_pv <- m7_disp$p.value
m7_out_pv <- m7_out$p.value
```

Validation of m7 confirmed the normality of residuals (Shapiro-Wilk test, `r fmt_p_val(m7_st_pv)`), with significant heteroscedasticity (`r fmt_p_val(m7_het_pv)`). Traditional QQ plots of the raw residuals do not properly account for the hierarchical structure and non-independence of mixed effect models, leading to overly optimistic results. The `{DHARMa}` package was specifically designed to address this problem by comparing observed residuals to a large number of simulations from the fitted model. This technique provides a more robust and accurate representation of model accuracy. Diagnostics produced by `testResiduals` indicate deviation from the expected distribution of residuals (Kolmogorov-Smirnov test, `r fmt_p_val(m7_ks_pv)`) without significant dispersion issues (`r fmt_p_val(m7_disp_pv)`) or outliers (`r fmt_p_val(m7_out_pv)`). Visual inspection of residual plots confirmed these findings (@fig-m7-dharma).

```{r fig-m7-dharma}
#| fig-cap: Residual Plots

DHARMa::plotQQunif(sim_m7, testDispersion = F, testUniformity = F, testOutliers = F)
```

Given these minor deviations from model assumptions, the model was refit using a robust linear mixed model designed to mitigate those issues and ensure reliable inference. Robust LMs (RLMs) use techniques such as M-estimation or MM-estimation to reduce the influence of outliers and high-leverage observations. They also employ weighted estimation to account for unequal variances (heteroscedasticity). This approach makes RLMs less sensitive to violations of normality assumptions in both fixed and random effects, and more resistant to departures from other common assumptions of linear mixed models. As a result, RLMs typically provide more stable and accurate parameter estimates when dealing with non-ideal data conditions. The result from `robustlmm::rlmer` is compared with the original fit in @tbl-h1b-m7-comp.

```{r fit-m7r}
#| warning: false
#| message: false

# robust fit of m7
m7r <- robustlmm::rlmer(
    tct ~ 1 + seq.fp + log(seq.fp + 1) + (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = F
)

cp_h1b_uc_r <- compare_performance(m7, m7r, metrics = "common", rank = T)

# validate result
m7r_het <- hetero_test(m7r)
m7r_het_pv <- m7r_het$p.value

residuals_m7r <- residuals(m7r)

# Shapiro-Wilk test for normality
m7r_st <- shapiro.test(residuals_m7r)
m7r_st_pv <- m7r_st$p.value

```

```{r tbl-h1b-m7-comp}
#| tbl-cap: "Model Performance Results"

cp_h1b_uc_r |> 
  select(!Model) |> 
  display()
```

These results, particularly the decrease in RMSE (`r round(cp_h1b_uc_r$RMSE[2], 2)`), indicate marginal improvement gains were achieved by the robust fit. However, the real benefit comes from improved model diagnostics. The robust model eliminates heteroscedasticity (`r fmt_p_val(m7r_het_pv)`) and down-weights extreme values as seen in @fig-m7r-resid. While the Shapiro-Wilk test still indicates non-normality of residuals (`r fmt_p_val(m7r_st_pv)`), this is less concerning for robust models, which are designed to perform well even when normality assumptions are violated.

```{r fig-m7r-resid}
#| fig-cap: Extreme values are assigned lower weights in robust model
plot(m7r, which = 1)
```

As a final model validation @fig-h1b-rlm-pred shows the predictions generated by the robust model for the first five participants in each treatment. In this figure, the black circles and lines represent observed TCT values for each participant, while the colored curves represent the RLM's predictions. Each row corresponds to a different treatment group (AR, MR, PAR, PWI), and each subplot shows data for an individual participant.

```{r h1b-insp-cond}
# Select the first 5 participants from each treatment
m7r_sample <- mdf |>
  mutate(pred = fitted(m7r)) |> 
  group_by(treatment) |>
  arrange(participant) |>
  filter(participant %in% unique(participant)[1:5]) |>
  ungroup() |> 
  mutate(facet_label = paste(treatment, participant, sep = " - "))

# Set a consistent maximum for y-axis
max_tct <- max(m7r_sample$tct) + 20
```


```{r fig-h1b-rlm-pred}
#| fig-cap: "Visualizing RLM model fit"

ggplot(m7r_sample, aes(x = seq.fp, y = tct)) + 
  geom_point(fill = "grey", pch = 21, size = 2, stroke = 1.25) + 
  geom_line(aes(group = participant)) + 
  geom_line(aes(y = pred, color = treatment), lwd = 1.5) +
  facet_wrap(~ facet_label, ncol = 5) +
  scale_y_continuous(limits = c(0, max_tct)) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "Model Fit by Participant and Treatment",
    subtitle = "RLM predictions for the first five participants in each treatment group"
  )
```

This visualization allows for a comparison between observed data and model predictions across different treatments and participants, and helps confirm that a good fit is obtained for a wide range of observation counts. The variation in TCT is evident here, with participants assigned to the PWI and PAR groups completing substantially more task iterations than their AR / MR peers.

```{r tbl-m7r-params}
#| tbl-cap: Fixed Effects Parameters for the Unconditioned RLM (m7r)
m7r_params <- parameters::model_parameters(m7r)

# Function to add significance stars
add_significance_stars <- function(p_value) {
  if (p_value < 0.001) {
    "***"
  } else if (p_value < 0.01) {
    "**"
  } else if (p_value < 0.05) {
    "*"
  } else if (p_value < 0.1) {
    "."
  } else {
    ""
  }
}


# Separate the count model coefficients (fixed effects)
m7r_params_fe <- m7r_params  |> 
  filter(Effects == "fixed")

m7r_params_fe |> 
  mutate(Sig = sapply(p, add_significance_stars)) |> 
  select(!Group) |> 
  display(digits = 3, ci_digits = 2)
```

A summary of parameters for the fixed effects in m7r is found in @tbl-m7r-params, revealing significant fixed effects for both linear ($\beta_1$ = 989.75, p < .001) and logarithmic ($\beta_2$ = -1345.31, p < .001) components of the learning curve. To ease interpretation, this can be expressed mathematically as:

$$
\small{
\text{TCT}_{ij} = 142.1 + 989.7 \times \text{SC}_{ij} - 1345.3 \times \log(\text{SC}_{ij} + 1) + \text{RE}_{ij} + \varepsilon_{ij}
}
$$ {#eq-m7r-form}

$TCT_{ij}$ is the TCT in seconds for participant $j$ on task $i$. SC is the continuous transform of the discrete task number, $\text{SC} = (\text{SEQ} - 1) / (\max(\text{SEQ}) - 1)$. This suggests rapid initial improvement ($\beta_{1}$), followed by a decelerating learning rate ($\beta_{2}$). In this equation, $\text{RE}_{ij}$ represents the random effects of the m7r model, which take the form:

$$
\small{
\text{RE}_{ij} = U_{0j} + U_{1j} \times \text{SC}_{ij} + U_{2j} \times \log(\text{SC}_{ij} + 1)
}
$$ {#eq-m7r-form-re}

where $U_{kj} \sim \mathcal{N}(0, \sigma_i^2)$, for $k = 0, 1, 2$ and $\sigma_{k}$ are the coefficients for each of the three random effect components. The $U_{kj}$ terms quantify the substantial between-participant variability evident in the model, particularly in learning rates ($\sigma_1$ = 1141.6 for linear, $\sigma_2$ = 1556.9 for logarithmic components, and $\sigma_0$ = 48.2 for the random intercept). A strong correlation is observed between the intercept and the linear (0.935) and log (-0.928) components of SC, suggesting that participants with higher initial task completion times (intercepts) tend to have steeper linear improvements (positive correlation with linear component) and a more rapid deceleration in their rate of improvement (negative correlation with log component). This pattern indicates that individuals starting with slower performance generally show more dramatic initial gains but also a faster tapering off of improvement over time.

The model's robustness is demonstrated by the down-weighting of 37 residuals and 6 random effects, addressing potential outliers or influential observations. Additionally, the relatively low residual error ($\varepsilon_{ij}$ = 9.25) aligns with the high $R^2_{cond}$ previously noted. Overall, this model captures the general learning trends while accounting for individual differences, providing a solid foundation for examining treatment effects in the subsequent conditioned model. 

#### Conditional Models and Hypothesis Testing

Add fixed treatment effect to best fit from above - does it improve the model? Use estimated slopes and intercepts to understand relationship between time, sequence, treatment, and TCT.

using treatment and interactions with `seq` and `seq ^ 2`

```{r h1b-fit-cond}
#| warning: true

# conditional models based on m7r

# add treatment as fixed effect
m7rc1 <- robustlmm::rlmer(
    tct ~ 
      1 + seq.fp + log(seq.fp + 1) + treatment + 
      (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = F
)

# include treatment interactions with linear term
m7rc2 <- robustlmm::rlmer(
    tct ~ 
      1 + seq.fp * treatment + log(seq.fp + 1) + treatment + 
      (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = F
)

# include treatment interactions with linear and log terms
m7rc3 <- robustlmm::rlmer(
    tct ~ 
      1 + seq.fp * treatment + log(seq.fp + 1) * treatment + 
      (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = F
)
```

observed issues during fitment with ord treatment:

- initial convergence issues for rc3 → singularities suggesting random effects are too complex
- likely due to high correlation of intercept, seq, and log(seq)
  - educational theory supports high correlation of seq with intercept - the farther you start from understanding the more rapid your initial improvement
  - seq and log(seq) are obviously correlated
- interpret random effects of rc3 very carefully

```{r}
compare_performance(m7r, m7rc1, m7rc2, m7rc3, metrics = "common", rank = T)
```

observations from performance::compare_performance:

- rc3 is 2nd best performing but does not outperform the unconditioned model; interpret with caution
- rc2, the most stable (no convergence issues), performs worse than the unconditioned model overall
- all conditioned models have substantially higher marginal R2 than m7r
  - this implies that fixed effects, and therefore treatment, play an important role
- all other statistics are relatively close
- m7r is still the top performer, suggesting that individual differences (REs) are the most important contributor to variance in TCT
- slightly lower ICC in conditioned models (e.g., 0.907 for m7rc3) indicates that some of the variance previously attributed to participant differences is now explained by the fixed effects (treatments)

plan:

- explore both 7r, 7rc3, and 7rc2 with consideration given to the stability / performance tradeoff
- focus on treatment effects as REs, esp for rc3, may be suspect
- look for consistent treatment findings between rc2 and rc3 and differences between 7rc2/3 and 7r
- mention rlmer strengths in dealing with singularity encountered (reset)
- include visualizations of treatment effect, e.g. comparing rc2 and rc3
- discuss limitations and future exploration - simplify REs to improve fit, try other modeling methods


```{r}
summary(m7r)
summary(m7rc3)
summary(m7rc2)
```


```{r}
library(emmeans)

# Create a sequence of seq.fp values
seq_values <- seq(0.05, 1.0, by = 0.05)

# Compute slopes (learning rates) for each treatment at these seq.fp values
emm_rc2_slopes <- emtrends(m7rc2, ~treatment, var = "seq.fp", 
                           at = list(seq.fp = seq_values))

# View the estimated slopes
summary(emm_rc2_slopes)

# Perform pairwise comparisons of slopes between treatments
slope_comparisons <- pairs(emm_rc2_slopes)

# View the slope comparisons
summary(slope_comparisons)

### visualize results
# Create a sequence of seq.fp values
seq_values <- seq(0.05, 1.0, by = 0.05)

# Compute slopes (learning rates) for each treatment at these seq.fp values
emm_rc2_slopes <- emtrends(m7rc2, ~treatment | seq.fp, var = "seq.fp", 
                           at = list(seq.fp = seq_values))

# Convert to data frame and add seq.fp column
slopes_df <- as.data.frame(emm_rc2_slopes) %>%
  mutate(seq.fp = rep(seq_values, each = 4))

# Create the plot
ggplot(slopes_df, aes(x = seq.fp, y = seq.fp.trend, color = treatment, fill = treatment)) +
  geom_line() +
  # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.2) +
  labs(x = "Sequence Progress", y = "Estimated Learning Rate",
       title = "Learning Rates by Treatment") +
  theme_minimal()
```

look at last claude session for latest interpretation and discussion
there are problems


OLD STUFF...



rough interpretation:

- all fixed effects are significant
- TCT changes with task iteration (seq.fp)
- TCT varies by group (treatment)
- Rate of change of TCT changes by task iteration (seq.fp)
- The effect of iteration (seq.fp) differs by group (treatment)
- The rate of change of TCT differs by group (treatment)

From Chet, after confirming above, elaborating:

- Task Iteration Effect $(\text{seq.fp})$: TCT significantly decreases as task iterations increase, indicating learning over time (p < 0.001).
- Treatment Effect: Baseline TCT varies significantly by treatment group, demonstrating that different treatments have distinct initial impacts on task completion time (p < 0.001).
- Non-Linear Task Iteration Effect (Quadratic Term $(I(\text{seq.fp}^2))): The rate of change in TCT is non-linear and significantly changes with task iterations (p < 0.001).
- Interaction Effects:
  - $\text{seq.fp} \times \text{treatment}$: The impact of task iterations on TCT significantly differs across treatment groups (p = 0.005), indicating that treatments influence how quickly participants improve over iterations.
  - $I(\text{seq.fp}^2) \times \text{treatment}$: The non-linear rate of change in TCT also varies significantly by treatment (p = 0.017), suggesting that the treatments affect the curvature of the learning trajectory.
  
These results highlight the importance of both linear and quadratic terms in modeling learning rates and emphasize the varying effects of different treatments on TCT dynamics over time. The significant interactions underscore how treatments modulate not just the initial performance but also the improvement trajectories.

```{r h1b-sum-cond, eval=FALSE}
summary(cm1)
```

rough interpretation:
- (Intercept), seq.fp, and I(seq.fp^2) are all significant and give the initial TCT value, initial slope, and rate of change of slope for the reference treatment, AR
- treatmentMR/PWI/PAR give the change in intercept from AR; only PWI and PAR are significant ("reliably different")
- seq.fp:treatmentMR/PWI/PAR give the treatment level difference from the initial slope; only interactions with PWI and PAR are significant
- treatmentMR/PWI/PAR:I(seq.fp^2) gives the treatment level difference from the rate of change of slope; only PWI and PAR are signficant

from this we can see that AR and MR are roughly equivalent in all regards, and both PWI and PAR are significantly different from them in all measures. it does not provide sufficient information to ascertain the differences between MR, PWI, and PAR

compared to AR, these differences are significant:
- PWI and PAR start faster (lower baseline intercept)
- PWI and PAR start learning slower (positive seq.fp:treatment values)
- PWI and PAR improve learning faster (negative treatment:I(seq.fp^2) values)

interpretation of above:
- the first suggests PWI and PAR are more intuitive than AR
- the second aligns with the intuition that learning rate is related to the difference in understanding, i.e. lower initial task times give lower learning potential to the asymptote (proficiency), so the rate is lower
- the third observation seems counterintuitive - despite slower initial learning, the PWI and PAR reduce task completion times more per iteration

This counterintuitive trend highlights the deep insights derived from advanced modeling and underscores the importance of considering both initial performance and subsequent learning trajectories when evaluating instructional methods.

The learning for PWI and PAR might follow a sigmoidal curve, where there's slow initial improvement that accelerates rapidly as participants become more familiar with the task. This can happen if the initial hurdles are high, but once overcome, participants experience more significant gains.

PWI and PAR participants might spend more time initially adapting to a new, potentially more complex instruction method. Once they adapt, they leverage this understanding more effectively, leading to faster improvements.

The instructional methods for PWI and PAR might be inherently more effective in the long run, even if they take longer to grasp initially. These methods might impart skills that lead to compounded improvements over time.

I'm reminded of OH's comment: 'learning takes effort' - perhaps the advanced instructional techniques are too helpful, to the point of diminishing the learning. Where PWI and PAR require greater participant engagement in the task, which requires greater effort initially but that investment pays off with better overall understanding.

Early challenges encourage deeper cognitive processing, which fosters a better overall understanding despite the initial slower progress. The concept of "desirable difficulties" in learning theory posits that learning tasks should be challenging enough to promote deeper cognitive engagement but not so difficult that they become discouraging. PWI and PAR might embody this principle by requiring more effort initially, which results in more substantial long-term gains.

Chet's Revised Interpretation:

Initial Performance:

PWI and PAR participants start with lower baseline TCT, suggesting these methods are more intuitive or easier to grasp initially.
Effort and Engagement:

Advanced instructional techniques like AR might not challenge participants enough initially, leading to faster initial gains but slower long-term improvement.
PWI and PAR require greater initial effort, aligning with the participant insight that "learning takes effort." This investment of effort requires deeper cognitive engagement, which ultimately pays off.
Learning Trajectories:

Initial Slow Learning: PWI and PAR participants initially learn slower because more cognitive effort and engagement are required.
Accelerated Improvement: Over time, the effort invested initially by PWI and PAR participants results in faster learning, leading to greater reductions in TCT per iteration.
Integration of Insight:

This model aligns with learning theories suggesting that greater initial effort and engagement foster deeper understanding and long-term gains in proficiency. The comment "learning takes effort" encapsulates the necessity of active engagement, which may explain why participants eventually achieve faster reductions in TCT despite slower initial learning.

Summary:

PWI and PAR methods promote greater initial effort and cognitive engagement, leading to deeper learning and accelerated improvement over time.
While these methods may initially show slower learning rates, the effort invested results in faster long-term TCT reductions, demonstrating the importance of desirable difficulties in instructional design.

Summarized Interpretation with Concept Integration:

Initial Performance:

Lower Baseline TCT: PWI and PAR start with lower initial TCT values, suggesting these methods are more naturally intuitive.
Effort and Engagement:

Desirable Difficulty: These methods incorporate a level of difficulty that requires greater cognitive effort initially, which is essential for deep learning.
Zone of Proximal Development: PWI and PAR engage participants within their ZPD, challenging them sufficiently to enhance skill acquisition over time.
Learning Trajectories:

Initial Slower Learning: Slower initial learning rates for PWI and PAR participants reflect the necessary investment in overcoming early challenges.
Accelerating Improvement: Over time, the learning rate accelerates as participants gain proficiency, aligning with the principles of Flow Theory where optimal challenge and skill balance lead to deeper engagement and faster learning improvements.
Conclusion and Insight:
The integration of educational theories with model outputs provides a comprehensive explanation for the observed learning patterns:

PWI and PAR methods require initial higher efforts (desirable difficulty) but lead to better long-term understanding and faster TCT reductions.
AR and MR methods may offer quicker, easier starts but do not sustain the accelerated improvement seen with PWI and PAR.
Closing Thoughts:

This nuanced interpretation underscores the importance of instructional design that balances initial challenge with long-term learning benefits, leveraging the principles of desirable difficulty, ZPD, and Flow Theory for optimal educational outcomes.

pairwise comparisons of intercept, seq, and seq^2 effect using emmeans

```{r h1b-pw-cond, eval=FALSE}
em_int <- emmeans::emmeans(cm1, ~ treatment)
em_slope <- emmeans::emmeans(cm1, ~ treatment | seq.fp)
em_quad <- emmeans::emmeans(cm1, ~ treatment | I(seq.fp ^ 2))

pw_int <- pairs(em_int, adjust = "holm")
pw_slope <- pairs(em_slope, adjust = "holm")
pw_quad <- pairs(em_quad, adjust = "holm")

summary(pw_int)
summary(pw_slope)
summary(pw_quad)
```

in each case, all contrasts are significant except for AR-MR and PWI-PAR

visualize predictions...

```{r h1b-plot-cond, eval=FALSE}
library(ggplot2)

# Predicted values for different seq.fp levels
seq_levels <- unique(mdf$seq.fp)
predicted_vals <- expand.grid(treatment = levels(mdf$treatment), seq.fp = seq_levels)
predicted_vals$pred_tct <- predict(cm1, newdata = predicted_vals, re.form = NA)

# Plot with both predicted lines and actual data points
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  geom_point(data = mdf, aes(x = seq.fp, y = tct, color = treatment, alpha = 0.4)) +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  facet_wrap(~ treatment) +
  theme_minimal()
```

model is overfitting / extrapolating beyond data
obviously TCT should not rapidly increase after 3-6 iterations of the task
BUT - model performance is only a measure of residuals (y error) of predicted vs real data, so the up trend only matters where there is existing data - min impact on AR, max on PWI

sufficient evidence for now to go with this model, understanding its limitations, but let's try other things - limits of model fit undermine model comparisons since we're comparing models of the data now, not the data itself

further attempts to address this with:

- explored inverse, cubic, logarithmic, logistic, and trigonometric terms as alternative to quadratic / spline in the random slopes; random intercepts LME
- attempted a non-linear mixed effects model based on exponential decay
- fit the model with valid subsets of data with balanced iterations, e.g. the first n observations for all participants that completed that many tasks

no success - worse results, convergence problems

but important to get good model fit as that is what we are comparing now, not the data itself - another layer of abstraction

continuing efforts with spline, bayesian methods moved to `h1b_mod_select.qmd`

::: content-hidden

#### Old Material

This formulation captures individual variability by allowing each participant to have a different initial TCT (intercept). It accounts for variations in TCT due to differences between participants (random effects) and within each participant's repeated measures (residuals). This design naturally incorporates all sources of participant-level variation, including imbalances due to the differing number of tasks completed by each participant, thereby addressing the data's inherent non-independence.

Variations of this model were fit using the residual maximum likelihood method (REML), as implemented by the `lmer` function from the `{lme4}` package. To address issues related to the non-linear nature of TCT, which follows an exponential decay pattern commonly observed in learning outcomes, models were fit to response variables transformed by log, square-root, and inverse (i.e., $1/\text{tct}$) functions. The results were then compared using the `compare_performance` function from the `{performance}` package. @tbl-h1b-results summarizes the results.

|Metric         |      `log(tct)` |     `sqrt(tct)` |  `inverse(tct)` |
|:--------------|:---------------:|:---------------:|:---------------:|
|R2 (cond.)     |            0.72 |            0.70 |            0.71 |
|R2 (marg.)     |            0.39 |            0.36 |            0.39 |
|ICC            |            0.54 |            0.53 |            0.52 |
|RMSE           |            0.20 |            1.03 |        2.32e-03 |
|Sigma          |            0.22 |            1.13 |        2.54e-03 |

: Comparison of Model Performance Metrics {#tbl-h1b-results .striped}

The values of ICC and $R^2$ (both conditional and marginal) are comparable across the three transformations. However, the inverse response resulted in substantially lower root mean square error (RMSE) and Sigma ($\sigma$), which represent the average and standard deviation of the residuals, respectively. With prediction errors that are both lower overall and have less variance, the inverse-transformed response, shown in @fig-inverse-tct, results in the best model performance.

```{r fig-inverse-tct, eval=FALSE}
#| fig-cap: Inverse transformation of the response
#| fig-height: 3.5
#| warning: false

# Create a data frame with the original and inverse transformed TCT values
pd <- learn_h1b |>
  mutate(inverse_tct = 1 / tct)

# Plot for tct
plot_tct <- ggplot(pd, aes(x = seq, y = tct, color = treatment)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "TCT vs. Sequence",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plot for inverse(tct)
plot_inverse_tct <- ggplot(pd, aes(x = seq, y = inverse_tct, color = treatment)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Inverse(TCT) vs. Sequence",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Display the plots
plot_tct + plot_inverse_tct +
  plot_annotation(
    title = "Task Completion Time and Inverse",
    subtitle = "Inverse transformation of TCT gives best model performance"
  )
```

:::
