#### try linear spline model

```{r}
library(glmmTMB)
library(splines)

cms <- glmmTMB(
  tct ~ 
    1 + ns(seq.fp, 2) * treatment + 
    (1 + ns(seq.fp, 2) | participant), 
  data = md, 
  family = gaussian()
)

summary(cms)
r2(cms)
```
would not fit with higher order (3/2 or 3/3) polynomial terms in spline; note: uses glmm model - probably drop this as "also tried"

visualize model fit of cms

```{r}
# Predicted values for different seq.fp levels
seq_levels <- unique(md$seq.fp)
predicted_vals <- expand.grid(treatment = levels(md$treatment), seq.fp = seq_levels)
predicted_vals$pred_tct <- predict(cms, newdata = predicted_vals, re.form = NA)

# Plot with both predicted lines and actual data points
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  geom_point(data = md, aes(x = seq.fp, y = tct, color = treatment, alpha = 0.4)) +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  facet_wrap(~ treatment) +
  theme_minimal()
```

similar to previous fit

#### try bayesian methods

seem much more robust - few convergence problems, take a while to fit

```{r}
library(brms)

cmb <- brm(
  tct ~ 1 + ns(seq.fp, 2) * treatment + 
         (1 + ns(seq.fp, 2) | participant), 
  data = md,
  family = gaussian(),
  prior = c(
    set_prior("normal(0, 10)", class = "b"),
    set_prior("normal(0, 10)", class = "sd", group = "participant")
  ),
  control = list(adapt_delta = 0.95)
)

summary(cmb)
r2(cmb)
emmeans::emmeans(cmb, pairwise ~ treatment)
pp_check(cmb)
plot(cmb, ask = F)

```

visualize fit for cmb

```{r}
seq_levels <- unique(md$seq.fp)
predicted_vals <- expand.grid(treatment = levels(md$treatment), seq.fp = seq_levels)

# Add dummy participant ID to prevent error, not used in prediction
predicted_vals$participant <- NA

# Use fitted() to get predictions, handling hierarchical structure
predictions <- fitted(cmb, newdata = predicted_vals, re_formula = NA)

# Add predictions to predicted_vals
predicted_vals$pred_tct <- as.vector(predictions[, "Estimate"])

# Convert treatment to a factor to ensure consistent plotting
predicted_vals$treatment <- factor(predicted_vals$treatment, levels = levels(md$treatment))

# Plot with both predicted lines and actual data points
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  geom_point(data = md, aes(x = seq.fp, y = tct, color = treatment), alpha = 0.4) +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  facet_wrap(~ treatment) +
  theme_minimal() +
  guides(alpha = "none") # Hide the alpha legend

```

shows potential - better in some ways (flatter on the right), worse in others (too flat on the left)

build alternative bayesian models

```{r}
# base model from above
cmb0 <- cmb

# increase polynomial degree (3,2)
cmb1 <- brm(
  tct ~ 1 + ns(seq.fp, 3) * treatment + 
         (1 + ns(seq.fp, 2) | participant), 
  data = md,
  family = gaussian(),
  prior = c(
    set_prior("normal(0, 10)", class = "b"),
    set_prior("normal(0, 10)", class = "sd", group = "participant")
  ),
  control = list(adapt_delta = 0.95)
)

# increase polynomial degree (3,3)
cmb2 <- brm(
  tct ~ 1 + ns(seq.fp, 3) * treatment + 
         (1 + ns(seq.fp, 3) | participant), 
  data = md,
  family = gaussian(),
  prior = c(
    set_prior("normal(0, 10)", class = "b"),
    set_prior("normal(0, 10)", class = "sd", group = "participant")
  ),
  control = list(adapt_delta = 0.95)
)

# # use b-spline instead
# cmb2 <- brm(
#   tct ~ 1 + bs(seq.fp, degree = 3) * treatment + 
#          (1 + bs(seq.fp, degree = 2) | participant), 
#   data = md,
#   family = gaussian(),
#   prior = c(
#     set_prior("normal(0, 10)", class = "b"),
#     set_prior("normal(0, 10)", class = "sd", group = "participant")
#   ),
#   control = list(adapt_delta = 0.95)
# )

# # use non-gaussian family
# cmb3 <- brm(
#   tct ~ 1 + ns(seq.fp, 2) * treatment + 
#          (1 + ns(seq.fp, 2) | participant), 
#   data = md,
#   family = student(),
#   prior = c(
#     set_prior("normal(0, 10)", class = "b"),
#     set_prior("normal(0, 10)", class = "sd", group = "participant")
#   ),
#   control = list(adapt_delta = 0.95)
# )

# add interaction term for early vs late iterations
md$early <- ifelse(md$seq.fp <= 0.25, 1, 0)
cmb4 <- brm(
  tct ~ 1 + ns(seq.fp, 2) * treatment * early + 
         (1 + ns(seq.fp, 2) | participant), 
  data = md,
  family = gaussian(),
  prior = c(
    set_prior("normal(0, 10)", class = "b"),
    set_prior("normal(0, 10)", class = "sd", group = "participant")
  ),
  control = list(adapt_delta = 0.95)
)

# use early term and higher order poly (3,3)
cmb5 <- brm(
  tct ~ 1 + ns(seq.fp, 3) * treatment * early + 
         (1 + ns(seq.fp, 3) | participant), 
  data = md,
  family = gaussian(),
  prior = c(
    set_prior("normal(0, 10)", class = "b"),
    set_prior("normal(0, 10)", class = "sd", group = "participant")
  ),
  control = list(adapt_delta = 0.95)
)


```

look at best model, cmb5

```{r}
m <- cmb5
summary(m)
r2(m)
emmeans::emmeans(m, pairwise ~ treatment)
pp_check(m)
plot(m, ask = F)# Predicted values for different seq.fp levels

seq_levels <- unique(md$seq.fp)
predicted_vals <- expand.grid(treatment = levels(md$treatment), seq.fp = seq_levels)

# Add dummy participant ID to prevent error, not used in prediction
predicted_vals$participant <- NA

# Add 'early' variable based on the same condition used for the model data
predicted_vals$early <- ifelse(predicted_vals$seq.fp <= 0.25, 1, 0)

# Use fitted() to get predictions, handling hierarchical structure
predictions <- fitted(m, newdata = predicted_vals, re_formula = NA)

# Add predictions to predicted_vals
predicted_vals$pred_tct <- as.vector(predictions[, "Estimate"])

# Convert treatment to a factor to ensure consistent plotting
predicted_vals$treatment <- factor(predicted_vals$treatment, levels = levels(md$treatment))

# Plot with both predicted lines and actual data points
ggplot(predicted_vals, aes(x = seq.fp, y = pred_tct, color = treatment, group = treatment)) +
  geom_line() +
  geom_point(data = md, aes(x = seq.fp, y = tct, color = treatment), alpha = 0.4) +
  labs(title = "Predicted TCT across Task Iterations by Treatment", x = "Task Iteration (seq.fp)", y = "Predicted TCT") +
  facet_wrap(~ treatment) +
  theme_minimal() +
  guides(alpha = "none") # Hide the alpha legend


```

```{r}
# can't compare directly lmer / glmm with brm fits
# lmer quad rsri, glmmTMB splines
cp1 <- compare_performance(cm1, cms, rank = T)
# b-ns2/2, b-ns3/2, b-ns3/3, be-ns2/2, be-ns3/3
cp2 <- compare_performance(cmb0, cmb1, cmb2, cmb4, cmb5, rank = T)
display(cp1, layout = "vertical")
display(cp2, layout = "vertical")
# cp1 best = cm1 (lmer quad rsri)
# cp2 best = cmb5 (be-ns3/3), cmb2 (b-ns3/3), cmb1 (b-ns3/2)
```

cmb5 appears to be the best model

need to do more exhaustive search with cross validation or other best practices to search the hyper-parameter space for: `ns`, `early` array


Hyperparameter search of the bayesian model to maximize fit.

```{r load-libs}
#| warning: false

library(tidyverse)
library(lmerTest)
library(brms)
library(tidymodels)
library(broom.mixed) # For tidymodels compatibility with brms

# library(pwr)
# library(emmeans)
# library(performance)
# library(gt)
# library(patchwork)
# library(moments)
# library(ggpubr)
# library(ggstatsplot)
```

```{r load-data}
# Load objects from the file
load("rdata/learn.RData")
```

```{r prep-h1b-data}
# disaggregated data required for rate of change of tct
# create data
learn_h1b <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj)

# mutate
learn_h1b <- learn_h1b |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

# transform seq to float from 0 to 1
md <- learn_h1b |>
  mutate(seq.fp = (seq - 1) / (max(seq) - 1))

summary(md$seq.fp)
glimpse(md)
```

## Unconditional Bayesian  Model

```{r}
umb <- brm(
  tct ~ 1 + ns(seq.fp, 2) + 
        (1 + ns(seq.fp, 2) | participant), 
  data = md,
  family = gaussian(),
  prior = c(
    set_prior("normal(0, 10)", class = "b"),
    set_prior("normal(0, 10)", class = "sd", group = "participant")
  ),
  control = list(adapt_delta = 0.95)
)
```

Tune unconditioned Bayesian model

$$
\begin{align}
\text{TCT}_{ij} &= \beta_0 + \beta_1 (\text{SEQ}_{ij}) + \beta_2 (\text{SEQ}_{ij}^{ns}) \\
                &\quad + \beta_3 (\text{SW}_{ij}) + \beta_4 (\text{SEQ}_{ij}) \times \text{SW}_{ij} + \\
                &\quad + (1 + \text{SEQ}_{ij} | \text{Part}_i) + \epsilon_{ij}
\end{align}
$$

unconditioned version of model definitions:

m4 (lmer): `tct ~ 1 + seq.fp + I(seq.fp ^ 2) + (1 + seq.fp + I(seq.fp ^ 2) | participant)`

$$
\text{TCT}_{ij} = \beta_0 + \beta_1 (\text{SEQ}_{ij}) + \beta_2 (\text{SEQ}_{ij}^2) + (1 + \beta_3 (\text{SEQ}_{ij}) + \beta_4 (\text{SEQ}_{ij}^2) \space | \space \text{PART}_{i}) + \epsilon_{ij}
$$

- Fixed Effects:
  - Intercept: 1
  - Linear Sequence: seq.fp
  - Quadratic Sequence: I(seq.fp ^ 2)
  - Interaction: *None*
- Random Effects:
  - Intercept: 1
  - Linear Sequence: seq.fp (random slope)
  - Quadratic Sequence: I(seq.fp ^ 2) (random slope)
  - Condition: random effects grouped by participant

cmb0 (brm): `tct ~ 1 + ns(seq.fp, 2) + (1 + ns(seq.fp, 2) | participant)`
- Fixed Effects:
  - Intercept: 1
  - Natural Splines of Sequence: ns(seq.fp, 2)
  - Linear Sequence: Included in ns(seq.fp, 2) but not explicit as term
  - Quadratic Sequence: Included in ns(seq.fp, 2) but not explicit as term
  - Interaction: *None*
- Random Effects:
  - Intercept: 1
  - Natural Splines of Sequence: ns(seq.fp, 2) (random slopes)
  - Condition: random effects grouped by participant

cmb5 (brm):

`tct ~ 1 + ns(seq.fp, 2) * early + (1 + ns(seq.fp, 2) | participant)`

$$
\begin{align*}
\text{TCT}_{ij} &= \beta_0 + \beta_1 (\text{SEQ}_{ij}^{\text{s1}} \times \text{E}_{ij}) + \beta_2 (\text{SEQ}_{ij}^{\text{s2}} \times \text{E}_{ij}) \\
                &\quad + \left( 1 + \beta_3 (\text{SEQ}_{ij}^{\text{s1}}) + \beta_4 (\text{SEQ}_{ij}^{\text{s2}}) \mid \text{PART}_i \right) + \epsilon_{ij}
\end{align*}
$$

where each interaction includes the components:

$$
A \times B = A + B + A \times B
$$

so:

$$
\beta_1 (\text{SEQ}_{ij}^{\text{s1}} \times \text{E}_{ij}) + \beta_2 (\text{SEQ}_{ij}^{\text{s2}} \times \text{E}_{ij}) \\
$$

is equivalent to: 

$$
\beta_1 (\text{SEQ}_{ij}^{\text{s1}}) + \beta_2 (\text{SEQ}_{ij}^{\text{s2}}) + \beta_3 (\text{SEQ}_{ij}^{\text{s1}} \times \text{E}_{ij}) + \beta_4 (\text{SEQ}_{ij}^{\text{s2}} \times \text{E}_{ij}) + \beta_5 (\text{E}_{ij})
$$

therefore:

$$
\begin{align*}
\text{TCT}_{ij} &= \beta_0 + \beta_1 (\text{SEQ}_{ij}^{\text{s1}}) + \beta_2 (\text{SEQ}_{ij}^{\text{s2}}) + \beta_3 (\text{SEQ}_{ij}^{\text{s1}} \times \text{E}_{ij}) + \beta_4 (\text{SEQ}_{ij}^{\text{s2}} \times \text{E}_{ij}) + \beta_5 (\text{E}_{ij}) \\
&\quad + \left( 1 + \beta_6 (\text{SEQ}_{ij}^{\text{s1}}) + \beta_7 (\text{SEQ}_{ij}^{\text{s2}}) \mid \text{PART}_i \right) + \epsilon_{ij}
\end{align*}
$$

- $\beta_0$: Intercept
- $\beta_1(\text{SEQ}_{ij}^{s1})$: Main effect for the first spline basis
- $\beta_2(\text{SEQ}_{ij}^{s2})$: Main effect for the second spline basis
- $\beta_3(\text{SEQ}_{ij}^{s1} \times \text{E}_{ij})$: Interaction of first spline basis with early phase
- $\beta_4(\text{SEQ}_{ij}^{s2} \times \text{E}_{ij})$: Interaction of second spline basis with early phase
- $\beta_5 (\text{E}_{ij})$: Main effect of the early phase
- $\left( 1 + \beta_6 (\text{SEQ}_{ij}^{s1}) + \beta_7 (\text{SEQ}_{ij}^{s2}) \mid \text{PART}_i \right)$: 

Random effects grouped by participant
- $\epsilon_{ij}$: Residual error
- Fixed Effects:
  - Intercept: 1
  - Natural Splines of Sequence: ns(seq.fp, 2)
  - Linear Sequence: Included in ns(seq.fp, 2) but not explicit as term
  - Quadratic Sequence: Included in ns(seq.fp, 2) but not explicit as term
  - Interaction: ns(seq.fp, 2) * early
- Random Effects:
  - Intercept: 1
  - Natural Splines of Sequence: ns(seq.fp, 2) (random slopes)
  - Condition: random effects grouped by participant


```{r}
# Create a recipe
umb_recipe <- recipe(tct ~ seq.fp + participant, data = md) |>
  step_ns(seq.fp, deg_free = tune()) |>
  step_mutate(early = ifelse(seq.fp <= tune(), 1, 0)) |>
  step_interact(terms = ~ ns(seq.fp, 2):early)

# Define the Bayesian model
bayesian_spec <- linear_reg() |>
  set_engine(
    "stan",
    prior_intercept = normal(0, 10),
    prior = normal(0, 10),
    prior_aux = cauchy(0, 5),
    adapt_delta = 0.95
  ) |>
  set_mode("regression")

# Define Workflow
workflow_bayesian <- workflow() |>
  add_recipe(umb_recipe) |>
  add_model(bayesian_spec)

# Define the parameter object for deg_free with a specific range
deg_free_param <- deg_free(range = c(2, 5))

# Define the parameter object for early
early_param <- threshold(range = c(0.1, 0.5))

# Create a parameter grid to be searched
tune_grid <- grid_regular(
  deg_free_param,   # Degrees of freedom for natural splines
  early_param,      # Cut-off for early variable
  levels = 4
)

# Grouped 5-fold cross-validation by participant
set.seed(123)
folds <- group_vfold_cv(md, group = participant, v = 5)

# Check balance within cross-validation folds
balance_check <- folds |>
  mutate(train_treat_dist = map(
    splits,
    ~ .x |> analysis() |>
      count(treatment) |>
      mutate(prop = n / sum(n))
  )) |>
  unnest(train_treat_dist) |>
  group_by(id) |>
  summarize(
    treatment_mean = mean(n),
    treatment_sd = sd(n),
    prop_mean = mean(prop),
    prop_sd = sd(prop)
  )

balance_check

```

pre-flight checks

```{r}
# Take a small sample of the data
md_sample <- md |> sample_n(10)

# Perform dry run with small data subset and limited iterations
set.seed(123)
folds_sample <- group_vfold_cv(md_sample, group = participant, v = 2)

# Dry run tuning with limited resamples (use a rapid backend if fitting models)
dry_run_results <- tune_grid(
  workflow_bayesian_sample,
  resamples = folds_sample,
  grid = tune_grid,
  control = control_grid(save_pred = TRUE, extract = partial(rmse, skip = TRUE))
)

# Check the results from the dry run
print(dry_run_results)

```

```{r}
# Tune the model using the workflow and predefined grid
tune_results <- tune_grid(
  workflow_bayesian,
  resamples = folds,
  grid = tune_grid,
  control = control_grid(save_pred = TRUE)
)

# Collect best tuning results
best_params <- select_best(tune_results, "rmse")  # You can use appropriate metric

```


```{r}
# Sample Data Preparation
set.seed(123)
md <- md %>% mutate(participant = as.factor(participant))

# Split Data
folds <- group_vfold_cv(md, group = participant, v = 5)

# Hyperparameter Grid for Tuning
deg_free_param <- tibble::tibble(deg_free = seq(2, 4, by = 1))
early_param <- tibble::tibble(early_cutoff = seq(0.1, 0.5, by = 0.1))

# For each Fold
results <- list()

for (i in seq_len(length(folds$splits))) {
  for (deg_free in deg_free_param$deg_free) {
    for (early_cutoff in early_param$early_cutoff) {
      
      # Training Data for the Fold
      train_data <- analysis(folds$splits[[i]])
      
      # Create the early variable
      train_data <- train_data %>% mutate(early = ifelse(seq.fp <= early_cutoff, 1, 0))
      
      # Fit the Model using brms
      formula <- brms::bf(
          tct ~ 1 + ns(seq.fp, deg_free) * early + (1 + ns(seq.fp, deg_free) | participant)
        )
      
      fit <- brm(
        formula = formula,
        data = train_data,
        family = gaussian(),
        prior = c(
          set_prior("normal(0, 10)", class = "b"),
          set_prior("normal(0, 10)", class = "sd", group = "participant")
        ),
        control = list(adapt_delta = 0.95),
        sample_prior = TRUE
      )
      
      # Store Results
      results[[paste0("fold", i, "_df", deg_free, "_early", early_cutoff)]] <- fit
    }
  }
}

# Processing Results
print(results)

```

