::: {.callout-note collapse=false}
# **To Do Items**
- Quarto Cleanup
  - [x] Convert `{r load data, echo=FALSE}` format to Quarto style
  - [ ] Add plot labels to code blocks
- Add / Refactor Code
  - [ ] Add tests for assumptions of ANOVA, etc.
  - [ ] Refactor box plot by treatment as function
  - [ ] Refactor ANOVA → Tukey → Tukey Plot as one function
  - [ ] Order treatment factors consistently in all charts
  - [ ] Show statistically different groups in TukeyHSD charts
- Questions to Answer
  - [ ] Explore impact of age on results
  - [ ] Can we assess learning rates by treatment type from phase 1 results?
  - [ ] Can we uncover relationships with underlying affordances of treatments?
- Other
  - [ ] Consider incomplete cars as partials?
  - [ ] Review analysis for NA drop issues (see note in PWI Ref Count)
:::

## Treatment Effects

### Elapsed Time

**Hypothesis:** elapsed times will differ

**Expectation:** PAR will be faster than PWI (in-context instruction). This benefit will be mitigated by some inconsistency in gesture recognition. AR and MR will take progressively longer than PAR, owing to similar benefits but more complex and less reliable interfaces.

Results in line with expectations. Surprised to see mean time with AR the highest. Note increased variance in results in the treatment progression. Surprised that PAR and PWI are not significantly different.

::: {.callout-note collapse=false}
# **To Do Items**
- Elapsed Time
  - [ ] Compare with adjusted versions - significant differences?
  - [ ] How would this differ if PAR used a discrete switch rather than a gesture sensor?
:::

```{r}
#| label: boxplot-et-by-treat

car_results |>
  filter(phase == 2, car_outcome == "complete") |>
  ggplot(aes(x = treatment, y = car_elapsed_time)) +
  geom_boxplot() +
  labs(
    title = "Car Elapsed Time by Treatment for Phase 2",
    x = "Treatment",
    y = "Car Elapsed Time (seconds)"
  )

```

```{r}
#| label: anova-et-by-treat

phase_two_data <- car_results |> filter(phase == 2, car_outcome == "complete")

time_anova_result <- aov(car_elapsed_time ~ treatment, data = phase_two_data)

time_anova_summary <- summary(time_anova_result)
print(time_anova_summary)
```

```{r}
#| label: tukey-et-by-treat

p_value <- time_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  time_post_hoc <- TukeyHSD(time_anova_result)
  print(time_post_hoc)
} else {
  print("not significant")
}
```

```{r}
#| label: plot-tukey-et

plot(time_post_hoc)
```

### Uncorrected Error Count

**Hypothesis:** error count will differ

**Expectation:** PWI will be highest, other treatments will show decreasing error counts in order (PWI, AR, MR) due to multimodal learning. Improvements in AR and especially MR may be confounded by added complexity / reduced reliability of added technology.

Results meet expectations. Note that the PWI are clearly problematic as we observed. Though the other treatments are designed around the same instructions, each afford a certain amount of built-in error checking - the operator sees where the part should go - that help avoid making errors and/or learning to complete the task incorrectly.

Surprised to see AR error count so low. Task isn't complex enough to show lower limit.

*ChatGPT: You might say that the results exhibit "floor effects," where the lower limit of the measure (in this case, zero) constrains the range of possible scores, especially for certain treatments. This can indicate that the measure may not be sensitive enough to detect differences at the lower end of the performance spectrum for these treatments.*

::: {.callout-note collapse=false}
# **To Do Items**
- Uncorrected Error Count
  - [ ] Error types? Patterns?
:::

```{r}
#| label: boxplot-uce-by-treat

car_results |>
  filter(phase == 2, car_outcome == "complete") |>
  ggplot(aes(x = treatment, y = uce)) +
  geom_boxplot() +
  labs(
    title = "Uncorrected Error Count by Treatment for Phase 2",
    x = "Treatment",
    y = "Count"
  )

```

```{r}
#| label: anova-uce-by-treat

error_anova_result <- aov(uce ~ treatment, data = phase_two_data)

error_anova_summary <- summary(error_anova_result)
print(error_anova_summary)
```

```{r}
#| label: tukey-uce-by-treat

p_value <- error_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  error_post_hoc <- TukeyHSD(error_anova_result)
  print(error_post_hoc)
} else {
  print("not significant")
}
```

```{r}
#| label: plot-tukey-uce

plot(error_post_hoc)
```

## PWI Reference Count

::: {.callout-note collapse=false}
# **To Do Items**
- Treatment Effects
  - [x] Add PWI Reference Count analysis
  - [x] Add PWI Reference Duration analysis
  - [ ] Add Average Time per Reference by treatment
:::

How many times did participants have to refer to the paper work instructions during phase 2, by treatment?

**Hypothesis:** PWI reference counts will differ.

**Expectation:** Assuming that better recall of the instructions results in fewer PWI references, expect to see PWI ref counts decrease in order PWI, PAR, AR, MR. This effect may be countered by the participants' lack of familiarity with those instructions, requiring more refs for treatments other than PWI.

In fact, the opposite is seen, with PWI having the lowest average count, followed by PAR, then AR, and finally MR. All treatments exhibit floor effect, which may contribute to this result. Significant differences are only seen in the PAR-AR, PWI-AR, and PWI-MR pairs.

::: {.callout-warning}
Very important realization here - NA is **not** a good replacement for 0 values in the data. In the original data prep I used NA in the `pwi_event_count` column where no pwi references were recorded. But dropping those rows led to eliminating meaningful data (not looking at the instructions).
:::

```{r}
#| label: boxplot-pwi-ref-by-treat

pwi_ref_count_data <- car_results |>
  filter(
    phase == 2,
    car_outcome == "complete"
  ) |>
  mutate(pwi_event_count = ifelse(is.na(pwi_event_count), 0, pwi_event_count))

pwi_ref_count_data |>
  ggplot(aes(x = treatment, y = pwi_event_count)) +
  geom_boxplot() +
  labs(
    title = "Number of PWI References During Phase 2",
    x = "Treatment",
    y = "Count"
  )

```

```{r}
#| label: anova-pwi-refs-by-treat

pwi_ref_anova_result <-
  aov(pwi_event_count ~ treatment, data = pwi_ref_count_data)

pwi_ref_anova_summary <- summary(pwi_ref_anova_result)
print(pwi_ref_anova_summary)
```

```{r}
#| label: tukey-pwi-ref-by-treat

p_value <- pwi_ref_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  pwi_ref_post_hoc <- TukeyHSD(pwi_ref_anova_result)
  print(pwi_ref_post_hoc)
} else {
  print("not significant")
}
```

```{r}
#| label: plot-tukey-pwi-ref

plot(pwi_ref_post_hoc)
```

## PWI Reference Duration

What is the total time that participants spent referring to the paper work instructions during phase 2, by treatment?

**Hypothesis:** PWI reference durations will differ.

**Expectation:** Assuming that better recall of the instructions results in less time spent reviewing the work instructions, expect to see times decrease in order PWI, PAR, AR, MR. This effect may be countered by the participants' lack of familiarity with those instructions, requiring more time for treatments other than PWI.

A pattern similar to what was observed for the PWI references is observed, including the floor effect. This is consistent if the average duration per PWI ref is similar. Significant differences are only seen in the PWI-AR and PWI-MR pairs.

```{r}
#| label: boxplot-pwi-dur-by-treat

pwi_ref_dur_data <- car_results |>
  filter(
    phase == 2,
    car_outcome == "complete"
  ) |>
  mutate(pwi_event_tot_time = ifelse(is.na(pwi_event_tot_time), 0, pwi_event_tot_time))

pwi_ref_dur_data |>
  ggplot(aes(x = treatment, y = pwi_event_tot_time)) +
  geom_boxplot() +
  labs(
    title = "Total Duration of PWI References During Phase 2",
    x = "Treatment",
    y = "Time (seconds)"
  )

```

```{r}
#| label: anova-pwi-dur-by-treat

pwi_tot_time_anova_result <-
  aov(pwi_event_tot_time ~ treatment, data = pwi_ref_dur_data)

pwi_tot_time_anova_summary <- summary(pwi_tot_time_anova_result)
print(pwi_tot_time_anova_summary)
```

```{r}
#| label: tukey-pwi-dur-by-treat

p_value <- pwi_tot_time_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  pwi_tot_time_post_hoc <- TukeyHSD(pwi_tot_time_anova_result)
  print(pwi_tot_time_post_hoc)
} else {
  print("not significant")
}
```

```{r}
#| label: plot-tukey-pwi-tot-time

plot(pwi_tot_time_post_hoc)
```

## Explore Difference in UX by Treatment

Load System Usability and TLX scores.

```{r}
#| label: load-sus-and-tlx
#| echo: false

sus_scores <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "system usability scores"
)

tlx_scores <- read_xlsx(
  "./data/combined_results.xlsx",
  na = c("", "N/A"),
  sheet = "tlx scores"
)
```

### System Usability Scores (SUS)

**Hypothesis:** There will be differences in system usability.

**Expectations:** Decreasing usability owing to increased complexity: PWI > PAR > AR > MR.

Surprised to see no significant difference in perceived usability between treatments.

::: {.callout-note collapse=false}
# **To Do Items**
- System Usability Scores
  - [ ] Convert chart to box plot?
  - [ ] Compare PWI treatment across phase 1 and 2?
  - [ ] Phase 2 PWI could serve as baseline for system usability. Use that for comparisons?
:::

```{r}
#| label: barplot-sus-by-treat

sus_summary <- sus_scores |>
  filter(phase == 1) |>
  group_by(treatment) |>
  summarise(
    mean_value = mean(sus, na.rm = TRUE),
    sd_value = sd(sus, na.rm = TRUE)
  )

ggplot(sus_summary, aes(x = treatment, y = mean_value)) +
  geom_bar(stat = "identity", position = "dodge", fill = base_color) +
  geom_errorbar(
    aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
    width = 0.25, position = position_dodge(0.9)
  ) +
  labs(
    title = "SUS by Treatment in Phase 1",
    y = "Mean SUS",
    x = "Treatment"
  ) +
  theme_minimal()

```

```{r}
#| label: anova-sus-by-treat

sus_anova_result <- aov(sus ~ treatment, data = sus_scores)
sus_anova_summary <- summary(sus_anova_result)
print(sus_anova_summary)
```

```{r}
#| label: tukey-sus-by-treat

p_value <- sus_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  post_hoc <- TukeyHSD(sus_anova_result)
  print(sus_post_hoc)
} else {
  print("not significant")
}
```

### Task Load Index Scores (TLX)

**Hypothesis:** There will be differences in perceived workload.

**Expectations:** Variable workload based on a combination of instructional quality and interface complexity: PWI high, PWI medium, AR medium, MR high

Surprised to see very even workload across treatments and no significant difference. Suggests other factors have larger effect across all treatment? Experiment design? Time constraints?

::: {.callout-note collapse=false}
# **To Do Items**
- Task Load Index
  - [ ] Convert chart to box plot?
  - [ ] Compare PWI treatment across phase 1 and 2?
  - [ ] Phase 2 PWI could serve as baseline for task workload. Use that for comparisons?
  - [ ] Underlying factors of TLX?
  - [ ] Look at difference of perceived workload between phase 1 and 2? Per Victoria's study. She calls it "normalized", really just difference.
:::

```{r}
#| label: barplot-tlx-by-treat

tlx_task_workload_summary <- tlx_scores |>
  filter(type == "task workload", phase == 1) |>
  group_by(treatment) |>
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )

ggplot(tlx_task_workload_summary, aes(x = treatment, y = mean_value)) +
  geom_bar(stat = "identity", position = "dodge", fill = base_color) +
  geom_errorbar(
    aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
    width = 0.25, position = position_dodge(0.9)
  ) +
  labs(
    title = "Task Workload by Treatment in Phase 1",
    y = "Mean Task Workload",
    x = "Treatment"
  ) +
  theme_minimal()

```

```{r}
#| label: anova-tlx-by-treat

tlx_anova_result <- aov(value ~ treatment, data = tlx_scores)
tlx_anova_summary <- summary(tlx_anova_result)
print(tlx_anova_summary)
```

```{r}
#| label: tukey-tlx-ty-treat

p_value <- tlx_anova_summary[[1]]$`Pr(>F)`[1]
if (p_value < 0.05) {
  post_hoc <- TukeyHSD(tlx_anova_result)
  print(tlx_post_hoc)
} else {
  print("not significant")
}
```

## Visualize Car Time Trends in Phase 1 (Learning Rate)

```{r}
#| label: car-time-trends

plot_data <- car_results |>
  filter(phase == 1, car_outcome == "complete") |>
  group_by(participant, treatment) |>
  arrange(sequence) |>
  mutate(car_number = row_number()) |>
  ungroup() |>
  arrange(participant, car_number) |>
  mutate(treatment = factor(treatment, levels = c("PWI", "PAR", "AR", "MR")))

plot_data |>
  ggplot(aes(x = as.factor(car_number), y = car_elapsed_time)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5, color = base_color) +
  labs(
    title = "Car Elapsed Time by Car Number for Phase 1",
    x = "Car Number",
    y = "Elapsed Time (seconds)"
  ) +
  facet_wrap(~treatment, ncol = 1)

```

### Estimate the Learning Rate

Fit an exponential decay model to estimate learning rate, $c$. The equation for this model is:

$$y(t) = a + b \times e^{-ct}$$

```{r}
# Function to fit nls model and return tidy summary
fit_model <- function(df) {
  fit <- tryCatch(
    {
      nls(car_elapsed_time ~ a + b * exp(-c * car_number),
        start = list(a = 200, b = 1000, c = 0.1), # Initial guesses
        data = df
      )
    },
    error = function(e) {
      print(paste("Error fitting model for treatment:", unique(df$treatment)[1]))
      print(e)
      return(NULL)
    }
  )

  if (!is.null(fit)) {
    tidy_fit <- tidy(fit)
    tidy_fit$treatment <- unique(df$treatment)[1]
    return(tidy_fit)
  } else {
    return(NULL)
  }
}

# Apply the function to each treatment group and collate results
results <- plot_data |>
  (\(df) split(df, df$treatment))() |>
  map_df(fit_model, .id = "treatment")


results

```

