# Results {#25-results}

```{r load-libs}
#| message: false

library(here)
library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(patchwork)
library(xtable)
library(broom)
library(report)
library(tools)

library(emmeans)
library(performance)
library(lmerTest)
library(modelsummary)
library(boot)
library(english)

# library(ggtext)
# library(glue) now included with tidyverse as str_glue
# library(gridExtra)

# currently unused:
# library(skimr)
# library(gtExtras)
# library(grid)
```

```{r load-data}

car_results <- read_xlsx(
  here("data", "combined_results.xlsx"),
  na = c("", "N/A"),
  sheet = "car results"
)

demographics <- read_xlsx(
  here("data", "combined_results.xlsx"),
  na = c("", "N/A"),
  sheet = "demographics"
)
```

```{r def-cust-theme}
base_color <- "#69b3a2"
outline_color <- "grey25"

# Define a custom theme for ggplot
custom_theme <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      # plot.subtitle = element_text(face = "italic"),
      # plot.caption = element_text(color = "grey50"),
      # axis.title.x = element_text(face = "bold"),
      # axis.title.y = element_text(face = "bold"),
      # axis.text = element_text(color = "black"),
      # axis.line = element_line(color = "grey80"),
      # legend.position = "bottom",
      # legend.background = element_rect(fill = "white", colour = "grey50"),
      # legend.key = element_rect(fill = "white", colour = "grey50"),
    )
}

# Set the custom theme as the default theme
theme_set(custom_theme())
```

```{r def-funcs}
# Function to create QQ plots and arrange them in a 2x2 grid
qq_plot_patchwork <-
  function(data,
           x_var,
           by_var) {
    levels <- unique(data[[by_var]])

    if (length(levels) != 4) {
      stop("This script expects exactly 4 levels in the grouping variable.")
    }

    plots <- lapply(levels, function(level) {
      p <- data |>
        filter((!!sym(by_var)) == level) |>
        ggpubr::ggqqplot(
          x = x_var,
          xlab = "",
          ylab = "",
          title = level,
          ggtheme = theme_minimal(base_size = 10) +
            theme(
              plot.margin = margin(0, 0, 0, 0),
              panel.spacing = unit(0, "lines"),
              panel.grid = element_blank(),
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              axis.line = element_blank(),
              plot.title = element_text(hjust = 0.5),
              panel.border = element_rect(
                color = outline_color,
                fill = NA,
                linewidth = 0.5
              ),
            )
        ) |>
        ggpubr::ggpar(ticks = FALSE, tickslab = FALSE)
    })

    # Arrange the plots in a 2x2 grid using patchwork
    plot_layout <-
      (plots[[1]] + plots[[2]]) / (plots[[3]] + plots[[4]])

    return(plot_layout)
  }

# convert a p-value into an appropriately formatted string
fmt_p_val <- function(p_value, alpha = 0.05) {
  # # Check if p-value is greater than alpha
  # if (p_value > alpha) {
  #   return(paste0("> ", alpha))
  # }

  if (p_value < 0.001) {
    return("p < 0.001")
  }

  # need more precision for 0.01 > p-val > 0.001
  if (p_value < 0.01) {
    return(paste0("p = ", round(p_value, digits = 3)))
  }

  return(paste0("p = ", round(p_value, digits = 2)))
}

# Function to calculate power for Kruskal-Wallis test via simulation
simulate_kruskal_power <- function(
    n, k, effect_size, sig_level = 0.05, num_simulations = 1000) {
  set.seed(42)
  p_values <- numeric(num_simulations)
  for (i in 1:num_simulations) {
    # Simulate data for each group with an effect size
    data <- data.frame(
      value = unlist(lapply(1:k, function(x) {
        rnorm(n, mean = x * effect_size)
      })),
      group = gl(k, n)
    )
    # Perform Kruskal-Wallis test
    p_values[i] <- kruskal.test(value ~ group, data = data)$p.value
  }
  # Proportion of tests rejecting null hypothesis gives power estimate
  power <- mean(p_values < sig_level)
  return(power)
}

# Function to run post-hoc power analysis based on observed epsilon squared
posthoc_power_analysis <- function(
    n, k, observed_effect_size, sig_level = 0.05, num_simulations = 1000) {
  # Convert epsilon squared back to standard effect size approximation for simulation
  std_effect_size <- sqrt(observed_effect_size / (1 - observed_effect_size))

  # Use simulation to estimate power as before
  power <- simulate_kruskal_power(
    n, k, std_effect_size, sig_level, num_simulations
  )
  return(power)
}

# convert a column of a table into a string with the format "a, b, and c"
# numbers are rounded to the specified number of digits
col_to_string <- function(data, column, digits = 2) {
  data |>
    mutate({{ column }} := round({{ column }}, digits)) |>
    pull({{ column }}) |>
    str_c(collapse = ", ") |>
    (\(x) sub(", ([^,]+)$", ", and \\1", x))()
}

# display selected model parameters
display_model_parameters <- function(model, type = "fixed", exp = F) {

  # Function to add significance stars
  add_significance_stars <- function(p_value) {
    if (p_value < 0.001) {
      "***"
    } else if (p_value < 0.01) {
      "**"
    } else if (p_value < 0.05) {
      "*"
    } else if (p_value < 0.1) {
      "."
    } else {
      ""
    }
  }

  # Get model parameters
  model_params <- parameters::model_parameters(model, exponentiate = exp)

  # Separate the fixed effects
  model_params_ef <- model_params |> 
    filter(Effects == type)

  # Create the table
  result_table <- model_params_ef |> 
    mutate(Sig = sapply(p, add_significance_stars)) |> 
    select(-Group) 
    
  display(result_table, digits = 3, ci_digits = 2)
}
```

```{r def-boot-funcs}
# Hodges-Lehmann Estimator - stat_func option
hl_estimator <- function(x) {
  n <- length(x)
  pairwise_means <- outer(x, x, FUN = function(a, b) (a + b) / 2)
  median(pairwise_means[lower.tri(pairwise_means, diag = TRUE)])
}

# Generalized function to perform bootstrap and get results
bootstrap_stat_diff <- function(data, group1, group2, column, stat_func) {
  # cat("Starting bootstrap_stat_diff for", group1, "vs", group2, "\n")

  stat_diff <- function(data, indices, group1, group2, column, stat_func) {
    # static_counter <- 0
    # static_counter <- static_counter + 1
    # cat("stat_diff called for the", static_counter, "time for this resample\n")

    d <- data[indices, ]
    g1_data <- d[d$treatment == group1, column, drop = TRUE]
    g2_data <- d[d$treatment == group2, column, drop = TRUE]

    # cat(group1, "-", paste(sprintf("%.2f", g1_data), collapse = " "), "\n")
    # cat(group2, "-", paste(sprintf("%.2f", g2_data), collapse = " "), "\n")

    # cat(
    #   "Subset sizes: ", group1, "=", length(g1_data), ", ",
    #   group2, "=", length(g2_data), "\n"
    # )

    stat_func(g1_data) - stat_func(g2_data)
  }

  # Define statistic function for bootstrap
  stat_wrapper <- function(data, indices) {
    stat_diff(data, indices, group1, group2, column, stat_func)
  }

  set.seed(42) # Set seed right before boot call

  # browser()

  results <- boot(
    data = data,
    statistic = stat_wrapper,
    R = 10000,
    strata = data$treatment,
    # parallel = "multicore"
  )

  # cat("\nBootstrap results for", group1, "vs", group2, ":\n")
  # print(results)

  # Get bootstrap confidence intervals
  ci <- boot.ci(results, type = "perc")

  # cat("\n\nConfidence intervals for", group1, "vs", group2, ":\n")
  # print(ci)

  # Create a summary of bootstrap results
  summary <- list(
    original = results$t0,
    bias = mean(results$t) - results$t0,
    std_error = sd(results$t),
    ci_lower = ci$percent[4],
    ci_upper = ci$percent[5]
  )

  # cat("Returning summary for", group1, "vs", group2, "\n")
  return(summary)
}

# Function to perform multiple replications of the bootstrap
multiple_bootstrap <- function(data, group1, group2, column, stat_func, num_reps) {
  # cat("Starting multiple_bootstrap for", group1, "and", group2, "\n")

  # # original parallel implementation, results in invalid results
  # all_results <- foreach(
  #     i = 1:num_reps,
  #     .combine = list,
  #     .packages = c("boot", "dplyr"),
  #     .errorhandling = "remove"  # was stop
  #   ) %dopar% {
  #     # cat("Starting replication", i, "\n")
  #     tryCatch({
  #       result <- bootstrap_stat_diff(data, group1, group2, column, stat_func)
  #       # cat("Completed replication", i, "\n")
  #       result
  #     }, error = function(e) {
  #       cat("*** Error in replication", i, ":", conditionMessage(e), "\n")
  #       NULL
  #     })
  #   }

  # cat("Num reps =", num_reps, "\n")

  # working non-parallel implementation
  all_results <- vector("list", num_reps) # Initialize an empty list of the correct length

  for (i in 1:num_reps) {
    # cat("\nStarting replication", i, "\n")
    tryCatch(
      {
        result <- bootstrap_stat_diff(data, group1, group2, column, stat_func)
        all_results[[i]] <- result
      },
      error = function(e) {
        cat("*** Error in replication", i, ":", conditionMessage(e), "\n")
        all_results[[i]] <- NULL
      }
    )
  }

  cat("Completed all replications. Number of results:", length(all_results), "\n")

  # Check for errors
  errors <- sapply(all_results, is.null)
  if (any(errors)) {
    cat("*** Errors occurred during bootstrap for the following replications:\n")
    cat(paste("Replication", which(errors), "\n", collapse = ""))
  }

  # Process successful results
  results_list <- lapply(seq_along(all_results), function(i) {
    res <- all_results[[i]]
    if (is.null(res)) {
      cat("*** Warning: Missing result for replication", i, "\n")
      return(NULL)
    }

    if (!is.list(res) || !all(c("original", "bias", "std_error", "ci_lower", "ci_upper") %in% names(res))) {
      cat("*** Warning: Invalid result for replication", i, "\n")
      return(NULL)
    }

    list(
      Replication = i,
      original = res$original,
      bias = res$bias,
      std_error = res$std_error,
      ci_lower = res$ci_lower,
      ci_upper = res$ci_upper
    )
  })

  results_list <- results_list[!sapply(results_list, is.null)]

  if (length(results_list) == 0) {
    stop("*** No valid results were produced. ***")
  }

  cat("Processed", length(results_list), "valid results.\n")

  return(do.call(rbind, results_list))
}

# Generalized pairwise tests function to capture results for each replication
bootstrap_pairs <- function(data, column, stat_func, num_reps = 5) {
  set.seed(42)
  cat("+++ Starting bootstrap for", CUR_STAT, "of", column, "...\n")
  msg <- paste("+++ Bootstrap for", column, CUR_STAT, "complete")
  tictoc::tic(msg)
  results <- data.frame(
    Group1 = character(),
    Group2 = character(),
    Replication = numeric(),
    Mean_Difference = numeric(),
    Bias = numeric(),
    Std_Error = numeric(),
    CI_Lower = numeric(),
    CI_Upper = numeric(),
    Sig = character(),
    stringsAsFactors = FALSE
  )

  for (i in 1:(length(TREATMENTS) - 1)) {
    for (j in (i + 1):length(TREATMENTS)) {
      group1 <- TREATMENTS[i]
      group2 <- TREATMENTS[j]

      cat("Starting bootstrap for", group1, "vs", group2, "\n")

      # Perform bootstrap test
      test_results <- multiple_bootstrap(data, group1, group2, column, stat_func, num_reps)

      # cat("Completed bootstrap for", group1, "vs", group2, "\n")

      # Add results to the dataframe
      for (k in 1:nrow(test_results)) {
        rep_result <- test_results[k, ]
        results <- rbind(results, data.frame(
          Group1 = group1,
          Group2 = group2,
          Replication = rep_result$Replication,
          Mean_Difference = rep_result$original,
          Bias = rep_result$bias,
          Std_Error = rep_result$std_error,
          CI_Lower = rep_result$ci_lower,
          CI_Upper = rep_result$ci_upper,
          Sig = ifelse(rep_result$ci_lower > 0 | rep_result$ci_upper < 0, "*", "")
        ))
      }
    }
  }

  tictoc::toc()
  return(results)
}

# Function to summarize pairwise results
summarize_pairwise_results <- function(pairwise_results) {
  result_summary <- pairwise_results |>
    filter(is.finite(Bias) & is.finite(Std_Error)) |>
    group_by(Group1, Group2) |>
    # summarize across all replications
    summarise(
      Diff = mean(Mean_Difference, na.rm = TRUE),
      Diff_L = mean(CI_Lower, na.rm = TRUE),
      Diff_U = mean(CI_Upper, na.rm = TRUE),
      Bias_mean = mean(Bias, na.rm = TRUE),
      Bias_sd = sd(Bias, na.rm = TRUE),
      SE_mean = mean(Std_Error, na.rm = TRUE),
      SE_sd = sd(Std_Error, na.rm = TRUE),
      Sig = ifelse(
        mean(CI_Lower > 0, na.rm = TRUE) | mean(CI_Upper < 0, na.rm = TRUE), "*", ""
      ),
      .groups = "drop"
    ) |>
    mutate(
      Pair = paste(Group1, Group2, sep = " - "),
      Pair = factor(Pair, levels = ORDERED_PAIRS)
    ) |>
    arrange(Pair) |>
    select(Pair, Diff:Sig)

  return(result_summary)
}

# Wrapper function to run bootstrap pairs for each stat function and combine the results
run_bootstrap_tests <- function(data, column, stat_funcs) {
  results <- map_dfr(names(stat_funcs), function(stat_name) {
    stat_func <- stat_funcs[[stat_name]]

    # set global for current stat
    assign("CUR_STAT", stat_name, envir = .GlobalEnv)

    # Perform bootstrap pairs
    pairwise_results <- bootstrap_pairs(data, column = column, stat_func = stat_func)

    # Summarize the results
    summary_results <- summarize_pairwise_results(pairwise_results)

    # Add the statistic column
    summary_results <- summary_results |>
      mutate(Stat = stat_name)

    return(summary_results)
  })

  # Ensure 'Stat' column correctly placed
  results <- results |>
    select(Pair, Stat, Diff:Sig)

  return(results)
}

create_boot_summary_table <- function(data, title, subtitle) {
  data |>
    filter(Sig == "*") |>
    select(-Sig) |>
    gt() |>
    tab_header(
      title = title,
      subtitle = subtitle
    ) |>
    cols_label(
      Pair = "Comparison",
      Diff = "OEE Difference",
      Diff_L = "Lower CI",
      Diff_U = "Upper CI",
      Bias_mean = "Bias",
      Bias_sd = "sd_Bias",
      SE_mean = "Std Error",
      SE_sd = "sd_SE"
    ) |>
    cols_merge(
      columns = c(Bias_mean, Bias_sd),
      pattern = "{1} ({2})"
    ) |>
    cols_merge(
      columns = c(SE_mean, SE_sd),
      pattern = "{1} ({2})"
    ) |>
    cols_merge(
      columns = c(Diff_L, Diff_U),
      pattern = "[{1}, {2}]"
    ) |>
    cols_label(
      .list = c(
        "Diff" = "Mean",
        "Diff_L" = "95% CI"
      )
    ) |>
    fmt_number(
      decimals = 3
    ) |>
    cols_align(align = "center") |>
    tab_options(
      table.font.size = "smaller",
      column_labels.font.weight = "bold",
      heading.title.font.weight = "bolder",
      heading.subtitle.font.size = px(13)
    ) |>
    tab_footnote(
      footnote = "Significance indicated by confidence intervals that do not include zero."
    )
}

```

```{r save-shared}
# saving for use elsewhere
save(
  qq_plot_patchwork,
  fmt_p_val,
  simulate_kruskal_power,
  posthoc_power_analysis,
  col_to_string,
  base_color,
  outline_color,
  custom_theme,
  file = here("rdata", "shared.RData")
)
```


```{r study-metadata}

participant_count <- demographics |>
  pull(participant) |>
  n_distinct()

# Convert the trial_datetime to Date type
demographics <- demographics |>
  mutate(
    trial_date = as.Date(
      as.POSIXct(trial_datetime, format = "%Y-%m-%d %H:%M:%S")
    )
  )

# Get the first and last date
first_trial_date <- demographics |>
  summarize(first_date = min(trial_date)) |>
  pull(first_date)

last_trial_date <- demographics |>
  summarize(last_date = max(trial_date)) |>
  pull(last_date)

trial_start_str <- str_glue(
  "{month(first_trial_date)}/{day(first_trial_date)}"
)
trial_end_str <- str_glue(
  "{month(last_trial_date)}/{day(last_trial_date)}"
)
trial_year_str <- str_glue("{year(last_trial_date)}")

# start, end = Feb 10, Apr 27 2023
```

## Study Administration and Participation

This study was administered during the Spring of 2023. Ultimately, `r participant_count` eligible participants were recruited. All completed the Learning and Recall experiments of the first session beginning February 10th. Progress of the conduct of these trials is illustrated by @fig-s1-progress.

```{r fig-s1-progress}
#| fig-cap: "Daily Trial Administration and Completion Percentage"
#| fig-height: 3.5

# Count experiments per day
daily_counts <- demographics |>
  group_by(trial_date) |>
  summarize(daily_count = n()) |>
  arrange(trial_date)

# Calculate the cumulative count of experiments
daily_counts <- daily_counts |>
  mutate(cumulative_count = cumsum(daily_count))

# Calculate remaining experiments
daily_counts <- daily_counts |>
  mutate(
    percentage_complete = (cumulative_count / participant_count) * 100
  )

pc_eng <- as.character(english::english(participant_count))
pc_eng <- str_replace(pc_eng, "^.", str_to_upper)
subtitle <- str_glue("{pc_eng} Learning and Recall Experiments Conducted between {trial_start_str} and {trial_end_str} of {trial_year_str}")

# Plotting
ggplot(daily_counts, aes(x = trial_date)) +
  geom_line(
    aes(y = daily_count, group = 1),
    color = base_color,
    linewidth = 1,
    alpha = 0.5
  ) +
  geom_line(
    aes(y = percentage_complete, group = 1),
    color = base_color,
    linewidth = 1.25,
    alpha = 1
  ) +
  scale_y_continuous(
    name = "Trials Administered",
    sec.axis = sec_axis(~., name = "Percentage Complete"),
    position = "right",
    limits = c(0, 100)
  ) +
  # Add annotation
  annotate("text", 
           x = as.Date("2023-02-28"), y = 30, 
           label = "Pilot Study (n=8)", 
           color = "red", hjust = 0.5,
           size = 3.5
  ) +
  annotate("segment", 
           x = as.Date("2023-02-10"), xend = as.Date("2023-03-15"), 
           y = 25, yend = 25, 
           color = "red", arrow = arrow(ends = "both", length = unit(0.1, "cm"))
  ) +
  labs(
    title = "First Session Progress",
    subtitle = subtitle,
    x = element_blank(),
    y = element_blank()
  ) +
  custom_theme()
```

Only eight trials were run before March 16th. That period was treated as a pilot study, during which the research team was trained, administration methods were refined, the self-service onboarding system was deployed, and version 2.1 of our IRB application was approved. Most significantly, the intake process was adjusted to collect more demographic data, increase the role of the TLX and SUS instruments, and gather BCS data for a companion study.

Otherwise, only minor tweaks were made during the pilot to streamline the experiment's administration. Although the changes outlined above were not believed to have a material effect on the results, the data collected during the pilot were omitted from further analysis. Following the pilot, steady progress was made until the completion of the study's first session on April 27th. The second session was conducted during the end of study event on April 29th. Twenty-four of the study's original participants volunteered to attend and complete the Retention experiment.

## Analysis of Participant Demographics

### Univariate Analysis

```{r drop-pilot-data}
# drop participants in pilot study (first eight)
demographics <- demographics |>
  filter(!(participant >= 1001 & participant <= 1010))

# drop event data from the pilot study
car_results <- car_results |>
  filter(!(participant >= 1001 & participant <= 1010))
```

```{r def-to-factors}

# safely convert data to ord / cat factors
# create levels / labels as required and identify / warn about mismatches

to_factors <- function(
    column,
    levels = NULL,
    labels = NULL,
    ordered = FALSE) {
  if (is.null(levels)) {
    levels <- unique(na.omit(column))
  }

  if (is.null(labels)) {
    labels <- levels
  }

  mismatched <- setdiff(unique(na.omit(column)), levels)

  if (length(mismatched) > 0) {
    # Append mismatched values to levels and labels with a specific prefix
    bad_levels <- paste("Bad -", mismatched)
    levels <- c(levels, mismatched)
    labels <- c(labels, bad_levels)
    warning(
      "Mismatched or unexpected values incorporated: ",
      paste(bad_levels, collapse = ", ")
    )
  }

  # Proceed to convert to factor using cleaned-up data
  factor_column <- factor(column,
    levels = levels, labels = labels,
    ordered = ordered
  )

  return(factor_column)
}
```

```{r factor-demographics}

# convert all categorical / ordinal data to factors

demo_f <- demographics |>
  mutate(
    # categorical, levels != labels
    # unordered with predefined levels
    gender = to_factors(
      gender,
      levels = c("F", "M", "O"),
      labels = c("Female", "Male", "Non-Binary")
    ),
    race = to_factors(
      race,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "American Indian or Alaska Native",
        "Asian",
        "Black or African American",
        "Native Hawaiian or Other Pacific Islander",
        "White",
        "More than one race",
        "Unknown or not reported"
      )
    ),
    ethnicity = to_factors(
      ethnicity,
      levels = c("1", "2", "3"),
      labels = c(
        "Hispanic or Latino",
        "Not Hispanic or Latino",
        "Unknown or not reported"
      )
    ),
    lenses = to_factors(
      lenses,
      levels = c("G", "C", "N", "Y"),
      labels = c(
        "Glasses",
        "Contacts",
        "No",
        "Yes"
      )
    ),

    # categorical, levels == labels
    # unordered with predefined levels
    language = to_factors(
      language,
      levels = c("English", "Other"),
    ),
    worn = to_factors(
      worn,
      levels = c("Y", "N")
    ),
    color_blind = to_factors(
      color_blind,
      levels = c("Y", "N")
    ),
    condition = to_factors(
      condition,
      levels = c("Y", "N")
    ),

    # ad-hoc
    # categorized in transcription, no predefined levels
    country = to_factors(country),
    college = to_factors(college),
    program = to_factors(program),
    referral = to_factors(referral),

    # ordinal, levels != labels
    # ordered with predefined levels
    education = to_factors(
      education,
      levels = c("1", "2", "3", "4", "5", "6", "7"),
      labels = c(
        "Pre-HS", # No HS diploma
        "HS", # High school diploma or equivalent
        "CND", # Some college but no degree
        "AS",
        "BS",
        "MS",
        "PhD"
      ),
      ordered = TRUE
    ),
    lego = to_factors(
      lego,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "Little to no experience",
        "Some experience",
        "Lots of experience",
        "Expert builder"
      ),
      ordered = TRUE
    ),
    manuf = to_factors(
      manuf,
      levels = c("1", "2", "3", "4"),
      labels = c(
        "No experience",
        "Taken one or more classes",
        "Held a part time or temporary position",
        "One or more year’s experience"
      ),
      ordered = TRUE
    ),
    runner = to_factors(
      runner,
      levels = c("N", "Y", "5", "10", "H", "M", "T"),
      labels = c(
        "None",
        "Yes - Unspecified",
        "5K",
        "10K",
        "Half-Marathon",
        "Marathon",
        "Triathalon"
      ),
      ordered = TRUE
    ),
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    )
  )
```

#### Numeric Data

```{r numeric-data}
nums <- demo_f |>
  select(where(is.numeric))

height_stats <- nums |>
  summarize(
    mean = mean(height, na.rm = TRUE),
    sd = sd(height, na.rm = TRUE),
    min = min(height, na.rm = TRUE),
    max = max(height, na.rm = TRUE),
    pct_missing = sum(is.na(height)) / n() * 100
  )

height_summary <- str_glue(
  "Mean height = {round(height_stats$mean, 1)}, SD = {round(height_stats$sd, 1)}, range: [{height_stats$min}, {height_stats$max}], {round(height_stats$pct_missing, 1)}% missing"
)
```

After excluding data from the pilot study, `r report_participants(nums, age = "age", participants = "participant")` were included in this analysis. Other than age, height (`r height_summary`) was the only numeric data collected. As discussed in @sec-res-challenges, height was first recorded following the pilot study, when its impact on user behavior and results was first observed. Both numeric characteristics are visualized with a combined box and scatter plot in @fig-numeric. Where present, the notches of these box plots show the 95% confidence interval around the median.

```{r fig-numeric}
#| fig-cap: "Distribution of Participant Age and Height"
#| fig-height: 3.5
#| warning: false

# Plotting age
age_plot <- demo_f |>
  filter(!is.na(age)) |>
  ggplot(aes(x = age)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    x = "Age (yr)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting height
height_plot <- demo_f |>
  filter(!is.na(height)) |>
  ggplot(aes(x = height)) +
  geom_bar(
    aes(y = ..count.. / max(..count..)),
    fill = I(base_color),
    alpha = 0.75,
    width = 0.55,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    x = "Height (in)",
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
age_plot + height_plot +
  # plot_layout(ncol = 2, widths = c(1,0.8)) +
  plot_annotation(
    title = "Numeric Sample Characteristics",
    subtitle = "Age reflects the student-oriented nature of the sample"
  )
```

#### Categorical Data

```{r categorical-data}
cats <- demo_f |>
  select(where(~ is.factor(.) && !is.ordered(.))) |>
  select(-c("referral")) |>
  mutate(across(everything(), \(x) fct_lump(x, prop = 0.1)))

# Summary table
cat_tbl <- cats |>
  select(gender:language) |>
  tbl_summary(
    label = setNames(str_to_title(names(cats)), names(cats)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

cat_names <- cats |>
  select(gender:country) |>
  names()

cats_long <- cats |>
  select(gender:country) |>
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "Value"
  ) |>
  drop_na() |>
  group_by(Category, Value) |>
  summarize(Count = n(), .groups = "drop") |>
  group_by(Category) |>
  mutate(Freq = Count / sum(Count))

country_levels <- demo_f |>
  filter(country != "USA") |>
  count(country) |>
  arrange(desc(n))
```

Of the sample's `r inline_text(cat_tbl, language, level = "English", pattern = "{N} participants, {p}% (n={n})")` reported that they mainly speak English at home. As shown in @fig-categorical, the gender composition was relatively balanced, with slightly more males `r inline_text(cat_tbl, gender, level = "Male", pattern = "(n={n}, {p}%)")` than females `r inline_text(cat_tbl, gender, level = "Female", pattern = "(n={n}, {p}%)")`. The racial makeup of the sample was predominantly White `r inline_text(cat_tbl, race, level = "White", pattern = "(n={n}, {p}%)")` and Asian `r inline_text(cat_tbl, race, level = "Asian", pattern = "(n={n}, {p}%)")`, with smaller representations from `r as.character(english::english(n_distinct(demo_f$race) - 2))` other groups. Fully `r inline_text(cat_tbl, ethnicity, level="Not Hispanic or Latino", pattern = "{p}% of participants (n={n})")` reported a non-Hispanic or Latino ethnicity. The majority of participants `r inline_text(cat_tbl, country, level = "USA", pattern = "(n={n}, {p}%)")` were from the United States, though `r as.character(english::english(n_distinct(demo_f$country, na.rm = TRUE) - 1))` other countries were represented, including `r paste(country_levels[1, ]$country, " (n=", country_levels[1, ]$n, ")", sep = "")` and `r paste(country_levels[2, ]$country, " (n=", country_levels[2, ]$n, ")", sep = "")`. `r paste(country_levels$country[3:(nrow(country_levels) - 1)], collapse = ", ")`, and `r tail(country_levels$country, 1)` were also listed as countries of origin by one participant each (n=1). All of this aligns with the study's recruitment focus.

```{r fig-categorical}
#| fig-cap: "Distribution of Categorical Characteristics"
#| fig-height: 3.5
#| results: hide

create_cat_plot <- function(data, category, base_color) {
  # specify alpha ramp end points
  min_alpha <- 0.3
  max_alpha <- 0.8

  # Filter and reorder data
  plot_data <- data |>
    filter(Category == category) |>
    mutate(
      Value = if_else(Value == "Not Hispanic or Latino",
        "Not Hispanic\nor Latino", Value
      ),
      Value = reorder(Value, Count, FUN = sum, decreasing = TRUE)
    )

  # Calculate alpha values based on Count linearly between min_alpha and max_alpha
  alpha_values <- plot_data |>
    mutate(alpha = scales::rescale(Count, to = c(min_alpha, max_alpha))) |>
    pull(alpha)

  # Set unique Value groups for setting colors and alphas
  unique_values <- unique(plot_data$Value)

  # Create the plot
  ggplot(
    plot_data,
    aes(x = Value, y = Count, fill = I(base_color), alpha = Value)
  ) +
    geom_bar(stat = "identity", position = "dodge", width = 0.8, color = outline_color) +
    scale_x_discrete(labels = function(x) tools::toTitleCase(x)) +
    labs(x = tools::toTitleCase(category), y = element_blank()) +
    scale_alpha_manual(values = setNames(alpha_values, unique_values)) +
    guides(fill = "none", alpha = "none") +
    ylim(0, max_count) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
}

# Calculate the maximum count to standardize the y-axis
max_count <- max(cats_long$Count)

# Create all plots
cat_plots <- map(
  cat_names,
  ~ create_cat_plot(cats_long, .x, base_color)
)

# Add y-axis text and ticks to first plot
first_plot <- cat_plots[[1]] +
  theme(
    axis.text.y = element_text(),
    axis.ticks.y = element_line()
  )

# Other plots without y-axis text and ticks, drop first
other_plots <- cat_plots[-1]

# combine!
all_plots <- c(list(first_plot), other_plots)

# Combine all plots
combined_plot <- reduce(all_plots, `+`) +
  plot_layout(ncol = length(all_plots), widths = c(1, 1, 0.667, 0.667)) +
  plot_annotation(
    title = "Categorical Sample Characteristics",
    subtitle = "Value Count for Gender, Race, Ethnicity, and Country"
  )

# Print the combined plot
print(combined_plot)
```

```{r make-vc-table}
vc_tbl <- cats |>
  select(lenses:worn) |>
  filter(!lenses %in% c("No", "Other", "Unknown")) |>
  filter(!(is.na(lenses) & is.na(worn))) |>
  droplevels() |>
  mutate(worn = recode(worn, "Y" = "Yes", "N" = "No")) |>
  tbl_cross(
    row = lenses,
    col = worn,
    label = list(lenses = "Lenses", worn = "Worn During Session"),
    percent = "row",
    missing_text = "NA"
  ) |>
  add_stat_label() |>
  bold_labels()
```

Of the 54 participants, 52% (n=28) reported having a vision prescription. Of those, `r inline_text(vc_tbl, row_level = "Contacts", col_level = "Total")` were for contact lenses , but `r inline_text(vc_tbl, row_level = "Glasses", col_level = "Total")` had glasses. Whereas all contact wearers reported they would wear them during the session, among glasses wearers only seven (47%) planned the same. Five (33%) did not intend to do so, and three (20%) gave no indication. This breakdown is summarized by @tbl-vc. Two participants also reported color-blindness (n=1) or other vision conditions (n=1). No participants indicated any other condition that might affect their performance.

```{r tbl-vc}
#| tbl-cap: Vision Correction Table

vc_tbl |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    # linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

Following the primary hypothesis testing, additional analyses may explore the differences in key outcomes across these demographic subgroups. That effort will aim to provide additional context and insights into potential underlying factors influencing the findings.

#### Ordinal Data

```{r ord-data}
ords <- demo_f |>
  select(where(~ is.factor(.) && is.ordered(.)))

# Summary table
ords_tbl <- ords |>
  tbl_summary(
    label = setNames(str_to_title(names(ords)), names(ords)),
    sort = list(everything() ~ "frequency")
  ) |>
  add_stat_label()

# ords_tbl

# Helper function to calculate mode
calculate_mode <- function(data, var_name) {
  count_table <- table(data[[var_name]])
  max_freq <- max(count_table)
  # pick first if more than one value has max count
  mode_value <- names(count_table[count_table == max_freq])[1]

  # Convert to an output that includes the level's name (mode_value) and its index
  mode_index <- which(levels(data[[var_name]]) == mode_value)
  return(list(value = mode_value, index = mode_index))
}

# Main function to calculate statistics and return formatted string
calculate_ord_summary_stats <- function(data, var_name) {
  # Ensure the variable is an ordered factor
  if (!is.ordered(data[[var_name]])) {
    stop("The variable must be an ordered factor.")
  }

  # Convert factor to integer codes
  factor_int <- as.integer(data[[var_name]])

  # Calculate median, mode, and IQR
  median_value <- median(factor_int, na.rm = TRUE)
  mode_results <- calculate_mode(data, var_name)
  iqr_value <- IQR(factor_int, na.rm = TRUE)

  # Format and return the result as a string
  result_string <- sprintf(
    "(Median = %.1f; Mode = %d, %s; IQR = %.1f)",
    median_value, mode_results$index, mode_results$value,
    iqr_value
  )
  return(result_string)
}
```

Unlike the data above, the categories analyzed in this section are inherently ordered. Responses for each participant's current Education level `r calculate_ord_summary_stats(ords, "education")` and their experience with Lego `r calculate_ord_summary_stats(ords, "lego")` and Manufacturing `r calculate_ord_summary_stats(ords, "manuf")` are visualized in increasing order by @fig-ordinal. The results are mostly in line with expectations based on the recruitment, though it is disappointing to see that `r inline_text(ords_tbl, lego, level = "Little to no experience", pattern = "{p}% of participants (n={n})")` claimed to have little to no experience with Lego. The great majority were undergraduate students working towards a degree `r inline_text(ords_tbl, education, level = "CND", pattern = "(n={n}, {p}%)")`, with no experience in manufacturing `r inline_text(ords_tbl, manuf, level = "No experience", pattern = "(n={n}, {p}%)")`.

```{r fig-ordinal}
#| fig-cap: "Distribution of Ordinal Characteristics"
#| fig-height: 3.5
#| results: hide

plot_ords <- function(data, var, title) {
  norm_data <- data |>
    count(var_name = {{ var }}) |>
    mutate(percent = n / sum(n) * 100)

  p <- norm_data |>
    ggplot(aes(y = var_name, x = percent)) +
    geom_bar(
      aes(alpha = percent / 100),
      fill = base_color,
      stat = "identity",
      width = 0.75,
      color = outline_color
    ) +
    scale_alpha_continuous(trans = "log10", range = c(0.2, 0.8), guide = "none") +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank()
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(size = 9, face = "bold"),
      legend.position = "none",
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
    )
  return(p)
}


educ_plot <- ords |>
  mutate(education = fct_recode(education,
    College = "CND",
    Associate = "AS",
    Bachelor = "BS",
    Master = "MS",
    Doctorate = "PhD"
  )) |>
  plot_ords(var = education, title = "Current Education Level")

lego_plot <- ords |>
  mutate(lego = fct_recode(lego,
    Little = "Little to no experience",
    Some = "Some experience",
    Lots = "Lots of experience",
    Expert = "Expert builder"
  )) |>
  plot_ords(lego, title = "Lego Experience")

manuf_plot <- ords |>
  mutate(manuf = fct_recode(manuf,
    None = "No experience",
    Training = "Taken one or more classes",
    PartTime = "Held a part time or temporary position",
    Some = "One or more year’s experience"
  )) |>
  plot_ords(manuf, title = "Manufacturing Experience")

# combine using patchwork
educ_plot / (lego_plot + manuf_plot) +
  # plot_layout(ncol = 2) +
  plot_annotation(
    title = "Ordinal Sample Characteristics",
    subtitle = "Reflects the Focus of Recruiting Efforts"
  )
```

### Multivariate Analysis

In order to assess the equivalence of treatment groups and validate the random assignment that is assumed by most statistical analysis, Age, Gender, Lego experience, and Education were compared across treatment groups. These variables were selected for their meaningful variation and potential relevance to the outcomes of interest. Other notable variables, including Ethnicity and Country of Origin were considered less critical due to their skewed distributions and small subgroup sizes.

```{r make-mv-table}

num_cols <- demographics |>
  select(age)

cat_cols <- demo_f |>
  select(c("gender", "lego", "education", "treatment"))

mv_data <- bind_cols(num_cols, cat_cols)

mv_tbl <- mv_data |>
  mutate(across(where(is.factor), ~ fct_drop(.))) |>
  mutate(
    lego = fct_recode(lego,
      "Little" = "Little to no experience",
      "Some" = "Some experience",
      "Lots" = "Lots of experience",
      "Expert" = "Expert builder"
    ),
    education = fct_recode(education,
      "College" = "CND",
      "Associate" = "AS",
      "Bachelor" = "BS",
      "Master" = "MS",
      "Doctorate" = "PhD"
    )
  ) |>
  tbl_summary(
    by = "treatment",
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    label = list(
      "age" ~ "Age",
      "gender" ~ "Gender",
      "lego" ~ "Lego",
      "education" ~ "Education"
    )
  ) |>
  add_p() |>
  # separate_p_footnotes() |>
  # add_stat_label() |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  italicize_levels() |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nN = {n} ({style_percent(p)}%)",
    p.value ~ "**p-val**"
  )

```

@tbl-demo-by-treatment shows the result of this analysis. The Kruskal-Wallis rank sum test was utilized for the numeric variable (Age) to compare distributions across groups, while Fisher's exact test was applied to the categorical and ordinal variables (Gender, Education, and Lego experience) to assess the independence of distributions from treatment assignments.

```{r tbl-demo-by-treatment}
#| tbl-cap: Participant Demographics by Treatment Group

mv_tbl |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    # linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )

```

The high p-values for Age, `r inline_text(mv_tbl, variable = age, column = "p.value")`; Gender, `r inline_text(mv_tbl, variable = gender, column = "p.value")`; Lego, `r inline_text(mv_tbl, variable = lego, column = "p.value")`; and Education, `r inline_text(mv_tbl, variable = lego, column = "p.value")` indicate that these observed characteristics are not strongly associated with treatment groupings. This result validates the group randomization effort, mitigating the influence of confounding variables, and reducing bias in the data. It is essential to the validity of most of the statistical tests employed by this study, which assume sample independence.

## $H_1$: Learning Phase Analysis

The results of the first phase of the experiment are tested with three hypotheses, each of which explore the affect of instructional method (treatment) on key measures of learning: average task completion time (TCT), rate of change of TCT, and the average number of uncorrected errors (UCE). Only completed tasks, the primary result of interest, are considered in this analysis. While excluding incomplete tasks may limit the generalizability of the findings to scenarios where time constraints are common, it ensures the analysis is based on objective, directly observed data and avoids introducing assumptions or biases that could arise from predicting results for unfinished tasks. 

```{r factor-car-results}

res_f <- car_results |>
  mutate(
    treatment = to_factors(
      treatment,
      levels = c("PWI", "PAR", "AR", "MR"),
      ordered = TRUE
    ),
    car_outcome = to_factors(car_outcome)
  )
```

```{r save-results-data}
# saving for use in broken out recall and retention analysis files
save(res_f, file = here("rdata", "res_f.RData"))
```

```{r learn-data}
learn <- res_f |>
  filter(
    phase == 1,
    car_outcome == "complete",
    # hack to remove car incorrectly labeled as complete
    uid != "1048-P1-C5"
  ) |>
  select(
    participant,
    car_elapsed_time,
    uce,
    treatment,
    drop_event_count,
    drop_event_tot_time_adj,
    repair_event_count,
    repair_event_tot_time,
    system_event_count,
    system_event_tot_time
  )

# add tot_sys_adj, tot_usr_adj, and adjusted TCT time to learn
# tot_sys_adj will be used to calculate OEE
# tot_usr_adj just to reduce TCT

learn <- learn |>
  mutate(
    # Replace NA with 0 using coalesce for safe addition
    tot_sys_adj = coalesce(drop_event_tot_time_adj, 0) +
      coalesce(system_event_tot_time, 0),
    tot_usr_adj = coalesce(repair_event_tot_time, 0),
    tot_adj = tot_sys_adj + tot_usr_adj,
    car_elapsed_time_adj = car_elapsed_time - tot_adj
  )

changes <- learn |>
  filter(car_elapsed_time != car_elapsed_time_adj)

total_tasks_completed <- nrow(learn)
total_tasks_adjusted <- nrow(changes)
```

One final step of data preparation was taken to account for drop-outs and other system events, along with time spent making repairs. This lost time was deducted from measured task times to give a more accurate record of participant performance. `r stringr::str_to_sentence(as.character(english::as.english(total_tasks_adjusted)))` of `r total_tasks_completed` completed tasks (`r round(total_tasks_adjusted / total_tasks_completed * 100, 1)`%) were affected by this adjustment.

### Descriptive Statistics for Granular Data

```{r make-report}
# report descriptive stats for cet and uce
learn_rpt <- learn |>
  rename(
    "Task Completion Time (sec)" = car_elapsed_time_adj,
    "Uncorrected Error Count" = uce
  ) |>
  select("Task Completion Time (sec)", "Uncorrected Error Count") |>
  report::report() # options not working!

rpt_table <- as.report_table(learn_rpt)
cv_tct <- rpt_table$SD[1] / rpt_table$Mean[1]
cv_uce <- rpt_table$SD[2] / rpt_table$Mean[2]
```

`r learn_rpt`

These statistics provide a high-level summary of performance across all tasks and participants, where each observation corresponds to a single repetition of the Learning task. Both Task Completion Time (TCT) and Uncorrected Error Count (UCE) show considerable variance, as depicted in @fig-learn.

```{r fig-learn}
#| fig-cap: "Distribution of Task Completion Time and Error Rates in Learning"
#| fig-height: 3.5
#| warning: false

# Plotting adjusted car_elapsed_time
cet_plot <- learn |>
  ggplot(aes(x = car_elapsed_time_adj)) +
  geom_histogram(
    aes(y = ..density.. * 65), # hack for relative size / scale
    binwidth = 10,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    notch = TRUE
  ) +
  labs(
    title = "Task Completion Time (sec)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# Plotting Uncorrected Error Counts
uce_plot <- learn |>
  ggplot(aes(x = uce)) +
  geom_histogram(
    aes(y = ..density.. * 1.5), # hack for relative size / scale
    binwidth = 1,
    fill = base_color,
    alpha = 0.75,
    color = outline_color
  ) +
  geom_boxplot(
    width = 0.075,
    position = position_nudge(y = -0.07),
    # notch = TRUE
  ) +
  labs(
    title = "Number of Uncorrected Errors per Task",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 9, face = "bold"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

# combine using patchwork
cet_plot / uce_plot +
  # plot_layout(ncol = 2, widths = c(1,1)) +
  plot_annotation(
    title = "Learning Phase - Observed Measures",
    subtitle = "Error Count Exhibits Greater Variability than Task Completion Time"
  )
```

Compared to TCT, with a coefficient of variation (CV) of `r round(cv_tct, 2)`, UCE shows substantially greater spread (CV=`r round(cv_uce, 2)`) and the presence of a floor effect. The later could potentially suggest limitations in the task's ability to capture a full range of participant performance. Further investigation is required to determine if this is due to task complexity or other factors.

### Aggregated Data

```{r agg-results}
# agg participant results
learn_agg <- learn |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    tot_cars = n(), # Count the number of tasks per participant
    avg_tct_adj = mean(car_elapsed_time_adj, na.rm = TRUE),
    avg_uce = mean(uce, na.rm = TRUE)
  ) |>
  ungroup()
```

@tbl-agg-results-by-treat presents participant-level metrics, including the number of tasks completed along with the average TCT and UCE per task. In addition to the overall results, the data is broken out by assigned treatment group to allow for comparisons across conditions. Note that the statistics presented are the overall (aka "grand") mean and median values. That is, they are the average or middle values of all individual participant averages. This provides overall summary statistics for each treatment group that reflect the central tendency and variability among the average outcomes reported by participants within each group.

```{r tbl-agg-results-by-treat}
#| tbl-cap: Avg. Participant Performance Data Summarized by Treatment Group

t <- learn_agg |>
  select(-participant) |>
  tbl_summary(
    type = all_continuous() ~ "continuous2",
    digits = all_continuous2() ~ 1,
    label = list(
      avg_tct_adj ~ "Avg TCT per Task (sec)",
      avg_uce ~ "Avg UCE per Task",
      tot_cars ~ "Number of Tasks Completed"
    ),
    by = treatment,
    statistic = list(
      all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} ({p25}, {p75})"
      )
    )
  ) |>
  # bold_labels() |>
  # add_p() |>
  # separate_p_footnotes() |>
  add_overall() |>
  # bold_p(t = 0.05) |>
  modify_header(
    all_stat_cols() ~ "**{level}**\nn = {n} ({style_percent(p)}%)",
    # p.value ~ "**p-val**"
  )

t |>
  modify_footnote(update = everything() ~ NA) |>
  as_kable_extra(
    booktabs = TRUE,
    escape = FALSE,
    # linesep = "\\addlinespace",
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down"),
    font_size = 11
  )
```

See @fig-perf-by-treat for a depiction of the grand means for all variables, with 95% confidence intervals for each. Observe that the previously noted variability and positive skew are also present in the grouped data, though both are less pronounced due to averaging effects. It appears that the PWI treatment group may have prioritized speed over accuracy.

```{r make-perf-data}
# Reshape data to long format
long_data <- learn_agg |>
  pivot_longer(
    cols = c(avg_tct_adj, tot_cars, avg_uce),
    names_to = "metric",
    values_to = "value"
  )

# Calculate means, sample size, and SE for each metric within each treatment
metrics_stats <- long_data |>
  group_by(treatment, metric) |>
  summarize(
    mean_value = mean(value, na.rm = TRUE),
    n = n(), # sample size
    se = sd(value, na.rm = TRUE) / sqrt(n()), # standard error
    .groups = "drop"
  )

# Find the highest mean for each metric across all treatments
max_means <- metrics_stats |>
  group_by(metric) |>
  summarize(max_mean = max(mean_value), .groups = "drop")

# Normalize means and calculate CI within their metrics
normalized_metrics_stats <- metrics_stats |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = mean_value / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  )

# Calculate overall mean, SE, and normalize
overall_stats <- long_data |>
  group_by(metric) |>
  summarize(
    overall_mean = mean(value, na.rm = TRUE),
    n = n(),
    se = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
  left_join(max_means, by = "metric") |>
  mutate(
    normalized_mean = overall_mean / max_mean,
    lower_ci = normalized_mean - qt(0.975, df = n - 1) * (se / max_mean),
    upper_ci = normalized_mean + qt(0.975, df = n - 1) * (se / max_mean)
  ) |>
  mutate(treatment = "ALL")

# Combine with other stats
final_stats <- bind_rows(normalized_metrics_stats, overall_stats)

# Factorize the 'treatment' column
final_stats$treatment <- factor(final_stats$treatment, levels = c("ALL", "PWI", "PAR", "AR", "MR"))

# Factorize and label 'metric' column
final_stats$metric <- factor(
  final_stats$metric,
  levels = c("avg_tct_adj", "avg_uce", "tot_cars"),
  labels = c("Average Task Time", "Average Error Count", "Tasks Completed")
)

# Predefined alpha values by treatment
alpha_values <- c(0.4, 0.55, 0.7, 0.85, 1.0)
names(alpha_values) <- levels(final_stats$treatment)
```


```{r fig-perf-by-treat}
#| fig-cap: "Aggregated Learning Performance Data by Treatment"
#| fig-height: 3.5
#| warning: false

ggplot(final_stats, aes(x = metric, y = normalized_mean, fill = treatment, label = treatment)) +
  geom_col(
    aes(alpha = normalized_mean),
    position = position_dodge(0.9),
    color = outline_color
  ) +
  geom_errorbar(
    aes(ymin = lower_ci, ymax = upper_ci),
    position = position_dodge(0.9),
    width = 0.25,
    alpha = 0.7
  ) +
  geom_text(
    aes(y = 0),
    position = position_dodge(width = 0.9),
    vjust = 1.65, # Adjust based on visual preference
    size = 3,
    alpha = 0.9
  ) +
  scale_fill_manual(values = rep(base_color, length(unique(final_stats$treatment)))) +
  scale_alpha_continuous(transform = "pseudo_log", range = c(0.3, 1.0)) +
  labs(
    title = "Comparison of Normalized Performance Metrics by Treatment",
    subtitle = "Includes Overall and Treatment Groups with 95% Confidence Intervals",
    y = element_blank(),
    x = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(
      margin = margin(b = 5, unit = "pt"),
      color = "black"
    )
  )

```

#### Assess outliers

Given the level of variation and spread seen above, the presence of outliers was analyzed. Participant-level outliers within treatment groups were the focus, rather than individual data points across the entire dataset. This method identifies annomolous participant level contributions to treatment variability.

Four different tests were used to assess each participant as an outlier in their treatment group: (1) Tukey's Fences at 1.5 times the IQR, (2) 2.5 times the Z-score, (3) outside the 5th and 95th percentiles, and (4) using the Mahalanobis distance. TCE and UCE were each tested separately using the first three methods. The fourth method is a multivariate test that considers TCE and UCE together. The results are summarized in @tbl-outlier-summary, for all participants that were identified by two or more tests.

```{r find-outliers}
# Function to detect outliers in a vector using Z-score
check_zscore <- function(x, threshold = 2.5) {
  z_scores <- (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
  abs(z_scores) > threshold
}

# Function to detect outliers in a vector
check_iqr <- function(x, mult) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower_bound <- q1 - mult * iqr
  upper_bound <- q3 + mult * iqr
  x < lower_bound | x > upper_bound
}

# Function to calculate Mahalanobis distance for a group
compute_mahalanobis <- function(df, cols) {
  cov_matrix <- cov(df[, cols], use = "complete.obs")
  center <- colMeans(df[, cols], na.rm = TRUE)
  mahal_dist <- mahalanobis(df[, cols], center, cov_matrix)
  return(mahal_dist)
}

outlier_summary <- learn_agg |>
  group_by(treatment) |>
  mutate(
    # find 1.5 and 3x IQR outliers using Tukey method
    out15_tct = check_iqr(avg_tct_adj, 1.5),
    out15_uce = check_iqr(avg_uce, 1.5),
    zs_25x_tct = check_zscore(avg_tct_adj, threshold = 2.5),
    zs_25x_uce = check_zscore(avg_uce, threshold = 2.5),
    # find by 5th and 95th percentile
    pct_tct = avg_tct_adj < quantile(avg_tct_adj, 0.05, na.rm = TRUE) |
      avg_tct_adj > quantile(avg_tct_adj, 0.95, na.rm = TRUE),
    pct_uce = avg_uce < quantile(avg_uce, 0.05, na.rm = TRUE) |
      avg_uce > quantile(avg_uce, 0.95, na.rm = TRUE),
    # find by Mahalanobis distance
    mahal_dist = compute_mahalanobis(
      pick(avg_tct_adj, avg_uce),
      c("avg_tct_adj", "avg_uce")
    ),
    md_out = mahal_dist > qchisq(0.975, df = 2)
  ) |>
  ungroup() |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    out15_tct = sum(out15_tct),
    out15_uce = sum(out15_uce),
    zs25_tct = sum(zs_25x_tct),
    zs25_uce = sum(zs_25x_uce),
    pct_tct = sum(pct_tct),
    pct_uce = sum(pct_uce),
    md_out = sum(md_out),
    .groups = "drop"
  )
```


```{r tbl-outlier-summary}
#| tbl-cap: Summary of Participants Identified by Two or More Outlier Tests
#| warning: false

outlier_summary |>
  rowwise() |>
  mutate(tot_out = sum(c_across(-c(participant, treatment)))) |>
  ungroup() |>
  filter(tot_out > 1) |>
  arrange(desc(tot_out)) |>
  select(participant, treatment, tot_out) |>
  inner_join(learn_agg) |>
  gt()

```

Further investigation of these four participants showed that only \#1063 experienced systematic problems that might warrant excluding their performance. The others were just poor performers, with unusually high overall TCT and/or UCE. Given the limited amount of data collected per treatment, it was decided to retain all but 1063's data for hypothesis testing, where additional outlier detection steps may be taken.

```{r drop-learn-outliers}
# drop 1063
learn_agg_out <- learn_agg |> filter(participant != 1063)
```

```{r save-learn-data}
# save learn objects with types, etc. as binary data file for external analysis
save(
  learn,
  learn_agg_out,
  outline_color,
  base_color,
  file = here("rdata", "learn.RData")
)
```


### $H_{1a}$: Average Time per Car

The first hypothesis of the learning phase:
$$H_{1a}\textrm{: Average task completion time varies with treatment}$$

will be tested by comparing the average TCT of each treatment group to understand the magnitude, direction, and significance of differences.

#### Applicable Statistical Methods

Methods based on one-way ANOVA, including Fisher's or Welch's parametric methods and their non-parametric equivalent, Kruskal-Wallis' test by ranks, are commonly used for such comparisons. All assume observations are independent and treatment effects are additive. The use of aggregated data reduces each participant to a single observation, addressing independence. The experimental design generally ensures the treatments themselves are additive, with independent effects on the response.

Parametric methods also require a normal distribution of data *within each treatment group*. As noted in the original assessment of the learning data, the overall TCT dataset exhibits skewness and kurtosis, suggesting deviations from a normal distribution. This must be revisited with group wise testing of the aggregated data. By averaging the observations and eliminating an outlier, the grouped data may better resemble a normal distribution.

#### Check Model Assumptions

A combination of quantitative and qualitative analysis is required to accurately assess normality of each treatment group. Statistical tests provide a formal but imperfect measure of the data's shape. Visual confirmation of the characteristic quantile-quantile (Q-Q) plot provides further support for the claim. @tbl-learn-norm-tct tabulates the results of the D'Agostino skewness test, Anscombe-Glynn kurtosis test, and the Shapiro-Wilk normality test. In each case, the null hypothesis ($H_0$) for these tests assumes that the data follows the characteristics of a normal distribution. Therefore, low p-values indicate evidence against $H_0$, suggesting that the data deviates from normality in the tested aspect.

```{r def-tab-norm}
# Function to tabulate normality test results for each treatment group
tabulate_normality_tests <-
  function(data, treatment_col, value_col) {
    levels <- levels(data[[treatment_col]])
    all_levels <- c("ALL", levels)

    results <- map(all_levels, function(level) {
      if (level == "ALL") {
        data_subset <- data |> pull(!!sym(value_col))
      } else {
        data_subset <-
          data |>
          filter((!!sym(treatment_col)) == level) |>
          pull(!!sym(value_col))
      }

      if (length(data_subset) > 0) {
        # Ensure there's data to process
        skew_test <- moments::agostino.test(data_subset)
        kurtosis_test <- moments::anscombe.test(data_subset)
        normality_test <- shapiro.test(data_subset)

        tibble(
          Group = level,
          Skewness = paste(
            round(skew_test$statistic[1], 2),
            round(skew_test$p.value, 3),
            sep = ", "
          ),
          Kurtosis = paste(
            round(kurtosis_test$statistic[1], 2),
            round(kurtosis_test$p.value, 3),
            sep = ", "
          ),
          Normality = paste(
            round(normality_test$statistic, 2),
            round(normality_test$p.value, 3),
            sep = ", "
          )
        )
      } else {
        tibble(
          Group = level,
          Skewness = NA,
          Kurtosis = NA,
          Normality = NA
        )
      }
    }) |> bind_rows()

    # Ensure the desired order, converting Group to a factor
    results <- results |>
      mutate(Group = factor(Group, levels = c("ALL", "MR", "AR", "PWI", "PAR"))) |>
      filter(!is.na(Group)) |>
      arrange(Group)

    results
  }
```

```{r tbl-learn-norm-tct}
#| tbl-cap: "Summary of TCT Test Statistics, p-Values for Normality Assessment"
#|
tct_norm <- tabulate_normality_tests(learn_agg_out, "treatment", "avg_tct_adj")
tct_norm |> gt()
```

From this we see that only PWI fails the normality test (p=0.021), despite showing marginally non-significant skewness (p=0.058). Similarly, MR exhibits significant positive skew (p=0.021) but the Shapiro-Wilk test gives evidence of normality (p=0.084). Both AR and PAR are consistent in confirming normality, though PAR's slightly positive skew is marginally non-significant (p=0.068).

To reconcile these findings for MR and PWI, and confirm those for AR and PAR, the set of four Q-Q plots in @fig-qq-h1a were generated. Observations for normal data should fall along or near the line, within the shaded confidence interval. These plots show that PAR and AR both meet the expectations of normality. The MR group mostly aligns with the line, except for a single point in the right tail. This aligns with the skewness and kurtosis test results and provides insufficient evidence to reject normality. On the other hand, the PWI group diverges substantially from the line, with several points leaving the confidence interval. This supports the statistical evidence that the PWI data is not normally distributed.

```{r fig-qq-h1a}
#| fig-cap: Q-Q Plots of Learning Times by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(learn_agg_out, x_var = "avg_tct_adj", by_var = "treatment")

p + plot_annotation(
  title = "Quantile-Quantile Plots of TCT Treatment Data",
  subtitle = "PWI Data is the Most Uncharacteristic of Normality"
)

```

Given these deviations from normality, particularly for the PWI group, a non-parametric approach was adopted. This provides a reliable and straightforward approach for hypothesis testing for data that are non-normal or heteroscedastic (have unequal variances). This decision also ensures robust statistical results in the scale of the original data, easing interpretation.

#### Analysis

@fig-h1a-plot shows the resulting comparison of average TCT within each treatment group. It combines elements of violin, scatter, and box plots to provide a comprehensive portrayal of the data. Treatment differences were analyzed using the Kruskal-Wallis test by ranks. Its test statistic, $H$, approximately follows a $\chi^{2}_{k-1}$ distribution, where $k$ is the number of groups. Here, $\chi^{2}_{Kruskal-Wallis}(3)=25.9$, with a p-value of less than 0.001. These results indicate a statistically significant difference between groups. The effect size of this difference is $\hat{\varepsilon}^{2}_{rank}=0.5$, with $CI_{95\%}\:[0.4,1.0]$, indicating a moderate to large treatment effect where 50% of the variance in TCT can be explained by the difference in groups. Together, these results provide strong evidence of overall differences in TCT that are both statistically and practically significant.

```{r fig-h1a-plot}
#| fig-cap: Task Completion Time Differences by Treatment
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

h1a_plot <- ggstatsplot::ggbetweenstats(
  data = learn_agg_out,
  x = treatment,
  y = avg_tct_adj,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "TCT Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between 'Fast' (PWI, PAR) and 'Slow' (AR, MR) Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 2,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

h1a_plot
```

Pairwise comparisons were conducted post-hoc using Dunn's non-parametric test for Kruskal-type ranked data. Holm's adjustment for multiple comparisons was preferred over the often cited Bonferroni method, which tends to be too conservative. The results show that PWI and PAR are both significantly faster than AR and MR, with all p-values less than 0.01.

```{r test-power-sim}
k <- length(unique(learn_agg_out$treatment))
n <- round(nrow(learn_agg_out) / k)
epsilon_sq <- 0.5

# Estimate post-hoc power
posthoc_power <- posthoc_power_analysis(n, k, epsilon_sq, sig_level = 0.05, num_simulations = 1000)
```

A post-hoc simulation was run to validate these results by estimating the statistical power. Power analysis measures the effectiveness of a statistical test in detecting true differences when they exist. High power (typically 0.8 or above) suggests that the test is appropriate for the given data and experimental conditions, therefore supporting factual decision-making. The simulation repeatedly applies the Kruskal-Wallis test to data generated based on the observed mean and effect size, and resulting power is the proportion of tests that correctly reject the null hypothesis. The outcome of `r posthoc_power` (`r posthoc_power * 100`% power), while uncommon, suggests that, for the conditions of this experiment (n=53, an average of 13.25 samples in each of the four groups, $\hat{\varepsilon}^{2}=0.5$, and $\alpha = 0.05$), the test is extremely effective at detecting the observed differences. The combination of a large effect size and a reasonable sample size is largely attributable to this value, which offers great confidence in the validity of the test results.

#### $H_{1a}$ Result

Based on analysis of the data and validation of the methods, it is determined that sufficient statistical evidence exists to accept $H_{1a}$: average task completion time varies by treatment. Specifically, the PWI and PAR instructional methods are both equivalent and faster than AR and MR, which are also equivalent to one another.


\pagebreak

### $H_{1b}$: Learning Rates

```{r prep-h1b-data}
# disaggregated data required for rate of change of tct
# create data
learn_h1b <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj)

# mutate
learn_h1b <- learn_h1b |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    # treatment = factor(as.character(treatment))
  ) |>
  ungroup() |>
  select(participant, treatment, seq, tct, uce)

# glimpse(learn_h1b)
```

The second hypothesis of the learning phase:
$$H_{1b}\textrm{: Learning rates vary with treatment}$$

will be tested by comparing how TCT changes with each repetition of the task based on treatment. Decreased TCT is used as a measure of the learning effect, so a change in TCT over time is a proxy for learning rate.

#### Applicable Statistical Methods

It is tempting to assess this using aggregated data as before. For example, the average change in TCT between consecutive tasks completed might give suitable measure. For clarity, this can be expressed as follows for each participant, $j$:

$$
\small{
\overline{\Delta TCT_{j}} = \frac{\sum_{i=1}^{N-1} (TCT_{i+1} - TCT_i)}{N - 1}
}
$$ {#eq-naive}

where $N$ is the total number of tasks completed by participant $j$, and $i$ indexes each of the $N-1$ differences in TCT. This metric can be compared by group using the approach employed by $H_{1a}$, as seen in @fig-h1b-naive. The results show that there is a significant difference in groups (p = 0.002), with small to moderate effect size $\hat{\varepsilon}^{2}_{rank}=0.29$. Significant pairwise differences were identified between PWI and both AR (p=0.027) and MR (p=0.039) as well as PAR and the same (AR p=0.015, MR p=0.025). 

```{r agg-h1b-data}
pd <- learn_h1b |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    num_tasks = n(),
    avg_change = mean(diff(tct))  # This calculates the average of consecutive differences
  ) |>
  ungroup()

```


```{r fig-h1b-naive}
#| fig-cap: Naive method
#| fig-height: 6
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

p <- ggstatsplot::ggbetweenstats(
  data = pd,
  x = treatment,
  y = avg_change,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "Avg Change in TCT per Task by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 3,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

p

```

While this method offers initial insights and highlights significant differences in learning rates across treatment groups, it has crucial limitations. Aggregating data by participant and within treatments masks both between and within-participant variability, discarding important nuances in learning rates. By ignoring the repeated measures nature of the data it fails to capture how learning rate changes over time, and how that differs between individuals. It also does not account for imbalanced repeated measures, which may lead to biased estimates. By reducing the number of observations, it may also limit the statistical power of the test. Finally, this method does not account for the non-linear nature of change in TCT over time, further limiting the validity of these findings. These simplifying assumptions may be reasonable for a gross estimate of overall differences, e.g., average TCT, but they are not appropriate for an in-depth analysis of the rate and change of learning in the presence of individual and treatment effects.

For analysis of this sort, particularly where repeated measures are nested within each group and participant variation is an important consideration, mixed-effects models are often used. These models provide a thorough and accurate analysis by accounting for individual differences in learning rates and leveraging all available repeated measures data. This approach preserves data granularity, explicitly models the dynamic aspect of learning, and is specifically designed to handle imbalanced data. It increases the statistical power and flexibility to detect effects and interactions in a complex data structure, and it can also account for non-linear response data. Overall, mixed-effects models provide a robust framework for examining the interaction between treatment and learning effect over time.

The general form of a suitable linear mixed-effects model that expresses TCT as a function of task sequence, ignoring treatment effects, is:

$$
\small{
TCT_{ij}=(\beta_0+U_{0j})+(\beta_1+U_{1j})(SEQ_{ij})+\varepsilon_{ij}
}
$$ {#eq-lme1}

where:

- $i$ and $j$ are the indices for each observation and participant, respectively.
- $\beta_0$ is the overall intercept, a baseline value of $TCT$.
- $U_{0j}$ is the participant variation on the intercept.
- $\beta_1$ is the overall slope.
- $U_{1j}$ is the participant variation on the slope.
- $SEQ_{ij}$ is the time value associated with each observation $i$ of participant $j$, i.e. the task number.
- $\varepsilon_{ij}$ represents the residuals for each observation---variation not accounted for by the model.

This formulation allows for random intercepts, $(\beta_0+U_{0j})$, and slopes $(\beta_1+U_{1j})$, each based on individual differences per participants $U_j$. In the context of this analysis those can be interpreted as individual starting points (intercept) and learning rates (slope). While @eq-lme1 makes it easy to understand how participant-level variation contributes to both slope and intercept, it is more commonly formulated as:

$$
\small{
TCT_{ij} = (\beta_0 + \beta_1 SEQ_{ij}) + (U_{0j} + U_{1j}SEQ_{ij}) + \varepsilon_{ij}
}
$$ {#eq-lme2}

where the terms are rearranged to differentiate *fixed* and *random* effects. The first term corresponds to group-level fixed-effects ($\beta_i$) while the second captures the random-effects of between- and within-participant variation ($U_j$). This is visualized in @fig-h1b-rsri, where the completion time for each task iteration is shown in all treatment groups. Black lines indicate the best fit within each group, the fixed-effect term, and colored lines show the same for each participant, reflecting the random-effect term.

```{r fig-h1b-rsri}
#| fig-cap: Random slopes and intercepts model visualized.
#| fig-height: 4
#| message: false

ggplot(learn_h1b, aes(x = seq, y = tct)) +
  geom_point(
    aes(fill = as.factor(participant)),
    pch = 21,
    size = 1,
    stroke = 0.5
  ) +
  stat_smooth(
    aes(group = participant, col = as.factor(participant)),
    se = FALSE,
    method = "lm",
    lwd = 0.5
  ) +
  stat_smooth(
    aes(group = treatment),
    col = "black",
    se = FALSE,
    method = "lm",
    lwd = 1
  ) +
  facet_wrap( ~ treatment) +
  scale_x_continuous(breaks = c(0:12)) +
  labs(
    title = "Completion Time in Seconds for each Completed Task",
    subtitle = "Linear fits for participants (color) vary from group averages (black)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_bw(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
    )
```

Using linear mixed effect models (LMEs) to assess treatment effect is a two-step process. First, a model using task number as the only predictor is fit to the data, including both fixed and random effects. Because this model is independent of the predictor of interest (treatment), it is referred to as the *unconditioned* model. Once selected and validated, this acts as a baseline for comparison with more complex models that incorporate the treatment effects. These so-called *conditioned* models are compared with the baseline to see which, if any, improve the model fit. Finally, the parameters of the model that best explains the data are interpreted to quantify the relationships between time, treatment, and the response. The following sections will elaborate on each step of this process.

#### Unconditional Models of Time

To find an unconditional model that best fits the data, several options of increasing complexity will be considered. In the model selection process, it is important to consider that learning is an inherently non-linear process. Typically, TCT starts high and decreases with each training iteration, but the rate of decrease is not constant. The pattern often resembles exponential decay, with rapid initial improvement followed by more gradual changes. When working with an LME, quadratic terms can be used to create a curvilinear approximation of this effect. This is depicted in @fig-h1b-qrsri, where each line is fit using a quadratic term for sequence: $TCT \approx \beta_{0} + \beta_{1}SEQ + \beta_{2}SEQ^2$. The result appears provide an overall improvement to fit at both participant and treatment levels compared to the previous linear formulation.

```{r fig-h1b-qrsri}
#| fig-cap: Quadratic random slopes and intercepts model visualized.
#| fig-height: 4
#| warning: false

learn_h1b |>
  # filter(treatment == "MR") |>
ggplot(aes(x = seq, y = tct)) +
  geom_point(
    aes(fill = as.factor(participant)),
    pch = 21,
    size = 1,
    stroke = 0.5
  ) +
  stat_smooth(
    aes(group = participant, col = as.factor(participant)),
    formula=y~x+I(x^2),
    se = FALSE,
    method = "lm",
    lwd = 0.5
  ) +
  stat_smooth(
    aes(group = treatment),
    formula=y~x+I(x^2),
    col = "black",
    se = FALSE,
    method = "lm",
    lwd = 1
  ) +
  facet_wrap( ~ treatment) +
  scale_x_continuous(breaks = c(0:12)) +
  labs(
    title = "Completion Time in Seconds for each Completed Task",
    subtitle = "Curvilinear fits for participants (color) vary from group averages (black)",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_bw(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "none",
    )
```

Seven models of increasing complexity were fit, as summarized in @tbl-h1b-fit-uncond. This table adopts the model formulation notation used by many R modeling packages, including `lmer`, which is used here. A fixed overall intercept is specified for all models with the first term (`1`). Random effects are described within the parentheses of the last term, where terms to the right of the bar (`|`) are grouping variables, and those to its left specify the random intercept and coefficients. Terms between the intercept and random effects, outside the parentheses, are fixed effects with constant coefficients. For brevity, Sequence and Participant terms have been abbreviated in this table as `s` and `p`, respectively. The "XF" column describes the non-linear transformation employed (quadratic or log), and "SE" identifies the sequence effect. SE is "Fixed" where SEQ is included only as a fixed effect, and "Both" where it also appears in the random effect.

| **\#** | **Model Formulation in `{lmer}` Notation**       | **XF** |  **SE**   |
| :----: | ------------------------------------------------ | :----: | :-------: |
|   1    | 1 + (1 \| p)                                     |  None  |   None    |
|   2    | 1 + s + (1 \| p)                                 |  None  |   Fixed   |
|   3    | 1 + s + (1 + s \| p)                             |  None  |   Both    |
|   4    | 1 + s + s^2 + (1 + s + s^2 \| p)                 |  Quad  |   Both    |
|   5    | 1 + log(s + 1) + (1 \| p)                        |  Log   |   Fixed   |
|   6    | 1 + log(s + 1) + (1 + log(s + 1) \| p)           |  Log   |   Both    |
|   7    | 1 + s + log(s + 1) + (1 + s + log(s + 1) \| p)   |  Log   |   Both    |

: Model Summary {#tbl-h1b-fit-uncond .striped}

For example, model 4 has a fixed overall intercept (`1`) with additional fixed effects for `s` and `s^2`. Its random effects (`1 + s + s^2 \| p`) can be interpreted as random intercepts and slopes for `s` and `s^2` for each participant. This can be expressed in the previous mathematical notation as:

\begin{align}
\small
TCT_{ij} =\:  &(\beta_{0} + \beta_1 SEQ_{ij} + \beta_2 SEQ_{ij}^2) \label{eq:h1b-m4} \\
&+ (U_{0j} + U_{1j} SEQ_{ij} + U_{2j} SEQ_{ij}^2) + \varepsilon_{ij} \notag
\end{align}

```{r h1b_mod_data}

mdf <- learn_h1b |>
  filter(seq <= 5) |> 
  mutate(
    seq.fp = (seq - 1) / (max(seq) - 1),
    treatment = factor(treatment, ordered = FALSE),
  )

```

Prior to fitting, the sequence predictor was transformed from discrete integers to continuous values in the range [0, 1]. This is common practice to prevent numerical issues, aid model convergence, and to make the results easier to interpret and generalize. Additionally, The treatment variable was converted to an unordered factor to remove any implied ranking among treatments that could bias the models. Treatment was originally ordered to ensure the PWI, PAR, AR, MR presentation of results in charts and tables. It has no predictive implications.

Finally, only a subset of the data is used. As seen in @fig-h1b-rsri and @fig-h1b-qrsri, the number of task iterations completed in the allotted time varied substantially by treatment. Numerous attempts to fit the full dataset using various modeling techniques all resulted in significant overfitting, particularly after the fifth iteration where available observations decline rapidly and treatments become imbalanced (see @tbl-h1b-obs). The overfit models produced poor extrapolations that were not reflected in performance diagnostics, as these measures only compare observed and predicted values. To address this, it was decided to use data only from the first five task iterations, where all treatments are still reasonably represented. This cutoff was chosen as the best trade-off between available data and extrapolation quality, after considering earlier (four) and later (six) alternatives.

```{r tbl-h1b-obs}
#| tbl-cap: "Observations per Treatment for all recorded task iterations"
table(learn_h1b$treatment, learn_h1b$seq)
```

Following those data transformations, the models were fit using maximum likelihood estimation (rather than the default restricted maximum likelihood) to enable valid comparisons between models with different fixed effects structures. Fitting was accomplished using the overloaded version of `lme4::lmer` from `{lmerTest}`, which adds p-values for fixed effects using Satterthwaite's method. The results were compared using `performance` from `{compare_performance}`, the output of which is summarized in @tbl-h1b-modperf-unc. 

Columns include Akaike (AIC) and Bayesian (BIC) Information Criterion, conditional and marginal $R^2$, Interclass Correlation Coefficient (ICC), and Root Mean Squared Error (RMSE). AIC and BIC are, respectively, frequentist and Bayesian measures of the amount of information lost by a model when approximating the true data generating process. Both account for the number and relevance of predictors, and lower scores are better. $R^2_{cond}$ represents the variance explained by both fixed and random effects, while $R^2_{marg}$ only accounts for the fixed effects. ICC measures the proportion of total variance that is accounted for by the grouping of data. In this context, high ICC values suggest that a larger portion of the variability in TCT is due to differences *between* participants, not within (i.e., across task iterations). Finally, RMSE is a measure of the average magnitude of the model's prediction error (residuals).

```{r h1b-fit-uncond}
#| warning: false

# random intercept
m1 <- lmer(tct ~ 1 + (1 | participant), data = mdf, REML = F)
# fixed slope, random intercept
m2 <- lmer(tct ~ 1 + seq.fp + (1 | participant), data = mdf, REML = F)
# random slope, random intercept
m3 <- lmer(tct ~ 1 + seq.fp + (1 + seq.fp | participant), data = mdf, REML = F)
# quadratic random slope, random intercept
m4 <- lmer(
    tct ~ 1 + seq.fp + I(seq.fp ^ 2) + (1 + seq.fp + I(seq.fp ^ 2) | participant),
    data = mdf, REML = F
)
# log-transformed time; fixed slope, random intercept
m5 <- lmer(tct ~ 1 + log(seq.fp + 1) + (1 | participant), data = mdf, REML = F)
# log-transformed time; random slope, random intercept
m6 <- lmer(tct ~ 1 + log(seq.fp + 1) + (1 + log(seq.fp + 1) | participant), data = mdf, REML = F)
# linear and log-transformed time; fixed slope, random intercept
m7 <- lmer(
    tct ~ 1 + (seq.fp) + log(seq.fp + 1) + (1 + (seq.fp) + log(seq.fp + 1) | participant),
    data = mdf, REML = F
)
```


```{r h1b-fit-uncond-load}
cp_h1b_uc <- compare_performance(m1, m2, m3, m4, m5, m6, m7, metrics = "common", rank = F)
```

```{r tbl-h1b-modperf-unc}
#| tbl-cap: "Model Performance Results"

cp_h1b_uc |>
  select(Name, AIC, BIC, R2_conditional, R2_marginal, ICC, RMSE) |>
  display()
```

The result of this comparison shows that m7 outperforms the other options in most measures, with the lowest AIC, BIC, and RMSE. Approximately 95% of the variance in TCT can be explained by this model ($R^2_{cond}$), of which only 28% is due to its fixed effects ($R^2_{marg}$). The high ICC (0.93) indicates that a large proportion of the total variance is due to differences between participants, suggesting that individual variation is a crucial factor in the model. The resulting predictions are plotted in @fig-m7-pred.

```{r fig-m7-pred}
# Create a data frame for predictions
pred_data <- data.frame(seq.fp = seq(0, 1, length.out = 100))

# Generate predictions
pred_data$TCT <- predict(m7, newdata = pred_data, re.form = NA)

# Calculate confidence interval
vcov_fixed <- vcov(m7)
X <- model.matrix(~ seq.fp + log(seq.fp + 1), data = pred_data)
se <- sqrt(diag(X %*% vcov_fixed %*% t(X)))
pred_data$lower <- pred_data$TCT - 1.96 * se
pred_data$upper <- pred_data$TCT + 1.96 * se

# Create the plot
ggplot(pred_data, aes(x = seq.fp, y = TCT)) +
  geom_point(data = mdf, aes(y = tct), alpha = 0.3) +  # Add actual data points
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  geom_line(color = "blue") +
  labs(title = "Predicted Learning Curve from Unconditioned Model (m7)",
       x = "Standardized Task Sequence (SC)",
       y = "Predicted Task Completion Time (seconds)") +
  theme_minimal()
```

Even with the subset of data, we see that the log-based model begins to exhibit a slight upward trend after the fourth iteration, which defies learning theory, expectations, and the limited data from subsequent iterations. Model fits based on exponential and hyperbolic forms, which better align with theoretical learning curves, led to more promising predictions but were plagued by convergence issues, especially as additional terms were added to account for treatment and its interactions. This is a limitation of the study that originates primarily from the imbalanced experimental design employed by the learning phase.

Inspecting the fixed effects parameters for m7, summarized in @tbl-m7-params, shows that all terms are significant. The intercept of 145.5 seconds establishes the baseline average performance for all participants on the first iteration. Estimated coefficients for the linear and log terms of sequence show that these terms compete for dominance, with the linear term increasing TCT at a rate of 246.5 seconds per unit of seq.fp, the continuous transformation of task iteration number. This is offset by the log term, which decreases TCT by 431.7 seconds per unit of $\log{(\text{seq.fp} + 1)}$. These coefficients represent the change in TCT for a one-unit change in their respective transformed sequence terms, resulting in a non-linear learning curve when combined. The strong negative correlation of -0.75 between the log term and intercept indicates that participants with higher initial TCT experience a stronger effect from the log term, which drives rapid initial improvement. In learning theory, this phenomenon is known as the *power law of practice*. It formalizes the idea that, all other things constant, the farther a learner starts from expected proficiency, the faster they will initially improve.

```{r tbl-m7-params}
#| tbl-cap: Fixed Effects Parameters for the Unconditioned RLM (m7)
display_model_parameters(m7)
```

```{r}
residuals_m7 <- residuals(m7)

# Shapiro-Wilk test for normality
m7_st <- shapiro.test(residuals_m7)
m7_st_pv <- m7_st$p.value

# Custom function for heteroscedasticity test
hetero_test <- function(model) {
  res <- residuals(model)
  fit <- fitted(model)
  aux_model <- lm(I(res^2) ~ fit)
  bp_statistic <- summary(aux_model)$r.squared * length(res)
  p_value <- 1 - pchisq(bp_statistic, df = 1)
  return(list(statistic = bp_statistic, p.value = p_value))
}

# Apply the test
m7_het <- hetero_test(m7)
m7_het_pv <- m7_het$p.value

# DHARMa diagnostics
sim_m7 <- DHARMa::simulateResiduals(m7)

invisible(capture.output({
  m7_dharma <- DHARMa::testResiduals(sim_m7, plot = FALSE)
}))

# Extract individual test results
m7_ks <- m7_dharma$uniformity
m7_disp <- m7_dharma$dispersion
m7_out <- m7_dharma$outliers

m7_ks_pv <- m7_ks$p.value
m7_disp_pv <- m7_disp$p.value
m7_out_pv <- m7_out$p.value
```

Validation of m7 confirmed that the residuals exhibit equal variance (`r fmt_p_val(m7_het_pv)`), but are not normally distributed (Shapiro-Wilk test, `r fmt_p_val(m7_st_pv)`). Traditional QQ plots of the raw residuals do not properly account for the hierarchical structure and non-independence of mixed effect models, leading to overly optimistic results. The `{DHARMa}` package was specifically designed to address this problem by comparing observed residuals to a large number of simulations from the fitted model. This technique provides a more robust and accurate representation of model accuracy. Diagnostics produced by `testResiduals` indicate deviation from the expected distribution of residuals (Kolmogorov-Smirnov test, `r fmt_p_val(m7_ks_pv)`) without significant dispersion issues (`r fmt_p_val(m7_disp_pv)`) or outliers (`r fmt_p_val(m7_out_pv)`). Visual inspection of residual plots confirmed these findings (@fig-m7-dharma).

```{r fig-m7-dharma}
#| fig-cap: Residual Plots

DHARMa::plotQQunif(sim_m7, testDispersion = F, testUniformity = F, testOutliers = F)
```

Given these minor deviations from model assumptions, the model was refit using a robust linear mixed model designed to mitigate those issues and ensure reliable inference. Robust LMs (RLMs) use techniques such as M-estimation or MM-estimation to reduce the influence of high-leverage observations, and weighted estimation to equalize variance. This approach makes RLMs less sensitive to violations of normality assumptions in both fixed and random effects, and more resistant to departures from other common assumptions of linear mixed models. As a result, RLMs typically provide more stable and accurate parameter estimates when dealing with non-ideal data conditions.

The result from `robustlmm::rlmer` is compared with the original fit in @tbl-h1b-m7-comp. Robust fits for two additional variations of m7 are also included. Both add the initial TCT as a predictor, first as a covariate, and then including interactions with the linear and log terms in the second. This was done to account for the impact of baseline performance on the model, as discussed above. As seen in @tbl-h1b-m7-comp, both models incorporating initial TCT (m7r_ti and m7r_ta) show much higher $R^2_{marg}$ values, indicating that initial performance explains a large portion of the fixed effects variance. The version with interactions also exhibits the best ICC and RMSE scores, suggesting it is the best choice for further development.

```{r update-mdf}
# First, calculate initial TCT for each participant
mdf <- mdf |>
  group_by(participant) |>
  mutate(initial_tct = first(tct)) |>
  ungroup()

```

```{r fit-m7r, eval=FALSE}
#| warning: false
#| message: false

# robust fit of m7
m7r <- robustlmm::rlmer(
    tct ~ 1 + seq.fp + log(seq.fp + 1) + (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = F
)

# add initial_tct as covariate predictor
m7r_ta <- robustlmm::rlmer(
    tct ~ 1 + seq.fp + log(seq.fp + 1) + initial_tct + 
         (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = FALSE
)

# add initial_tct as interactive predictor
m7r_ti <- robustlmm::rlmer(
    tct ~ 1 + seq.fp * initial_tct + log(seq.fp + 1) * initial_tct + 
         (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = FALSE
)

save(m7r, m7r_ta, m7r_ti, file = here("rdata", "h1b-fit-robust.RData"))
```

```{r tbl-h1b-m7-comp}
#| tbl-cap: "Model Performance Results"

load(file = here("rdata", "h1b-fit-robust.RData"))

cp_h1b_uc_r <- compare_performance(m7, m7r, m7r_ta, m7r_ti, metrics = "common", rank = F)

cp_h1b_uc_r |> 
  select(Name, R2_conditional, R2_marginal, ICC, RMSE) |> 
  display()
```

```{r valid-m7r_ti}
# validate result
m7r_ti_het <- hetero_test(m7r_ti)
m7r_ti_het_pv <- m7r_ti_het$p.value

residuals_m7r_ti <- residuals(m7r_ti)

# Shapiro-Wilk test for normality
m7r_ti_st <- shapiro.test(residuals_m7r_ti)
m7r_ti_st_pv <- m7r_ti_st$p.value

```

These results, particularly the decrease in RMSE (`r round(cp_h1b_uc_r$RMSE[4], 2)`), indicate meaningful performance gains were achieved by the robust fit. However, the real benefit comes from improved model diagnostics. The robust model reduces heteroscedasticity (`r fmt_p_val(m7r_ti_het_pv)`) and down-weights extreme values as seen in @fig-m7r_ti-resid. Fourty-one residuals and six random effects were affected, addressing potential outliers or overly influential observations. The low residual error achieved (SD = 4.26) aligns with the high $R^2_{cond}$ previously noted. While the Shapiro-Wilk test still indicates non-normality of residuals (`r fmt_p_val(m7r_ti_st_pv)`), this is less concerning for robust models, which are designed to perform well even when normality assumptions are violated.

```{r fig-m7r_ti-resid}
#| fig-cap: Extreme values are assigned lower weights in robust model
plot(m7r_ti, which = 1)
```
As a final model validation, @fig-h1b-rlm-pred shows the predictions generated by the robust model for the first five participants in each treatment. In this figure, the black circles and lines represent observed TCT values for each participant, while the colored curves represent the RLM's predictions. Each row corresponds to a different treatment group (AR, MR, PAR, PWI), and each subplot shows data for an individual participant.

```{r h1b-insp-cond}
# Select the first 5 participants from each treatment
m7r_ti_sample <- mdf |>
  mutate(pred = fitted(m7r_ti)) |> 
  group_by(treatment) |>
  arrange(participant) |>
  filter(participant %in% unique(participant)[1:5]) |>
  ungroup() |> 
  mutate(facet_label = paste(treatment, participant, sep = " - "))

# Set a consistent maximum for y-axis
max_tct <- max(m7r_ti_sample$tct) + 20
```


```{r fig-h1b-rlm-pred}
#| fig-cap: "Visualizing RLM model fit"

ggplot(m7r_ti_sample, aes(x = seq.fp, y = tct)) + 
  geom_point(fill = "grey", pch = 21, size = 2, stroke = 1.25) + 
  geom_line(aes(group = participant)) + 
  geom_line(aes(y = pred, color = treatment), lwd = 1.5) +
  facet_wrap(~ facet_label, ncol = 5) +
  scale_y_continuous(limits = c(0, max_tct)) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "Model Fit by Participant and Treatment",
    subtitle = "RLM predictions for the first five participants in each treatment group"
  )
```

This visualization allows for a comparison between observed data and model predictions across different treatments and participants, and helps confirm that a good fit is obtained for the specified range of observation counts. The variation in TCT is evident here, with participants assigned to the PWI and PAR groups completing substantially more task iterations than their AR / MR peers.

```{r tbl-m7r-ti-fe}
#| tbl-cap: Fixed Effects Parameters for the Unconditioned RLM including initial TCT interactions (m7r_ti)
display_model_parameters(m7r_ti)
```

@tbl-m7r-ti-fe summarizes the fixed effects parameters for m7r_ti, revealing a complex interplay of factors influencing task completion time (TCT). The intercept term is near-zero (0.189) and not likely significant[^25-results-1] (t = 0.10). This counter-intuitive result is a direct consequence of including initial TCT as an explanatory variable, centering predictions around each participant's starting point and calculating deviations from initial performance rather than absolute TCT values. The initial_tct coefficient (0.996, t = 83.54) is highly significant and nearly 1, indicating that initial performance is the dominating predictor of subsequent performance, a role played by the intercept in prior formulations. The linear (seq.fp: -224.6, t = -3.70) and logarithmic (log(seq.fp + 1): 354.0, t = 4.13) terms for sequence are both significant and compete for dominance, with signs reversed from the m7 model. This sign reversal is another consequence of including initial TCT. The larger magnitude of the logarithmic term suggests a strong non-linear component to learning, aligning with theory, expectations, and observed values. Significant interactions between initial TCT and both the linear (3.36, t = 7.72) and logarithmic (-5.54, t = -9.24) terms reveal how baseline performance affects TCT changes over time. These interactions suggest that participants with higher initial TCT tend to have a slower linear decrease but stronger logarithmic decrease in TCT over time.

[^25-results-1]: For this interpretation, t values with a magnitude greater than 2.0 are taken as likely significant.

Finally, the model's random effects structure allows for participant-level variations in both the intercept and slope for the linear and log terms. Together, this sophisticated design incorporates the power law of practice in a nuanced, participant-specific manner. The combination of linear and logarithmic terms, along with their interactions with initial performance, allows it to represent complex, dynamic learning patterns that vary based on sequence, starting proficiency, and participant idiosyncrasies. 

To ease its interpretation, this model can be expressed mathematically as:

\begin{align}
\small
\text{TCT}_{ij} &=\: 0.189 + (0.996 \times \text{TCT}_{0i}) - (224.6 \times \text{SC}_{ij}) + \left( 354.0 \times \log(\text{SC}_{ij} + 1) \right) \label{eq:m7r-ti-form} \\
&\quad + (3.36 \times \text{SC}_{ij} \times \text{TCT}_{0i}) - \left(5.54 \times \log(\text{SC}_{ij} + 1) \times \text{TCT}_{0i} \right) \notag \\
&\quad + \text{RE}_{ij} + \varepsilon_{ij} \notag
\end{align}


Where:

- $i$ and $j$ are indices for participant and task iteration 
- $TCT_{ij}$ is the Task Completion Time in seconds
- $TCT_{0i}$ is the initial Task Completion Time
- $SC_{ij}$ is the standardized sequence count, $\text{SC} = (\text{SEQ} - 1) / (\max(\text{SEQ}) - 1)$
- $RE_{ij}$ represents the random effects
- $\varepsilon_{ij}$ is the residual error term

Note that $RE_{ij}$ can be further broken down as:

$$
\small{
\text{RE}_{ij} = U_{INTi} + (U_{LINi} \times \text{SC}_{ij}) + \left( U_{LOGi} \times \log(\text{SC}_{ij} + 1) \right)
}
$$ {#eq-m7r-ti-re}

Where the participant-specific random effects for intercept, linear, and log terms are:

- $U_{INTi} \sim \mathcal{N}(0, \sigma_{INT}^2)$ and $\sigma_{INT} = 0.749$
- $U_{LINi} \sim \mathcal{N}(0, \sigma_{LIN}^2)$ and $\sigma_{LIN} = 108.34$
- $U_{LOGi} \sim \mathcal{N}(0, \sigma_{LOG}^2)$ and $\sigma_{LOG} = 160.37$

These terms are participant-specific (as noted by the subscript $i$), and have a complex correlation structure. Notably, there is a strong negative correlation (-0.99) between the random effects for seq.fp and log(seq.fp + 1), indicating these terms strongly counteract each other at the participant level. The intercept has a moderate negative correlation with seq.fp (-0.68) and a moderate positive correlation with log(seq.fp + 1) (0.79). Finally, the residual error, $\varepsilon_{ij} \sim \mathcal{N}(0, 4.256^2)$. Together, these specifications provide a complete picture of the model's structure and variability.

#### Conditional Models and Hypothesis Testing

To understand the treatment effects and test $H_{1b}$, a model conditioned on treatment must be fit. Using m7r_ti as a base, three models of increasing complexity were again fit. The first includes treatment as a covariate predictor, the second includes interactions with the linear term, and the final model adds interactions with the log term. As before, these were fit with `robustlmm::rlmer`, and performance of the resulting models is compared in @tbl-h1b-fit-cond.

```{r h1b-fit-cond, eval=FALSE}

# include treatment as a covariate
m7r_ti_c1 <- robustlmm::rlmer(
    tct ~ 1 + seq.fp * initial_tct + log(seq.fp + 1) * initial_tct + treatment +
         (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = FALSE
)

# include treatment interaction with linear term
m7r_ti_c2 <- robustlmm::rlmer(
    tct ~ 1 + seq.fp * initial_tct + seq.fp * treatment + log(seq.fp + 1) * initial_tct +
         (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = FALSE
)

# include treatment interactions with both linear and log terms
m7r_ti_c3 <- robustlmm::rlmer(
    tct ~ 1 + seq.fp * initial_tct + seq.fp * treatment + log(seq.fp + 1) * initial_tct + log(seq.fp + 1) * treatment + 
         (1 + seq.fp + log(seq.fp + 1) | participant),
    data = mdf, REML = FALSE
)

save(m7r_ti_c1, m7r_ti_c2, m7r_ti_c3, file = here("rdata", "h1b-fit-cond.RData"))

```

```{r tbl-h1b-fit-cond}
load(here("rdata", "h1b-fit-cond.RData"))
cp_h1b_cn_r <- compare_performance(m7r, m7r_ti_c1, m7r_ti_c2, m7r_ti_c3, metrics = "common", rank = F)

cp_h1b_cn_r |> 
  select(Name, R2_conditional, R2_marginal, ICC, RMSE) |> 
  display()
```

All models incorporating $TCT_{0}$ as a predictor show a substantial increase in $R^2_{marg}$. This suggests that adding treatment as a fixed effect improves the model's explanatory power. Model m7r_ti_c2, which includes treatment interactions with the linear term, performs better than the alternatives in every metric ($R^2_{marg}$ = 0.93, ICC = 0.85, RMSE = 5.93) except $R^2_{cond}$, where it matches the most complex matter (0.99). Furthermore, where the other models produced singularity warnings during the fit, c2 did not. These warnings did not invalidate the results for c1 or c3, as `robustlmm` automatically negotiated to an alternate optimizer. But the added stability provides additional comfort.

```{r fig-plot-make}

# Create a data frame for predictions
# include predictions for m7r_ti (no treatment effect)
pred_data_m7r_ti <- data.frame(seq.fp = seq(0, 1, length.out = 100))
pred_data_m7r_ti$initial_tct <- mean(mdf$initial_tct, na.rm = TRUE)  # Use mean initial_tct
pred_data_m7r_ti$TCT <- predict(m7r_ti, newdata = pred_data_m7r_ti, re.form = NA)

# use average initial TCT for each treatment
avg_tct_0 <- aggregate(initial_tct ~ treatment, data = mdf, FUN = mean)
names(avg_tct_0)[2] <- "avg_initial_tct"

# Create a data frame for predictions
pred_data <- expand.grid(
  seq.fp = seq(0, 1, length.out = 100),
  treatment = unique(mdf$treatment)
)

# Merge the average initial TCT values into pred_data
pred_data <- merge(pred_data, avg_tct_0, by = "treatment")

# Rename avg_initial_tct to initial_tct to match the model
names(pred_data)[names(pred_data) == "avg_initial_tct"] <- "initial_tct"

# Generate predictions
pred_data$TCT <- predict(m7r_ti_c2, newdata = pred_data, re.form = NA)

# Calculate confidence interval
vcov_fixed <- vcov(m7r_ti_c2)
X <- model.matrix(~ seq.fp * initial_tct + seq.fp * treatment + log(seq.fp + 1) * initial_tct, data = pred_data)
se <- sqrt(diag(X %*% vcov_fixed %*% t(X)))
pred_data$lower <- pred_data$TCT - 1.96 * se
pred_data$upper <- pred_data$TCT + 1.96 * se
```

```{r fig-plot-show}
#| fig-cap: "Learning Curves by Treatment"

# Create the plot
ggplot() +
  # Add points for actual data
  geom_point(data = mdf, aes(x = seq.fp, y = tct, color = treatment), alpha = 0.5) +
  # Add lines for m7r_ti_c2 predictions
  geom_line(data = pred_data, aes(x = seq.fp, y = TCT, color = treatment)) +
  # Add line for m7r_ti predictions
  geom_line(data = pred_data_m7r_ti, aes(x = seq.fp, y = TCT), color = "black", size = 1) +
  labs(title = "Predicted Learning Curves by Treatment",
       subtitle = "Black line represents overall prediction from m7r_ti",
       x = "Standardized Task Sequence (SC)",
       y = "Predicted Task Completion Time (seconds)") +
  scale_color_brewer(palette = "Set1") +  # Use a color-blind friendly palette
  theme_minimal() +
  theme(legend.position = "bottom")
```

@fig-plot-show plots the predicted learning curves for each treatment based on their mean $TCT_{0}$. The black line shows the predictions from the prior model, without treatment effects. All treatments show a general non-linear decrease in TCT with sequence, demonstrating learning. PWI and PAR perform better than average, while AR and MR are slower. This ranking persists through all iterations, despite clear differences in initial performance. To quantify these trends in statistical terms, @tbl-m7r-ti-c2-params summarizes the parameters for the model's fixed effects.

```{r tbl-m7r-ti-c2-params}
display_model_parameters(m7r_ti_c2)
```

These results are similar to those for the unconditioned model (@tbl-m7r-ti-fe). All terms previously found significant remain so, including $TCT_{0}$, the linear and log terms, and their interactions with $TCT_{0}$. The relevant interpretations stand, though the magnitude of the effects vary slightly. Importantly, all interactions between treatment and seq.fp are identified as significant. This directly addresses $H_{1b}$, confirming that treatment varies TCT over time, when $TCT_{0}$ varies. To test if these findings hold when all participants begin with equal proficiency (equalizing the effects of the power law of practice), a second set of predictions is generated using equal values of $TCT_{0}$ for all treatments. The resulting learning curves are plotted in @fig-plot-stdtct.

```{r fit-with-std-tct0}
# Use the overall mean initial TCT for all treatments
mean_initial_tct <- mean(mdf$initial_tct, na.rm = TRUE)

# Create a new prediction data frame
pred_data_equal_start <- expand.grid(
  seq.fp = seq(0, 1, length.out = 100),
  treatment = unique(mdf$treatment),
  initial_tct = mean_initial_tct
)

# Generate predictions
pred_data_equal_start$TCT <- predict(m7r_ti_c2, newdata = pred_data_equal_start, re.form = NA)
```

```{r fig-plot-stdtct}
#| fig-cap: "Compare Learning Curves with Equal Initial Proficiency"

ggplot(pred_data_equal_start, aes(x = seq.fp, y = TCT, color = treatment)) +
  geom_line() +
  labs(title = "Predicted Learning Curves by Treatment",
       subtitle = "Assuming equal initial proficiency",
       x = "Standardized Task Sequence (SC)",
       y = "Predicted Task Completion Time (seconds)") +
  scale_color_brewer(palette = "Set1") +
  theme_minimal()
```

This plot shows that, even when the effects of $TCT_{0}$ are standardized, the instructional method influences the learning rate. To quantify this effect, `emtrends` from the `emmeans` package is used to calculate the rate of change in the response variable with respect to specified predictors. In the context of this analysis, it is used to calculate the slope of TCT with respect to sequence for each treatment while holding $TCT_{0}$ constant. Pairwise comparisons of the resulting slopes allow for statistical tests, as seen in @tbl-slopes-fixedtct. These results show significant differences between the learning rates for PWI and both AR (p = 0.007) and MR (p < 0.001) as well as PAR and MR (p = 0.04). The overall ranking of effectiveness can be expressed as PWI ≈ PAR, PWI > (AR ≈ MR), and PAR > MR.

```{r tbl-slopes-fixedtct}
#| tbl-cap: "Average Slope of TCT over Task Sequence by Treatment"

# Compute slopes
emm_slopes <- emtrends(m7r_ti_c2, ~ treatment, var = "seq.fp", 
                       at = list(initial_tct = mean_initial_tct))

# Pairwise comparisons of slopes
pairs(emm_slopes)

```

While @fig-plot-stdtct clearly illustrates the differences in learning curves, and @tbl-slopes-fixedtct provides compelling statistical evidence of these effects, both approaches focus on absolute changes in task completion time. As a complementary perspective, @fig-learn-cont expresses progress as cumulative percentage improvements. This approach offers some advantages:

1. It allows for comparison of relative progress across treatments, which can be informative when participants start at different performance levels.
2. Percentage improvements can sometimes highlight subtle differences in learning patterns that are less apparent in absolute time measures.
3. It provides a way to visualize learning gains that is independent of the scale of the original task completion times.

While the result may appear similar to the inverse of earlier plots, it offers a standardized view of learning progress across treatments that aligns closely with how learning is often conceptualized in cognitive and educational theories. This representation helps confirm and clarify the trends observed in our previous analyses, reinforcing our understanding of how each treatment impacts learning over time.

```{r fig-learn-cont}
#| fig-cap: "The last damn plot"

# Create a grid of seq.fp values that matches observed values
seq_grid <- seq(0, 1, by = 0.25)

emm <- emmeans(
  m7r_ti_c2, 
  ~ treatment | seq.fp, 
  at = list(
    seq.fp = seq_grid,
    initial_tct = mean_initial_tct
    )
  )

emm_data <- as.data.frame(emm)

emm_data_cumulative <- emm_data |>
  arrange(treatment, seq.fp) |>
  group_by(treatment) |>
  mutate(
    chg_emm = lag(emmean) - emmean,
    pct_chg = chg_emm / lag(emmean),
    cumulative_learning = c(0, cumprod(1 + pct_chg[-1]) - 1)
  ) |>
  ungroup()

# Plot cumulative learning
ggplot(emm_data_cumulative, aes(x = seq.fp, y = cumulative_learning, color = treatment)) +
  geom_line() +
  geom_point() +
  labs(x = "Sequence Progress", 
       y = "Cumulative Learning Progress", 
       title = "Cumulative Learning Progress by Treatment",
       subtitle = "All treatments start at 0; higher values indicate more learning") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  coord_cartesian(ylim = c(0, NA))  # Start y-axis at 0 and let the upper limit adjust automatically


```

#### $H_{1b}$ Results

Based on comprehensive analysis of learning curves and statistical modeling, there is strong evidence to accept $H_{1b}$: the rate of improvement in task completion time varies significantly by treatment. Specifically, PWI demonstrates the steepest learning curve, followed closely by PAR. Both AR and MR show slower rates of improvement. The relationship can be summarized as: PWI ≈ PAR, PWI > (AR ≈ MR), and PAR > MR. This pattern holds true even when controlling for initial task completion time, suggesting that the treatment effects on learning rates are robust. The differences in learning rates are most pronounced in the early stages of task repetition and tend to converge over time, highlighting the importance of instructional method particularly in the initial phases of skill acquisition. However, these findings should be interpreted with the described limitations in mind, primarily those related to the limited extrapolation capabilities resulting from an imbalanced data set and statistical model formulation. In light of these considerations, great effort was invested to ensure the robust and reliable findings outlined above.

### $H_{1c}$: Average Error Count per Car

```{r prep-h1c-data}
# create data
learn_h1c <- learn |>
  filter(participant != 1063) |>
  rename(tct = car_elapsed_time_adj) |>
  group_by(participant) |>
  mutate(
    seq = row_number(),
    treatment = factor(as.character(treatment))
  ) |>
  summarize(
    treatment = first(treatment),
    avg_uce = mean(uce, na.rm = T)
  )

```

The third and final hypothesis of the learning phase tests the third commonly assessed aspect of learning:
$$H_{1b}\textrm{: Average error count per car varies with treatment}$$ 

Here again, we are interested in finding statistically significant differences in the average error count under different treatment conditions to identify which instructional method leads to the lowest defect rate. Lower error rates signal a higher quality of learning outcomes. Combined with the previous measures of efficiency (TCT) and learning rate (change in TCT per task), these provide a balanced and comprehensive view of training performance.

#### Descriptive Statistics

As with previous measures, the uncorrected error count (UCE) is neither independent nor balanced. Additionally, it is a discrete (counted) variable. All of these have implications for the statistical method used, but where average counts are used, many are addressed. The standard five number summary for UCE is presented in @tbl_h1c_ss. 

```{r tbl_h1c_ss}
#| tbl-cap: "Summary Statistic for Dependent Variables, H1c Quality"

datasummary(
  Heading("Average Uncorrected Errors") * avg_uce ~
    Mean + SD + Min + Median + Max,
  data = learn_h1c,
  output = "kableExtra"
)
```

```{r h1c-tests}
# check linear model assumptions for avgUCE

# normality
h1c_st <- shapiro.test(learn_h1c$avg_uce)
h1c_st_pv <- fmt_p_val(h1c_st$p.value)

h1c_st_treat <- by(learn_h1c$avg_uce, learn_h1c$treatment, shapiro.test)

# homogeneity of variance, aka homoscedasticity
h1c_lt_tct <- car::leveneTest(avg_uce ~ treatment, data = learn_h1c)
h1c_lt_tct_pv <- fmt_p_val(h1c_lt_tct$`Pr(>F)`[1])

h1c_skew <- moments::agostino.test(learn_h1c$avg_uce)
h1c_skew_pv <- fmt_p_val(h1c_skew$p.value)

h1c_kurt <- moments::anscombe.test(learn_h1c$avg_uce)
h1c_kurt_pv <- fmt_p_val(h1c_kurt$p.value)

# outliers
h1c_out <- check_iqr(learn_h1c$avg_uce, 1.5)
h1c_out_count <- sum(h1c_out)  # 3 found: 6, 7, 45 → participants 1016, 1017, and 1056
```

As this data is aggregated by participant, there is only one observation for each ($n$ = 53). As is common in this study, variation is high (SD = 3.15) compared to the mean value (2.19). Many participants made no errors, but one averaged 12 errors per task iteration, resulting in an overall median value of 0.33. The distribution of UCE is illustrated by @fig-h1c-hist. As seen before, the overall average error count data exhibits floor effects that contribute to a strongly positive skew. Obvious non-normality (Shapiro-Wilk, `r h1c_st_pv`), skew (D'Agostino, `r h1c_skew_pv`), and kurtosis (Anscombe-Glynn, `r h1c_kurt_pv`) are all confirmed in the overall dataset using the appropriate tests.

```{r fig-h1c-hist}
#| fig-cap: Histogram of Average Error Count in Learning
#| fig-height: 3.5
#| warning: false
#| message: false

# Plotting uce
learn_h1c |>
  filter(!is.na(avg_uce)) |>
  ggplot(aes(x = avg_uce)) +
  geom_histogram(
    bins = 18,
    fill = I(base_color),
    alpha = 0.75,
    width = 0.25,
    color = outline_color
  ) +
  geom_boxplot(
    width = 1.5,
    position = position_nudge(y = -1.25),
    notch = FALSE
  ) +
  labs(
    x = element_blank(),
    y = element_blank(),
    title = "Distribution of UCE Values",
    subtitle = "Broad range of result quality centered near zero"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
    # legend.position = "none",
    # axis.text.y = element_blank(),
    # axis.ticks.y = element_blank(),
    # axis.text.x = element_text(margin = margin(b = 3, unit = "pt"))
  )

```

Finally, Shapiro-Wilk tests are performed for each treatment group, revealing strong evidence that PAR (`r fmt_p_val(h1c_st_treat$PAR$p.value)`), AR (`r fmt_p_val(h1c_st_treat$AR$p.value)`), and MR (`r fmt_p_val(h1c_st_treat$MR$p.value)`) are not normally distributed, but PWI is (`r fmt_p_val(h1c_st_treat$PWI$p.value)`).  These findings are supported by the Q--Q plots in @fig-qq-h1c. Furthermore, it is important to note that this is counted data. Despite the fact that means / grand means may appear continuous in this aggregated form, the underlying distribution remains discrete. Traditional parametric tests are not recommended in this case. 

```{r fig-qq-h1c}
#| fig-cap: Q-Q Plots of UCE by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(learn_h1c, x_var = "avg_uce", by_var = "treatment")

p + plot_annotation(
  title = "Quantile-Quantile Plots of UCE Treatment Data",
  subtitle = "Only PWI data appears normally distributed"
)
```

#### Analysis

The Kruskal-Wallis non-parametric test is used to compare average UCE by treatment, the results of which are shown in @fig-h1c-plot. Large differences in the overall mean are observed, along with some floor effects in the augmented treatments. The test statistic $\chi^{2}_{KW}(3)=25.5$, with a p-value < 0.001, indicates a statistically significant difference between groups. The effect is $\hat{\epsilon}^{2}_{rank}=0.49$, with a $CI_{95\%}\:[0.32,1.00]$, reflecting the data's variability. The results of Dunn's Holm-adjusted pairwise comparisons show statistically significant differences between PWI and all other treatments (p << 0.01 in all cases), while PAR, AR, and MR treatments are statistically similar. These results formally quantify the obvious reduction in error count experienced with any form of augmented instruction.

```{r fig-h1c-plot}
#| fig-cap: Differences in Average Uncorrected Error per Task by Treatment
#| fig-height: 5
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

x <- ggstatsplot::ggbetweenstats(
  data = learn_h1c,
  x = treatment,
  y = avg_uce,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "Avg UCE per Task by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant pairwise differences identified between PWI and Augmented Treatmentts",
  digits = 3,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
      ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
    )
)

```

```{r h1c-outliers}

co <- check_outliers(learn_h1c$avg_uce, method = "iqr", threshold = 1.5)

# use co as boolean mask to get relevant participants
co_res <- learn_h1c[as.vector(co), ]

# tabulates the number of tasks completed by outlier participants
car_counts <- learn |>
  filter(participant %in% co_res$participant) |> 
  group_by(participant) |> 
  summarize(count = n()
)

# col_to_string(co_res, participant)
# col_to_string(co_res, avg_uce, digits = 1)

# check test results without outliers
without_outliers <- learn_h1c[!learn_h1c$participant %in% co_res$participant, ]
kt_h1b_no_outliers <- kruskal.test(avg_uce ~ treatment, data = without_outliers)
```

KW is a robust test that does not assume normality or homogeneity of variances. It can be sensitive to the presence of outliers, three of which were detected in the UCE data based on a combination of IQR and Z-score. Participants `r col_to_string(co_res, participant)` were all assigned to the PWI treatment and averaged `r col_to_string(co_res, avg_uce, digits = 1)` errors per task, respectively. A second Kruskal-Wallis test of the data without those outliers showed the difference remained significant (`r fmt_p_val(kt_h1b_no_outliers$p.val)`). This confirms that the outliers did not impacting the outcome of the test.


#### $H_{1c}$ Results

$H_{1c}$ is accepted - all augmented instruction methods result in error rates that are drastically lower in magnitude than PWI, the statistical significance of which is confirmed. Floor effects may imply that the task complexity was insufficent to capture the full range of participant performance. Further analysis using a generalized linear mixed model with poisson or negative binomial distribution for count data could be used to better account for participant variation but that seems unnecessary given the strong observed differences. While this method was deemed insufficient for capturing the nuances of learning rate analysis, it is entirely appropriate for comparing average error counts.

## $H_2$: Recall Phase Analysis

```{r define-constants}
TREATMENTS <- levels(res_f$treatment)

ORDERED_PAIRS <- c()
for (i in 1:(length(TREATMENTS) - 1)) {
  for (j in (i + 1):length(TREATMENTS)) {
    ORDERED_PAIRS <- c(ORDERED_PAIRS, paste(TREATMENTS[i], "-", TREATMENTS[j]))
  }
}

# set global flag variable used for logging current statistic
CUR_STAT <- NULL
```

```{r recall-data}

# drop outlier from source
res_f <- res_f |>
  filter(participant != 1063)

recall <- res_f |>
  filter(
    phase == 2,
    car_outcome == "complete",
  ) |>
  select(
    participant,
    car_elapsed_time,
    uce,
    treatment,
    drop_event_count,
    drop_event_tot_time_adj,
    repair_event_count,
    repair_event_tot_time,
    system_event_count,
    system_event_tot_time,
    pwi_event_count,
    pwi_event_tot_time
  )

# add tot_sys_adj, tot_usr_adj, and adjusted TCT time to learn
# tot_sys_adj will be used to calculate OEE
# tot_usr_adj just to reduce TCT

recall <- recall |>
  mutate(
    # Replace NA with 0 using coalesce for safe addition
    tot_sys_adj = coalesce(drop_event_tot_time_adj, 0) +
      coalesce(system_event_tot_time, 0),
    tot_usr_adj = coalesce(repair_event_tot_time, 0),
    tot_adj = tot_sys_adj + tot_usr_adj,
    car_elapsed_time_adj = car_elapsed_time - tot_adj,
    pwi_event_count = replace_na(pwi_event_count, 0),
    pwi_event_tot_time = replace_na(pwi_event_tot_time, 0)
  )


changes <- recall |>
  filter(car_elapsed_time != car_elapsed_time_adj)

# only 1057 had deductions for system events; none for user events
# what were the source of these in recall, which is PWI only?
# hololens menu popping up inadvertantly - lost time dismissing it

total_tasks_completed <- nrow(recall) # 212
total_tasks_adjusted <- nrow(changes) # 2

recall <- recall |>
  select(
    participant,
    treatment,
    car_elapsed_time_adj,
    uce,
    pwi_event_count,
    pwi_event_tot_time,
    tot_sys_adj,
    tot_usr_adj
  )

# save(recall, file = here("rdata", "recall.RData"))
```

```{r make-h2a-data}
# calculate OEE from filtered res_f data
oee_data <- res_f |>
  filter(
    phase == 2,
    car_outcome == "complete",
  ) |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    session_start = min(car_start_time, na.rm = TRUE),
    session_end = max(car_end_time, na.rm = TRUE),
    session_duration = session_end - session_start,
    units_produced = n(),
    units_defective = sum(uce > 0, na.rm = TRUE),
    tot_sys_adj = sum(
      coalesce(drop_event_tot_time_adj, 0)
      + coalesce(system_event_tot_time, 0),
      na.rm = TRUE
    ),
    prod = units_produced * 60 / session_duration,
    qual = (units_produced - units_defective) / units_produced,
    avail = (session_duration - tot_sys_adj) / session_duration,
    oee_100 = 1 * prod * qual,
    oee_85 = 0.85 * prod * qual,
    oee_avail = avail * prod * qual
  )
```

```{r make-h2b-data}
recall_agg <- recall |>
  group_by(participant) |>
  summarize(
    treatment = first(treatment),
    tot_pwi_count = sum(pwi_event_count, na.rm = TRUE),
    tot_pwi_time = sum(pwi_event_tot_time, na.rm = TRUE),
    task_num = n()
  ) |>
  mutate(
    avg_pwi_count = tot_pwi_count / task_num,
    avg_pwi_time = tot_pwi_time / task_num,
    pwi_time_per_ref_agg = ifelse(avg_pwi_count > 0, avg_pwi_time / avg_pwi_count, 0)
  )
```

As with the learning phase results, only completed tasks are considered for recall analysis. Paper work instructions were available for reference, and participants wore the HL2 only for recording purposes. This almost entirely eliminated system-related interruptions, though participant 1057 inadvertently triggered the HL2 menu twice during their trial. The time lost to those events was deducted from measured task times as before. `r total_tasks_completed` observations were recorded during this phase, as expected for `r length(unique(recall$participant))` participants, each with four iterations.

### $H_{2a}$: Overall Equipment Effectiveness

The first hypothesis of the recall phase:
$$H_{2a}\textrm{: OEE varies with treatment}$$

is designed to determine if the instructional method has an impact on overall performance after the initial training period. As described in @sec-calc-meas, OEE is the product of performance (task rate relative to takt time), quality (percentage of units produced without defects), and availability (percentage of system up time). For the purposes of this analysis, 100% availability is assumed. Takt time is set to 60 seconds in accordance with the line policy and instructional design.

#### Descriptive Statistics

```{r make-h2a-report}
# report descriptive stats for cet and uce
h2a_rpt <- oee_data |>
  select(prod, qual, oee_100) |>
  rename(
    "Productivity" = prod,
    "Quality" = qual,
    "Overall Equipment Effectiveness" = oee_100
  ) |>
  report::report() # options not working!

# Extract and split the string by the pattern "\n - ", remove first line
h2a_rpt_list <- strsplit(h2a_rpt[1], "\n  - ")[[1]][-1]

# Add hyphen to first element and rejoin
h2a_rpt_list[1] <- paste0("- ", h2a_rpt_list[1])
h2a_rpt_list <- paste(h2a_rpt_list, collapse = "\n- ")
h2a_rpt_list <- gsub(", 0 missing", "", h2a_rpt_list)
h2a_rpt_list <- gsub("n = 54, ", "", h2a_rpt_list)

# cat(h2a_rpt_list)
```

The calculated measures are summarized below:

`r h2a_rpt_list`

As this is aggregated data, there is one observation for each participant, and no missing values. Productivity and quality are slightly skewed right (0.20) and left (-0.25), respectively. Kurtosis measures suggest that productivity is close to normal (1.28), but quality has a flatter distribution (-1.96). Productivity exhibits relatively low variability and similar mean (0.84) and median (0.82), suggesting consistency among participants. For quality, the median (1.00) and its absolute deviation (MAD = 0.00) show that many participants achieved perfect quality during recall. However, the range (0, 1) and high standard deviation (0.49) relative to its mean (0.56) indicates considerable variability. In fact, several participants produced *only* defective assemblies. Note that quality is essentially a discrete variable, as it can only take on values of $n/4$, where $n = 0, 1, \dots, 4$, corresponding to each of the four task iterations.

OEE shows a notably asymmetric distribution with a mean of 0.48 and a higher median of 0.62, indicating a significant number of low values dragging down the mean. A high concentration of zeros are observed in the OEE histogram, @fig-oee-hist, suggesting that many participants experience zero effectiveness due to high defect rates. The standard deviation (sd = 0.43) and median absolute deviation (MAD = 0.58) reflect substantial variability in effectiveness across participants. Despite the approximately symmetric skewness (-0.07), the range (0 to 1.14) indicates diverse levels of effectiveness, from complete failures to high performance. These findings highlight the variability in participants' abilities to achieve consistent and effective output.

```{r fig-oee-hist}
#| fig-cap: Histogram of Overall Equipment Effectiveness in Recall
#| fig-height: 3.5
#| warning: false
#| message: false

oee_data |>
  ggplot(aes(x = oee_100)) +
  geom_histogram(
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color,
    binwidth = 0.1
  ) +
  labs(
    title = "Distribution of OEE Values",
    subtitle = "High defect rates have an obvious effect on effectiveness",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
  )
```

```{r oee-tests}
st_oee_treat <- by(oee_data$oee_100, oee_data$treatment, shapiro.test)

# no outliers are detected (output nrow = 0)
outliers <- oee_data |>
  group_by(treatment) |>
  mutate(z_oee = (oee_100 - mean(oee_100, na.rm = TRUE)) / sd(oee_100, na.rm = TRUE)) |>
  filter(abs(z_oee) > 2.5) |>
  nrow()

lt_oee_treat <- car::leveneTest(oee_100 ~ treatment, data = oee_data)
pv <- lt_oee_treat$`Pr(>F)`[1]
```

Results from the Shapiro-Wilk test provide further evidence against the normality hypothesis for OEE, both overall (p `r fmt_p_val(shapiro.test(oee_data$oee_100)$p.value)`) and by treatment (p < 0.05 in all cases). Q-Q plots in @fig-qq-h2a suggest that OEE is close to normally distributed for all augmented treatments, but PWI is clearly not. No outliers were detected in the OEE data based on the Z-score test. Finally, Levene's test indicated no significant heterogeneity of variance between treatment groups (`r fmt_p_val(pv)`). These findings again dictate a non-parametric approach to testing $H_{2a}$.

```{r fig-qq-h2a}
#| fig-cap: Q-Q Plots of OEE by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(oee_data, x_var = "oee_100", by_var = "treatment")

p + plot_annotation(
  title = "Quantile-Quantile Plots of OEE Treatment Data",
  subtitle = "PWI effectiveness is not normally distributed"
)
```

#### Analysis

```{r test-oee-power}
k <- length(unique(oee_data$treatment))
n <- round(nrow(oee_data) / k)
epsilon_sq <- 0.19

# Estimate post-hoc power
oee_power <- posthoc_power_analysis(
  n, k, epsilon_sq,
  sig_level = 0.05, num_simulations = 1000
)
```

In @fig-h2a-plot, group-wise significant differences are found (p = 0.02), with with small to moderate effect size $\hat{\epsilon}^{2}_{rank}=0.19$ and a broad $CI_{95\%}\:[0.08,1.0]$. Pair-wise tests show only PWI and PAR are significantly different (also p = 0.02). A post-hoc simulation was performed to assess the power of the analysis. The result of `r round(oee_power * 100, 1)`% indicates the analysis had sufficient power to detect significant differences, which lends further support to the reliability of this outcome.

```{r fig-h2a-plot}
#| fig-cap: $H_{2a}$ Test - OEE Differences by Treatment
#| fig-height: 5
#| warning: false
#| message: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

h2a_plot <- ggstatsplot::ggbetweenstats(
  data = oee_data,
  x = treatment,
  y = oee_100,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "OEE Differences by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "Significant Differences Identified between PWI and PAR Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 3,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

h2a_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "9.79", italic(p) ==
#     "0.02", widehat(epsilon)["ordinal"]^2 == "0.19", CI["95%"] ~
#     "[" * "0.08", "1.00" * "]", italic("n")["obs"] == "53")
```

```{r oee_summary}
ost <- oee_data |>
  group_by(treatment) |>
  summarize(
    mean = mean(oee_100, na.rm = T),
    sd = sd(oee_100, na.rm = T),
    med = median(oee_100, na.rm = T),
    iqr = IQR(oee_100, na.rm = T),
    min = min(oee_100, na.rm = T),
    max = max(oee_100, na.rm = T),
  ) |>
  mutate(
    cv = sd / mean,
    rng = max - min,
    iqr_ratio = iqr / rng
  )

# Extract the iqr_ratio for PAR
pr <- function(trt) {
  iqr_r <- ost |>
    filter(treatment == trt) |>
    pull(iqr_ratio)
  return(round(iqr_r, 2))
}
```

This unexpected result indicates that, despite *substantial observed differences*, the PWI, AR, and MR treatments have *statistically equivalent* OEE. High within-group variability likely contributes to this finding by obscuring genuine differences in group means. Kruskal-Wallis test is somewhat vulnerable to this effect due to its rank-based design and underlying assumptions. The ratio of interquartile range to overall range for PAR (`r pr("PAR")`), AR (`r pr("AR")`), and MR (`r pr("MR")`) show that the central 50% of observations in these treatments is spread over more than half of its overall range. This wide variation is easily observed in the box plots of @fig-h2a-plot.

To further investigate the observed variability and validate these unexpected findings, bootstrap resampling analysis was employed. This flexible technique samples with replacement to construct robust estimates of the distribution for the statistics of interest (e.g. mean or median) and their confidence intervals. Bootstrapping relies on the empirical distribution of the data to accurately capture variability with few assumptions. Additionally, the replication process effectively increases the sample size and statistical power of the analysis, enhancing the accuracy and reliability of its results.

A bootstrap analysis was used to estimate pairwise differences in three measures of centrality for OEE--mean, median, and the Hodges-Lehmann estimator. The later is a non-parametric approach to estimate population median shifts that is robust to outliers. It provides a consistent estimate of the true shift in medians, even when the underlying distributions are dissimilar, by considering the differences between all possible observation pairs. This process was replicated five times, each with 10,000 resamples, and the results were averaged to determine grand means and related statistics. The significant results of this complementary analysis, where the 95% confidence interval does not include zero, are tabulated in @tbl-bootstrap-oee.

```{r bootstrap-oee, eval=FALSE}
# registerDoParallel(cores = parallel::detectCores())

stat_funcs <- list(
  Mean = mean,
  Median = median,
  HL_Est = hl_estimator
)

h2a_oee_boot_res <- run_bootstrap_tests(
  oee_data,
  column = "oee_100",
  stat_funcs = stat_funcs
)

h2a_oee_boot_res
save(h2a_oee_boot_res, file = here("rdata", "h2a_oee_boot_res.RData"))
```

```{r tbl-boot-oee}
#| tbl-cap: $H_{2a}$ Bootstrap Analysis of Pairwise OEE Differences

load(here("rdata", "h2a_oee_boot_res.RData"))

title <- "Significant OEE Differences in Centrality: Bootstrapped Estimates"
subtitle <- "Five Replications of 10,000 Resamples for Mean, Median, and the Hodges-Lehmann Estimator"


create_boot_summary_table(h2a_oee_boot_res, title = title, subtitle = subtitle)
```

Compared to the initial findings, bootstrapping identifies a number of significant pairwise differences. Even when testing OEE median values, as done with Kruskal-Wallis, this method finds both PWI-PAR and PWI-AR significant. The mean tests, which are more sensitive to variance and outliers, also include PWI-MR. Hodges-Lehmann matches the mean findings, despite its reduced sensitivity. 

This result supports the previous findings of significant overall differences in OEE, as well as between the PWI and PAR conditions for all measures of centrality. In addition, the bootstrap analysis, with its robust estimation methods, provided strong evidence of significant differences between PWI and both the AR (for all measures) and MR groups (except for median). These results show that the OEE of PWI is substantially lower than all other treatments, with average estimated differences of 0.687 (PAR), 0.585 (AR), and 0.307 (MR). The negligible bias values and low standard error reported across all replicates reinforce the reliability of these findings.

#### $H_{2a}$ Result

As both approaches identified a significant treatment effect on OEE, there is ample evidence to accept $H_{2a}$: OEE does vary with instructional method. However, the conflicting pairwise analysis results must be reconciled. While the Kruskal-Wallis test only identified significant differences between the PWI and PAR treatments, subsequent bootstrapped analysis revealed additional differences between PWI and the AR and MR treatments using a variety of measures. Considering the bootstrap method's advantages when faced with high between-group variance, limited sample size, and non-normally distributed data, the consistent differences and robust confidence intervals it produced provide compelling evidence that the observed differences are reliable. Ultimately, it is determined that PAR, AR, and MR treatments all result in statistically significant and practically meaningful improvements in OEE compared to PWI, which was hampered by a high defect rate.

### $H_{2b}$: Paper Work Instruction References

```{r h2b-data}
recall_h2b <- recall |>
  mutate(
    pwi_time_per_ref_all = ifelse(
      pwi_event_count > 0,
      pwi_event_tot_time / pwi_event_count,
      0
    )
  )

recall_h2b_agg <- recall_agg
```

The second hypothesis of the recall phase assesses an alternative measure of training effectiveness:
$$H_{2b}\textrm{: PWI reliance varies with treatment}$$

Reliance is quantified by the number of times each participant refers to the paper work instructions (PWI count), and the duration of each event (PWI time). PWI count is considered a measure of the *frequency* of the participant's task uncertainty, while PWI time is associated with the *degree* of uncertainty experienced.

#### Descriptive Statistics

```{r make-h2b-report}
#| warning: false

# report descriptive stats for pwi count and pwi time (for each car, n = 212)
h2b_rpt <- recall_h2b |>
  select(pwi_event_count, pwi_event_tot_time, pwi_time_per_ref_all) |>
  rename(
    "PWI Count per Task" = pwi_event_count,
    "Total PWI Time per Task" = pwi_event_tot_time,
    "PWI Avg Time per Reference" = pwi_time_per_ref_all
  ) |>
  report::report() # options not working!

# Extract and split the string by the pattern "\n - ", remove first line
h2b_rpt_list <- strsplit(h2b_rpt[1], "\n  - ")[[1]][-1]

# Add hyphen to first element and rejoin
h2b_rpt_list[1] <- paste0("- ", h2b_rpt_list[1])
h2b_rpt_list <- paste(h2b_rpt_list, collapse = "\n- ")
h2b_rpt_list <- gsub(", 0 missing", "", h2b_rpt_list)
h2b_rpt_list <- gsub("n = 212, ", "", h2b_rpt_list)

# convert back into list of strings
h2b_rpt_list <- strsplit(h2b_rpt_list, "\n")[[1]]

# Calculate and test Spearman's correlation
h2b_cor_test_spearman <- cor.test(recall_h2b$pwi_event_count, recall_h2b$pwi_event_tot_time, method = "spearman")

# print(h2b_rpt_list[1:2])
```

Overall statistics for the measures of interest (n = `r nrow(recall_h2b)`) are summarized below:

`r paste(h2b_rpt_list[1:2], collapse = "\n")`

Despite low mean values for count and total time (1.47 and 2.86, respectively), both exhibit substantial variability, with standard deviations of 4.12 and 9.43. High skewness (4.51 for count, 5.64 for time) and kurtosis (23.27, 36.73) confirm their strong right skew and heavy tails. Together with a median and MAD of zero for both measures, these statistics suggest that many participants rarely referred to the instructions, but a small number relied on them heavily.

The statistical similarity between count and time suggests a high correlation, confirmed by Spearman's rank correlation test (rho = `r round(h2b_cor_test_spearman$estimate,2)`, `r fmt_p_val(h2b_cor_test_spearman$p.value)`). This is natural and expected for two dependent measures, but must be accounted for. To create a stable measure of overall reliance for each participant, a composite measure was created by averaging the time per PWI reference across all four tasks. This approach reduces the impact of task-specific factors and eliminates the correlation between reference count and duration without diminishing either aspect of task uncertainty. Using a single, composite measure simplifies the analysis and interpretation of the results by reducing dimensionality. It also enables a traditional univariate approach to the comparison, which benefits from reductions to both within-- and between--participant variance due to averaging. The distribution of average time per PWI reference is visualized in @fig-pwi-hist.

```{r fig-pwi-hist}
#| fig-cap: Histogram of Composite PWI Measure in Recall
#| fig-height: 3.5
#| warning: false
#| message: false

# # calculate bin widths based on theory... don't think any work well
# x <- recall_h2b_agg$pwi_time_per_ref_agg
# bin_width_fd <- 2 * IQR(x, na.rm = TRUE) / (length(x[!is.na(x)])^(1/3))
# bin_width_sturges <- (max(x) - min(x)) / (log2(length(x)) + 1)
# bin_width_scott <- 3.5 * sd(x) / (length(x)^(1/3))

recall_h2b_agg |>
  ggplot(aes(x = pwi_time_per_ref_agg)) +
  geom_histogram(
    fill = I(base_color),
    alpha = 0.75,
    color = outline_color,
    binwidth = 0.5
  ) +
  labs(
    title = "Average Time per PWI Reference in Recall",
    subtitle = "Reliance is strongly right-skewed with high variability and long tails",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"),
  )
```

```{r h2b_tests}
st_rely_all <- shapiro.test(recall_h2b_agg$pwi_time_per_ref_agg)
st_rely_treat <- by(recall_h2b_agg$pwi_time_per_ref_agg, recall_h2b_agg$treatment, shapiro.test)

lt_rely_treat <- car::leveneTest(pwi_time_per_ref_agg ~ treatment, data = recall_h2b_agg)
lt_pv <- lt_rely_treat$`Pr(>F)`[1]
```

From the descriptive statistics and visualization, this data does not appear normal. To confirm, the Shapiro-Wilk test was administered for the overall data (`r fmt_p_val(st_rely_all$p.value)`) and within each treatment group. The results show that neither PWI (`r fmt_p_val(st_rely_treat$PWI$p.value)`) nor PAR (`r fmt_p_val(st_rely_treat$PAR$p.value)`) are normally distributed, but AR is (`r fmt_p_val(st_rely_treat$AR$p.value)`), and MR shows marginal evidence (`r fmt_p_val(st_rely_treat$MR$p.value)`) of non-normality. Visual inspection of the Q-Q plots for each treatment, provided in @fig-qq-h2b, supports these findings. Levene's test (`r fmt_p_val(lt_pv)`) confirms equal variance by treatment.

```{r fig-qq-h2b}
#| fig-cap: Q-Q Plots of PWI Reliance by Treatment
#| fig-height: 3.5

p <- qq_plot_patchwork(recall_h2b_agg, x_var = "pwi_time_per_ref_agg", by_var = "treatment")

p + plot_annotation(
  title = "Quantile-Quantile Plots of PWI Treatment Data",
  subtitle = "subtitle..."
)
```

```{r drop-h2b-outliers}
outliers <- recall_h2b_agg |>
  group_by(treatment) |>
  mutate(
    z_rely = abs((pwi_time_per_ref_agg - mean(pwi_time_per_ref_agg, na.rm = T))
    / sd(pwi_time_per_ref_agg, na.rm = T)),
    z_count = abs((avg_pwi_count - mean(avg_pwi_count, na.rm = T))
    / sd(avg_pwi_count, na.rm = T)),
    z_time = abs((avg_pwi_time - mean(avg_pwi_time, na.rm = T))
    / sd(avg_pwi_time, na.rm = T))
  ) |>
  filter(z_rely > 2.5 | z_count > 2.5 | z_time > 2.5) |>
  # filter(z_rely > 3.0 | z_count > 3.0 | z_time > 3.0) |>
  ungroup()

recall_h2b_agg_out <- recall_h2b_agg |>
  anti_join(outliers, by = "participant")
```

Outliers were identified within each treatment group for all measures of task uncertainty. Data for `r as.character(english(nrow(outliers)))` participants with Z-scores of 2.5 or higher are summarized in @tbl-h2b-outliers. These participants were removed from the primary analysis to ensure the data represents typical participant behavior and to maintain the integrity of the comparative analysis.

```{r tbl-h2b-outliers}
#| tbl-cap: Outliers in Measures of Reliance
#| warning: false

c_range <- c(2.5, max(outliers |> select(starts_with("z_")), na.rm = T))
# Define the custom subset of the palette (e.g., first half of the red_material palette)
full_palette <- ggsci::pal_material("deep-orange")(10)
subset_palette <- full_palette[5:10]

h2b_out_tbl <- outliers |>
  select(participant, treatment, z_count, z_time, z_rely) |>
  gt() |>
  tab_header(
    title = "|Z-scores| for Reliance Metrics of Participant Outliers",
    subtitle = "Participants with |Z-score| >= 2.5 in reliance components"
  ) |>
  cols_label(
    participant = "Participant",
    treatment = "Treatment",
    z_count = "Count", # "Avg PWI Count / Task"
    z_time = "Time", # "Avg Total PWI Time / Task"
    z_rely = "Reliance" # "Avg Time / PWI Ref"
  ) |>
  fmt_number(
    columns = c(z_rely, z_count, z_time),
    decimals = 2
  ) |>
  cols_align(align = "center") |>
  tab_options(
    table.font.size = "smaller",
    # table.width = pct(90),
    column_labels.font.weight = "bold",
    heading.title.font.weight = "bolder",
    heading.subtitle.font.size = px(13)
  )
# data_color(
#   columns = starts_with("z_"),
#   method = "numeric",
#   domain = c_range,
#   palette = subset_palette,
#   # reverse = T,
#   na_color = "transparent"
# )

h2b_out_tbl

# Calculate mean PWI time per reference with and without outliers
overall_mean_with_outliers <- mean(recall_h2b_agg$pwi_time_per_ref_agg, na.rm = TRUE)
overall_mean_without_outliers <- mean(recall_h2b_agg_out$pwi_time_per_ref_agg, na.rm = TRUE)
```

```{r h2b-test-out}
# does removing outliers change the normality or levene test results?

st_rely_all_out <- shapiro.test(recall_h2b_agg_out$pwi_time_per_ref_agg)
st_rely_treat_out <- by(recall_h2b_agg_out$pwi_time_per_ref_agg, recall_h2b_agg_out$treatment, shapiro.test)

lt_rely_treat_out <- car::leveneTest(pwi_time_per_ref_agg ~ treatment, data = recall_h2b_agg_out)
lt_pv_out <- lt_rely_treat_out$`Pr(>F)`[1]

h2b_group_sizes <- summary(recall_h2b_agg_out$treatment)
h2b_gs_txt <- glue::glue_collapse(glue::glue("{h2b_group_sizes} {names(h2b_group_sizes)}"), ", ", last = ", and ")

```

```{r h2b-test-out-check, eval=FALSE}
# Store the results in a data frame
h2b_compare_out <- data.frame(
  Test = c(
    "Levene",
    "SW: All",
    "SW: PWI",
    "SW: PAR",
    "SW: AR",
    "SW: MR"
  ),
  p_all = c(
    fmt_p_val(lt_pv),
    fmt_p_val(st_rely_all$p.value),
    fmt_p_val(st_rely_treat$PWI$p.value),
    fmt_p_val(st_rely_treat$PAR$p.value),
    fmt_p_val(st_rely_treat$AR$p.value),
    fmt_p_val(st_rely_treat$MR$p.value)
  ),
  p_out = c(
    fmt_p_val(lt_pv_out),
    fmt_p_val(st_rely_all_out$p.value),
    fmt_p_val(st_rely_treat_out$PWI$p.value),
    fmt_p_val(st_rely_treat_out$PAR$p.value),
    fmt_p_val(st_rely_treat_out$AR$p.value),
    fmt_p_val(st_rely_treat_out$MR$p.value)
  )
)

h2b_compare_out

# as of 6/10 6:40pm...only change in significance at 0.05: Levene test
#      Test     p_all     p_out
# 1  Levene  p = 0.95  p = 0.01 becomes significant, variance differs w/out outliers
# 2 SW: All p < 0.001 p < 0.001 n/c, not normal
# 3 SW: PWI p < 0.001 p < 0.001 n/c, not normal
# 4 SW: PAR p < 0.001 p = 0.004 n/c, not normal in either case
# 5  SW: AR  p = 0.16  p = 0.12 n/c, normal
# 6  SW: MR  p = 0.05  p = 0.05 n/c, borderline significance, no MR obs dropped
```

```{r h2b-comp-out-overall, eval=FALSE}
# does removing the outliers significantly change the mean of the overall data?

# Wilcoxon signed-rank test
overall_wilcoxon <- wilcox.test(
  recall_h2b_agg$pwi_time_per_ref_agg,
  recall_h2b_agg_out$pwi_time_per_ref_agg,
  exact = FALSE
)

print(overall_wilcoxon)

# p-value = 0.5391 n/c, no significant difference in overall data
```

```{r h2b-comp-out-treat, eval=FALSE}
# does removing the outliers significantly change the mean of any treatment group?

# Function to compare means within each treatment group using Wilcoxon rank-sum test (unpaired)
compare_by_treatment <- function(treatment) {
  with_outliers <- recall_h2b_agg |> filter(treatment == !!treatment)
  without_outliers <- recall_h2b_agg_out |> filter(treatment == !!treatment)

  mean_with_outliers <- mean(with_outliers$pwi_time_per_ref_agg, na.rm = TRUE)
  mean_without_outliers <- mean(without_outliers$pwi_time_per_ref_agg, na.rm = TRUE)

  test_result <- wilcox.test(
    with_outliers$pwi_time_per_ref_agg, without_outliers$pwi_time_per_ref_agg,
    exact = FALSE
  )

  list(
    treatment = treatment,
    mean_with_outliers = mean_with_outliers,
    mean_without_outliers = mean_without_outliers,
    p_value = test_result$p.value,
    test_result = test_result
  )
}

# Apply the comparison function to each treatment group
comparison_results <- lapply(TREATMENTS, compare_by_treatment)

# Print results
for (result in comparison_results) {
  cat("Treatment:", result$treatment, "\n")
  cat("  Mean with outliers:", result$mean_with_outliers, "\n")
  cat("  Mean without outliers:", result$mean_without_outliers, "\n")
  cat("  p-value:", result$p_value, "\n")
  print(result$test_result)
  cat("\n")
}

# all p-values > 0.493 (PAR) n/c, not significantly difference within any treatment
```

Several checks were then repeated to assess the impact of outlier removal. The Shapiro-Wilk test showed no change in significance for data overall or by treatment group. While Levene's test found significant difference in variance between groups (`r fmt_p_val(lt_pv_out)`) after the change, Kruskall-Wallis does not assume otherwise. A Wilcoxon signed-rank test was applied to compare overall and group means with and without outliers. No significant differences were found in average PWI time per reference in the the full data set or by treatment group (p > 0.40 for all cases). It is determined that, despite a reduction in the overall mean value from `r round(overall_mean_with_outliers, 2)` to `r round(overall_mean_without_outliers, 2)`, the character of the underlying data is not significantly changed by the outlier removal. Given that most treatments are not normally distributed, and variance differs between groups, a non-parametric approach is chosen for testing $H_{2b}$. The total number of observations (n = `r nrow(recall_h2b_agg_out)`) and group sizes (`r h2b_gs_txt`) remain sufficient for these methods, though statistical power is reduced.

#### Analysis

```{r test-pwi-power}
k <- length(unique(recall_h2b_agg_out$treatment))
n <- round(nrow(recall_h2b_agg_out) / k)
epsilon_sq <- 0.17

# Estimate post-hoc power
pwi_power <- posthoc_power_analysis(
  n, k, epsilon_sq,
  sig_level = 0.05, num_simulations = 1000
)
```

Per @fig-h2b-out-kw, despite a nominal p-value (0.05) the Kruskal-Wallis test fails to identify significant between-group differences for the average PWI reference duration. A small effect size $\hat{\epsilon}^{2}_{rank}=0.17$ with broad $CI_{95\%}\:[0.07,1.0]$, along with a post-hoc power simulation (`r round(pwi_power * 100, 1)`%) all support this result.

```{r fig-h2b-out-kw}
#| fig-cap: $H_{2b}$ Test - Average Time per PWI Reference by Treatment
#| fig-height: 4
#| warning: false

# since ggstatsplot computes confidence intervals for the effect sizes
# using bootstrapping, it is important to set a seed for reproducibility
set.seed(42)

pwi_rely_plot <- ggstatsplot::ggbetweenstats(
  data = recall_h2b_agg_out,
  x = treatment,
  y = pwi_time_per_ref_agg,
  type = "nonparametric",
  xlab = "",
  ylab = "",
  title = "PWI Reliance by Treatment: Kruskal-Wallis non-parametric rank sum test",
  # results.subtitle = FALSE,
  subtitle = "No Significant Differences Identified between Groups using Dunn's Test, Holm-adjusted post-hoc",
  digits = 3,
  # centrality.label.args = list(point.padding = 30),
  ggtheme = theme_minimal(base_size = 10) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      axis.text.y.right = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.text.x = element_text(color = "black")
      # legend.position = "none",
      # axis.text.y = element_blank(),
      # axis.ticks.y = element_blank(),
    ),
  # remove side label
  ggplot.component = list(
    ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = NULL))
  )
)

pwi_rely_plot

# list(chi["Kruskal-Wallis"]^2 * "(" * 3 * ")" == "7.8097", italic(p) ==
#     "0.0501", widehat(epsilon)["ordinal"]^2 == "0.1662", CI["95%"] ~
#     "[" * "0.0730", "1.0000" * "]", italic("n")["obs"] == "48")
```

Given the marginal p-value and high variance data, bootstrap analysis was again performed to provide complementary insights. As before, five replications of 10,000 resamples were conducted for all pairs using mean, median, and the Hodges-Lehmann estimator. Significant results are summarized in @tbl-boot-pwi. As with the Kurskal-Wallis test, no significant differences were identified using median, but the mean and Hodges-Lehmann estimator both identified 95% CI based differences in PWI-AR, PWI-MR, and PAR-AR pairs. A significant difference in mean was also indicated for PAR-MR.

```{r bootstrap-pwi, eval=FALSE}
stat_funcs <- list(
  Mean = mean,
  Median = median,
  HL_Est = hl_estimator
)

h2b_pwi_boot_res <- run_bootstrap_tests(
  recall_h2b_agg_out,
  column = "pwi_time_per_ref_agg",
  stat_funcs = stat_funcs
)

h2b_pwi_boot_res
save(h2b_pwi_boot_res, file = here("rdata", "h2b_pwi_boot_res.RData"))
```

```{r tbl-boot-pwi}
#| tbl-cap: $H_{2b}$ Bootstrap Analysis of Pairwise PWI Differences

load(here("rdata", "h2b_pwi_boot_res.RData"))
title <- "Significant Reliance Differences in Centrality: Bootstrapped Estimates"
subtitle <- "Five Replications of 10,000 Resamples for Mean, Median, and the Hodges-Lehmann Estimator"
create_boot_summary_table(h2b_pwi_boot_res, title = title, subtitle = subtitle)
```

On average, AR and MR reduce reliance by 0.686 and 0.612 seconds per reference, respectively, relative to the PWI instructional method. This finding is statistically and practically meaningful, given that the overall average reference duration is `r round(mean(recall_h2b_agg_out$pwi_time_per_ref_agg), 2)` seconds. In summary, the bootstrap analysis finds AR and MR are equivalent to one another but reduce reliance more than than PAR and PWI, which are themselves equivalent. The negligible bias values and low standard error reported across all replicates once again reinforce the reliability of these findings.

#### $H_{2b}$ Result

The marginal p-value returned by the Kruskall-Wallis comparison of median reliance, together with consistent and robust estimates of the differences provided by bootstrap analysis give sufficient evidence to accept $H_{2b}$: reliance does vary with instructional method. However, the conflicting results of pairwise comparisons must be interpreted with caution. Like other results in this study, they suggest that AR/MR are superior to PWI/PAR but the limited data and high variation make this somewhat inconclusive. Ultimately, it is determined that statistically significant differences in reliance exist between treatments, but specific pair-wise differences are not clear enough to declare.





::: content-hidden
TODO: recall add. analysis

### Additional Analyses

- what treatment group had the highest OEE during recall?

#### Efficiency vs. Quality

#### PWI Reference Frequency and Length

#### Error Types and Task Completion Rate

#### Ongoing Learning
:::

## $H_3$: Retention Phase Analysis

### Descriptive Statistics

### $H_{3a}$: Change in OEE

::: content-hidden
TODO: retention add. analysis

### Additional Analyses

#### Absolute and Percentage Change in OEE

#### Primary Driver of OEE Change
:::

## Workload, Usability, and User Experience

### NASA Task Load Index Results

### System Usability Scale Results

### Qualitative Feedback Analysis

::: content-hidden
TODO: exploratory analysis

## Exploratory Analyses

### Relationship between TLX, SUS, and Performance

- first task performance in learning as a measure of initial comprehension, intuitiveness of each treatment, effectiveness of instructional design, etc.; may correlate with SUS / TLX

### Impact of Participant Demographics on Performance

### Within-Group Performance Variance

### TLX Components and Performance

### Interactions between Workload, Usability, and Treatment
:::

## Challenges and Unexpected Observations {#sec-res-challenges}

Discussion of any challenges encountered during the conduct of the protocol  Modifications made to the protocol in response to these challenges

demoing AR/MR: The challenge of this choreography was compounded by significant latency in streamed video, which lagged the live action by approximately 2 seconds. With practice

> Technical Issues with Equipment: Discussion of any technical challenges encountered with the HL2 or other devices, and how these were addressed.
>
> Participant Scheduling and No-Shows: Challenges related to participant scheduling, no-shows, or last-minute cancellations, and the strategies developed to mitigate these issues.
>
> Data Collection Consistency: Any issues encountered in maintaining consistency in data collection across different trials and days, and how these were resolved.

limitations

- participant behavior
    -   few participants assigned HMDMR took advantage of the affordance
- analysis
    -   does not account for interactions between affordances
    - cross validation or similar models not  employeed
    - imbalanced data in learning

noticed improvising - discourage as occurs

sleeves, long hair problems for HL2

height and "down-lookers"; tracking dropouts;

battery life and performance; brightness and accuracy;

complexity of MR, less reliable tracking, crashes

## Summary of Key Findings
